---
title: "Thesis Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
execute:
  warning: false
  message: false
format:
  pdf:
    output-file: "results.pdf"
    keep-tex: false
    fig-pos: "htbp"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
# --- set root BEFORE any relative path is used ---
proj_dir <- "/workspaces/thesis"
setwd(proj_dir)

data_dir <- "/workspaces/thesis/data"
raw_dir <- file.path(data_dir, "data_raw")
clean_dir <- file.path(data_dir, "cleaned_data")
if (!dir.exists(clean_dir)) dir.create(clean_dir, recursive = TRUE)

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)
library(patchwork)
library(ggraph)
library(ggrepel)
library(tidygraph)
library(maps)
library(countrycode)
library(fixest)
library(knitr)
library(parallel)
library(cowplot)

source(file.path(proj_dir, "thesis_helpers_refactored.r"))

set.seed(1861)
options(readr.show_col_types = FALSE)

EPS_EDGE <- 1e-5

out_dir <- file.path(data_dir, "output") # was file.path(proj_dir, "output")
out_rerun <- file.path(data_dir, "output_rerun") # gitignored scratch
dir.create(out_rerun, showWarnings = FALSE)

read_dir <- if (Sys.getenv("USE_RERUN") == "true") out_rerun else out_dir

fig_dir <- file.path(proj_dir, "assets/figures")
tab_dir <- file.path(proj_dir, "assets/tables")
for (d in c(fig_dir, tab_dir)) if (!dir.exists(d)) dir.create(d, recursive = TRUE)

add_note <- function(k, note = "SAMPLE NOTE") {
  kableExtra::footnote(
    k,
    general = note, general_title = "Notes:",
    threeparttable = TRUE, footnote_as_chunk = TRUE, escape = FALSE
  )
}
```

```{r}
#| label: load-best
#| include: false
best <- readRDS(file.path(read_dir, "best_per_panel.rds"))
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
diag(W_q) <- 0
W_q[abs(W_q) <= EPS_EDGE] <- 0
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)
diag(W_e) <- 0
W_e[abs(W_e) <= EPS_EDGE] <- 0

sync_eu <- readRDS(file.path(clean_dir, "sync_eu.rds"))
country_labels <- sort(sync_eu)
rownames(W_q) <- colnames(W_q) <- country_labels
rownames(W_e) <- colnames(W_e) <- country_labels
```

```{r}
#| label: load-best-lagged
#| include: false
# The fully-lagged BDF specification reuses the SAME network recovered from the
# contemporaneous De Paula GMM; no separate lagged recovery is loaded.
# Aliases keep downstream "_lag" references pointing at the standard recovered
# network: only the BDF equation is lagged, never the network.
W_q_lag <- W_q
W_e_lag <- W_e
```

```{r}
#| label: dyadic-build-and-exogenous-build
#| echo: false
#| include: false
N <- length(country_labels)
cl_iso3 <- countrycode::countrycode(
  country_labels,
  origin = "country.name", destination = "iso3c", warn = FALSE
)
names(cl_iso3) <- country_labels
iso_to_label <- setNames(names(cl_iso3), cl_iso3)

data("world.cities", package = "maps")
capitals <- world.cities |>
  dplyr::filter(capital == 1) |>
  dplyr::mutate(iso3 = countrycode::countrycode(
    country.etc,
    origin = "country.name", destination = "iso3c", warn = FALSE
  )) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::group_by(iso3) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::transmute(country, iso3, lat, lon = long) |>
  dplyr::arrange(match(country, country_labels))

D_geo <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  for (j in seq_len(N)) {
    D_geo[i, j] <- haversine_km(
      capitals$lat[i], capitals$lon[i],
      capitals$lat[j], capitals$lon[j]
    )
  }
}
log_dist <- log(D_geo + 1)
diag(log_dist) <- 0

hofstede_path <- file.path(raw_dir, "6-dimensions-for-website-2015-08-16(1).csv")
hof_raw <- tryCatch(
  readr::read_csv2(hofstede_path, show_col_types = FALSE) |> janitor::clean_names(),
  error = function(e) readr::read_csv(hofstede_path, show_col_types = FALSE) |> janitor::clean_names()
) |>
  dplyr::mutate(dplyr::across(dplyr::everything(), \(x) ifelse(as.character(x) == "#NULL!", NA, x)))

candidates <- c("country", "ctr", "iso", "iso3")
country_col <- candidates[candidates %in% names(hof_raw)][1]
to_iso3 <- function(x) {
  x <- as.character(x)
  out <- suppressWarnings(countrycode::countrycode(x, "country.name", "iso3c", warn = FALSE))
  if (mean(is.na(out)) > 0.5) {
    out <- suppressWarnings(countrycode::countrycode(x, "iso2c", "iso3c", warn = FALSE))
  }
  if (mean(is.na(out)) > 0.5) out <- toupper(x)
  out
}
hof <- hof_raw |>
  dplyr::mutate(iso3 = to_iso3(.data[[country_col]])) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::transmute(iso3,
    pdi = suppressWarnings(as.numeric(pdi)), idv = suppressWarnings(as.numeric(idv)),
    mas = suppressWarnings(as.numeric(mas)), uai = suppressWarnings(as.numeric(uai))
  ) |>
  dplyr::filter(!is.na(pdi) & !is.na(idv) & !is.na(mas) & !is.na(uai)) |>
  dplyr::distinct(iso3, .keep_all = TRUE)
if (!"CYP" %in% hof$iso3 && "GRC" %in% hof$iso3 && "CYP" %in% cl_iso3) {
  hof <- dplyr::bind_rows(hof, hof |> dplyr::filter(iso3 == "GRC") |> dplyr::mutate(iso3 = "CYP"))
}
missing_hof <- setdiff(cl_iso3, hof$iso3)
if (length(missing_hof) > 0) {
  hof_mean <- hof |> dplyr::summarise(dplyr::across(c(pdi, idv, mas, uai), \(x) mean(x, na.rm = TRUE)))
  for (iso in missing_hof) {
    hof <- dplyr::bind_rows(hof, tibble::tibble(
      iso3 = iso, pdi = hof_mean$pdi, idv = hof_mean$idv,
      mas = hof_mean$mas, uai = hof_mean$uai
    ))
  }
}
hof <- hof |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::filter(!is.na(country) & country %in% country_labels) |>
  dplyr::arrange(match(country, country_labels))
H_mat <- as.matrix(hof[, c("pdi", "idv", "mas", "uai")])
rownames(H_mat) <- hof$country
D_cul <- as.matrix(dist(H_mat, method = "euclidean"))[country_labels, country_labels]

D_gdppc <- build_dist_wdi(
  file.path(raw_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000")
)

envtax_raw <- readr::read_csv(file.path(raw_dir, "estat_env_ac_taxind2_en.csv.gz"),
  show_col_types = FALSE
)
envtax <- envtax_raw |>
  dplyr::filter(tax == "ENV", unit == "MIO_EUR", nace_r2 == "TOTAL", TIME_PERIOD %in% 1996:2000) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(iso3 = iso2_to_iso3(geo), country = iso_to_label[iso3]) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))
if (nrow(envtax) < N) {
  envtax <- dplyr::bind_rows(
    envtax,
    data.frame(
      geo = NA, envtax = mean(envtax$envtax, na.rm = TRUE), iso3 = NA,
      country = setdiff(country_labels, envtax$country), stringsAsFactors = FALSE
    )
  )
}
envtax <- envtax |> dplyr::arrange(match(country, country_labels))

lab_raw <- readr::read_csv(file.path(raw_dir, "API_SL.TLF.TOTL.IN_DS2_en_csv_v2_196.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    pop = rowMeans(dplyr::across(
      dplyr::all_of(c("x1996", "x1997", "x1998", "x1999", "x2000")),
      \(z) suppressWarnings(as.numeric(z))
    ), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(pop), is.finite(pop)) |>
  dplyr::mutate(country = iso_to_label[iso3])

envtax_pw <- envtax |>
  dplyr::select(country, envtax) |>
  dplyr::filter(!is.na(envtax)) |>
  dplyr::inner_join(lab_raw, by = "country") |>
  dplyr::mutate(envtax_per_worker = envtax / pop) |>
  dplyr::arrange(match(country, country_labels))
if (nrow(envtax_pw) < N) {
  envtax_pw <- dplyr::bind_rows(
    envtax_pw,
    data.frame(
      country = setdiff(country_labels, envtax_pw$country), envtax = NA, iso3 = NA, pop = NA,
      envtax_per_worker = mean(envtax_pw$envtax_per_worker, na.rm = TRUE), stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}
pw_vec <- setNames(log(envtax_pw$envtax_per_worker + 1e-12), envtax_pw$country)
D_envtax_pw <- abs(outer(pw_vec, pw_vec, FUN = "-"))
dimnames(D_envtax_pw) <- list(country_labels, country_labels)

fdi_raw <- readr::read_csv(file.path(raw_dir, "estat_bop_fdi6_pos_en.csv.gz"), show_col_types = FALSE)
fdi <- fdi_raw |>
  dplyr::filter(
    counterp == "IMM", nace_r2 == "FDI", stk_flow == "ASS", entity == "TOTAL",
    fdi_item == "DI__D__F", currency == "MIO_EUR", TIME_PERIOD %in% 2013:2020
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi_value = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(geo_iso3 = iso2_to_iso3(geo), partner_iso3 = iso2_to_iso3(partner)) |>
  dplyr::filter(geo_iso3 %in% cl_iso3, partner_iso3 %in% cl_iso3) |>
  dplyr::mutate(sender = iso_to_label[geo_iso3], receiver = iso_to_label[partner_iso3])
F_fdi <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(fdi))) F_fdi[fdi$sender[k], fdi$receiver[k]] <- pmax(fdi$fdi_value[k], 0)
log_fdi <- log1p(F_fdi)
diag(log_fdi) <- 0

baci_path <- file.path(raw_dir, "BACI_HS92_Y1996_V202601.csv")
ccodes_path <- file.path(raw_dir, "country_codes_V202601.csv")
TOP_N <- 3
baci <- readr::read_csv(baci_path, show_col_types = FALSE)
ccodes <- readr::read_csv(ccodes_path, show_col_types = FALSE)
iso_map <- ccodes |>
  dplyr::select(country_code, country_iso3) |>
  dplyr::filter(country_iso3 %in% cl_iso3)
trade <- baci |>
  dplyr::select(i, j, v) |>
  dplyr::inner_join(iso_map, by = c("i" = "country_code")) |>
  dplyr::rename(exporter_iso = country_iso3) |>
  dplyr::inner_join(iso_map, by = c("j" = "country_code")) |>
  dplyr::rename(importer_iso = country_iso3) |>
  dplyr::mutate(exporter = iso_to_label[exporter_iso], importer = iso_to_label[importer_iso]) |>
  dplyr::group_by(exporter, importer) |>
  dplyr::summarise(value = sum(v, na.rm = TRUE), .groups = "drop")
T_trade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(trade))) T_trade[trade$exporter[k], trade$importer[k]] <- trade$value[k]

log_trade <- log1p(T_trade)
diag(log_trade) <- 0

W_geo <- top_n_network(D_geo, top_n = TOP_N, similarity = FALSE)
W_trade <- top_n_network(T_trade, top_n = TOP_N, similarity = TRUE)
W_cul <- top_n_network(D_cul, top_n = TOP_N, similarity = FALSE)
saveRDS(list(geo = W_geo, trade = W_trade, culture = W_cul), file.path(out_rerun, "exogenous_networks.rds"))

W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs] <- 1 / length(nbrs)
}
W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs] <- 1 / length(nbrs)
}
```

# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Descriptive Statistics

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
f9001 <- readRDS(file.path(clean_dir, "f9001.rds"))
f14001 <- readRDS(file.path(clean_dir, "f14001.rds"))

desc_years <- c(1993, 1999, 2017)
desc_tab <- function(df, countries, years) {
  df |>
    dplyr::filter(country %in% countries, year %in% years) |>
    dplyr::group_by(year) |>
    dplyr::summarise(
      total_cert = sum(n_cert, na.rm = TRUE), mean_cert = mean(n_cert, na.rm = TRUE),
      sd_cert = sd(n_cert, na.rm = TRUE), min_cert = min(n_cert, na.rm = TRUE),
      max_cert = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries = dplyr::n(), .groups = "drop"
    )
}
d9 <- desc_tab(f9001, country_labels, desc_years)
d14 <- desc_tab(f14001, country_labels, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

desc_stats_note <- "This table summarises the distribution of certifications across the 27 EU countries in the three selected years. The first row reports the number of countries with non-missing data in each year, which is the basis for all subsequent statistics. The second row reports the total number of certifications across all countries, while the third and fourth rows report the mean and standard deviation of certifications per country, respectively. The fifth row reports the minimum and maximum number of certifications across countries. Finally, the sixth row reports the mean certification intensity, defined as the number of certifications per thousand dollars of GDP per capita, averaged across countries. Environment 1993 is empty as its adoption started in 1997. Sources: ISO Survey, World Bank."

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Descriptive statistics}",
  "\\label{tab:desc_stats}",
  "\\begin{tabular}{lcccccc}\\hline\\hline",
  "& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\",
  "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
  "& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\\hline",
  sprintf(
    "Countries Number & %d & %d & %d & & %d & %d \\\\",
    d9$n_countries[1], d9$n_countries[2], d9$n_countries[3], d14$n_countries[1], d14$n_countries[2]
  ),
  sprintf(
    "Total & %s & %s & %s & & %s & %s \\\\",
    f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]), f0(d14$total_cert[1]), f0(d14$total_cert[2])
  ),
  sprintf(
    "Mean & %s & %s & %s & & %s & %s \\\\",
    f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]), f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
  ),
  sprintf(
    "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\",
    f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]), f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
  ),
  sprintf(
    "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\",
    f0(d9$min_cert[1]), f0(d9$max_cert[1]), f0(d9$min_cert[2]), f0(d9$max_cert[2]),
    f0(d9$min_cert[3]), f0(d9$max_cert[3]), f0(d14$min_cert[1]), f0(d14$max_cert[1]),
    f0(d14$min_cert[2]), f0(d14$max_cert[2])
  ),
  sprintf(
    "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\",
    f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
    f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
  ),
  "\\hline\\hline",
  sprintf("\\multicolumn{7}{p{0.92\\textwidth}}{\\footnotesize \\textit{Notes:} %s} \\\\", desc_stats_note),
  "\\end{tabular}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "desc_stats.tex"))
cat(tex, sep = "\n")
```

## Network Recovery and Dyadic Analysis

\subsection*{Methodology}

De Paula, Rasul \& Souza (2025) extend the peer-effects literature (Manski 1993; Bramoullé et al., 2009) to the case in which the network of interactions is unknown. Their structural model
\begin{equation}
y_{it} = \rho_0 \sum_{j=1}^{N} W_{0,ij}\, y_{jt} + \beta_0 x_{it} + \gamma_0 \sum_{j=1}^{N} W_{0,ij}\, x_{jt} + \alpha_i + \alpha_t + \varepsilon_{it}
\label{eq:structural}
\end{equation}
admits the compact form $y_t = \rho_0 W_0 y_t + \beta_0 x_t + \gamma_0 W_0 x_t + \varepsilon_t$, which rearranges into the reduced form
\begin{equation}
y_t = \Pi_0 x_t + \nu_t, \qquad \Pi_0 = (I - \rho_0 W_0)^{-1}(\beta_0 I + \gamma_0 W_0).
\label{eq:reduced}
\end{equation}
If $W_0$ were known, equation~\eqref{eq:reduced} would directly identify the structural parameters $(\rho_0,\beta_0,\gamma_0)$ from a reduced-form regression of outcomes on covariates (Bramoullé et al., 2009). The contribution of De Paula, Rasul \& Souza is to show that, when $W_0$ is unknown, $\Pi_0$ still contains enough information to recover both the network and the parameters jointly.

The intuition is the one that underlies any diffusion process. If a contagious disease spreads through a population, the people who fall ill at the same time tend to be the ones who interact: knowing who got sick when, and what their characteristics are, one can work backwards to who must have been in contact with whom. The same logic applies here. If four countries adopt a standard in synchronous patterns that cannot be explained by their own characteristics alone, the synchrony itself is evidence of a link between them. Cross-country covariation in $y$, conditional on $x$, identifies who is connected to whom, and the strength of the covariation, suitably disentangled, identifies how strong the link is.

In principle, $\Pi_0$ is identified as the linear projection of $y$ on $x$ and could be recovered by ordinary least squares whenever the panel is long enough relative to the number of parameters. In typical short panels this is infeasible. De Paula, Rasul \& Souza recover the network $\hat W$ jointly with the structural parameters $(\hat\rho, \hat\beta, \hat\gamma)$ via an adaptive elastic-net GMM (Caner \& Zhang 2014) defined directly on the moment conditions implied by the structural model. Sparsity in $\hat W$ regularises the high-dimensional problem and restores consistency in short panels, estimating the network from observed spillovers without requiring a priori commitment to any specific weighting scheme.

```{r}
#| label: network-stats-table
#| echo: false
#| results: asis
s_q <- net_stats(W_q)
s_e <- net_stats(W_e)
B_q <- (W_q != 0) * 1
B_e <- (W_e != 0) * 1
common <- sum(B_q & B_e)
only_q <- sum(B_q & !B_e)
only_e <- sum(!B_q & B_e)
f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")
rho_q <- best[["9001_eu"]]$rn$unpenalisedgmm$rho
rho_e <- best[["14001_eu"]]$rn$unpenalisedgmm$rho
lam_q <- best[["9001_eu"]]$lambda
lam_e <- best[["14001_eu"]]$lambda
bic_q <- best[["9001_eu"]]$bic
bic_e <- best[["14001_eu"]]$bic

network_stats_note <- "This table reports summary statistics of the two recovered networks, selected by BIC. The first row reports the number of edges in each network, while the second and third rows report how many edges are shared by both networks and how many are unique to each network, respectively. A shared link is intended when there is a link between the same two nodes in the same direction in both networks. The fourth and fifth rows report the clustering coefficient (share of closed triplets) and reciprocity (bidirectional edges) of each network, respectively. The sixth to ninth rows report the mean and standard deviation of out-degrees and in-degrees across nodes (countries) for each network. Finally, the tenth to twelfth rows report the GMM estimates of the structural parameters $\\hat{\\rho}$ (endogenous peer effects) and $\\hat{\\lambda}$ (penalties for the Adaptive Elastic Net) and the BIC values for each network. Sources: Elaboration on ISO Survey Data, the World Bank Data and the Worldwide Governance Indicators. The network estimation relies on GDP per capita (WDI: NY.GDP.PCAP.KD), Trade (\\% of GDP; WDI: NE.TRD.GNFS.ZS), Rule of Law estimate (WGI: RL.EST)*, Total Natural Resources Rents (\\% of GDP; WDI: NY.GDP.TOTL.RT.ZS), and Share of Renewables in Total Final Energy Consumption (\\% ; WDI: EG.FEC.RNEW.ZS). All the covariates were log linearised (log (1+x))."

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{BIC-selected networks}\\label{tab:network_eu}",
  "\\begin{tabular}{lcc}\\hline\\hline",
  "& Quality & Environment \\\\\\hline",
  sprintf("Number of edges & %d & %d \\\\", s_q$edges, s_e$edges),
  sprintf("Edges in both networks & %d & %d \\\\", common, common),
  #sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\", only_q),
  #sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\", only_e),
  sprintf("Clustering & %s & %s \\\\", f3(s_q$clustering), f3(s_e$clustering)),
  sprintf("Reciprocated edges & %s & %s \\\\", p1(s_q$reciprocity), p1(s_e$reciprocity)),
  "Degree distribution across nodes (countries) & & \\\\",
  sprintf("\\quad Out-degree & %s & %s \\\\", dg(s_q$out_mean, s_q$out_sd), dg(s_e$out_mean, s_e$out_sd)),
  sprintf("\\quad In-degree & %s & %s \\\\", dg(s_q$in_mean, s_q$in_sd), dg(s_e$in_mean, s_e$in_sd)),
  "\\hline",
  sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\",
    formatC(rho_q, format = "f", digits = 4),
    formatC(rho_e, format = "f", digits = 4)
  ),
  sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\",
    paste(formatC(lam_q, format = "f", digits = 3), collapse = ", "),
    paste(formatC(lam_e, format = "f", digits = 3), collapse = ", ")
  ),
  sprintf(
    "BIC & %s & %s \\\\",
    formatC(bic_q, format = "f", digits = 2),
    formatC(bic_e, format = "f", digits = 2)
  ),
  "\\hline\\hline",
  sprintf(
    "\\multicolumn{3}{p{0.95\\textwidth}}{\\footnotesize \\textit{Notes:} %s} \\\\",
    network_stats_note
  ),
  "\\end{tabular}\\end{table}"
)
writeLines(tex, file.path(tab_dir, "network_stats.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: fig-degree-hist
#| echo: false
#| fig-width: 9
#| fig-height: 3.5
#| fig-cap: "Unweighted degree distributions of recovered networks"
B_q_d <- (W_q != 0) * 1
diag(B_q_d) <- 0
B_e_d <- (W_e != 0) * 1
diag(B_e_d) <- 0
deg_df <- dplyr::bind_rows(
  data.frame(standard = "Quality (ISO 9001)", direction = "Out-degree", value = rowSums(B_q_d)),
  data.frame(standard = "Quality (ISO 9001)", direction = "In-degree", value = colSums(B_q_d)),
  data.frame(standard = "Environment (ISO 14001)", direction = "Out-degree", value = rowSums(B_e_d)),
  data.frame(standard = "Environment (ISO 14001)", direction = "In-degree", value = colSums(B_e_d))
) |>
  dplyr::mutate(
    standard  = factor(standard, levels = c("Quality (ISO 9001)", "Environment (ISO 14001)")),
    direction = factor(direction, levels = c("Out-degree", "In-degree"))
  )
p_deg <- ggplot(deg_df, aes(x = value, fill = standard)) +
  geom_histogram(position = "dodge", alpha = 0.55, colour = "white", linewidth = 0.3, bins = 12) +
  geom_rug(aes(colour = standard), sides = "b", length = unit(0.03, "npc"), alpha = 0.7) +
  facet_wrap(~direction, ncol = 2) +
  scale_fill_grey(start = 0.2, end = 0.65, name = NULL) +
  scale_colour_grey(start = 0.1, end = 0.55, name = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02))) +
  labs(x = "Degree", y = "Count", fill = NULL, colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"), panel.grid.minor = element_blank())
print(p_deg)
save_figure("degree_hist", width = 9, height = 3.5)
```

```{r}
#| label: jaccard-compute
#| include: false
B_q <- (W_q != 0) * 1
B_e <- (W_e != 0) * 1
diag(B_q) <- diag(B_e) <- 0
n_q <- sum(B_q)
n_e <- sum(B_e)
n_inter <- sum(B_q & B_e)
n_union <- sum(B_q | B_e)
jaccard <- n_inter / n_union
jaccard_val <- formatC(jaccard, format = "f", digits = 3)
```

### Network similarity

To assess how closely the two recovered networks $\hat W^Q$ and $\hat W^E$ overlap, we summarise their structural similarity through the **Jaccard index** (Jaccard 1901; Wasserman \& Faust 1994, Sec.\ 11.4):
\begin{equation}
J(\hat W^Q, \hat W^E) = \frac{|E_Q \cap E_E|}{|E_Q \cup E_E|},
\label{eq:jaccard}
\end{equation}
where $E_Q = \{(i, j) : |\hat w^Q_{ij}| > 10^{-5}\}$ and $E_E$ is defined analogously. The index ranges in $[0, 1]$, with $J = 0$ for disjoint networks and $J = 1$ for identical edge sets.

For the BIC-selected recovered networks, $|E_Q| = `r n_q`$, $|E_E| = `r n_e`$, $|E_Q \cap E_E| = `r n_inter`$, and $|E_Q \cup E_E| = `r n_union`$, yielding $J = `r jaccard_val`$.


```{r}
#| label: network-jaccard-matrices
#| echo: false
#| results: asis
jac <- function(A, B) {
  a <- as.numeric(A[row(A) != col(A)] != 0)
  b <- as.numeric(B[row(B) != col(B)] != 0)
  u <- sum(a | b)
  if (u == 0) {
    return(0)
  }
  sum(a & b) / u
}

nets <- list(
  "$\\hat{W}^Q$ (Recovered)" = W_q,
  "$\\hat{W}^E$ (Recovered)" = W_e,
  "Geography" = W_geo,
  "Trade" = W_trade,
  "Cultural dist." = W_cul,
  "Geo + Trade" = W_geotrade,
  "Geo + Culture" = W_geoculture
)

n_nets <- length(nets)
jac_mat <- matrix(NA_real_, n_nets, n_nets,
  dimnames = list(names(nets), names(nets))
)
for (i in seq_len(n_nets)) {
  for (j in seq_len(n_nets)) {
    jac_mat[i, j] <- jac(nets[[i]], nets[[j]])
  }
}

# Format: diagonal as "1.000", off-diagonal to 3 dp
jac_fmt <- matrix(sprintf("%.3f", jac_mat), n_nets, n_nets,
  dimnames = dimnames(jac_mat)
)
diag(jac_fmt) <- "1.000"

tab_jac_mat <- kbl(
  as.data.frame(jac_fmt),
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = paste(rep("c", n_nets + 1), collapse = ""),
  row.names = TRUE,
  caption = "Pairwise Jaccard similarity matrix of recovered and benchmark networks.",
  label = "jaccard_matrix"
) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 8
  ) |>
  add_note(paste(
  "$J(A,B) = |E_A \\\\cap E_B| / |E_A \\\\cup E_B|$ on binarised directed edge sets",
  "($w_{ij} \\\\neq 0$). Diagonal entries are 1 by definition. The recovered networks",
  "($\\\\hat W^Q$, $\\\\hat W^E$) are the BIC-selected networks. Benchmark networks",
  "follow Albuquerque et al. (2007)."
))

save_table(tab_jac_mat, "jaccard_matrix")
tab_jac_mat |> print()
```

### Network visualisation

```{r}
#| label: fig-networks-weighted-full
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Recovered networks"
make_full_weighted_graph <- function(W, labels) {
  g <- igraph::graph_from_adjacency_matrix(W, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- labels
  g
}
g_q_full <- make_full_weighted_graph(W_q, country_labels)
g_e_full <- make_full_weighted_graph(W_e, country_labels)
w_max <- max(c(abs(igraph::E(g_q_full)$weight), abs(igraph::E(g_e_full)$weight)), 1e-9)
layout_q <- igraph::layout_with_fr(g_q_full)
layout_e <- igraph::layout_with_fr(g_e_full)
rownames(layout_q) <- rownames(layout_e) <- country_labels
plot_net_gg <- function(g, lo, title, w_max) {
  lo_ordered <- lo[igraph::V(g)$name, , drop = FALSE]
  ggraph(g, layout = lo_ordered) +
    geom_edge_arc(aes(alpha = abs(weight), colour = abs(weight) > 0.9),
      width = 0.5,
      arrow = arrow(length = unit(2, "mm"), type = "closed", angle = 22),
      end_cap = circle(5, "mm"), start_cap = circle(5, "mm"), strength = 0.12
    ) +
    geom_node_point(aes(colour = igraph::degree(g) > 0), size = 4.5, show.legend = FALSE) +
    geom_node_text(aes(label = name),
      repel = TRUE, size = 2.6,
      bg.colour = "white", bg.r = 0.18, segment.colour = "grey55", segment.size = 0.25,
      segment.alpha = 0.7, max.overlaps = Inf, force = 3, force_pull = 0.5,
      point.padding = unit(0.6, "lines"), box.padding = unit(0.45, "lines"), min.segment.length = 0
    ) +
    scale_edge_alpha(range = c(0.35, 0.9), limits = c(0, w_max), guide = "none") +
    scale_edge_colour_manual(
      values = c(`TRUE` = "black", `FALSE` = "grey65"),
      labels = c(`TRUE` = "Strong influence", `FALSE` = "Weak influence"),
      guide = guide_legend(title = NULL, override.aes = list(width = 2, alpha = 0.7), position = "bottom")
    ) +
    scale_colour_manual(values = c(`TRUE` = "grey15", `FALSE` = "grey85")) +
    ggtitle(title) + theme_void(base_size = 11) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "bottom",
      plot.margin = margin(8, 8, 8, 8)
    )
}
p_qq <- plot_net_gg(g_q_full, layout_q, "Quality (own layout)", w_max)
p_eq <- plot_net_gg(g_e_full, layout_q, "Environment (Quality layout)", w_max)
p_qe <- plot_net_gg(g_q_full, layout_e, "Quality (Environment layout)", w_max)
p_ee <- plot_net_gg(g_e_full, layout_e, "Environment (own layout)", w_max)

panel <- (p_qq | p_eq) / (p_qe | p_ee)
print(panel)
save_figure("networks_weighted_full", width = 13, height = 13)
```

### Dyadic regression


```{r}
#| label: dyadic-fullperiod-robustness
#| echo: false
#| results: asis
dict_full <- c(
  d_econ = "Economic distance", d_envir = "Env. policy asymmetry", d_cult = "Cultural distance",
  d_geo = "Geographic distance", l_fdi = "Bilateral FDI", l_trade = "Bilateral Trade (log)",
  sender = "Sender FE ($\\alpha_i^{\\text{out}}$)", receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)
full_years_wdi <- paste0("x", 1993:2017)
full_years_estat <- 1993:2017
D_gdppc_fp <- build_dist_wdi(
  file.path(raw_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"), full_years_wdi
)
envtax_fp <- envtax_raw |>
  dplyr::filter(tax == "ENV", unit == "MIO_EUR", nace_r2 == "TOTAL", TIME_PERIOD %in% full_years_estat) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(iso3 = iso2_to_iso3(geo), country = iso_to_label[iso3]) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))
if (nrow(envtax_fp) < N) {
  envtax_fp <- dplyr::bind_rows(
    envtax_fp,
    data.frame(
      geo = NA, envtax = mean(envtax_fp$envtax, na.rm = TRUE), iso3 = NA,
      country = setdiff(country_labels, envtax_fp$country), stringsAsFactors = FALSE
    )
  )
}
envtax_fp <- envtax_fp |> dplyr::arrange(match(country, country_labels))
lab_raw_fp <- readr::read_csv(file.path(raw_dir, "API_SL.TLF.TOTL.IN_DS2_en_csv_v2_196.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    pop = rowMeans(dplyr::across(dplyr::any_of(full_years_wdi), \(z) suppressWarnings(as.numeric(z))), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(pop), is.finite(pop)) |>
  dplyr::mutate(country = iso_to_label[iso3])
envtax_pw_fp <- envtax_fp |>
  dplyr::select(country, envtax) |>
  dplyr::filter(!is.na(envtax)) |>
  dplyr::inner_join(lab_raw_fp, by = "country") |>
  dplyr::mutate(envtax_per_worker = envtax / pop) |>
  dplyr::arrange(match(country, country_labels))
if (nrow(envtax_pw_fp) < N) {
  envtax_pw_fp <- dplyr::bind_rows(
    envtax_pw_fp,
    data.frame(
      country = setdiff(country_labels, envtax_pw_fp$country), envtax = NA, iso3 = NA, pop = NA,
      envtax_per_worker = mean(envtax_pw_fp$envtax_per_worker, na.rm = TRUE), stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}
pw_vec_fp <- setNames(log(envtax_pw_fp$envtax_per_worker + 1e-12), envtax_pw_fp$country)
D_envtax_pw_fp <- abs(outer(pw_vec_fp, pw_vec_fp, FUN = "-"))
dimnames(D_envtax_pw_fp) <- list(country_labels, country_labels)

build_dyadic_dz_fp <- function(W_outcome) {
  expand.grid(sender = seq_len(N), receiver = seq_len(N), KEEP.OUT.ATTRS = FALSE) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc_fp[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = D_envtax_pw_fp[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))],
      l_trade  = log_trade[cbind(as.integer(sender), as.integer(receiver))]
    )
}
dz_q_fp <- build_dyadic_dz_fp(W_q)
dz_e_fp <- build_dyadic_dz_fp(W_e)
f_ols_fe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi + l_trade | sender + receiver
fit_ols_q_rec_fe <- feols(f_ols_fe, data = dz_q_fp, cluster = ~ sender + receiver)
fit_ols_e_rec_fe <- feols(f_ols_fe, data = dz_e_fp, cluster = ~ sender + receiver)
etable(list(fit_ols_q_rec_fe, fit_ols_e_rec_fe),
  depvar = FALSE,
  headers = list(" " = c("Quality", "Environment"), "Specification" = c("OLS", "OLS")),
  style.tex = style.tex("base", model.title = "", fixef.suffix = ":", tablefoot = FALSE),
  dict = dict_full, signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10), fitstat = ~ n + r2, tex = TRUE,
  title = "Dyadic regression of recovered network weights with sender and receiver fixed effects, full-period averages (1993--2017).",
  label = "tab:dyadic_recovered_fp_fe",
  notes = "*Notes* OLS estimates of $\\hat w_{ij}$ on dyadic covariates. Sample: 702 directed pairs, 27 EU countries, full-period (1993--2017) averages. Two-way clustered SE (sender $\\times$ receiver). Significance: $^{***}$ $p<0.01$, $^{**}$ $p<0.05$, $^{*}$ $p<0.10$.",
  file = file.path(tab_dir, "dyadic_recovered_fp_fe.tex")
)
```

## Application of Bramoullé, Djebbari & Fortin (2009)

```{r}
#| label: bdf-prereqs
#| echo: false
#| include: false
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")
id_to_country <- setNames(country_labels, seq_along(country_labels))
load_sbs_2017 <- function() {
  raw <- readxl::read_excel(file.path(raw_dir, "enterprises_eu.xlsx"),
    sheet = "Sheet 1", skip = 9, col_names = FALSE
  )
  tibble::tibble(country = raw[[1]], n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))) |>
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) |>
    dplyr::mutate(country = dplyr::case_match(country,
      "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
      .default = country
    )) |>
    dplyr::distinct(country, .keep_all = TRUE)
}
firm17 <- load_sbs_2017()
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs] <- 1 / length(nbrs)
}
W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs] <- 1 / length(nbrs)
}

# Add at the top of bdf-prereqs, before the W_geotrade/W_geoculture blocks:
drop_country <- "Slovenia"
keep_idx <- which(country_labels != drop_country)
country_keep <- country_labels[keep_idx]

W_geotrade_sub <- W_geotrade[keep_idx, keep_idx]
W_geoculture_sub <- W_geoculture[keep_idx, keep_idx]

# Lagged recovered networks (subsetted, Slovenia dropped)
W_q_lag_sub <- W_q_lag[keep_idx, keep_idx]
W_e_lag_sub <- W_e_lag[keep_idx, keep_idx]
```

```{r}
#| label: bdf-formulas
#| echo: false
#| include: false
rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2 <- paste(paste0("G2_", x_vars), collapse = " + ")
inst3 <- paste(paste0("G3_", x_vars), collapse = " + ")
f_nofe <- as.formula(sprintf("y ~ %s + %s | Gy ~ %s + %s", rhs, exo, inst2, inst3))
f_cfe <- as.formula(sprintf("y ~ %s + %s | country | Gy ~ %s + %s", rhs, exo, inst2, inst3))
```

```{r}
#| label: bdf-logdiff-lag-noslo
#| echo: false
#| results: asis
drop_country <- "Slovenia"
keep_idx <- which(country_labels != drop_country)
country_keep <- country_labels[keep_idx]

W_q_sub <- W_q[keep_idx, keep_idx]
W_e_sub <- W_e[keep_idx, keep_idx]
W_geotrade_sub <- W_geotrade[keep_idx, keep_idx]
W_geoculture_sub <- W_geoculture[keep_idx, keep_idx]
W_geo_sub <- W_geo[keep_idx, keep_idx]
W_trade_sub <- W_trade[keep_idx, keep_idx]
W_cul_sub <- W_cul[keep_idx, keep_idx]

row_normalise <- function(W) {
  rs <- rowSums(W)
  rs[rs == 0] <- 1
  W / rs
}
W_geo_sub <- row_normalise(W_geo_sub)
W_trade_sub <- row_normalise(W_trade_sub)
W_cul_sub <- row_normalise(W_cul_sub)

keep_full_local <- function(d, n_target) {
  yrs <- d |>
    dplyr::count(year) |>
    dplyr::filter(n == n_target) |>
    dplyr::pull(year)
  dplyr::filter(d, year %in% yrs)
}
fit_one_sub <- function(panel, W, B = 999) {
  d <- build_bdf(panel, W, x_vars, use_lag = TRUE)
  list(
    nofe = extract_nofe(feols(f_nofe, data = d, cluster = ~country)),
    cfe = extract_cfe(cluster_boot(d, f_cfe, "country", B = B))
  )
}
get_wr2 <- function(panel, W) {
  d <- build_bdf(panel, W, x_vars, use_lag = TRUE)
  m <- tryCatch(feols(f_cfe, data = d, cluster = ~country), error = \(e) NULL)
  if (is.null(m)) {
    return(NA_real_)
  }
  rv <- tryCatch(fixest::r2(m), error = \(e) NULL)
  if (is.null(rv)) {
    return(NA_real_)
  }
  unname(rv["wr2"])
}

panel_q <- attach_y_logdiff(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))
panel_e <- attach_y_logdiff(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))

q_rec <- fit_one_sub(panel_q, W_q_sub)
e_rec <- fit_one_sub(panel_e, W_e_sub)
q_abc2 <- fit_one_sub(panel_q, W_geotrade_sub)
e_abc2 <- fit_one_sub(panel_e, W_geoculture_sub)
q_geo <- fit_one_sub(panel_q, W_geo_sub)
e_geo <- fit_one_sub(panel_e, W_geo_sub)
q_trade <- fit_one_sub(panel_q, W_trade_sub)
e_trade <- fit_one_sub(panel_e, W_trade_sub)
q_cul <- fit_one_sub(panel_q, W_cul_sub)
e_cul <- fit_one_sub(panel_e, W_cul_sub)

wr2_vals <- c(
  get_wr2(panel_q, W_q_sub),       get_wr2(panel_q, W_geotrade_sub),
  get_wr2(panel_q, W_geo_sub),     get_wr2(panel_q, W_trade_sub),
  get_wr2(panel_q, W_cul_sub),
  get_wr2(panel_e, W_e_sub),       get_wr2(panel_e, W_geoculture_sub),
  get_wr2(panel_e, W_geo_sub),     get_wr2(panel_e, W_trade_sub),
  get_wr2(panel_e, W_cul_sub)
)

tab <- as.data.frame(rbind(
  rb("Quality ", "Recovered", q_rec),
  rb("Quality ", "Exogenous, Geo + Trade", q_abc2),
  rb("Quality ", "Exogenous, Geography", q_geo),
  rb("Quality ", "Exogenous, Trade", q_trade),
  rb("Quality ", "Exogenous, Cultural dist.", q_cul),
  rb("Environment ", "Recovered", e_rec),
  rb("Environment ", "Exogenous, Geo + Culture", e_abc2),
  rb("Environment ", "Exogenous, Geography", e_geo),
  rb("Environment ", "Exogenous, Trade", e_trade),
  rb("Environment ", "Exogenous, Cultural dist.", e_cul)
), stringsAsFactors = FALSE)

# --- CHANGES START HERE ---
# 1. Remove the 3rd column (no FE) immediately after binding
tab <- tab[, -3]

# 2. Update column names to reflect the removed column
colnames(tab) <- c("Panel", "Network", "$\\hat\\beta$ (country FE)")
tab[["$R^2$ (within)"]] <- ifelse(is.na(wr2_vals), "---", sprintf("%.3f", wr2_vals))

# 3. Adjust your plotting helper dataframe structure to match
tab_plot <- tab
colnames(tab_plot)[3] <- "beta_cfe"

# 4. Changed align from "llccc" to "llcc" since we have 4 columns total now
kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE, align = "llcc", row.names = FALSE,
  # caption = "Peer effect on log-differences: recovered vs Albuquerque networks",
  label = "tab:bdf_logdiff_lag_noslo"
) |>
  kable_styling(latex_options = "hold_position", font_size = 9) |>
  add_note(paste(
    " mmm."
  )) |>
  print()
```


```{r}
#| label: fig-gmm-ci
#| echo: false
#| fig-width: 9
#| fig-height: 5
K <- 5
x_names <- c("GDP p.c.", "Trade", "Rule of Law", "Renewables", "Nat. Res. Rents")

extract_ci <- function(rn, panel_label) {
  av <- rn$unpenalisedgmm$asyvar
  est <- c(rn$unpenalisedgmm$rho, rn$unpenalisedgmm$beta)
  se <- sqrt(av[1:(K + 1)])
  pars <- c(
    "Peer effect",
    "GDP p.c.",
    "Trade (% GDP)",
    "Rule of Law",
    "Renewables (%)",
    "Nat. Res. Rents (% GDP)"
  )
  data.frame(
    panel = panel_label, param = pars, estimate = est, se = se,
    lo95 = est - 1.96 * se, hi95 = est + 1.96 * se, stringsAsFactors = FALSE
  )
}

ci_df <- dplyr::bind_rows(
  extract_ci(best[["9001_eu"]]$rn, "Quality (ISO 9001)"),
  extract_ci(best[["14001_eu"]]$rn, "Environment (ISO 14001)")
) |>
  dplyr::mutate(
    param = factor(param, levels = rev(unique(param))),
    panel = factor(panel, levels = c("Quality (ISO 9001)", "Environment (ISO 14001)")),
    # GMM has no within-R^2; label the point estimate only
    est_label = sprintf("%.3f", estimate)
  )

dodge_width <- 0.55
pd <- position_dodge(width = dodge_width)

p_ci <- ggplot(ci_df, aes(x = estimate, y = param, colour = panel, group = panel)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_linerange(aes(xmin = lo95, xmax = hi95), position = pd, linewidth = 0.8) +
  geom_point(position = pd, size = 2.8) +
  geom_text(
    aes(label = est_label),
    position = pd, vjust = -1.2, size = 2.3,
    show.legend = FALSE
  ) +
  scale_colour_manual(values = c(
    "Quality (ISO 9001)" = "black",
    "Environment (ISO 14001)" = "grey55"
  )) +
  coord_cartesian(clip = "off") +
  labs(x = "Estimate", y = NULL, colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(size = 7, colour = "grey40", hjust = 0)
  )

print(p_ci)
save_figure("gmm_ci", width = 9, height = 5)
```

```{r}
#| label: bdf-adjusted-tables
#| echo: false
#| results: asis
fit_spec <- function(panel, W, lag, B = 999) {
  d <- build_bdf(panel, W, x_vars, use_lag = lag)
  list(
    nofe = extract_nofe(feols(f_nofe, data = d, cluster = ~country)),
    cfe  = extract_cfe(cluster_boot(d, f_cfe, "country", B = B))
  )
}

get_wr2_adjusted <- function(panel, W, lag) {
  d <- build_bdf(panel, W, x_vars, use_lag = lag)
  m <- tryCatch(feols(f_cfe, data = d, cluster = ~country), error = \(e) NULL)
  if (is.null(m)) {
    return(NA_real_)
  }
  rv <- tryCatch(fixest::r2(m), error = \(e) NULL)
  if (is.null(rv)) {
    return(NA_real_)
  }
  unname(rv["wr2"])
}

get_nobs <- function(panel, W, lag) {
  d <- build_bdf(panel, W, x_vars, use_lag = lag)
  nrow(tidyr::drop_na(d, Gy))
}

# Network lists — recovered network switches by spec
make_networks <- function(lag) {
  list(
    q = list(
      "Recovered"                 = W_q_sub,
      "Exogenous, Geo + Trade"    = W_geotrade_sub,
      "Exogenous, Geography"      = W_geo_sub,
      "Exogenous, Trade"          = W_trade_sub,
      "Exogenous, Cultural dist." = W_cul_sub
    ),
    e = list(
      "Recovered"                 = W_e_sub,
      "Exogenous, Geo + Culture"  = W_geoculture_sub,
      "Exogenous, Geography"      = W_geo_sub,
      "Exogenous, Trade"          = W_trade_sub,
      "Exogenous, Cultural dist." = W_cul_sub
    )
  )
}

run_all <- function(lag) {
  nets <- make_networks(lag)
  rows <- NULL
  for (nm in names(nets$q)) {
    rows <- rbind(rows, rb("Quality", nm, fit_spec(panel_q, nets$q[[nm]], lag)))
  }
  for (nm in names(nets$e)) {
    rows <- rbind(rows, rb("Environment", nm, fit_spec(panel_e, nets$e[[nm]], lag)))
  }
  as.data.frame(rows, stringsAsFactors = FALSE)
}

tab_nolag <- run_all(lag = FALSE)
tab_lag <- run_all(lag = TRUE)

# Within-R2
make_wr2 <- function(lag) {
  nets <- make_networks(lag)
  c(
    get_wr2_adjusted(panel_q, nets$q[["Recovered"]], lag),
    get_wr2_adjusted(panel_q, nets$q[["Exogenous, Geo + Trade"]], lag),
    get_wr2_adjusted(panel_q, nets$q[["Exogenous, Geography"]], lag),
    get_wr2_adjusted(panel_q, nets$q[["Exogenous, Trade"]], lag),
    get_wr2_adjusted(panel_q, nets$q[["Exogenous, Cultural dist."]], lag),
    get_wr2_adjusted(panel_e, nets$e[["Recovered"]], lag),
    get_wr2_adjusted(panel_e, nets$e[["Exogenous, Geo + Culture"]], lag),
    get_wr2_adjusted(panel_e, nets$e[["Exogenous, Geography"]], lag),
    get_wr2_adjusted(panel_e, nets$e[["Exogenous, Trade"]], lag),
    get_wr2_adjusted(panel_e, nets$e[["Exogenous, Cultural dist."]], lag)
  )
}

wr2_nolag_vals <- make_wr2(lag = FALSE)
wr2_lag_vals <- make_wr2(lag = TRUE)

nobs_q_nolag <- get_nobs(panel_q, W_q_sub, lag = FALSE)
nobs_e_nolag <- get_nobs(panel_e, W_e_sub, lag = FALSE)
nobs_q_lag <- get_nobs(panel_q, W_q_sub, lag = TRUE)
nobs_e_lag <- get_nobs(panel_e, W_e_sub, lag = TRUE)

print_bdf_tab <- function(tab, wr2_vec, caption, label, note, nobs_q, nobs_e) {
  tab <- tab[, -3]
  colnames(tab) <- c("Panel", "Network", "$\\hat\\beta$ (country FE)")
  odd_idx <- seq(1, nrow(tab), by = 2)
  r2_col <- rep("", nrow(tab))
  obs_col <- rep("", nrow(tab))
  r2_col[odd_idx] <- ifelse(is.na(wr2_vec), "---", sprintf("%.3f", wr2_vec))
  obs_col[odd_idx[1:5]] <- formatC(nobs_q, format = "d", big.mark = ",")
  obs_col[odd_idx[6:10]] <- formatC(nobs_e, format = "d", big.mark = ",")
  tab[["$R^2$ (within)"]] <- r2_col
  tab[["$N$"]] <- obs_col
  kbl(tab,
    booktabs = TRUE, format = "latex", escape = FALSE,
    align = "llccc", row.names = FALSE, caption = caption, label = label
  ) |>
    kable_styling(latex_options = "hold_position", font_size = 9) |>
    add_note(note) |>
    print()
}

print_bdf_tab(
  tab_nolag, wr2_nolag_vals,
  caption = "Peer-effect estimates — simultaneous specification. $\\Delta\\log y_{it}$ on $x_{it}$, $Gx_{it}$; $Gy_{it}$ instrumented by $G^2x_{it}$, $G^3x_{it}$. Network recovered from simultaneous De Paula et al.\\ (2025) specification.",
  label = "tab:bdf_adjusted_nolag",
  note = "2SLS with country FE. Cluster-bootstrap SEs ($B=999$, country clusters). Slovenia excluded.",
  nobs_q = nobs_q_nolag, nobs_e = nobs_e_nolag
)

print_bdf_tab(
  tab_lag, wr2_lag_vals,
  caption = "Peer-effect estimates — lagged specification. $\\Delta\\log y_{it}$ on $x_{i,t-1}$, $Gx_{i,t-1}$; $Gy_{i,t-1}$ instrumented by $G^2x_{i,t-1}$, $G^3x_{i,t-1}$. Network recovered from lagged De Paula et al.\\ (2025) specification.",
  label = "tab:bdf_adjusted_lag",
  note = "2SLS with country FE. All regressors dated $t-1$. Cluster-bootstrap SEs ($B=999$, country clusters). Slovenia excluded.",
  nobs_q = nobs_q_lag, nobs_e = nobs_e_lag
)
```

```{r}
#| label: fig-bdf-forest-nolag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "BDF peer-effect estimates and 95\\% confidence intervals — simultaneous specification"
parse_est <- function(x) as.numeric(gsub("\\$\\^\\{.*?\\}\\$|[[:space:]]", "", x))
parse_se <- function(x) as.numeric(gsub("[()[:space:]]", "", x))

build_ci <- function(tab, spec_label) {
  est_rows <- tab[seq(1, nrow(tab), by = 2), ]
  se_rows <- tab[seq(2, nrow(tab), by = 2), ]
  data.frame(
    panel = est_rows$V1, network = est_rows$V2,
    estimate = parse_est(est_rows$V4), se = parse_se(se_rows$V4),
    stringsAsFactors = FALSE
  ) |>
    dplyr::mutate(
      lo95 = estimate - 1.96 * se,
      hi95 = estimate + 1.96 * se,
      network = dplyr::recode(network,
        "Exogenous, Geo + Trade"    = "Geo + Trade",
        "Exogenous, Geo + Culture"  = "Geo + Culture",
        "Exogenous, Geography"      = "Geography",
        "Exogenous, Trade"          = "Trade",
        "Exogenous, Cultural dist." = "Culture"
      ),
      network = forcats::fct_relevel(
        factor(network, levels = c(
          "Recovered", "Geo + Trade", "Geo + Culture",
          "Geography", "Trade", "Culture"
        )),
        "Recovered",
        after = Inf
      ),
      panel = factor(panel, levels = c("Quality", "Environment")),
      spec = spec_label,
      est_label = sprintf("%.3f", estimate)
    )
}

spec_vals <- c("Simultaneous" = "black", "Lagged" = "grey55")
pd <- position_dodge(width = 0.55, preserve = "single")

forest_plot <- function(df) {
  df <- dplyr::mutate(df, spec = factor(spec, levels = c("Simultaneous", "Lagged")))
  ggplot(df, aes(x = estimate, y = network, colour = spec, group = spec)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
    geom_linerange(aes(xmin = lo95, xmax = hi95), position = pd, linewidth = 0.8, na.rm = TRUE) +
    geom_point(position = pd, size = 2.6, na.rm = TRUE) +
    geom_text(aes(label = est_label),
      position = pd, vjust = -1.1, size = 2.3,
      show.legend = FALSE, na.rm = TRUE
    ) +
    scale_colour_manual(values = spec_vals, drop = FALSE) +
    facet_wrap(~panel, ncol = 2) +
    coord_cartesian(clip = "off") +
    labs(x = expression("Estimate (" * hat(beta) * ")"), y = NULL, colour = NULL) +
    theme_minimal(base_size = 10) +
    theme(
      legend.position = "bottom", panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(), axis.text.y = element_text(size = 9),
      strip.text = element_text(face = "bold")
    )
}

ci_nolag <- build_ci(tab_nolag, "Simultaneous")

p_bdf_nolag <- forest_plot(ci_nolag)
print(p_bdf_nolag)
save_figure("bdf_forest_nolag", width = 12, height = 6)
```

```{r}
#| label: fig-bdf-forest-lag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "BDF peer-effect estimates and 95\\% confidence intervals — lagged specification"
ci_lag <- build_ci(tab_lag, "Lagged")

p_bdf_lag <- forest_plot(ci_lag)
print(p_bdf_lag)
save_figure("bdf_forest_lag", width = 12, height = 6)
```

```{r}
#| label: fig-bdf-forest-combined
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "BDF peer-effect estimates and 95\\% confidence intervals — simultaneous and lagged specifications"
ci_combined <- dplyr::bind_rows(ci_nolag, ci_lag) |>
  dplyr::mutate(
    network = forcats::fct_relevel(network, "Recovered", after = Inf),
    spec    = factor(spec, levels = c("Simultaneous", "Lagged"))
  )

p_bdf_combined <- forest_plot(ci_combined)
print(p_bdf_combined)
save_figure("bdf_forest_combined", width = 12, height = 6)

# R² data frames and plot function — used by fig-bdf-r2-nolag and fig-bdf-r2-lag
net_q <- c("Recovered", "Geo + Trade", "Geography", "Trade", "Culture")
net_e <- c("Recovered", "Geo + Culture", "Geography", "Trade", "Culture")

relevel_net <- function(df) {
  df |>
    dplyr::mutate(
      network = forcats::fct_relevel(
        factor(network, levels = c(
          "Recovered", "Geo + Trade", "Geo + Culture",
          "Geography", "Trade", "Culture"
        )),
        "Recovered",
        after = Inf
      ),
      panel = factor(panel, levels = c("Quality", "Environment"))
    )
}

r2_nolag <- data.frame(
  panel = c(rep("Quality", 5), rep("Environment", 5)),
  network = c(net_q, net_e),
  wr2 = wr2_nolag_vals,
  stringsAsFactors = FALSE
) |>
  relevel_net() |>
  dplyr::mutate(spec = "Simultaneous")

r2_lag <- data.frame(
  panel = c(rep("Quality", 5), rep("Environment", 5)),
  network = c(net_q, net_e),
  wr2 = wr2_lag_vals,
  stringsAsFactors = FALSE
) |>
  relevel_net() |>
  dplyr::mutate(spec = "Lagged")

spec_vals <- c("Simultaneous" = "grey20", "Lagged" = "grey65")
pd_r2 <- position_dodge(width = 0.7, preserve = "single")

r2_plot <- function(df) {
  df <- dplyr::mutate(df, spec = factor(spec, levels = c("Simultaneous", "Lagged")))
  ggplot(df, aes(x = wr2, y = network, fill = spec)) +
    geom_col(position = pd_r2, width = 0.6, na.rm = TRUE) +
    geom_text(aes(label = ifelse(is.na(wr2), "", sprintf("%.3f", wr2)), group = spec),
      position = pd_r2, hjust = -0.15, size = 2.3, na.rm = TRUE
    ) +
    scale_fill_manual(values = spec_vals, drop = FALSE) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
    facet_wrap(~panel, ncol = 2) +
    labs(x = expression("Within-" * R^2), y = NULL, fill = NULL) +
    theme_minimal(base_size = 10) +
    theme(
      legend.position = "bottom", panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(), strip.text = element_text(face = "bold")
    )
}
```

```{r}
#| label: fig-bdf-r2-nolag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Within-$R^2$ — simultaneous specification"
p_r2_nolag <- r2_plot(r2_nolag)
print(p_r2_nolag)
save_figure("bdf_r2_nolag", width = 12, height = 6)
```

```{r}
#| label: fig-bdf-r2-lag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Within-$R^2$ — lagged specification"
p_r2_lag <- r2_plot(r2_lag)
print(p_r2_lag)
save_figure("bdf_r2_lag", width = 12, height = 6)
```

```{r}
#| label: grid-search-summary
#| echo: false
#| results: asis
grid_meta <- readRDS(file.path(read_dir, "grid_search_meta.rds"))

grid_df <- do.call(rbind, lapply(grid_meta, function(r) {
  data.frame(
    Panel = r$panel,
    L1 = sprintf("%.3f", r$lambda[1]),
    L2 = sprintf("%.3f", r$lambda[2]),
    L3 = sprintf("%.3f", r$lambda[3]),
    Edges = ifelse(is.na(r$edges), "---", as.character(r$edges)),
    rho_GMM = ifelse(is.na(r$rho_gmm), "---", sprintf("%.3f", r$rho_gmm)),
    BIC = ifelse(is.na(r$bic), "---", sprintf("%.2f", r$bic)),
    bic_num = suppressWarnings(as.numeric(r$bic)),
    stringsAsFactors = FALSE
  )
})) |>
  dplyr::mutate(Panel = dplyr::recode(Panel,
    "9001_eu"  = "Quality",
    "14001_eu" = "Environment"
  )) |>
  dplyr::group_by(Panel) |>
  dplyr::slice_min(bic_num, n = 5, with_ties = FALSE) |>
  dplyr::ungroup() |>
  dplyr::arrange(Panel, bic_num)

# BIC-best row per panel (rank 1 within each panel = bold)
best_rows <- which(!duplicated(grid_df$Panel))

grid_df <- grid_df |> dplyr::select(-bic_num)

tab_grid <- kbl(grid_df,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "lcccrrrr", row.names = FALSE,
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat\\rho_{GMM}$", "BIC"
  ),
  caption = "Top-5 BIC grid search results per standard. BIC-selected specification in bold.",
  label = "tab:grid_search"
) |>
  kable_styling(latex_options = "hold_position", font_size = 9) |>
  row_spec(best_rows, bold = TRUE) |>
    add_note("Top 5 penalisation-parameter combinations by BIC, from a grid search over the adaptive elastic-net GMM estimator of De Paula, Rasul and Souza (2025). The BIC-selected specification per standard is shown in bold.")

save_table(tab_grid, "grid_search_summary")
tab_grid |> print()
```

# Year FE

```{r}
#| label: bdf-twfe-tables
#| echo: false
#| results: asis
f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy ~ %s + %s", rhs, exo, inst2, inst3
))
# Drop-in replacement: same as fit_spec but with f_twfe instead of f_cfe
fit_spec_twfe <- function(panel, W, lag, B = 999) {
  d <- build_bdf(panel, W, x_vars, use_lag = lag)
  list(
    nofe = extract_nofe(feols(f_nofe, data = d, cluster = ~country)),
    cfe  = extract_cfe(cluster_boot(d, f_twfe, "country", B = B))
  )
}

get_wr2_twfe <- function(panel, W, lag) {
  d <- build_bdf(panel, W, x_vars, use_lag = lag)
  m <- tryCatch(feols(f_twfe, data = d, cluster = ~country), error = \(e) NULL)
  if (is.null(m)) {
    return(NA_real_)
  }
  rv <- tryCatch(fixest::r2(m), error = \(e) NULL)
  if (is.null(rv)) {
    return(NA_real_)
  }
  unname(rv["wr2"])
}

run_all_twfe <- function(lag) {
  nets <- make_networks(lag)
  rows <- NULL
  for (nm in names(nets$q)) {
    rows <- rbind(rows, rb("Quality", nm, fit_spec_twfe(panel_q, nets$q[[nm]], lag)))
  }
  for (nm in names(nets$e)) {
    rows <- rbind(rows, rb("Environment", nm, fit_spec_twfe(panel_e, nets$e[[nm]], lag)))
  }
  as.data.frame(rows, stringsAsFactors = FALSE)
}

make_wr2_twfe <- function(lag) {
  nets <- make_networks(lag)
  c(
    get_wr2_twfe(panel_q, nets$q[["Recovered"]], lag),
    get_wr2_twfe(panel_q, nets$q[["Exogenous, Geo + Trade"]], lag),
    get_wr2_twfe(panel_q, nets$q[["Exogenous, Geography"]], lag),
    get_wr2_twfe(panel_q, nets$q[["Exogenous, Trade"]], lag),
    get_wr2_twfe(panel_q, nets$q[["Exogenous, Cultural dist."]], lag),
    get_wr2_twfe(panel_e, nets$e[["Recovered"]], lag),
    get_wr2_twfe(panel_e, nets$e[["Exogenous, Geo + Culture"]], lag),
    get_wr2_twfe(panel_e, nets$e[["Exogenous, Geography"]], lag),
    get_wr2_twfe(panel_e, nets$e[["Exogenous, Trade"]], lag),
    get_wr2_twfe(panel_e, nets$e[["Exogenous, Cultural dist."]], lag)
  )
}

tab_twfe_nolag <- run_all_twfe(lag = FALSE)
tab_twfe_lag <- run_all_twfe(lag = TRUE)
wr2_twfe_nolag <- make_wr2_twfe(lag = FALSE)
wr2_twfe_lag <- make_wr2_twfe(lag = TRUE)

print_bdf_tab(
  tab_twfe_nolag, wr2_twfe_nolag,
  caption = "Peer-effect estimates — simultaneous specification, country $+$ year FE.",
  label = "tab:bdf_twfe_nolag",
  note = "2SLS with country and year FE. Cluster-bootstrap SEs ($B=999$, country clusters). Slovenia excluded.",
  nobs_q = nobs_q_nolag, nobs_e = nobs_e_nolag
)

print_bdf_tab(
  tab_twfe_lag, wr2_twfe_lag,
  caption = "Peer-effect estimates — lagged specification, country $+$ year FE.",
  label = "tab:bdf_twfe_lag",
  note = "2SLS with country and year FE. All regressors dated $t-1$. Cluster-bootstrap SEs ($B=999$, country clusters). Slovenia excluded.",
  nobs_q = nobs_q_lag, nobs_e = nobs_e_lag
)
```

```{r}
#| label: fig-bdf-forest-twfe
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "BDF peer-effect estimates and 95\\% CI — country $+$ year FE, simultaneous and lagged"
ci_twfe_nolag <- build_ci(tab_twfe_nolag, "Simultaneous")
ci_twfe_lag <- build_ci(tab_twfe_lag, "Lagged")

ci_twfe_combined <- dplyr::bind_rows(ci_twfe_nolag, ci_twfe_lag) |>
  dplyr::mutate(
    network = forcats::fct_relevel(network, "Recovered", after = Inf),
    spec    = factor(spec, levels = c("Simultaneous", "Lagged"))
  )

p_twfe <- forest_plot(ci_twfe_combined)
print(p_twfe)
save_figure("bdf_forest_twfe", width = 12, height = 6)
```


```{r}
#| label: bdf-fullag-standalone
#| echo: false
#| results: asis
# Fully-lagged: x_{t-1}, Gx_{t-1}, y_{t-1}, Wy_{t-1}. OLS-FE, no IV.
build_bdf_fullag <- function(panel, W, x_vars) {
  build_bdf(panel, W, x_vars, use_lag = TRUE) |> # x_{t-1}, Gx_{t-1}; Gy is Wy_t
    dplyr::arrange(country, year) |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      y_lag  = dplyr::lag(y, 1, order_by = year), # theta: y_{i,t-1}
      Gy_lag = dplyr::lag(Gy, 1, order_by = year) # beta1: (Wy)_{i,t-1}
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(Gy_lag, y_lag)
}
f_fullag <- as.formula(sprintf("y ~ %s + %s + y_lag + Gy_lag | country + year", rhs, exo))
fit_fullag <- function(panel, W) feols(f_fullag, data = build_bdf_fullag(panel, W, x_vars), cluster = ~country)

nets_fa <- make_networks(lag = TRUE) # "Recovered" = contemporaneous recovered W; lag applies to the equation only
m_q <- fit_fullag(panel_q, nets_fa$q[["Recovered"]])
m_e <- fit_fullag(panel_e, nets_fa$e[["Recovered"]])

lr <- function(m) {
  b <- coef(m)
  unname(b["Gy_lag"] / (1 - b["y_lag"]))
}

etable(
  list(m_q, m_e),
  depvar = FALSE,
  headers = c("Quality", "Environment"),
  order = c("Gy_lag", "y_lag"), keep = c("%Gy_lag", "%y_lag"),
  dict = c(
    Gy_lag = "$\\beta_1$: peers' past adoption $Wy_{t-1}$",
    y_lag = "$\\theta$: own persistence $y_{t-1}$"
  ),
  fitstat = ~ n + wr2, signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10), tex = TRUE,
  title = "Fully-lagged specification (recovered networks): covariates, own and peer outcomes all dated $t-1$. OLS with country and year FE; no instrument.",
  label = "tab:bdf_fullag",
  notes = sprintf(
    paste(
      "Lagged-peer restriction of BDF Eq.~9 ($\\beta_0=0$): $Wy_{t-1}$ is predetermined, so no reflection problem and no IV",
      "(Bramoull\\'e, Djebbari \\& Fortin 2009; Manski 1993). Network recovered with lagged covariates.",
      "$\\theta$ carries $O(1/T)$ Nickell (1981) bias. Long-run propagation $\\beta_1/(1-\\theta)$: Quality %.3f, Environment %.3f. Slovenia excluded."
    ),
    lr(m_q), lr(m_e)
  ),
  file = file.path(tab_dir, "bdf_fullag.tex")
)

# Forest of beta1 across all networks (built from fitted objects, no text parsing)
ci_fa <- function(panel, fits) {
  do.call(rbind, lapply(names(fits), function(nm) {
    ct <- summary(fits[[nm]])$coeftable
    if (!"Gy_lag" %in% rownames(ct)) {
      return(NULL)
    }
    b <- ct["Gy_lag", 1]
    se <- ct["Gy_lag", 2]
    data.frame(panel = panel, network = nm, estimate = b, lo95 = b - 1.96 * se, hi95 = b + 1.96 * se)
  }))
}
fa_all <- dplyr::bind_rows(
  ci_fa("Quality", lapply(nets_fa$q, function(W) fit_fullag(panel_q, W))),
  ci_fa("Environment", lapply(nets_fa$e, function(W) fit_fullag(panel_e, W)))
) |>
  dplyr::mutate(
    network = dplyr::recode(network,
      "Exogenous, Geo + Trade" = "Geo + Trade", "Exogenous, Geo + Culture" = "Geo + Culture",
      "Exogenous, Geography" = "Geography", "Exogenous, Trade" = "Trade", "Exogenous, Cultural dist." = "Culture"
    ),
    network = forcats::fct_relevel(factor(network), "Recovered", after = Inf),
    panel = factor(panel, levels = c("Quality", "Environment")),
    est_label = sprintf("%.3f", estimate)
  )

p_fa <- ggplot(fa_all, aes(estimate, network)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_linerange(aes(xmin = lo95, xmax = hi95), linewidth = 0.8) +
  geom_point(size = 2.6) +
  geom_text(aes(label = est_label), vjust = -1.1, size = 2.3) +
  facet_wrap(~panel, ncol = 2) +
  coord_cartesian(clip = "off") +
  labs(x = expression("Lagged peer effect (" * beta[1] * ")"), y = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(), panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold")
  )
print(p_fa)
save_figure("bdf_fullag_forest", width = 12, height = 6)
```

```{r}
#| label: bdf-fullag-splice
#| echo: false
#| results: asis
# Run AFTER bdf-adjusted-tables (needs panels, helpers, build_bdf_fullag/fit_fullag from chunk A).
# Overrides the "Lagged" objects so every combined figure now means fully-lagged-peer.

star_lp <- function(p) if (is.na(p)) "" else if (p < 0.01) "$^{***}$" else if (p < 0.05) "$^{**}$" else if (p < 0.10) "$^{*}$" else ""
rb_lp <- function(panel, network, m, coef = "Gy_lag") { # mirrors rb()'s 2-row, V1..V4 layout
  ct <- summary(m)$coeftable
  if (!coef %in% rownames(ct)) {
    est <- "---"
    se <- ""
  } else {
    b <- ct[coef, 1]
    s <- ct[coef, 2]
    p <- ct[coef, 4]
    est <- sprintf("%.3f%s", b, star_lp(p))
    se <- sprintf("(%.3f)", s)
  }
  rbind(c(panel, network, "", est), c("", "", "", se))
}
run_all_fullag <- function() {
  nets <- make_networks(lag = TRUE)
  rows <- NULL
  for (nm in names(nets$q)) rows <- rbind(rows, rb_lp("Quality", nm, fit_fullag(panel_q, nets$q[[nm]])))
  for (nm in names(nets$e)) rows <- rbind(rows, rb_lp("Environment", nm, fit_fullag(panel_e, nets$e[[nm]])))
  as.data.frame(rows, stringsAsFactors = FALSE) # columns auto-named V1..V4
}
wr2_fullag <- function() {
  nets <- make_networks(lag = TRUE)
  c(
    sapply(nets$q, function(W) unname(fixest::r2(fit_fullag(panel_q, W), "wr2"))),
    sapply(nets$e, function(W) unname(fixest::r2(fit_fullag(panel_e, W), "wr2")))
  )
}

# Override the lagged objects the downstream figures consume
tab_lag <- run_all_fullag()
wr2_lag_vals <- wr2_fullag()
ci_lag <- build_ci(tab_lag, "Lagged")
nobs_q_lag <- nrow(build_bdf_fullag(panel_q, W_q_sub, x_vars))
nobs_e_lag <- nrow(build_bdf_fullag(panel_e, W_e_sub, x_vars))

print_bdf_tab(
  tab_lag, wr2_lag_vals,
  caption = "Peer-effect estimates — fully-lagged specification. $\\beta_1$ on peers' past adoption $Wy_{i,t-1}$; covariates, own outcome and peer outcome all dated $t-1$. OLS with country and year FE, no IV.",
  label = "tab:bdf_fullag_panel",
  note = "Lagged-peer restriction ($\\beta_0=0$): predetermined RHS, no reflection problem, no instrument (Bramoull\\'e, Djebbari \\& Fortin 2009). Recovered = contemporaneous recovered network; only the equation is lagged. Slovenia excluded.",
  nobs_q = nobs_q_lag, nobs_e = nobs_e_lag
)

# Recompute combined forest / R²: "Simultaneous" = beta0 (IV), "Lagged" = beta1 (lagged peer)
ci_combined <- dplyr::bind_rows(ci_nolag, ci_lag) |>
  dplyr::mutate(
    network = forcats::fct_relevel(network, "Recovered", after = Inf),
    spec = factor(spec, levels = c("Simultaneous", "Lagged"))
  )
print(forest_plot(ci_combined))
save_figure("bdf_forest_combined", width = 12, height = 6)

r2_lag <- data.frame(
  panel = c(rep("Quality", 5), rep("Environment", 5)),
  network = c(net_q, net_e), wr2 = wr2_lag_vals
) |>
  relevel_net() |>
  dplyr::mutate(spec = "Lagged")
print(r2_plot(dplyr::bind_rows(r2_nolag, r2_lag)))
save_figure("bdf_r2_lag", width = 12, height = 6)
```

```{r}
#| label: fig-bdf-r2-twfe
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Within-$R^2$ — simultaneous specification, country $+$ year FE"
# net_q / net_e / relevel_net / r2_plot defined in fig-bdf-forest-combined; wr2_twfe_nolag in bdf-twfe-tables.
r2_twfe_nolag <- data.frame(
  panel = c(rep("Quality", 5), rep("Environment", 5)),
  network = c(net_q, net_e),
  wr2 = wr2_twfe_nolag,
  stringsAsFactors = FALSE
) |>
  relevel_net() |>
  dplyr::mutate(spec = "Simultaneous")

p_r2_twfe <- r2_plot(r2_twfe_nolag) + theme(legend.position = "none")
print(p_r2_twfe)
save_figure("bdf_r2_twfe", width = 12, height = 6)
```
```{r}
#| label: fig-bdf-r2-fullag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Within-$R^2$ — fully-lagged-peer specification"
# wr2_fullag() is defined in bdf-fullag-splice; this chunk must run after it.
# Names come back as the full "Exogenous, ..." labels -> recode to match net_q/net_e.
wr2_fa <- wr2_fullag()
r2_fullag <- data.frame(
  panel = c(rep("Quality", 5), rep("Environment", 5)),
  network = c(net_q, net_e), # same network order make_networks() iterates
  wr2 = unname(wr2_fa),
  stringsAsFactors = FALSE
) |>
  relevel_net() |>
  dplyr::mutate(spec = "Fully lagged")

p_r2_fullag <- r2_plot(r2_fullag) + theme(legend.position = "none")
print(p_r2_fullag)
save_figure("bdf_r2_fullag", width = 12, height = 6)
```

```{r}
#| label: fig-bdf-forest-twfe-sim-vs-fullag
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Peer-effect estimates, country $+$ year FE: simultaneous ($\\beta_0$ on $Wy_t$, IV) vs fully-lagged ($\\beta_1$ on $Wy_{t-1}$, OLS)"
# Needs: ci_twfe_nolag (fig-bdf-forest-twfe), forest_plot/spec_vals (fig-bdf-forest-combined),
#        fit_fullag/nets_fa (bdf-fullag-standalone).

# Pairs cluster bootstrap (country clusters, B = 999) for the fully-lagged beta_1,
# mirroring cluster_boot() on the simultaneous side so both series in the figure
# use cluster-bootstrap SEs, as stated in the figure note.
boot_se_fullag <- function(panel, W, B = 999, seed = 42) {
  d <- build_bdf_fullag(panel, W, x_vars)
  ids <- unique(d$country)
  set.seed(seed)
  draws <- replicate(B, {
    samp <- sample(ids, length(ids), replace = TRUE)
    db <- do.call(rbind, lapply(seq_along(samp), function(k) {
      dk <- d[d$country == samp[k], , drop = FALSE]
      dk$country <- paste0(samp[k], "_", k) # duplicates get distinct FE ids
      dk
    }))
    m <- tryCatch(feols(f_fullag, data = db), error = \(e) NULL)
    if (is.null(m) || !"Gy_lag" %in% names(coef(m))) NA_real_ else unname(coef(m)["Gy_lag"])
  })
  stats::sd(draws, na.rm = TRUE)
}

ci_fullag <- function(panel_label, panel_data, nets) {
  do.call(rbind, lapply(names(nets), function(nm) {
    m <- fit_fullag(panel_data, nets[[nm]])
    ct <- summary(m)$coeftable
    if (!"Gy_lag" %in% rownames(ct)) {
      return(NULL)
    }
    b <- ct["Gy_lag", 1]
    se <- boot_se_fullag(panel_data, nets[[nm]]) # cluster-bootstrap SE (country, B = 999)
    data.frame(
      panel = panel_label, network = nm, estimate = b, se = se,
      lo95 = b - 1.96 * se, hi95 = b + 1.96 * se, stringsAsFactors = FALSE
    )
  }))
}

ci_lagpeer <- dplyr::bind_rows(
  ci_fullag("Quality", panel_q, nets_fa$q),
  ci_fullag("Environment", panel_e, nets_fa$e)
) |>
  dplyr::mutate(
    network = dplyr::recode(network,
      "Exogenous, Geo + Trade" = "Geo + Trade", "Exogenous, Geo + Culture" = "Geo + Culture",
      "Exogenous, Geography" = "Geography", "Exogenous, Trade" = "Trade", "Exogenous, Cultural dist." = "Culture"
    ),
    network = forcats::fct_relevel(
      factor(network, levels = c("Recovered", "Geo + Trade", "Geo + Culture", "Geography", "Trade", "Culture")),
      "Recovered",
      after = Inf
    ),
    panel = factor(panel, levels = c("Quality", "Environment")),
    spec = "Lagged",
    est_label = sprintf("%.3f", estimate)
  )

ci_twfe_sim_vs_lag <- dplyr::bind_rows(ci_twfe_nolag, ci_lagpeer) |>
  dplyr::mutate(
    network = forcats::fct_relevel(network, "Recovered", after = Inf),
    spec    = factor(spec, levels = c("Simultaneous", "Lagged"))
  )

p_twfe_sim_vs_fullag <- forest_plot(ci_twfe_sim_vs_lag)
print(p_twfe_sim_vs_fullag)
save_figure("bdf_forest_twfe_sim_vs_fullag", width = 12, height = 6)
```