---
title: "ISOs Diffusion in the EU: an endogenous network recovery for Environmental and Quality Management Systems"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "main.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
data_dir <- "/workspaces/thesis/data"
clean_dir <- "/workspaces/thesis/data/cleaned_data"

# install.packages(
#  c("maps", "countrycode")
# )

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)
library(patchwork)
library(ggraph)
library(ggrepel)
library(tidygraph)
library(maps)
library(countrycode)
set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
fig_dir <- file.path(out_dir, "fig")
tab_dir <- file.path(out_dir, "tab")
for (d in c(fig_dir, tab_dir)) if (!dir.exists(d)) dir.create(d, recursive = TRUE)
```

```{r}
#| label: load-best
#| include: false
#| cache: true
best <- readRDS(file.path(out_dir, "best_per_panel.rds"))
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
diag(W_q) <- 0
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)
diag(W_e) <- 0

sync_eu <- readRDS(file.path(clean_dir, "sync_eu.rds"))
country_labels <- sort(sync_eu)
rownames(W_q) <- colnames(W_q) <- country_labels
rownames(W_e) <- colnames(W_e) <- country_labels
```

# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Descriptive Statistics, Network Recoverey and Grid Search

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
#| cache: true 
f9001 <- readRDS(file.path(clean_dir, "f9001.rds"))
f14001 <- readRDS(file.path(clean_dir, "f14001.rds"))

desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, countries, years) {
  df |>
    dplyr::filter(country %in% countries, year %in% years) |>
    dplyr::group_by(year) |>
    dplyr::summarise(
      total_cert     = sum(n_cert, na.rm = TRUE),
      mean_cert      = mean(n_cert, na.rm = TRUE),
      sd_cert        = sd(n_cert, na.rm = TRUE),
      min_cert       = min(n_cert, na.rm = TRUE),
      max_cert       = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries    = dplyr::n(),
      .groups        = "drop"
    )
}

d9 <- desc_tab(f9001, country_labels, desc_years)
d14 <- desc_tab(f14001, country_labels, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Descriptive statistics: certification counts and adoption intensity across EU countries}",
  "\\label{tab:desc_stats}",
  "\\begin{tabular}{lcccccc}\\hline\\hline",
  "& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\",
  "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
  "& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\\hline",
  sprintf(
    "N countries & %d & %d & %d & & %d & %d \\\\",
    d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
    d14$n_countries[1], d14$n_countries[2]
  ),
  sprintf(
    "Total & %s & %s & %s & & %s & %s \\\\",
    f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
    f0(d14$total_cert[1]), f0(d14$total_cert[2])
  ),
  sprintf(
    "Mean & %s & %s & %s & & %s & %s \\\\",
    f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
    f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
  ),
  sprintf(
    "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\",
    f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
    f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
  ),
  sprintf(
    "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\",
    f0(d9$min_cert[1]), f0(d9$max_cert[1]),
    f0(d9$min_cert[2]), f0(d9$max_cert[2]),
    f0(d9$min_cert[3]), f0(d9$max_cert[3]),
    f0(d14$min_cert[1]), f0(d14$max_cert[1]),
    f0(d14$min_cert[2]), f0(d14$max_cert[2])
  ),
  sprintf(
    "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\",
    f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
    f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
  ),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Note:} Cert./GDP p.c.\\ is the mean across countries of certificates per 1{,}000 USD of GDP per capita (constant 2015 USD). Environment (ISO 14001) data begin in 1999.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "desc_stats.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: grid-summary
#| echo: false
summary_df <- readRDS(file.path(out_dir, "grid_search_summary.rds"))

summary_top5 <- summary_df |>
  dplyr::mutate(
    Panel = dplyr::case_match(Panel,
      "Quality (9001) eu" ~ "Quality",
      "Environment (14001) eu" ~ "Environment",
      .default = Panel
    ),
    BIC_num = suppressWarnings(as.numeric(BIC))
  ) |>
  dplyr::group_by(Panel) |>
  dplyr::slice_min(BIC_num, n = 5) |>
  dplyr::arrange(Panel, BIC_num) |>
  dplyr::ungroup() |>
  dplyr::select(-BIC_num)

tab_top5 <- kbl(summary_top5,
  booktabs = TRUE,
  caption = "Grid search results: five lowest-BIC specifications per standard",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE,
  align = "lrrrccccc",
  format = "latex"
) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 9
  ) |>
  footnote(
    general = "Five specifications per standard with lowest BIC, sorted ascending within panel, out of $\\\\{0.05, 0.10, 0.15, 0.20\\\\}^3 = 64$ grid points. The penalty vector $\\\\lambda = (\\\\lambda_1, \\\\lambda_2, \\\\lambda_1^*)$ governs the elastic-net regularisation in the first stage and the adaptive elastic-net in the second, following De Paula, Rasul \\\\& Souza (2025). $\\\\hat{\\\\rho}_{EN}$, $\\\\hat{\\\\rho}_{Ada}$, $\\\\hat{\\\\rho}_{GMM}$ are the elastic-net, adaptive elastic-net, and post-LASSO GMM estimates of the endogenous peer-effect parameter; the latter is the unpenalised re-estimation on the support recovered by the adaptive step (Caner \\\\& Zhang, 2014). Edges count directed links with $|\\\\hat w_{ij}| > 10^{-5}$. Full grid in Appendix.",
    general_title = "Notes:",
    threeparttable = TRUE,
    escape = FALSE,
    footnote_as_chunk = TRUE
  )

writeLines(as.character(tab_top5), file.path(tab_dir, "grid_search_top5.tex"))
tab_top5
```

## Network recovery

```{r}
#| label: network-stats-table
#| echo: false
#| results: asis
#| cache: true 
net_stats <- function(W) {
  B <- (abs(W) > 1e-5) * 1
  diag(B) <- 0
  g <- igraph::graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = igraph::transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(igraph::degree(g, mode = "out")),
    out_sd      = sd(igraph::degree(g, mode = "out")),
    in_mean     = mean(igraph::degree(g, mode = "in")),
    in_sd       = sd(igraph::degree(g, mode = "in"))
  )
}
s_q <- net_stats(W_q)
s_e <- net_stats(W_e)

B_q <- (abs(W_q) > 1e-5) * 1
B_e <- (abs(W_e) > 1e-5) * 1
common <- sum(B_q & B_e)
only_q <- sum(B_q & !B_e)
only_e <- sum(!B_q & B_e)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

rho_q <- best[["9001_eu"]]$rn$unpenalisedgmm$rho
rho_e <- best[["14001_eu"]]$rn$unpenalisedgmm$rho
lam_q <- best[["9001_eu"]]$lambda
lam_e <- best[["14001_eu"]]$lambda
bic_q <- best[["9001_eu"]]$bic
bic_e <- best[["14001_eu"]]$bic

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{BIC-selected networks: Quality versus Environment (EU, $N = 27$)}",
  "\\label{tab:network_eu}",
  "\\begin{tabular}{lcc}\\hline\\hline",
  "& Quality & Environment \\\\\\hline",
  sprintf("Number of edges & %d & %d \\\\", s_q$edges, s_e$edges),
  sprintf("Edges in both networks & %d & %d \\\\", common, common),
  sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\", only_q),
  sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\", only_e),
  sprintf("Clustering & %s & %s \\\\", f3(s_q$clustering), f3(s_e$clustering)),
  sprintf("Reciprocated edges & %s & %s \\\\", p1(s_q$reciprocity), p1(s_e$reciprocity)),
  "Degree distribution across nodes (countries) & & \\\\",
  sprintf(
    "\\quad Out-degree & %s & %s \\\\",
    dg(s_q$out_mean, s_q$out_sd), dg(s_e$out_mean, s_e$out_sd)
  ),
  sprintf(
    "\\quad In-degree & %s & %s \\\\",
    dg(s_q$in_mean, s_q$in_sd), dg(s_e$in_mean, s_e$in_sd)
  ),
  "\\hline",
  sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\",
    formatC(rho_q, format = "f", digits = 4),
    formatC(rho_e, format = "f", digits = 4)
  ),
  sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\",
    paste(formatC(lam_q, format = "f", digits = 3), collapse = ", "),
    paste(formatC(lam_e, format = "f", digits = 3), collapse = ", ")
  ),
  sprintf(
    "BIC & %s & %s \\\\",
    formatC(bic_q, format = "f", digits = 2),
    formatC(bic_e, format = "f", digits = 2)
  ),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:}
  Top-5 BIC selected networks estimated via grid search. Following de Paula et al. (2024) methodology, a sensitivity analysis of .05 was performed on the penalty vector $\\lambda$ (1.0, 1.0, 1.0) to determine the final network structure
  Reciprocated edges is the frequency of in-edges that are reciprocated by out-edges. The clustering coefficient is the frequency of fully connected triplets over the total number of triplets. The degree distribution counts the average number of connections (sd in parentheses). $\\hat{\\rho}$ (GMM) is the post-LASSO endogenous peer-effect estimate. $\\lambda = (\\lambda_1, \\lambda_2, \\lambda_1^*)$ are the elastic net penalty parameters selected by BIC.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "network_stats.tex"))
cat(tex, sep = "\n")
```

## Distribution of recovered edge weights

```{r}
#| label: fig-weights-density
#| echo: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-cap: "Distribution of nonzero recovered edge weights $|\\hat w_{ij}|$ for the BIC-selected networks. Bars show counts (binwidth = 0.05); rug ticks mark individual edges. Quality and Environment plotted on a common horizontal scale."
w_df <- dplyr::bind_rows(
  data.frame(
    standard = "Quality (ISO 9001)",
    w = abs(W_q[abs(W_q) > 1e-5])
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    w = abs(W_e[abs(W_e) > 1e-5])
  )
) |>
  dplyr::mutate(standard = factor(standard,
    levels = c("Quality (ISO 9001)", "Environment (ISO 14001)")
  ))

p <- ggplot(w_df, aes(x = w, fill = standard)) +
  geom_histogram(
    binwidth = 0.05, boundary = 0,
    colour = "white", linewidth = 0.3
  ) +
  geom_rug(aes(colour = standard),
    sides = "b",
    length = unit(0.04, "npc"), alpha = 0.7
  ) +
  facet_wrap(~standard, scales = "free_y") +
  scale_fill_manual(values = c(
    "Quality (ISO 9001)" = "#3182bd",
    "Environment (ISO 14001)" = "#31a354"
  )) +
  scale_colour_manual(values = c(
    "Quality (ISO 9001)" = "#08519c",
    "Environment (ISO 14001)" = "#006d2c"
  )) +
  scale_x_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.02))) +
  labs(x = expression("|" * hat(w)[ij] * "|"), y = "Count") +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_dir, "weights_density.pdf"), p,
  width = 8, height = 3.5
)
print(p)
```

## Network similarity

The two networks are recovered on the same set of nodes, so the
estimated weight matrices $\hat W^Q, \hat W^E \in \mathbb R^{N\times N}$
are directly comparable element-wise. Define the dyad-level discrepancy
matrix

\begin{equation}
D_{ij} = \hat w^Q_{ij} - \hat w^E_{ij}, \qquad i \neq j,
\label{eq:Dij}
\end{equation}

and summarise it through the **mean absolute discrepancy**

\begin{equation}
\mathrm{MAD}(\hat W^Q, \hat W^E) =
\frac{1}{N(N-1)} \sum_{i\neq j} \bigl| \hat w^Q_{ij} - \hat w^E_{ij} \bigr|,
\label{eq:mad}
\end{equation}

i.e.\ the entry-wise $\ell_1$ distance between the two off-diagonal
weight matrices, normalised by the number of ordered dyads
(Donnet and Robin, 2021). MAD is zero when the two recovered networks
are identical and increases with the average per-dyad weight gap.
A complementary scalar is the Pearson correlation
$\rho(\mathrm{vec}(\hat W^Q), \mathrm{vec}(\hat W^E))$ across
off-diagonal entries, which measures co-movement irrespective of scale.

```{r}
#| label: similarity-stats
#| echo: false
#| results: asis
#| cache: true 
N <- nrow(W_q)
off <- which(row(W_q) != col(W_q))
D <- W_q - W_e
mad <- mean(abs(D[off]))
rmse <- sqrt(mean(D[off]^2))
corr <- cor(W_q[off], W_e[off])
mean_q <- mean(abs(W_q[off]))
mean_e <- mean(abs(W_e[off]))

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Network similarity between $\\hat W^Q$ and $\\hat W^E$ (EU, $N = 27$)}",
  "\\label{tab:similarity}",
  "\\begin{tabular}{lc}\\hline\\hline",
  "Statistic & Value \\\\\\hline",
  sprintf("Mean absolute discrepancy, $\\mathrm{MAD}$ & %s \\\\", formatC(mad, format = "f", digits = 4)),
  sprintf("Root mean squared discrepancy, $\\mathrm{RMSD}$ & %s \\\\", formatC(rmse, format = "f", digits = 4)),
  sprintf("Pearson correlation, $\\rho(\\mathrm{vec}\\hat W^Q, \\mathrm{vec}\\hat W^E)$ & %s \\\\", formatC(corr, format = "f", digits = 4)),
  "\\hline",
  sprintf("Mean $|\\hat w^Q_{ij}|$ (off-diagonal) & %s \\\\", formatC(mean_q, format = "f", digits = 4)),
  sprintf("Mean $|\\hat w^E_{ij}|$ (off-diagonal) & %s \\\\", formatC(mean_e, format = "f", digits = 4)),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} Discrepancy statistics computed on the $N(N-1) = 702$ off-diagonal entries of the BIC-selected weight matrices. MAD reported in equation \\eqref{eq:mad}; RMSD is the analogous $\\ell_2$ statistic. The last two rows give the average absolute weight in each network as a benchmark for interpreting MAD.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "similarity.tex"))
cat(tex, sep = "\n")
```

## Network visualisation

```{r}
#| label: fig-networks-weighted-full
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "..."
make_full_weighted_graph <- function(W, labels, eps = 1e-5) {
  W2 <- W
  W2[abs(W2) < eps] <- 0
  g <- igraph::graph_from_adjacency_matrix(W2, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(W_q, country_labels)
g_e_full <- make_full_weighted_graph(W_e, country_labels)

w_max <- max(c(
  abs(igraph::E(g_q_full)$weight),
  abs(igraph::E(g_e_full)$weight)
), 1e-9)

set.seed(1861)
layout_q <- igraph::layout_with_fr(g_q_full)
set.seed(1861)
layout_e <- igraph::layout_with_fr(g_e_full)
rownames(layout_q) <- rownames(layout_e) <- country_labels

plot_net_gg <- function(g, lo, title, w_max) {
  lo_ordered <- lo[igraph::V(g)$name, , drop = FALSE]

  ggraph(g, layout = lo_ordered) +
    geom_edge_arc(
      aes(
        alpha = abs(weight),
        colour = abs(weight) > 0.3
      ),
      width = 0.5, # spessore costante e sottile
      arrow = arrow(
        length = unit(2, "mm"),
        type = "closed", angle = 22
      ),
      end_cap = circle(5, "mm"),
      start_cap = circle(5, "mm"),
      strength = 0.12
    ) +
    geom_node_point(aes(colour = igraph::degree(g) > 0),
      size = 4.5, show.legend = FALSE
    ) +
    geom_node_text(aes(label = name),
      repel = TRUE,
      size = 2.6,
      bg.colour = "white", bg.r = 0.18,
      segment.colour = "grey55",
      segment.size = 0.25,
      segment.alpha = 0.7,
      max.overlaps = Inf,
      force = 3, force_pull = 0.5,
      point.padding = unit(0.6, "lines"),
      box.padding = unit(0.45, "lines"),
      min.segment.length = 0
    ) +
    scale_edge_alpha(range = c(0.35, 0.9), limits = c(0, w_max), guide = "none") +
    scale_edge_colour_manual(
      values = c(
        `TRUE` = "#67000D", # dark red
        `FALSE` = "grey60"
      ), # light grey
      guide = "none"
    ) +
    scale_colour_manual(values = c(`TRUE` = "steelblue", `FALSE` = "grey85")) +
    ggtitle(title) +
    theme_void(base_size = 11) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "none",
      plot.margin = margin(8, 8, 8, 8)
    )
}

p_qq <- plot_net_gg(g_q_full, layout_q, "Quality — Quality layout", w_max)
p_eq <- plot_net_gg(g_e_full, layout_q, "Environment — Quality layout", w_max)
p_qe <- plot_net_gg(g_q_full, layout_e, "Quality — Environment layout", w_max)
p_ee <- plot_net_gg(g_e_full, layout_e, "Environment — Environment layout", w_max)

panel <- (p_qq | p_eq) / (p_qe | p_ee)

ggsave(file.path(fig_dir, "networks_weighted_full.pdf"), panel,
  width = 13, height = 13
)
print(panel)
```


```{r}
#| label: exogenous-networks-grid
#| echo: false
#| fig-width: 14
#| fig-height: 14
#| fig-cap: "Exogenous networks for the EU 27 sample, constructed following Albuquerque, Bronnenberg and Corbett (2007). For each country, the top-3 nearest neighbors are connected with uniform weight $1/3$. The neighbor-set size is reduced from 5 (used by Albuquerque on a sample of 56 countries) to 3, in light of the smaller sample size ($N = 27$). \\textbf{Geographic}: great-circle distance between capitals. \\textbf{Trade}: largest export shares from BACI 1996, total bilateral trade aggregated across all products. \\textbf{Cultural}: Euclidean distance on Hofstede's four cultural dimensions (PDI, IDV, MAS, UAI), with Cyprus assigned Greece's scores. The networks are directed: $i$'s neighbor set is not symmetric to $j$'s. The 3$\\times$3 grid fixes the Fruchterman--Reingold layout by row (computed on the network on the diagonal); columns then plot all three networks on that row's layout, allowing direct visual comparison of edge structure across networks."
# ============================================================================
# 0. PATHS
# ============================================================================
hofstede_path <- "/workspaces/thesis/data/6-dimensions-for-website-2015-08-16(1).csv"
baci_path <- "/workspaces/thesis/data/BACI_HS92_Y1996_V202601.csv"
ccodes_path <- "/workspaces/thesis/data/country_codes_V202601.csv"

N <- length(country_labels)
TOP_N <- 3

# ============================================================================
# 0bis. ISO3 mapping (country_labels are full names)
# ============================================================================
cl_iso3 <- countrycode::countrycode(
  country_labels,
  origin = "country.name", destination = "iso3c", warn = FALSE
)
names(cl_iso3) <- country_labels
iso_to_label <- setNames(names(cl_iso3), cl_iso3)
stopifnot(!anyNA(cl_iso3))

# ============================================================================
# 1. CAPITALS via maps + countrycode
# ============================================================================
data("world.cities", package = "maps")

capitals <- world.cities |>
  dplyr::filter(capital == 1) |>
  dplyr::mutate(iso3 = countrycode::countrycode(
    country.etc,
    origin = "country.name", destination = "iso3c", warn = FALSE
  )) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::group_by(iso3) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::transmute(country, iso3, lat, lon = long) |>
  dplyr::arrange(match(country, country_labels))

stopifnot(nrow(capitals) == N)

haversine_km <- function(lat1, lon1, lat2, lon2) {
  R <- 6371
  to_rad <- pi / 180
  dlat <- (lat2 - lat1) * to_rad
  dlon <- (lon2 - lon1) * to_rad
  a <- sin(dlat / 2)^2 + cos(lat1 * to_rad) * cos(lat2 * to_rad) * sin(dlon / 2)^2
  2 * R * asin(pmin(1, sqrt(a)))
}

D_geo <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  for (j in seq_len(N)) {
    D_geo[i, j] <- haversine_km(
      capitals$lat[i], capitals$lon[i],
      capitals$lat[j], capitals$lon[j]
    )
  }
}

# ============================================================================
# 2. TRADE — BACI 1996
# ============================================================================
baci <- readr::read_csv(baci_path, show_col_types = FALSE)
ccodes <- readr::read_csv(ccodes_path, show_col_types = FALSE)

iso_map <- ccodes |>
  dplyr::select(country_code, country_iso3) |>
  dplyr::filter(country_iso3 %in% cl_iso3)

trade <- baci |>
  dplyr::select(i, j, v) |>
  dplyr::inner_join(iso_map, by = c("i" = "country_code")) |>
  dplyr::rename(exporter_iso = country_iso3) |>
  dplyr::inner_join(iso_map, by = c("j" = "country_code")) |>
  dplyr::rename(importer_iso = country_iso3) |>
  dplyr::mutate(
    exporter = iso_to_label[exporter_iso],
    importer = iso_to_label[importer_iso]
  ) |>
  dplyr::group_by(exporter, importer) |>
  dplyr::summarise(value = sum(v, na.rm = TRUE), .groups = "drop")

T_trade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(trade))) {
  T_trade[trade$exporter[k], trade$importer[k]] <- trade$value[k]
}

# ============================================================================
# 3. CULTURE — robust to format and missing countries
# ============================================================================
# Try semicolon delimiter first, fall back to comma
hof_raw <- tryCatch(
  {
    x <- readr::read_csv2(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
    if (ncol(x) < 3) stop("too few columns")
    x
  },
  error = function(e) {
    readr::read_csv(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
  }
)

# Hofstede uses "#NULL!" for missing
hof_raw <- hof_raw |>
  dplyr::mutate(dplyr::across(
    dplyr::everything(),
    \(x) ifelse(as.character(x) == "#NULL!", NA, x)
  ))

# Find country identifier column (varies by release)
candidates <- c("country", "ctr", "iso", "iso3")
country_col <- candidates[candidates %in% names(hof_raw)][1]
if (is.na(country_col)) {
  stop(
    "Hofstede CSV has no recognisable country column. Got: ",
    paste(names(hof_raw), collapse = ", ")
  )
}

# Convert identifier to ISO3 with progressive fallbacks
to_iso3 <- function(x) {
  x <- as.character(x)
  out <- suppressWarnings(countrycode::countrycode(
    x,
    origin = "country.name", destination = "iso3c", warn = FALSE
  ))
  if (mean(is.na(out)) > 0.5) {
    out <- suppressWarnings(countrycode::countrycode(
      x,
      origin = "iso2c", destination = "iso3c", warn = FALSE
    ))
  }
  if (mean(is.na(out)) > 0.5) out <- toupper(x) # assume already ISO3
  out
}

hof <- hof_raw |>
  dplyr::mutate(iso3 = to_iso3(.data[[country_col]])) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::transmute(
    iso3,
    pdi = suppressWarnings(as.numeric(pdi)),
    idv = suppressWarnings(as.numeric(idv)),
    mas = suppressWarnings(as.numeric(mas)),
    uai = suppressWarnings(as.numeric(uai))
  ) |>
  dplyr::filter(!is.na(pdi) & !is.na(idv) & !is.na(mas) & !is.na(uai)) |>
  dplyr::distinct(iso3, .keep_all = TRUE)

# Cyprus = Greece (standard practice; absent from Hofstede)
if (!"CYP" %in% hof$iso3 && "GRC" %in% hof$iso3 && "CYP" %in% cl_iso3) {
  hof <- dplyr::bind_rows(
    hof,
    hof |> dplyr::filter(iso3 == "GRC") |> dplyr::mutate(iso3 = "CYP")
  )
}

# Any remaining gaps → EU sample mean (with explicit message)
missing_hof <- setdiff(cl_iso3, hof$iso3)
if (length(missing_hof) > 0) {
  message(
    "Hofstede missing for: ",
    paste(iso_to_label[missing_hof], collapse = ", "),
    " -- imputed with EU sample mean."
  )
  hof_mean <- hof |>
    dplyr::summarise(dplyr::across(
      c(pdi, idv, mas, uai),
      \(x) mean(x, na.rm = TRUE)
    ))
  for (iso in missing_hof) {
    hof <- dplyr::bind_rows(
      hof,
      tibble::tibble(
        iso3 = iso, pdi = hof_mean$pdi, idv = hof_mean$idv,
        mas = hof_mean$mas, uai = hof_mean$uai
      )
    )
  }
}

hof <- hof |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::filter(!is.na(country) & country %in% country_labels) |>
  dplyr::arrange(match(country, country_labels))

stopifnot(nrow(hof) == N)

H_mat <- as.matrix(hof[, c("pdi", "idv", "mas", "uai")])
rownames(H_mat) <- hof$country
D_cul <- as.matrix(dist(H_mat, method = "euclidean"))
D_cul <- D_cul[country_labels, country_labels]

# ============================================================================
# 4. TOP-3 DIRECTED NETWORKS, UNIFORM 1/3 WEIGHTS
# ============================================================================
top_n_network <- function(M, top_n = TOP_N, similarity = FALSE) {
  W <- matrix(0, nrow(M), ncol(M), dimnames = dimnames(M))
  for (i in seq_len(nrow(M))) {
    v <- M[i, ]
    v[i] <- if (similarity) -Inf else Inf
    idx <- if (similarity) {
      order(v, decreasing = TRUE)[seq_len(top_n)]
    } else {
      order(v, decreasing = FALSE)[seq_len(top_n)]
    }
    W[i, idx] <- 1 / top_n
  }
  W
}

W_geo <- top_n_network(D_geo, similarity = FALSE)
W_trade <- top_n_network(T_trade, similarity = TRUE)
W_cul <- top_n_network(D_cul, similarity = FALSE)

saveRDS(
  list(geo = W_geo, trade = W_trade, culture = W_cul),
  file.path(out_dir, "exogenous_networks.rds")
)

# ============================================================================
# 5. 3x3 GRID PLOT, ROW-FIXED FR LAYOUTS
# ============================================================================
to_graph <- function(W) {
  g <- igraph::graph_from_adjacency_matrix(W, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- rownames(W)
  g
}
g_geo <- to_graph(W_geo)
g_trade <- to_graph(W_trade)
g_cul <- to_graph(W_cul)

set.seed(1861)
lo_geo <- igraph::layout_with_fr(g_geo)
set.seed(1861)
lo_trade <- igraph::layout_with_fr(g_trade)
set.seed(1861)
lo_cul <- igraph::layout_with_fr(g_cul)
rownames(lo_geo) <- rownames(lo_trade) <- rownames(lo_cul) <- country_labels

plot_one <- function(g, lo, title) {
  lo_ord <- lo[igraph::V(g)$name, , drop = FALSE]
  ggraph(g, layout = lo_ord) +
    geom_edge_arc(
      width = 0.4, alpha = 0.55, colour = "grey40",
      arrow = arrow(length = unit(1.8, "mm"), type = "closed", angle = 22),
      end_cap = circle(4, "mm"),
      start_cap = circle(4, "mm"),
      strength = 0.12
    ) +
    geom_node_point(size = 3.5, colour = "steelblue") +
    geom_node_text(aes(label = name),
      size = 2.4, repel = TRUE,
      bg.colour = "white", bg.r = 0.18,
      segment.colour = "grey55", segment.size = 0.25,
      max.overlaps = Inf, force = 3, force_pull = 0.5,
      point.padding = unit(0.5, "lines"),
      box.padding = unit(0.35, "lines"),
      min.segment.length = 0
    ) +
    ggtitle(title) +
    theme_void(base_size = 10) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.margin = margin(6, 6, 6, 6)
    )
}

p11 <- plot_one(g_geo, lo_geo, "Geography -- geo layout")
p12 <- plot_one(g_trade, lo_geo, "Trade -- geo layout")
p13 <- plot_one(g_cul, lo_geo, "Culture -- geo layout")
p21 <- plot_one(g_geo, lo_trade, "Geography -- trade layout")
p22 <- plot_one(g_trade, lo_trade, "Trade -- trade layout")
p23 <- plot_one(g_cul, lo_trade, "Culture -- trade layout")
p31 <- plot_one(g_geo, lo_cul, "Geography -- culture layout")
p32 <- plot_one(g_trade, lo_cul, "Trade -- culture layout")
p33 <- plot_one(g_cul, lo_cul, "Culture -- culture layout")

panel_exo <- (p11 | p12 | p13) /
  (p21 | p22 | p23) /
  (p31 | p32 | p33)

ggsave(file.path(fig_dir, "exogenous_networks_grid.pdf"), panel_exo,
  width = 14, height = 14, device = cairo_pdf
)
print(panel_exo)
```

```{r}
#| label: bdf-final-summary
#| echo: false
#| results: asis
#| cache: true
# Self-contained chunk: BDF (2009) G2SLS with one-year lag on the peer outcome,
# applied to the De Paula, Rasul & Souza (2025) recovered networks and the
# Albuquerque, Bronnenberg & Corbett (2007, Eq. 9) combined exogenous network.
# Two specifications: (i) no FE, asymptotic country-clustered SE;
# (ii) country FE, pairs cluster bootstrap SE (Cameron & Miller 2015).

library(fixest)
library(dplyr)
library(tidyr)
library(kableExtra)

set.seed(1861)
B <- 999

# ---- Helpers ---------------------------------------------------------------
row_normalise <- function(W, eps = 1e-5) {
  W[abs(W) < eps] <- 0
  diag(W) <- 0
  rs <- rowSums(W)
  rs[rs == 0] <- 1
  W / rs
}

attach_y <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |> dplyr::transmute(country, year, y = log1p(n_cert)),
      by = c("country", "year")
    ) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

keep_full <- function(d) {
  full <- d |>
    dplyr::count(year) |>
    dplyr::filter(n == length(country_labels)) |>
    dplyr::pull(year)
  dplyr::filter(d, year %in% full)
}

build_bdf_panel_lag <- function(panel, W, x_vars) {
  G <- row_normalise(W)
  G2 <- G %*% G
  G3 <- G2 %*% G
  d <- panel[order(panel$year, match(panel$country, rownames(G))), ]
  yrs <- sort(unique(d$year))
  d$Gy <- unlist(lapply(yrs, function(t) {
    yt <- d$y[d$year == t]
    as.numeric(G %*% yt)
  }))
  for (v in x_vars) {
    Xt <- matrix(d[[v]], ncol = length(yrs))
    d[[paste0("G_", v)]] <- as.numeric(G %*% Xt)
    d[[paste0("G2_", v)]] <- as.numeric(G2 %*% Xt)
    d[[paste0("G3_", v)]] <- as.numeric(G3 %*% Xt)
  }
  d <- d[order(d$country, d$year), ]
  d <- d |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      Gy_lag = dplyr::lag(Gy, n = 1, order_by = year),
      dplyr::across(
        dplyr::all_of(c(paste0("G2_", x_vars), paste0("G3_", x_vars))),
        \(z) dplyr::lag(z, n = 1, order_by = year),
        .names = "{.col}_lag"
      )
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(Gy_lag)
  d
}

cluster_boot <- function(data, fit_formula, cluster_var = "country", B = 999) {
  clusters <- unique(data[[cluster_var]])
  G <- length(clusters)
  est0 <- coef(fixest::feols(fit_formula, data = data, warn = FALSE, notes = FALSE))
  k <- length(est0)
  boot_mat <- matrix(NA_real_, nrow = B, ncol = k, dimnames = list(NULL, names(est0)))
  for (b in seq_len(B)) {
    idx <- sample(clusters, G, replace = TRUE)
    boot_df <- do.call(rbind, lapply(seq_along(idx), function(j) {
      d <- data[data[[cluster_var]] == idx[j], ]
      d[[cluster_var]] <- paste0(idx[j], "_", j)
      d
    }))
    fit_b <- tryCatch(
      fixest::feols(fit_formula, data = boot_df, warn = FALSE, notes = FALSE),
      error = function(e) NULL
    )
    if (!is.null(fit_b)) boot_mat[b, ] <- coef(fit_b)
  }
  list(
    estimate = est0,
    se = apply(boot_mat, 2, sd, na.rm = TRUE),
    p_val = sapply(seq_len(k), function(j) {
      mean(abs(boot_mat[, j] - est0[j]) >= abs(est0[j]), na.rm = TRUE)
    })
  )
}

sig_stars <- function(p) {
  if (is.na(p)) {
    return("")
  }
  if (p < 0.01) {
    "$^{***}$"
  } else if (p < 0.05) {
    "$^{**}$"
  } else if (p < 0.10) "$^{*}$" else ""
}

# ---- Build data ------------------------------------------------------------
id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(attach_y(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001))
panel_e <- keep_full(attach_y(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001))

# Combined ABC network
W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0), which(W_cul[i, ] > 0)))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

# BDF panels with one-year lag
bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

# ---- Formulas --------------------------------------------------------------
rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2_lag <- paste(paste0("G2_", x_vars, "_lag"), collapse = " + ")
inst3_lag <- paste(paste0("G3_", x_vars, "_lag"), collapse = " + ")

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE: asymptotic country-clustered SE ----------------------------
extract_nofe <- function(fit) {
  s <- summary(fit, cluster = ~country)
  ct <- s$coeftable
  i <- which(rownames(ct) == "fit_Gy_lag")
  list(b = ct[i, "Estimate"], se = ct[i, "Std. Error"], p = ct[i, "Pr(>|t|)"])
}

nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# ---- (ii) Country FE: pairs cluster bootstrap ------------------------------
boot_q_rec <- cluster_boot(bdf_q_lag, f_cfe, "country", B = B)
boot_q_abc <- cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B)
boot_e_rec <- cluster_boot(bdf_e_lag, f_cfe, "country", B = B)
boot_e_abc <- cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B)

extract_cfe <- function(b) {
  i <- which(names(b$estimate) == "fit_Gy_lag")
  list(b = b$estimate[i], se = b$se[i], p = b$p_val[i])
}

cfe_q_rec <- extract_cfe(boot_q_rec)
cfe_q_abc <- extract_cfe(boot_q_abc)
cfe_e_rec <- extract_cfe(boot_e_rec)
cfe_e_abc <- extract_cfe(boot_e_abc)

# ---- Build table -----------------------------------------------------------
fmt_cell <- function(b, se, p) {
  c(sprintf("%.3f%s", b, sig_stars(p)), sprintf("(%.3f)", se))
}

# ---- Build table (estimate above, SE below) --------------------------------
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = sprintf("Endogenous peer effect $\\hat\\beta$ from Bramoull\\'e, Djebbari \\& Fortin (2009) G2SLS with one-year lag on the peer outcome. Comparison between the De Paula, Rasul \\& Souza (2025) recovered networks and the combined exogenous network of Albuquerque, Bronnenberg \\& Corbett (2007, Eq. 9). The no-FE specification uses asymptotic country-clustered standard errors; the country-FE specification uses pairs cluster bootstrap on country, $B = %d$ replications (Cameron \\& Miller 2015, J. Human Resources). Standard errors in parentheses below estimates.", B),
  label = "tab:bdf_final_summary"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10) |>
  kableExtra::footnote(
    general = "Diffusion timing follows Albuquerque, Bronnenberg \\\\& Corbett (2007, Eq. 1) and Bass (1969). The combined exogenous network is the union of top-3 nearest neighbours from geography, bilateral trade, and Hofstede cultural distance, with uniform row-stochastic weights. Stability of $\\\\hat\\\\beta$ across the two specifications is diagnostic of correlated-effect contamination. $^{*}p<0.10$, $^{**}p<0.05$, $^{***}p<0.01$.",
    general_title = "Notes:", threeparttable = TRUE,
    escape = FALSE, footnote_as_chunk = TRUE
  )
```

```{r}
#| label: bdf-final-summary-twfe-countrycluster
#| echo: false
#| results: asis
# BDF (2009) G2SLS with one-year lag, country + year FE.
# Cluster on country (Cameron & Miller 2015): even with country FE,
# residual within-country serial correlation may persist (idiosyncratic
# national cycles, sectoral waves, regulatory transitions). Country
# clustering is the standard adjustment for this residual dependence.
# On N = 27 clusters, the bootstrap is at the lower end of the recommended
# regime but more reliable than year clustering on T = 18-24 clusters
# (MacKinnon, Nielsen & Webb 2023, J. Econometrics).
#
# Reuses bdf_q_lag, bdf_e_lag, bdf_comb_q_lag, bdf_comb_e_lag,
# rhs, exo, inst2_lag, inst3_lag, cluster_boot, sig_stars, extract_cfe
# from previous chunks.

set.seed(1861)
B <- 999

f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# Pairs cluster bootstrap on country
boot_q_rec_tw <- cluster_boot(bdf_q_lag, f_twfe, "country", B = B)
boot_q_abc_tw <- cluster_boot(bdf_comb_q_lag, f_twfe, "country", B = B)
boot_e_rec_tw <- cluster_boot(bdf_e_lag, f_twfe, "country", B = B)
boot_e_abc_tw <- cluster_boot(bdf_comb_e_lag, f_twfe, "country", B = B)

tw_q_rec <- extract_cfe(boot_q_rec_tw)
tw_q_abc <- extract_cfe(boot_q_abc_tw)
tw_e_rec <- extract_cfe(boot_e_rec_tw)
tw_e_abc <- extract_cfe(boot_e_abc_tw)

fmt_cell <- function(b, se, p) {
  c(sprintf("%.3f%s", b, sig_stars(p)), sprintf("(%.3f)", se))
}

tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", tw_q_rec$b, sig_stars(tw_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", tw_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", tw_q_abc$b, sig_stars(tw_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", tw_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", tw_e_rec$b, sig_stars(tw_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", tw_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", tw_e_abc$b, sig_stars(tw_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", tw_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (country + year FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llr", row.names = FALSE,
  caption = sprintf("Endogenous peer effect $\\hat\\beta$ from Bramoull\\'e, Djebbari \\& Fortin (2009) G2SLS with one-year lag on the peer outcome and country + year fixed effects. Inference based on pairs cluster bootstrap on country, $B = %d$ replications (Cameron \\& Miller 2015, J. Human Resources). Standard errors in parentheses below estimates.", B),
  label = "tab:bdf_twfe_countrycluster"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10) |>
  kableExtra::footnote(
    general = "Diffusion timing follows Albuquerque, Bronnenberg \\\\& Corbett (2007, Eq. 1) and Bass (1969). Country fixed effects absorb persistent national heterogeneity; year fixed effects absorb common EU-wide temporal shocks. For ISO 14001, observed predominantly in its take-off phase, year FE may absorb part of the diffusion-driven temporal commonality alongside common shocks (Bertrand, Duflo \\\\& Mullainathan 2004, QJE; Neumayer \\\\& Perkins 2004, Environment and Planning A). $^{*}p<0.10$, $^{**}p<0.05$, $^{***}p<0.01$.",
    general_title = "Notes:", threeparttable = TRUE,
    escape = FALSE, footnote_as_chunk = TRUE
  )
```








```{r}
#| label: bdf-final-summary-abc-model5
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS with one-year lag, applied to Albuquerque, Bronnenberg
# & Corbett (2007, Table 3, Model 5) preferred two-channel networks:
# - Quality (ISO 9001):     geography + bilateral trade
# - Environment (ISO 14001): geography + cultural similarity
#
# Two specifications:
# (i)  no FE, asymptotic country-clustered SE
# (ii) country + year FE, pairs cluster bootstrap on country (B = 999)
#
# Reuses panel_q, panel_e, x_vars, rhs, exo, inst2_lag, inst3_lag,
# build_bdf_panel_lag, cluster_boot, sig_stars, extract_cfe, extract_nofe,
# W_geo, W_trade, W_cul from previous chunks.

set.seed(1861)
B <- 999
N <- length(country_labels)

# ---- Build ABC Model 5 networks --------------------------------------------
# Quality: union of top-3 from geography + bilateral trade
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs_i] <- 1 / length(nbrs_i)
}

# Environment: union of top-3 from geography + Hofstede culture
W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs_i] <- 1 / length(nbrs_i)
}

# ---- Build BDF panels with one-year lag ------------------------------------
bdf_geotrade_q_lag <- build_bdf_panel_lag(panel_q, W_geotrade, x_vars)
bdf_geoculture_e_lag <- build_bdf_panel_lag(panel_e, W_geoculture, x_vars)

# ---- Formulas --------------------------------------------------------------
f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE: asymptotic country-clustered SE ----------------------------
nofe_q_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geotrade_q_lag, cluster = ~country))
nofe_e_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geoculture_e_lag, cluster = ~country))

# ---- (ii) Two-way FE: pairs cluster bootstrap on country -------------------
boot_q_abc5_tw <- cluster_boot(bdf_geotrade_q_lag, f_twfe, "country", B = B)
boot_e_abc5_tw <- cluster_boot(bdf_geoculture_e_lag, f_twfe, "country", B = B)

twfe_q_abc5 <- extract_cfe(boot_q_abc5_tw)
twfe_e_abc5 <- extract_cfe(boot_e_abc5_tw)

# ---- Build table -----------------------------------------------------------
fmt_cell <- function(b, se, p) {
  c(sprintf("%.3f%s", b, sig_stars(p)), sprintf("(%.3f)", se))
}

# ---- Build table (estimate above, SE below) --------------------------------
tab <- rbind(
  c(
    "Quality", "Exogenous: Geo + Trade",
    sprintf("%.3f%s", nofe_q_abc5$b, sig_stars(nofe_q_abc5$p)),
    sprintf("%.3f%s", twfe_q_abc5$b, sig_stars(twfe_q_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc5$se),
    sprintf("(%.3f)", twfe_q_abc5$se)
  ),
  c(
    "Environment", "Exogenous: Geo + Culture",
    sprintf("%.3f%s", nofe_e_abc5$b, sig_stars(nofe_e_abc5$p)),
    sprintf("%.3f%s", twfe_e_abc5$b, sig_stars(twfe_e_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc5$se),
    sprintf("(%.3f)", twfe_e_abc5$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country + year FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = sprintf("Endogenous peer effect $\\hat\\beta$ from Bramoull\\'e, Djebbari \\& Fortin (2009) G2SLS with one-year lag on the peer outcome, applied to the preferred two-channel exogenous networks of Albuquerque, Bronnenberg \\& Corbett (2007, Table 3, Model 5). The no-FE specification uses asymptotic country-clustered standard errors; the country + year FE specification uses pairs cluster bootstrap on country, $B = %d$ replications (Cameron \\& Miller 2015, J. Human Resources). Standard errors in parentheses below estimates.", B),
  label = "tab:bdf_abc_model5"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10) |>
  kableExtra::footnote(
    general = "Each network is the union of two top-3 nearest neighbour sets, with uniform row-stochastic weights. Diffusion timing follows Albuquerque et al.\\\\ (2007, Eq. 1) and Bass (1969). Country FE absorb time-invariant national heterogeneity; year FE absorb common EU-wide shocks. $|\\\\beta| < 1$ is required for SAR equilibrium (Lee 2007). $^{*}p<0.10$, $^{**}p<0.05$, $^{***}p<0.01$.",
    general_title = "Notes:", threeparttable = TRUE,
    escape = FALSE, footnote_as_chunk = TRUE
  )
```






































































```{r}
#| label: grid-summary-appendix
#| echo: false
#| output: false
summary_df <- readRDS(file.path(out_dir, "grid_search_summary.rds"))

tab_full <- kbl(summary_df,
  booktabs = TRUE,
  longtable = TRUE,
  caption = "Grid search results",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE,
  align = "lrrrccccc",
  format = "latex"
) |>
  kable_styling(
    latex_options = c("hold_position", "repeat_header"),
    font_size = 8
  ) |>
  footnote(
    general = "Penalty triplet $(\\\\lambda_1, \\\\lambda_2, \\\\lambda_1^*)$ from De Paula, Rasul \\\\& Souza (2025). $\\\\hat{\\\\rho}_{EN}$, $\\\\hat{\\\\rho}_{Ada}$, $\\\\hat{\\\\rho}_{GMM}$ are the elastic-net, adaptive elastic-net, and post-LASSO GMM estimates of the peer-effect parameter. BIC selects the preferred specification per panel.",
    general_title = "Notes:",
    threeparttable = TRUE,
    escape = FALSE,
    footnote_as_chunk = TRUE
  )

writeLines(as.character(tab_full), file.path(tab_dir, "grid_search_full.tex"))
tab_full
```

