---
title: "Thesis Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "results.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
data_dir <- "/workspaces/thesis/data"
clean_dir <- "/workspaces/thesis/data/cleaned_data"

# install.packages(
# c("maps", "countrycode")
# )

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)
library(patchwork)
library(ggraph)
library(ggrepel)
library(tidygraph)
library(maps)
library(countrycode)
library(fixest)


set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
fig_dir <- file.path(out_dir, "fig")
tab_dir <- file.path(out_dir, "tab")
for (d in c(fig_dir, tab_dir)) if (!dir.exists(d)) dir.create(d, recursive = TRUE)
```

```{r}
#| label: load-best
#| include: false
#| cache: true
best <- readRDS(file.path(out_dir, "best_per_panel.rds"))
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
diag(W_q) <- 0
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)
diag(W_e) <- 0

sync_eu <- readRDS(file.path(clean_dir, "sync_eu.rds"))
country_labels <- sort(sync_eu)
rownames(W_q) <- colnames(W_q) <- country_labels
rownames(W_e) <- colnames(W_e) <- country_labels
```

# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Descriptive Statistics

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
#| cache: true 
f9001 <- readRDS(file.path(clean_dir, "f9001.rds"))
f14001 <- readRDS(file.path(clean_dir, "f14001.rds"))

desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, countries, years) {
  df |>
    dplyr::filter(country %in% countries, year %in% years) |>
    dplyr::group_by(year) |>
    dplyr::summarise(
      total_cert     = sum(n_cert, na.rm = TRUE),
      mean_cert      = mean(n_cert, na.rm = TRUE),
      sd_cert        = sd(n_cert, na.rm = TRUE),
      min_cert       = min(n_cert, na.rm = TRUE),
      max_cert       = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries    = dplyr::n(),
      .groups        = "drop"
    )
}

d9 <- desc_tab(f9001, country_labels, desc_years)
d14 <- desc_tab(f14001, country_labels, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Descriptive statistics}",
  "\\label{tab:desc_stats}",
  "\\begin{tabular}{lcccccc}\\hline\\hline",
  "& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\",
  "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
  "& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\\hline",
  sprintf(
    "N countries & %d & %d & %d & & %d & %d \\\\",
    d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
    d14$n_countries[1], d14$n_countries[2]
  ),
  sprintf(
    "Total & %s & %s & %s & & %s & %s \\\\",
    f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
    f0(d14$total_cert[1]), f0(d14$total_cert[2])
  ),
  sprintf(
    "Mean & %s & %s & %s & & %s & %s \\\\",
    f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
    f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
  ),
  sprintf(
    "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\",
    f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
    f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
  ),
  sprintf(
    "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\",
    f0(d9$min_cert[1]), f0(d9$max_cert[1]),
    f0(d9$min_cert[2]), f0(d9$max_cert[2]),
    f0(d9$min_cert[3]), f0(d9$max_cert[3]),
    f0(d14$min_cert[1]), f0(d14$max_cert[1]),
    f0(d14$min_cert[2]), f0(d14$max_cert[2])
  ),
  sprintf(
    "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\",
    f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
    f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
  ),
  "\\hline\\hline\\end{tabular}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "desc_stats.tex"))
cat(tex, sep = "\n")
```

## Network Recovery and Dyadic Analysis

\subsection*{Methodology}

De Paula, Rasul \& Souza (2025) extend the peer-effects literature (Manski 1993; Bramoullé, Djebbari \& Fortin 2009) to the case in which the network of interactions is unknown. Their structural model
\begin{equation}
y_{it} = \rho_0 \sum_{j=1}^{N} W_{0,ij}\, y_{jt} + \beta_0 x_{it} + \gamma_0 \sum_{j=1}^{N} W_{0,ij}\, x_{jt} + \alpha_i + \alpha_t + \varepsilon_{it}
\label{eq:structural}
\end{equation}
admits the compact form $y_t = \rho_0 W_0 y_t + \beta_0 x_t + \gamma_0 W_0 x_t + \varepsilon_t$, which rearranges into the reduced form
\begin{equation}
y_t = \Pi_0 x_t + \nu_t, \qquad \Pi_0 = (I - \rho_0 W_0)^{-1}(\beta_0 I + \gamma_0 W_0).
\label{eq:reduced}
\end{equation}
Under their six assumptions (A1--A6, see Appendix~\ref{app:assumptions}), the structural map $(W_0,\rho_0,\beta_0,\gamma_0) \mapsto \Pi_0$ is one-to-one generically: since $\Pi_0$ and the covariates $x_t$ are observable, all structural unknowns can be recovered jointly by inverting this map.

The 3-node example on p.~2700 of De Paula, Rasul \& Souza (2025) makes the intuition concrete. Suppose the reduced-form matrix is estimated as
\begin{equation}
\Pi_0 = \frac{1}{455}
\begin{pmatrix}
275 & 310 & 0 \\
310 & 275 & 0 \\
0 & 0 & 182
\end{pmatrix}.
\label{eq:pihat}
\end{equation}
The third row and column are zero off the diagonal, so under (A3) individual 3 is isolated --- her covariate does not affect anyone else's outcome and vice versa. Individuals 1 and 2 cannot be isolated, since their covariates are correlated with the other's outcome (A5); the row-sum normalisation (A4) combined with (A1) forces $(W_0)_{12}=1$, and symmetry of $\Pi_0$ then gives $(W_0)_{21}=1$. The latent network is fully recovered, and the structural parameters $(\beta_0,\rho_0,\gamma_0)$ follow mechanically from the matrix identity $(I-\rho_0 W_0)\Pi_0 = \beta_0 I + \gamma_0 W_0$.

The takeaway is that, once $\Pi_0$ is in hand, the zero pattern of the reduced form identifies the topology of the latent network, while the magnitudes pin down the structural parameters. This is the result that justifies treating the recovered network as a primitive object in the second-stage analysis below: it is identified from observable cross-sectional spillovers under transparent assumptions, without requiring any a priori commitment to a specific weighting scheme (geographic, trade, cultural).


```{r}
#| label: network-stats-table
#| echo: false
#| results: asis
#| cache: true 
net_stats <- function(W) {
  B <- (abs(W) > 1e-5) * 1
  diag(B) <- 0
  g <- igraph::graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = igraph::transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(igraph::degree(g, mode = "out")),
    out_sd      = sd(igraph::degree(g, mode = "out")),
    in_mean     = mean(igraph::degree(g, mode = "in")),
    in_sd       = sd(igraph::degree(g, mode = "in"))
  )
}
s_q <- net_stats(W_q)
s_e <- net_stats(W_e)

B_q <- (abs(W_q) > 1e-5) * 1
B_e <- (abs(W_e) > 1e-5) * 1
common <- sum(B_q & B_e)
only_q <- sum(B_q & !B_e)
only_e <- sum(!B_q & B_e)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

rho_q <- best[["9001_eu"]]$rn$unpenalisedgmm$rho
rho_e <- best[["14001_eu"]]$rn$unpenalisedgmm$rho
lam_q <- best[["9001_eu"]]$lambda
lam_e <- best[["14001_eu"]]$lambda
bic_q <- best[["9001_eu"]]$bic
bic_e <- best[["14001_eu"]]$bic

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{BIC-selected networks}",
  "\\label{tab:network_eu}",
  "\\begin{tabular}{lcc}\\hline\\hline",
  "& Quality & Environment \\\\\\hline",
  sprintf("Number of edges & %d & %d \\\\", s_q$edges, s_e$edges),
  sprintf("Edges in both networks & %d & %d \\\\", common, common),
  sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\", only_q),
  sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\", only_e),
  sprintf("Clustering & %s & %s \\\\", f3(s_q$clustering), f3(s_e$clustering)),
  sprintf("Reciprocated edges & %s & %s \\\\", p1(s_q$reciprocity), p1(s_e$reciprocity)),
  "Degree distribution across nodes (countries) & & \\\\",
  sprintf(
    "\\quad Out-degree & %s & %s \\\\",
    dg(s_q$out_mean, s_q$out_sd), dg(s_e$out_mean, s_e$out_sd)
  ),
  sprintf(
    "\\quad In-degree & %s & %s \\\\",
    dg(s_q$in_mean, s_q$in_sd), dg(s_e$in_mean, s_e$in_sd)
  ),
  "\\hline",
  sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\",
    formatC(rho_q, format = "f", digits = 4),
    formatC(rho_e, format = "f", digits = 4)
  ),
  sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\",
    paste(formatC(lam_q, format = "f", digits = 3), collapse = ", "),
    paste(formatC(lam_e, format = "f", digits = 3), collapse = ", ")
  ),
  sprintf(
    "BIC & %s & %s \\\\",
    formatC(bic_q, format = "f", digits = 2),
    formatC(bic_e, format = "f", digits = 2)
  ),
  "\\hline\\hline\\end{tabular}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "network_stats.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: fig-degree-hist
#| echo: false
#| fig-width: 9
#| fig-height: 3.5
#| fig-cap: "Unweighted degree distributions of recovered networks"
eps <- 1e-5
# Use signed threshold: edges are entries with W_ij > eps.
# Negative entries (if any) are NOT counted as edges.
B_q <- (W_q > eps) * 1
diag(B_q) <- 0
B_e <- (W_e > eps) * 1
diag(B_e) <- 0

# Quick check: report if any negative entries exist in the recovered W
cat(sprintf(
  "W_q: %d negative off-diagonal entries (min = %.4f)\n",
  sum(W_q < -eps & row(W_q) != col(W_q)),
  min(W_q[row(W_q) != col(W_q)])
))
cat(sprintf(
  "W_e: %d negative off-diagonal entries (min = %.4f)\n",
  sum(W_e < -eps & row(W_e) != col(W_e)),
  min(W_e[row(W_e) != col(W_e)])
))

deg_df <- dplyr::bind_rows(
  data.frame(
    standard = "Quality (ISO 9001)",
    direction = "Out-degree", value = rowSums(B_q)
  ),
  data.frame(
    standard = "Quality (ISO 9001)",
    direction = "In-degree", value = colSums(B_q)
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    direction = "Out-degree", value = rowSums(B_e)
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    direction = "In-degree", value = colSums(B_e)
  )
) |>
  dplyr::mutate(
    standard = factor(standard,
      levels = c("Quality (ISO 9001)", "Environment (ISO 14001)")
    ),
    direction = factor(direction, levels = c("Out-degree", "In-degree"))
  )

p <- ggplot(deg_df, aes(x = value, fill = standard)) +
  geom_histogram(
    position = "identity", alpha = 0.55,
    colour = "white", linewidth = 0.3, bins = 12
  ) +
  geom_rug(aes(colour = standard),
    sides = "b",
    length = unit(0.03, "npc"), alpha = 0.7
  ) +
  facet_wrap(~direction, scales = "free", ncol = 2) +
  scale_fill_manual(values = c(
    "Quality (ISO 9001)"      = "#3182bd",
    "Environment (ISO 14001)" = "#31a354"
  )) +
  scale_colour_manual(values = c(
    "Quality (ISO 9001)"      = "#08519c",
    "Environment (ISO 14001)" = "#006d2c"
  )) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02))) +
  labs(x = "Degree", y = "Count", fill = NULL, colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = "bottom",
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(fig_dir, "degree_hist.pdf"), p, width = 9, height = 3.5)
print(p)
```


```{r}
#| label: jaccard-compute
#| include: false
eps <- 1e-5
B_q <- (abs(W_q) > eps) * 1
B_e <- (abs(W_e) > eps) * 1
diag(B_q) <- diag(B_e) <- 0

n_q <- sum(B_q)
n_e <- sum(B_e)
n_inter <- sum(B_q & B_e)
n_union <- sum(B_q | B_e)
jaccard <- n_inter / n_union

# Pre-formatted strings to insert inline in the text
jaccard_pct <- formatC(100 * jaccard, format = "f", digits = 1)
jaccard_val <- formatC(jaccard, format = "f", digits = 3)
```

### Network similarity

To assess how closely the two recovered networks $\hat W^Q$ and $\hat W^E$ overlap, we summarise their structural similarity through the **Jaccard index** (Jaccard 1901; Wasserman \& Faust 1994, Sec.\ 11.4), the share of directed edges shared by both networks relative to the union of their edge sets:

\begin{equation}
J(\hat W^Q, \hat W^E) = \frac{|E_Q \cap E_E|}{|E_Q \cup E_E|},
\label{eq:jaccard}
\end{equation}

where $E_Q = \{(i, j) : |\hat w^Q_{ij}| > 10^{-5}\}$ and $E_E$ is defined analogously. Edges are treated as directed, so that a link $i \to j$ and the reciprocal $j \to i$ count as distinct elements of the edge sets. The index ranges in $[0, 1]$, with $J = 0$ for disjoint networks (no edges in common) and $J = 1$ for identical edge sets.

For the BIC-selected recovered networks, $|E_Q| = `r n_q`$, $|E_E| = `r n_e`$, $|E_Q \cap E_E| = `r n_inter`$, and $|E_Q \cup E_E| = `r n_union`$, yielding $J = `r jaccard_val`$.


### Network visualisation

```{r}
#| label: fig-networks-weighted-full
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Recovered networks"
make_full_weighted_graph <- function(W, labels, eps = 1e-5) {
  W2 <- W
  W2[abs(W2) < eps] <- 0
  g <- igraph::graph_from_adjacency_matrix(W2, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(W_q, country_labels)
g_e_full <- make_full_weighted_graph(W_e, country_labels)

w_max <- max(c(
  abs(igraph::E(g_q_full)$weight),
  abs(igraph::E(g_e_full)$weight)
), 1e-9)

set.seed(1861)
layout_q <- igraph::layout_with_fr(g_q_full)
set.seed(1861)
layout_e <- igraph::layout_with_fr(g_e_full)
rownames(layout_q) <- rownames(layout_e) <- country_labels

plot_net_gg <- function(g, lo, title, w_max) {
  lo_ordered <- lo[igraph::V(g)$name, , drop = FALSE]

  ggraph(g, layout = lo_ordered) +
    geom_edge_arc(
      aes(
        alpha = abs(weight),
        colour = abs(weight) > 0.9
      ),
      width = 0.5,
      arrow = arrow(
        length = unit(2, "mm"),
        type = "closed", angle = 22
      ),
      end_cap = circle(5, "mm"),
      start_cap = circle(5, "mm"),
      strength = 0.12
    ) +
    geom_node_point(aes(colour = igraph::degree(g) > 0),
      size = 4.5, show.legend = FALSE
    ) +
    geom_node_text(aes(label = name),
      repel = TRUE,
      size = 2.6,
      bg.colour = "white", bg.r = 0.18,
      segment.colour = "grey55",
      segment.size = 0.25,
      segment.alpha = 0.7,
      max.overlaps = Inf,
      force = 3, force_pull = 0.5,
      point.padding = unit(0.6, "lines"),
      box.padding = unit(0.45, "lines"),
      min.segment.length = 0
    ) +
    scale_edge_alpha(range = c(0.35, 0.9), limits = c(0, w_max), guide = "none") +
    scale_edge_colour_manual(
      values = c(
        `TRUE` = "#67000D", # dark red
        `FALSE` = "grey60"
      ), # light grey
      guide = "none"
    ) +
    scale_colour_manual(values = c(`TRUE` = "steelblue", `FALSE` = "grey85")) +
    ggtitle(title) +
    theme_void(base_size = 11) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "none",
      plot.margin = margin(8, 8, 8, 8)
    )
}

p_qq <- plot_net_gg(g_q_full, layout_q, "Quality — Quality layout", w_max)
p_eq <- plot_net_gg(g_e_full, layout_q, "Environment — Quality layout", w_max)
p_qe <- plot_net_gg(g_q_full, layout_e, "Quality — Environment layout", w_max)
p_ee <- plot_net_gg(g_e_full, layout_e, "Environment — Environment layout", w_max)

panel <- (p_qq | p_eq) / (p_qe | p_ee)

ggsave(file.path(fig_dir, "networks_weighted_full.pdf"), panel,
  width = 13, height = 13
)
print(panel)
```

### Dyadic regression

```{r}
#| label: dyadic-correlation-check
#| echo: false
#| results: asis
#| cache: true
#| include: false
# Pairwise correlation matrix among candidate dyadic homophily covariates,
# computed on the N(N-1) = 702 ordered dyads.

# Helper: ISO2 -> ISO3 with Eurostat quirks
iso2_to_iso3 <- function(x) {
  x <- as.character(x)
  x <- dplyr::case_match(x, "EL" ~ "GR", "UK" ~ "GB", .default = x)
  countrycode::countrycode(x,
    origin = "iso2c",
    destination = "iso3c", warn = FALSE
  )
}

# We need cl_iso3 / iso_to_label / N / D_geo / D_cul which are built later in
# the application chunk. To keep ordering, we replicate the minimal mapping
# needed here (same code as in the exogenous-networks-grid chunk).
N <- length(country_labels)
cl_iso3 <- countrycode::countrycode(
  country_labels,
  origin = "country.name", destination = "iso3c", warn = FALSE
)
names(cl_iso3) <- country_labels
iso_to_label <- setNames(names(cl_iso3), cl_iso3)

# Build D_geo (capitals + haversine) — same logic as later chunk
data("world.cities", package = "maps")
capitals <- world.cities |>
  dplyr::filter(capital == 1) |>
  dplyr::mutate(iso3 = countrycode::countrycode(
    country.etc,
    origin = "country.name", destination = "iso3c", warn = FALSE
  )) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::group_by(iso3) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::transmute(country, iso3, lat, lon = long) |>
  dplyr::arrange(match(country, country_labels))

haversine_km <- function(lat1, lon1, lat2, lon2) {
  R <- 6371
  to_rad <- pi / 180
  dlat <- (lat2 - lat1) * to_rad
  dlon <- (lon2 - lon1) * to_rad
  a <- sin(dlat / 2)^2 + cos(lat1 * to_rad) * cos(lat2 * to_rad) * sin(dlon / 2)^2
  2 * R * asin(pmin(1, sqrt(a)))
}
D_geo <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  for (j in seq_len(N)) {
    D_geo[i, j] <- haversine_km(
      capitals$lat[i], capitals$lon[i],
      capitals$lat[j], capitals$lon[j]
    )
  }
}

# Build D_cul (Hofstede)
hofstede_path <- "/workspaces/thesis/data/6-dimensions-for-website-2015-08-16(1).csv"
hof_raw <- tryCatch(
  {
    x <- readr::read_csv2(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
    if (ncol(x) < 3) stop("too few columns")
    x
  },
  error = function(e) {
    readr::read_csv(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
  }
)
hof_raw <- hof_raw |>
  dplyr::mutate(dplyr::across(
    dplyr::everything(),
    \(x) ifelse(as.character(x) == "#NULL!", NA, x)
  ))
candidates <- c("country", "ctr", "iso", "iso3")
country_col <- candidates[candidates %in% names(hof_raw)][1]
to_iso3 <- function(x) {
  x <- as.character(x)
  out <- suppressWarnings(countrycode::countrycode(
    x,
    origin = "country.name", destination = "iso3c", warn = FALSE
  ))
  if (mean(is.na(out)) > 0.5) {
    out <- suppressWarnings(countrycode::countrycode(
      x,
      origin = "iso2c", destination = "iso3c", warn = FALSE
    ))
  }
  if (mean(is.na(out)) > 0.5) out <- toupper(x)
  out
}
hof <- hof_raw |>
  dplyr::mutate(iso3 = to_iso3(.data[[country_col]])) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::transmute(
    iso3,
    pdi = suppressWarnings(as.numeric(pdi)),
    idv = suppressWarnings(as.numeric(idv)),
    mas = suppressWarnings(as.numeric(mas)),
    uai = suppressWarnings(as.numeric(uai))
  ) |>
  dplyr::filter(!is.na(pdi) & !is.na(idv) & !is.na(mas) & !is.na(uai)) |>
  dplyr::distinct(iso3, .keep_all = TRUE)
if (!"CYP" %in% hof$iso3 && "GRC" %in% hof$iso3 && "CYP" %in% cl_iso3) {
  hof <- dplyr::bind_rows(
    hof,
    hof |> dplyr::filter(iso3 == "GRC") |> dplyr::mutate(iso3 = "CYP")
  )
}
missing_hof <- setdiff(cl_iso3, hof$iso3)
if (length(missing_hof) > 0) {
  hof_mean <- hof |>
    dplyr::summarise(dplyr::across(
      c(pdi, idv, mas, uai),
      \(x) mean(x, na.rm = TRUE)
    ))
  for (iso in missing_hof) {
    hof <- dplyr::bind_rows(
      hof,
      tibble::tibble(
        iso3 = iso, pdi = hof_mean$pdi, idv = hof_mean$idv,
        mas = hof_mean$mas, uai = hof_mean$uai
      )
    )
  }
}
hof <- hof |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::filter(!is.na(country) & country %in% country_labels) |>
  dplyr::arrange(match(country, country_labels))
H_mat <- as.matrix(hof[, c("pdi", "idv", "mas", "uai")])
rownames(H_mat) <- hof$country
D_cul <- as.matrix(dist(H_mat, method = "euclidean"))
D_cul <- D_cul[country_labels, country_labels]

# Generic builder: WDI-style country-year wide CSV -> distance matrix in logs
build_dist_wdi <- function(path, year_cols, take_log = TRUE) {
  raw <- readr::read_csv(path, skip = 4, show_col_types = FALSE) |>
    janitor::clean_names()
  available <- intersect(year_cols, names(raw))
  if (length(available) == 0) {
    available <- grep("^x[0-9]{4}$", names(raw), value = TRUE)[1:5]
  }
  d <- raw |>
    dplyr::transmute(
      iso3 = country_code,
      val = rowMeans(
        dplyr::across(
          dplyr::all_of(available),
          \(z) suppressWarnings(as.numeric(z))
        ),
        na.rm = TRUE
      )
    ) |>
    dplyr::filter(iso3 %in% cl_iso3, !is.na(val), is.finite(val)) |>
    dplyr::mutate(country = iso_to_label[iso3]) |>
    dplyr::arrange(match(country, country_labels))
  if (nrow(d) < N) {
    miss <- setdiff(country_labels, d$country)
    d <- dplyr::bind_rows(
      d,
      data.frame(
        iso3 = NA, val = mean(d$val, na.rm = TRUE),
        country = miss, stringsAsFactors = FALSE
      )
    ) |>
      dplyr::arrange(match(country, country_labels))
  }
  v <- if (take_log) log(d$val + 1) else d$val
  names(v) <- d$country
  M <- abs(outer(v, v, FUN = "-"))
  dimnames(M) <- list(country_labels, country_labels)
  M
}

# ---- Build all candidate distance / dyadic matrices -----------------------

# 1) GDP per capita
D_gdppc <- build_dist_wdi(
  file.path(data_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000")
)

# 2) Population (proxy: labour force total)
D_pop <- build_dist_wdi(
  file.path(data_dir, "API_SL.TLF.TOTL.IN_DS2_en_csv_v2_196.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000")
)

# 3) Government effectiveness (WDI GE.EST)
D_gov <- build_dist_wdi(
  file.path(data_dir, "API_GE.EST_DS2_en_csv_v2_213352.csv"),
  c("x1996", "x1998", "x2000"),
  take_log = FALSE
)

# 4) R&D / patents (proxy)
D_rd <- build_dist_wdi(
  file.path(data_dir, "API_IP.PAT.RESD_DS2_en_csv_v2_213364.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000")
)

# 5) Trade openness (% GDP)
D_trade_open <- build_dist_wdi(
  file.path(data_dir, "API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000"),
  take_log = FALSE
)

# 6) Renewables share
D_renew <- build_dist_wdi(
  file.path(data_dir, "API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000"),
  take_log = FALSE
)

# 7) PM2.5
D_pm25 <- build_dist_wdi(
  file.path(data_dir, "API_EN.ATM.PM25.MC.M3_DS2_en_csv_v2_316.csv"),
  c("x1998", "x1999", "x2000", "x2001", "x2002")
)

# 8) Environmental tax revenue (Eurostat)
envtax_raw <- readr::read_csv(
  file.path(data_dir, "estat_env_ac_taxind2_en.csv.gz"),
  show_col_types = FALSE
)
envtax <- envtax_raw |>
  dplyr::filter(
    tax == "ENV", unit == "MIO_EUR", nace_r2 == "TOTAL",
    TIME_PERIOD %in% 1996:2000
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    iso3 = iso2_to_iso3(geo),
    country = iso_to_label[iso3]
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))
if (nrow(envtax) < N) {
  miss <- setdiff(country_labels, envtax$country)
  envtax <- dplyr::bind_rows(
    envtax,
    data.frame(
      geo = NA, envtax = mean(envtax$envtax, na.rm = TRUE),
      iso3 = NA, country = miss, stringsAsFactors = FALSE
    )
  )
}
envtax <- envtax |> dplyr::arrange(match(country, country_labels))
envtax_vec <- setNames(log(envtax$envtax + 1), envtax$country)
D_envtax <- abs(outer(envtax_vec, envtax_vec, FUN = "-"))
dimnames(D_envtax) <- list(country_labels, country_labels)

# 9) FDI bilateral (Eurostat) - log(1 + FDI_ij)
fdi_raw <- readr::read_csv(
  file.path(data_dir, "estat_bop_fdi6_pos_en.csv.gz"),
  show_col_types = FALSE
)
fdi <- fdi_raw |>
  dplyr::filter(
    counterp == "IMM", nace_r2 == "FDI", stk_flow == "ASS",
    entity == "TOTAL", fdi_item == "DI__D__F",
    currency == "MIO_EUR", TIME_PERIOD %in% 2013:2020
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi_value = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    geo_iso3 = iso2_to_iso3(geo),
    partner_iso3 = iso2_to_iso3(partner)
  ) |>
  dplyr::filter(geo_iso3 %in% cl_iso3, partner_iso3 %in% cl_iso3) |>
  dplyr::mutate(
    sender = iso_to_label[geo_iso3],
    receiver = iso_to_label[partner_iso3]
  )
F_fdi <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(fdi))) {
  F_fdi[fdi$sender[k], fdi$receiver[k]] <- pmax(fdi$fdi_value[k], 0)
}
log_fdi <- log1p(F_fdi)
diag(log_fdi) <- 0

# 10) Cultural distance (already computed: D_cul)
# 11) Geographic distance (already computed: D_geo)
log_dist <- log(D_geo + 1)
diag(log_dist) <- 0

# ---- Build long dyadic dataset of all candidates --------------------------
mats <- list(
  d_gdppc      = D_gdppc,
  d_pop        = D_pop,
  d_gov        = D_gov,
  d_rd         = D_rd,
  d_tradeopen  = D_trade_open,
  d_renew      = D_renew,
  d_pm25       = D_pm25,
  d_envtax     = D_envtax,
  d_cult       = D_cul,
  d_geo        = log_dist,
  log_fdi      = log_fdi
)

dyad_idx <- expand.grid(i = seq_len(N), j = seq_len(N)) |>
  dplyr::filter(i != j)

X <- sapply(mats, function(M) M[cbind(dyad_idx$i, dyad_idx$j)])
X <- as.data.frame(X)

# ---- Pairwise correlation matrix ------------------------------------------
cor_mat <- cor(X, use = "pairwise.complete.obs")
cor_disp <- formatC(cor_mat, format = "f", digits = 2)
cor_disp[abs(cor_mat) >= 0.5 & abs(cor_mat) < 1] <-
  paste0("\\textbf{", cor_disp[abs(cor_mat) >= 0.5 & abs(cor_mat) < 1], "}")

cor_df <- as.data.frame(cor_disp)
cor_df <- cbind(Variable = rownames(cor_disp), cor_df)

dict_labels <- c(
  d_gdppc      = "$|\\Delta|$ log GDPpc",
  d_pop        = "$|\\Delta|$ log Pop",
  d_gov        = "$|\\Delta|$ Gov eff.",
  d_rd         = "$|\\Delta|$ log Patents",
  d_tradeopen  = "$|\\Delta|$ Trade open.",
  d_renew      = "$|\\Delta|$ Renewables",
  d_pm25       = "$|\\Delta|$ log PM2.5",
  d_envtax     = "$|\\Delta|$ log Env. tax",
  d_cult       = "Cultural dist.",
  d_geo        = "log Geo dist.",
  log_fdi      = "log(1 + FDI$_{ij}$)"
)
cor_df$Variable <- dict_labels[cor_df$Variable]
colnames(cor_df)[-1] <- dict_labels[colnames(cor_df)[-1]]

kbl(cor_df,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = c("l", rep("c", ncol(cor_df) - 1)), row.names = FALSE,
  caption = "Pairwise correlation among candidate covariates",
  label = "tab:dyadic_corr"
) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 8
  )
```

```{r}
#| label: dyadic-regression-homophily-new
#| echo: false
#| results: asis
#| cache: true
# OLS dyadic regression of recovered network weights w_hat_ij on five
# bilateral covariates retained after pairwise-correlation screening
# (all pairwise |rho| < 0.3 among the five):
# (1) economic asymmetry:    log GDP pc_i - log GDP pc_j        (signed)
# (2) environmental asym.:   log Env tax_i - log Env tax_j      (signed)
# (3) cultural distance:     Euclidean distance on Hofstede PDI/IDV/MAS/UAI (D_cul)
# (4) geographic distance:   log great-circle distance between capitals
# (5) bilateral FDI:         log(1 + FDI stock_ij)
#
# Rationale for signed gaps on (1) and (2): w_hat_ij encodes a directed
# influence link i -> j with non-negative magnitude. The absolute value
# specification tests homophily (Lazarsfeld & Merton 1954; McPherson,
# Smith-Lovin & Cook 2001) and mechanically pools opposite-sign effects;
# the signed (sender - receiver) specification tests directional
# asymmetry (Fafchamps & Gubert 2007 J. Dev. Econ.; Graham 2017
# Econometrica), which is what the leader/follower interpretation
# requires. Geographic and cultural distances remain symmetric metrics
# by construction.
#
# No fixed effects. Two-way clustered SE on sender and receiver
# (Cameron & Miller 2014, J. Human Resources).

# Helper: ISO2 -> ISO3 with Eurostat quirks
iso2_to_iso3 <- function(x) {
  x <- as.character(x)
  x <- dplyr::case_match(x, "EL" ~ "GR", "UK" ~ "GB", .default = x)
  countrycode::countrycode(x,
    origin = "iso2c",
    destination = "iso3c", warn = FALSE
  )
}

# ---- 1) Economic asymmetry: signed log GDP per capita gap -----------------
gdp_path <- file.path(data_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv")
gdp_raw <- readr::read_csv(gdp_path, skip = 4, show_col_types = FALSE) |>
  janitor::clean_names()

gdp <- gdp_raw |>
  dplyr::transmute(
    iso3 = country_code,
    gdppc = rowMeans(
      dplyr::across(
        c(x1996, x1997, x1998, x1999, x2000),
        \(z) suppressWarnings(as.numeric(z))
      ),
      na.rm = TRUE
    )
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(gdppc)) |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::arrange(match(country, country_labels))

if (nrow(gdp) < N) {
  miss <- setdiff(country_labels, gdp$country)
  gdp <- dplyr::bind_rows(
    gdp,
    data.frame(
      iso3 = NA, gdppc = mean(gdp$gdppc, na.rm = TRUE),
      country = miss, stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}
gdp_vec <- setNames(log(gdp$gdppc), gdp$country)
# Signed gap: sender (row) - receiver (column)
D_gdppc <- outer(gdp_vec, gdp_vec, FUN = "-")
dimnames(D_gdppc) <- list(country_labels, country_labels)
diag(D_gdppc) <- 0

# ---- 2) Environmental asymmetry: signed log Env Tax gap -------------------
envtax_path <- file.path(data_dir, "estat_env_ac_taxind2_en.csv.gz")
envtax_raw <- readr::read_csv(envtax_path, show_col_types = FALSE)

envtax <- envtax_raw |>
  dplyr::filter(
    tax == "ENV",
    unit == "MIO_EUR",
    nace_r2 == "TOTAL",
    TIME_PERIOD %in% 1996:2000
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    iso3    = iso2_to_iso3(geo),
    country = iso_to_label[iso3]
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))

if (nrow(envtax) < N) {
  miss <- setdiff(country_labels, envtax$country)
  envtax <- dplyr::bind_rows(
    envtax,
    data.frame(
      geo = NA, envtax = mean(envtax$envtax, na.rm = TRUE),
      iso3 = NA, country = miss, stringsAsFactors = FALSE
    )
  )
}
envtax <- envtax |> dplyr::arrange(match(country, country_labels))
envtax_vec <- setNames(log(envtax$envtax + 1), envtax$country)
# Signed gap: sender (row) - receiver (column)
D_envtax <- outer(envtax_vec, envtax_vec, FUN = "-")
dimnames(D_envtax) <- list(country_labels, country_labels)
diag(D_envtax) <- 0

# ---- 3) Cultural distance: Hofstede (D_cul, already computed) -------------

# ---- 4) Geographic distance: log great-circle distance --------------------
log_dist <- log(D_geo + 1)
diag(log_dist) <- 0

# ---- 5) Bilateral FDI: log(1 + FDI stock_ij) ------------------------------
fdi_path <- file.path(data_dir, "estat_bop_fdi6_pos_en.csv.gz")
fdi_raw <- readr::read_csv(fdi_path, show_col_types = FALSE)

fdi <- fdi_raw |>
  dplyr::filter(
    counterp == "IMM",
    nace_r2 == "FDI",
    stk_flow == "ASS",
    entity == "TOTAL",
    fdi_item == "DI__D__F",
    currency == "MIO_EUR",
    TIME_PERIOD %in% 2013:2020
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi_value = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    geo_iso3     = iso2_to_iso3(geo),
    partner_iso3 = iso2_to_iso3(partner)
  ) |>
  dplyr::filter(geo_iso3 %in% cl_iso3, partner_iso3 %in% cl_iso3) |>
  dplyr::mutate(
    sender_country   = iso_to_label[geo_iso3],
    receiver_country = iso_to_label[partner_iso3]
  )

F_fdi <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(fdi))) {
  F_fdi[fdi$sender_country[k], fdi$receiver_country[k]] <-
    pmax(fdi$fdi_value[k], 0)
}
log_fdi <- log1p(F_fdi)
diag(log_fdi) <- 0

# ---- 6) Build dyadic dataset (702 ordered dyads) --------------------------
build_dyadic_df <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      d_econ   = D_gdppc[cbind(as.integer(sender), as.integer(receiver))],
      d_envir  = D_envtax[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}
dyadic_q <- build_dyadic_df(W_q)
dyadic_e <- build_dyadic_df(W_e)

# ---- 7) Estimate OLS dyadic, two-way clustering ---------------------------
fit_q <- feols(w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi,
  data = dyadic_q, cluster = ~ sender + receiver
)
fit_e <- feols(w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi,
  data = dyadic_e, cluster = ~ sender + receiver
)

etable(
  list(
    "Quality $\\hat w^Q_{ij}$"     = fit_q,
    "Environment $\\hat w^E_{ij}$" = fit_e
  ),
  depvar = FALSE,
  style.tex = style.tex(
    "base",
    model.title = ""
  ),
  dict = c(
    d_econ  = "Economic asymmetry ($\\log \\text{GDPpc}_i - \\log \\text{GDPpc}_j$)",
    d_envir = "Environmental asymmetry ($\\log \\text{EnvTax}_i - \\log \\text{EnvTax}_j$)",
    d_cult  = "Cultural distance ($D^{cul}_{ij}$)",
    d_geo   = "Geographic distance ($\\log \\text{dist}_{ij}$)",
    l_fdi   = "Bilateral FDI ($\\log(1 + \\text{FDI}_{ij})$)"
  ),
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Dyadic regression of recovered weights",
  label = "tab:dyadic_homophily"
)
```

## Application of Bramoullé, Djebbari & Fortin (2009)

```{r}
#| label: exogenous-networks-grid
#| echo: false
#| fig-width: 14
#| fig-height: 14
#| fig-cap: "Exogenous networks (Albuquerque, Bronnenberg & Corbett 2007)"
#| include: false
# ============================================================================
# 0. PATHS
# ============================================================================
hofstede_path <- "/workspaces/thesis/data/6-dimensions-for-website-2015-08-16(1).csv"
baci_path <- "/workspaces/thesis/data/BACI_HS92_Y1996_V202601.csv"
ccodes_path <- "/workspaces/thesis/data/country_codes_V202601.csv"

N <- length(country_labels)
TOP_N <- 3

# ============================================================================
# 0bis. ISO3 mapping (country_labels are full names)
# ============================================================================
cl_iso3 <- countrycode::countrycode(
  country_labels,
  origin = "country.name", destination = "iso3c", warn = FALSE
)
names(cl_iso3) <- country_labels
iso_to_label <- setNames(names(cl_iso3), cl_iso3)
stopifnot(!anyNA(cl_iso3))

# ============================================================================
# 1. CAPITALS via maps + countrycode
# ============================================================================
data("world.cities", package = "maps")

capitals <- world.cities |>
  dplyr::filter(capital == 1) |>
  dplyr::mutate(iso3 = countrycode::countrycode(
    country.etc,
    origin = "country.name", destination = "iso3c", warn = FALSE
  )) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::group_by(iso3) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::transmute(country, iso3, lat, lon = long) |>
  dplyr::arrange(match(country, country_labels))

stopifnot(nrow(capitals) == N)

haversine_km <- function(lat1, lon1, lat2, lon2) {
  R <- 6371
  to_rad <- pi / 180
  dlat <- (lat2 - lat1) * to_rad
  dlon <- (lon2 - lon1) * to_rad
  a <- sin(dlat / 2)^2 + cos(lat1 * to_rad) * cos(lat2 * to_rad) * sin(dlon / 2)^2
  2 * R * asin(pmin(1, sqrt(a)))
}

D_geo <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  for (j in seq_len(N)) {
    D_geo[i, j] <- haversine_km(
      capitals$lat[i], capitals$lon[i],
      capitals$lat[j], capitals$lon[j]
    )
  }
}

# ============================================================================
# 2. TRADE — BACI 1996
# ============================================================================
baci <- readr::read_csv(baci_path, show_col_types = FALSE)
ccodes <- readr::read_csv(ccodes_path, show_col_types = FALSE)

iso_map <- ccodes |>
  dplyr::select(country_code, country_iso3) |>
  dplyr::filter(country_iso3 %in% cl_iso3)

trade <- baci |>
  dplyr::select(i, j, v) |>
  dplyr::inner_join(iso_map, by = c("i" = "country_code")) |>
  dplyr::rename(exporter_iso = country_iso3) |>
  dplyr::inner_join(iso_map, by = c("j" = "country_code")) |>
  dplyr::rename(importer_iso = country_iso3) |>
  dplyr::mutate(
    exporter = iso_to_label[exporter_iso],
    importer = iso_to_label[importer_iso]
  ) |>
  dplyr::group_by(exporter, importer) |>
  dplyr::summarise(value = sum(v, na.rm = TRUE), .groups = "drop")

T_trade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(trade))) {
  T_trade[trade$exporter[k], trade$importer[k]] <- trade$value[k]
}

# ============================================================================
# 3. CULTURE — robust to format and missing countries
# ============================================================================
# Try semicolon delimiter first, fall back to comma
hof_raw <- tryCatch(
  {
    x <- readr::read_csv2(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
    if (ncol(x) < 3) stop("too few columns")
    x
  },
  error = function(e) {
    readr::read_csv(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
  }
)

# Hofstede uses "#NULL!" for missing
hof_raw <- hof_raw |>
  dplyr::mutate(dplyr::across(
    dplyr::everything(),
    \(x) ifelse(as.character(x) == "#NULL!", NA, x)
  ))

# Find country identifier column (varies by release)
candidates <- c("country", "ctr", "iso", "iso3")
country_col <- candidates[candidates %in% names(hof_raw)][1]
if (is.na(country_col)) {
  stop(
    "Hofstede CSV has no recognisable country column. Got: ",
    paste(names(hof_raw), collapse = ", ")
  )
}

# Convert identifier to ISO3 with progressive fallbacks
to_iso3 <- function(x) {
  x <- as.character(x)
  out <- suppressWarnings(countrycode::countrycode(
    x,
    origin = "country.name", destination = "iso3c", warn = FALSE
  ))
  if (mean(is.na(out)) > 0.5) {
    out <- suppressWarnings(countrycode::countrycode(
      x,
      origin = "iso2c", destination = "iso3c", warn = FALSE
    ))
  }
  if (mean(is.na(out)) > 0.5) out <- toupper(x) # assume already ISO3
  out
}

hof <- hof_raw |>
  dplyr::mutate(iso3 = to_iso3(.data[[country_col]])) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::transmute(
    iso3,
    pdi = suppressWarnings(as.numeric(pdi)),
    idv = suppressWarnings(as.numeric(idv)),
    mas = suppressWarnings(as.numeric(mas)),
    uai = suppressWarnings(as.numeric(uai))
  ) |>
  dplyr::filter(!is.na(pdi) & !is.na(idv) & !is.na(mas) & !is.na(uai)) |>
  dplyr::distinct(iso3, .keep_all = TRUE)

# Cyprus = Greece (standard practice; absent from Hofstede)
if (!"CYP" %in% hof$iso3 && "GRC" %in% hof$iso3 && "CYP" %in% cl_iso3) {
  hof <- dplyr::bind_rows(
    hof,
    hof |> dplyr::filter(iso3 == "GRC") |> dplyr::mutate(iso3 = "CYP")
  )
}

# Any remaining gaps → EU sample mean (with explicit message)
missing_hof <- setdiff(cl_iso3, hof$iso3)
if (length(missing_hof) > 0) {
  message(
    "Hofstede missing for: ",
    paste(iso_to_label[missing_hof], collapse = ", "),
    " -- imputed with EU sample mean."
  )
  hof_mean <- hof |>
    dplyr::summarise(dplyr::across(
      c(pdi, idv, mas, uai),
      \(x) mean(x, na.rm = TRUE)
    ))
  for (iso in missing_hof) {
    hof <- dplyr::bind_rows(
      hof,
      tibble::tibble(
        iso3 = iso, pdi = hof_mean$pdi, idv = hof_mean$idv,
        mas = hof_mean$mas, uai = hof_mean$uai
      )
    )
  }
}

hof <- hof |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::filter(!is.na(country) & country %in% country_labels) |>
  dplyr::arrange(match(country, country_labels))

stopifnot(nrow(hof) == N)

H_mat <- as.matrix(hof[, c("pdi", "idv", "mas", "uai")])
rownames(H_mat) <- hof$country
D_cul <- as.matrix(dist(H_mat, method = "euclidean"))
D_cul <- D_cul[country_labels, country_labels]

# ============================================================================
# 4. TOP-3 DIRECTED NETWORKS, UNIFORM 1/3 WEIGHTS
# ============================================================================
top_n_network <- function(M, top_n = TOP_N, similarity = FALSE) {
  W <- matrix(0, nrow(M), ncol(M), dimnames = dimnames(M))
  for (i in seq_len(nrow(M))) {
    v <- M[i, ]
    v[i] <- if (similarity) -Inf else Inf
    idx <- if (similarity) {
      order(v, decreasing = TRUE)[seq_len(top_n)]
    } else {
      order(v, decreasing = FALSE)[seq_len(top_n)]
    }
    W[i, idx] <- 1 / top_n
  }
  W
}

W_geo <- top_n_network(D_geo, similarity = FALSE)
W_trade <- top_n_network(T_trade, similarity = TRUE)
W_cul <- top_n_network(D_cul, similarity = FALSE)

saveRDS(
  list(geo = W_geo, trade = W_trade, culture = W_cul),
  file.path(out_dir, "exogenous_networks.rds")
)

# ============================================================================
# 5. 3x3 GRID PLOT, ROW-FIXED FR LAYOUTS
# ============================================================================
to_graph <- function(W) {
  g <- igraph::graph_from_adjacency_matrix(W, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- rownames(W)
  g
}
g_geo <- to_graph(W_geo)
g_trade <- to_graph(W_trade)
g_cul <- to_graph(W_cul)

set.seed(1861)
lo_geo <- igraph::layout_with_fr(g_geo)
set.seed(1861)
lo_trade <- igraph::layout_with_fr(g_trade)
set.seed(1861)
lo_cul <- igraph::layout_with_fr(g_cul)
rownames(lo_geo) <- rownames(lo_trade) <- rownames(lo_cul) <- country_labels

plot_one <- function(g, lo, title) {
  lo_ord <- lo[igraph::V(g)$name, , drop = FALSE]
  ggraph(g, layout = lo_ord) +
    geom_edge_arc(
      width = 0.4, alpha = 0.55, colour = "grey40",
      arrow = arrow(length = unit(1.8, "mm"), type = "closed", angle = 22),
      end_cap = circle(4, "mm"),
      start_cap = circle(4, "mm"),
      strength = 0.12
    ) +
    geom_node_point(size = 3.5, colour = "steelblue") +
    geom_node_text(aes(label = name),
      size = 2.4, repel = TRUE,
      bg.colour = "white", bg.r = 0.18,
      segment.colour = "grey55", segment.size = 0.25,
      max.overlaps = Inf, force = 3, force_pull = 0.5,
      point.padding = unit(0.5, "lines"),
      box.padding = unit(0.35, "lines"),
      min.segment.length = 0
    ) +
    ggtitle(title) +
    theme_void(base_size = 10) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.margin = margin(6, 6, 6, 6)
    )
}

p11 <- plot_one(g_geo, lo_geo, "Geography -- geo layout")
p12 <- plot_one(g_trade, lo_geo, "Trade -- geo layout")
p13 <- plot_one(g_cul, lo_geo, "Culture -- geo layout")
p21 <- plot_one(g_geo, lo_trade, "Geography -- trade layout")
p22 <- plot_one(g_trade, lo_trade, "Trade -- trade layout")
p23 <- plot_one(g_cul, lo_trade, "Culture -- trade layout")
p31 <- plot_one(g_geo, lo_cul, "Geography -- culture layout")
p32 <- plot_one(g_trade, lo_cul, "Trade -- culture layout")
p33 <- plot_one(g_cul, lo_cul, "Culture -- culture layout")

panel_exo <- (p11 | p12 | p13) /
  (p21 | p22 | p23) /
  (p31 | p32 | p33)

ggsave(file.path(fig_dir, "exogenous_networks_grid.pdf"), panel_exo,
  width = 14, height = 14, device = cairo_pdf
)
print(panel_exo)
```

```{r}
#| label: bdf-final-summary-combined
#| echo: false
#| results: asis
#| cache: true
# Self-contained: BDF (2009) G2SLS with one-year lag on the peer outcome,
# applied to the De Paula, Rasul & Souza (2025) recovered networks and the
# Albuquerque, Bronnenberg & Corbett (2007, Eq. 9) combined exogenous network.
# Three specifications:
# (i)   no FE, asymptotic country-clustered SE
# (ii)  country FE, pairs cluster bootstrap on country (B = 999)
# (iii) country + year FE, pairs cluster bootstrap on country (B = 999)


set.seed(1861)
B <- 999

# ---- Helpers ---------------------------------------------------------------
row_normalise <- function(W, eps = 1e-5) {
  W[abs(W) < eps] <- 0
  diag(W) <- 0
  rs <- rowSums(W)
  rs[rs == 0] <- 1
  W / rs
}

attach_y <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |> dplyr::transmute(country, year, y = log1p(n_cert)),
      by = c("country", "year")
    ) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

keep_full <- function(d) {
  full <- d |>
    dplyr::count(year) |>
    dplyr::filter(n == length(country_labels)) |>
    dplyr::pull(year)
  dplyr::filter(d, year %in% full)
}

build_bdf_panel_lag <- function(panel, W, x_vars) {
  G <- row_normalise(W)
  G2 <- G %*% G
  G3 <- G2 %*% G
  d <- panel[order(panel$year, match(panel$country, rownames(G))), ]
  yrs <- sort(unique(d$year))
  d$Gy <- unlist(lapply(yrs, function(t) {
    yt <- d$y[d$year == t]
    as.numeric(G %*% yt)
  }))
  for (v in x_vars) {
    Xt <- matrix(d[[v]], ncol = length(yrs))
    d[[paste0("G_", v)]] <- as.numeric(G %*% Xt)
    d[[paste0("G2_", v)]] <- as.numeric(G2 %*% Xt)
    d[[paste0("G3_", v)]] <- as.numeric(G3 %*% Xt)
  }
  d <- d[order(d$country, d$year), ]
  d <- d |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      Gy_lag = dplyr::lag(Gy, n = 1, order_by = year),
      dplyr::across(
        dplyr::all_of(c(paste0("G2_", x_vars), paste0("G3_", x_vars))),
        \(z) dplyr::lag(z, n = 1, order_by = year),
        .names = "{.col}_lag"
      )
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(Gy_lag)
  d
}

cluster_boot <- function(data, fit_formula, cluster_var = "country", B = 999) {
  clusters <- unique(data[[cluster_var]])
  G <- length(clusters)
  est0 <- coef(fixest::feols(fit_formula, data = data, warn = FALSE, notes = FALSE))
  k <- length(est0)
  boot_mat <- matrix(NA_real_, nrow = B, ncol = k, dimnames = list(NULL, names(est0)))
  for (b in seq_len(B)) {
    idx <- sample(clusters, G, replace = TRUE)
    boot_df <- do.call(rbind, lapply(seq_along(idx), function(j) {
      d <- data[data[[cluster_var]] == idx[j], ]
      d[[cluster_var]] <- paste0(idx[j], "_", j)
      d
    }))
    fit_b <- tryCatch(
      fixest::feols(fit_formula, data = boot_df, warn = FALSE, notes = FALSE),
      error = function(e) NULL
    )
    if (!is.null(fit_b)) boot_mat[b, ] <- coef(fit_b)
  }
  list(
    estimate = est0,
    se = apply(boot_mat, 2, sd, na.rm = TRUE),
    p_val = sapply(seq_len(k), function(j) {
      mean(abs(boot_mat[, j] - est0[j]) >= abs(est0[j]), na.rm = TRUE)
    })
  )
}

sig_stars <- function(p) {
  if (is.na(p)) {
    return("")
  }
  if (p < 0.01) {
    "$^{***}$"
  } else if (p < 0.05) {
    "$^{**}$"
  } else if (p < 0.10) "$^{*}$" else ""
}

extract_nofe <- function(fit) {
  s <- summary(fit, cluster = ~country)
  ct <- s$coeftable
  i <- which(rownames(ct) == "fit_Gy_lag")
  list(b = ct[i, "Estimate"], se = ct[i, "Std. Error"], p = ct[i, "Pr(>|t|)"])
}

extract_cfe <- function(b) {
  i <- which(names(b$estimate) == "fit_Gy_lag")
  list(b = b$estimate[i], se = b$se[i], p = b$p_val[i])
}

# ---- Build data ------------------------------------------------------------
id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(attach_y(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001))
panel_e <- keep_full(attach_y(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001))

# Combined ABC network
W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0), which(W_cul[i, ] > 0)))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

# BDF panels with one-year lag
bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

# ---- Formulas --------------------------------------------------------------
rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2_lag <- paste(paste0("G2_", x_vars, "_lag"), collapse = " + ")
inst3_lag <- paste(paste0("G3_", x_vars, "_lag"), collapse = " + ")

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE -------------------------------------------------------------
nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# ---- (ii) Country FE -------------------------------------------------------
boot_q_rec_cfe <- cluster_boot(bdf_q_lag, f_cfe, "country", B = B)
boot_q_abc_cfe <- cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B)
boot_e_rec_cfe <- cluster_boot(bdf_e_lag, f_cfe, "country", B = B)
boot_e_abc_cfe <- cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B)

cfe_q_rec <- extract_cfe(boot_q_rec_cfe)
cfe_q_abc <- extract_cfe(boot_q_abc_cfe)
cfe_e_rec <- extract_cfe(boot_e_rec_cfe)
cfe_e_abc <- extract_cfe(boot_e_abc_cfe)

# ---- Build table (estimates above, SE below) -------------------------------
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrrr", row.names = FALSE,
  caption = "Endogenous peer effect: recovered vs combined exogenous network",
  label = "tab:bdf_combined_summary"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-final-summary-abc-model5
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS with one-year lag, applied to Albuquerque, Bronnenberg
# & Corbett (2007, Table 3, Model 5) preferred two-channel networks:
# - Quality (ISO 9001):     geography + bilateral trade
# - Environment (ISO 14001): geography + cultural similarity
#
# Two specifications:
# (i)  no FE, asymptotic country-clustered SE
# (ii) country + year FE, pairs cluster bootstrap on country (B = 999)
#
# Reuses panel_q, panel_e, x_vars, rhs, exo, inst2_lag, inst3_lag,
# build_bdf_panel_lag, cluster_boot, sig_stars, extract_cfe, extract_nofe,
# W_geo, W_trade, W_cul from previous chunks.

set.seed(1861)
B <- 999
N <- length(country_labels)

# ---- Build ABC Model 5 networks --------------------------------------------
# Quality: union of top-3 from geography + bilateral trade
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs_i] <- 1 / length(nbrs_i)
}

# Environment: union of top-3 from geography + Hofstede culture
W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs_i] <- 1 / length(nbrs_i)
}

# ---- Build BDF panels with one-year lag ------------------------------------
bdf_geotrade_q_lag <- build_bdf_panel_lag(panel_q, W_geotrade, x_vars)
bdf_geoculture_e_lag <- build_bdf_panel_lag(panel_e, W_geoculture, x_vars)

# ---- Formulas --------------------------------------------------------------
f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE: asymptotic country-clustered SE ----------------------------
nofe_q_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geotrade_q_lag, cluster = ~country))
nofe_e_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geoculture_e_lag, cluster = ~country))

# ---- (ii) Two-way FE: pairs cluster bootstrap on country -------------------
boot_q_abc5_tw <- cluster_boot(bdf_geotrade_q_lag, f_twfe, "country", B = B)
boot_e_abc5_tw <- cluster_boot(bdf_geoculture_e_lag, f_twfe, "country", B = B)

twfe_q_abc5 <- extract_cfe(boot_q_abc5_tw)
twfe_e_abc5 <- extract_cfe(boot_e_abc5_tw)

# ---- Build table -----------------------------------------------------------
fmt_cell <- function(b, se, p) {
  c(sprintf("%.3f%s", b, sig_stars(p)), sprintf("(%.3f)", se))
}

# ---- Build table (estimate above, SE below) --------------------------------
tab <- rbind(
  c(
    "Quality", "Exogenous: Geo + Trade",
    sprintf("%.3f%s", nofe_q_abc5$b, sig_stars(nofe_q_abc5$p)),
    sprintf("%.3f%s", twfe_q_abc5$b, sig_stars(twfe_q_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc5$se),
    sprintf("(%.3f)", twfe_q_abc5$se)
  ),
  c(
    "Environment", "Exogenous: Geo + Culture",
    sprintf("%.3f%s", nofe_e_abc5$b, sig_stars(nofe_e_abc5$p)),
    sprintf("%.3f%s", twfe_e_abc5$b, sig_stars(twfe_e_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc5$se),
    sprintf("(%.3f)", twfe_e_abc5$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country + year FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect: ABC two-channel networks",
  label = "tab:bdf_abc_model5"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: dyadic-regression-dzemski
#| echo: false
#| results: asis
#| cache: true
# Dzemski (2018) two-sided FE specification.
# Signed gaps log X_i - log X_j are perfectly absorbed by sender + receiver FE
# (additively separable). We replace them with absolute gaps |log X_i - log X_j|,
# which are non-separable in (i,j) and therefore identified within FE.
# This matches McPherson, Smith-Lovin & Cook (2001) homophily and standard
# practice in the gravity literature (Head & Mayer 2014).

build_dyadic_dz <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = abs(D_envtax[cbind(as.integer(sender), as.integer(receiver))]),
      d_cult   =     D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    =     log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    =     log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q <- build_dyadic_dz(W_q)
dz_e <- build_dyadic_dz(W_e)

f_logit <- edge  ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver
f_ols <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver

fit_logit_q <- feglm(f_logit,
  data = dz_q,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)
fit_logit_e <- feglm(f_logit,
  data = dz_e,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)

fit_ols_q <- feols(f_ols, data = dz_q, cluster = ~ sender + receiver)
fit_ols_e <- feols(f_ols, data = dz_e, cluster = ~ sender + receiver)

etable(
  list(
    "Logit: $g^Q_{ij}$"     = fit_logit_q,
    "Logit: $g^E_{ij}$"     = fit_logit_e,
    "OLS: $\\hat w^Q_{ij}$" = fit_ols_q,
    "OLS: $\\hat w^E_{ij}$" = fit_ols_e
  ),
  depvar = FALSE,
  style.tex = style.tex("base", model.title = "", fixef.suffix = ":"),
  dict = c(
    d_econ   = "Economic distance ($|\\log \\text{GDPpc}_i - \\log \\text{GDPpc}_j|$)",
    d_envir  = "Environmental distance ($|\\log \\text{EnvTax}_i - \\log \\text{EnvTax}_j|$)",
    d_cult   = "Cultural distance ($D^{cul}_{ij}$)",
    d_geo    = "Geographic distance ($\\log \\text{dist}_{ij}$)",
    l_fdi    = "Bilateral FDI ($\\log(1 + \\text{FDI}_{ij})$)",
    sender   = "Sender FE ($\\alpha_i^{\\text{out}}$)",
    receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
  ),
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2 + pr2,
  tex = TRUE,
  title = "Dyadic regression of recovered weights with sender and receiver FE (Dzemski 2018)",
  label = "tab:dyadic_dzemski"
)
```

```{r}
#| label: dyadic-regression-dzemski-signed
#| echo: false
#| results: asis
#| cache: true
# Dzemski (2018) two-sided FE on signed gaps. Note: signed gaps
# log X_i - log X_j are additively separable in (i,j) and therefore
# absorbed by sender + receiver FE. fixest will drop them automatically
# (collinearity with FE). Reported anyway for transparency: only
# non-separable covariates (cultural, geographic, FDI) are identified.

build_dyadic_dz_signed <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = D_gdppc[cbind(as.integer(sender), as.integer(receiver))],
      d_envir  = D_envtax[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q_s <- build_dyadic_dz_signed(W_q)
dz_e_s <- build_dyadic_dz_signed(W_e)

f_logit <- edge  ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver
f_ols <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver

fit_logit_q <- feglm(f_logit,
  data = dz_q_s,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)
fit_logit_e <- feglm(f_logit,
  data = dz_e_s,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)

fit_ols_q <- feols(f_ols, data = dz_q_s, cluster = ~ sender + receiver)
fit_ols_e <- feols(f_ols, data = dz_e_s, cluster = ~ sender + receiver)

etable(
  list(
    "Logit: $g^Q_{ij}$"     = fit_logit_q,
    "Logit: $g^E_{ij}$"     = fit_logit_e,
    "OLS: $\\hat w^Q_{ij}$" = fit_ols_q,
    "OLS: $\\hat w^E_{ij}$" = fit_ols_e
  ),
  depvar = FALSE,
  style.tex = style.tex("base", model.title = "", fixef.suffix = ":"),
  dict = c(
    d_econ   = "Economic asymmetry ($\\log \\text{GDPpc}_i - \\log \\text{GDPpc}_j$)",
    d_envir  = "Environmental asymmetry ($\\log \\text{EnvTax}_i - \\log \\text{EnvTax}_j$)",
    d_cult   = "Cultural distance ($D^{cul}_{ij}$)",
    d_geo    = "Geographic distance ($\\log \\text{dist}_{ij}$)",
    l_fdi    = "Bilateral FDI ($\\log(1 + \\text{FDI}_{ij})$)",
    sender   = "Sender FE ($\\alpha_i^{\\text{out}}$)",
    receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
  ),
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2 + pr2,
  tex = TRUE,
  title = "Dyadic regression with sender and receiver FE, signed asymmetries",
  label = "tab:dyadic_dzemski_signed"
)
```

```{r}
#| label: dyadic-regression-dzemski-abs-nofe
#| echo: false
#| results: asis
#| cache: true
# Pooled specification (no FE) on absolute-value asymmetries.
# Same covariates as Table tab:dyadic_dzemski (Dzemski 2018 with FE), but
# without sender/receiver FE — so all five coefficients are identified.
# Two-way clustered SE on sender and receiver.

build_dyadic_dz <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = abs(D_envtax[cbind(as.integer(sender), as.integer(receiver))]),
      d_cult   =     D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    =     log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    =     log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q <- build_dyadic_dz(W_q)
dz_e <- build_dyadic_dz(W_e)

f_logit_nofe <- edge  ~ d_econ + d_envir + d_cult + d_geo + l_fdi
f_ols_nofe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi

fit_logit_q_nf <- feglm(f_logit_nofe,
  data = dz_q,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)
fit_logit_e_nf <- feglm(f_logit_nofe,
  data = dz_e,
  family = binomial("logit"),
  cluster = ~ sender + receiver
)

fit_ols_q_nf <- feols(f_ols_nofe, data = dz_q, cluster = ~ sender + receiver)
fit_ols_e_nf <- feols(f_ols_nofe, data = dz_e, cluster = ~ sender + receiver)

etable(
  list(
    "Logit: $g^Q_{ij}$"     = fit_logit_q_nf,
    "Logit: $g^E_{ij}$"     = fit_logit_e_nf,
    "OLS: $\\hat w^Q_{ij}$" = fit_ols_q_nf,
    "OLS: $\\hat w^E_{ij}$" = fit_ols_e_nf
  ),
  depvar = FALSE,
  style.tex = style.tex("base", model.title = ""),
  dict = c(
    d_econ  = "Economic distance ($|\\log \\text{GDPpc}_i - \\log \\text{GDPpc}_j|$)",
    d_envir = "Environmental distance ($|\\log \\text{EnvTax}_i - \\log \\text{EnvTax}_j|$)",
    d_cult  = "Cultural distance ($D^{cul}_{ij}$)",
    d_geo   = "Geographic distance ($\\log \\text{dist}_{ij}$)",
    l_fdi   = "Bilateral FDI ($\\log(1 + \\text{FDI}_{ij})$)"
  ),
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2 + pr2,
  tex = TRUE,
  title = "Dyadic regression of recovered weights, pooled specification without fixed effects",
  label = "tab:dyadic_dzemski_abs_nofe"
)
```

```{r}
#| label: bdf-final-summary-combined-diff
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS with one-year lag on the peer outcome.
# Outcome: y_it = log(1 + n_it) - log(1 + n_{i,t-1}) — same log-difference
# transformation used in the recoverNetwork stage (consistent scale).
# Two specifications:
# (i)  no FE, asymptotic country-clustered SE
# (ii) country FE, pairs cluster bootstrap on country (B = 999)

set.seed(1861)
B <- 999

# ---- Outcome builder: log-difference with one-year lag --------------------
attach_y_diff <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::arrange(country, year) |>
        dplyr::group_by(country) |>
        dplyr::mutate(
          y = log1p(n_cert) -
            log1p(dplyr::lag(n_cert, n = 1, order_by = year))
        ) |>
        dplyr::ungroup() |>
        dplyr::transmute(country, year, y),
      by = c("country", "year")
    ) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

# ---- Build data ------------------------------------------------------------
id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(
  attach_y_diff(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001)
)
panel_e <- keep_full(
  attach_y_diff(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001)
)

# Combined ABC network
W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(
    which(W_geo[i, ] > 0),
    which(W_trade[i, ] > 0),
    which(W_cul[i, ] > 0)
  ))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

# BDF panels with one-year lag on Gy
bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

# ---- Formulas --------------------------------------------------------------
rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2_lag <- paste(paste0("G2_", x_vars, "_lag"), collapse = " + ")
inst3_lag <- paste(paste0("G3_", x_vars, "_lag"), collapse = " + ")

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE -------------------------------------------------------------
nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# ---- (ii) Country FE -------------------------------------------------------
boot_q_rec_cfe <- cluster_boot(bdf_q_lag, f_cfe, "country", B = B)
boot_q_abc_cfe <- cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B)
boot_e_rec_cfe <- cluster_boot(bdf_e_lag, f_cfe, "country", B = B)
boot_e_abc_cfe <- cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B)

cfe_q_rec <- extract_cfe(boot_q_rec_cfe)
cfe_q_abc <- extract_cfe(boot_q_abc_cfe)
cfe_e_rec <- extract_cfe(boot_e_rec_cfe)
cfe_e_abc <- extract_cfe(boot_e_abc_cfe)

# ---- Build table -----------------------------------------------------------
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log-differences: recovered vs combined exogenous network",
  label = "tab:bdf_combined_summary_diff"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```  

```{r}
#| label: bdf-summary-combined-diff-yfe
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on log-differences, recovered + combined exogenous network.
# Four FE specifications:
# (i)   no FE                  -> asymptotic country-clustered SE
# (ii)  country FE             -> pairs cluster bootstrap on country (B = 999)
# (iii) year FE only           -> pairs cluster bootstrap on country (B = 999)
# (iv)  country + year FE      -> pairs cluster bootstrap on country (B = 999)

set.seed(1861)
B <- 999

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_yfe <- as.formula(sprintf(
  "y ~ %s + %s | year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_twfe <- as.formula(sprintf(
  "y ~ %s + %s | country + year | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# (i) No FE - asymptotic
nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# (ii) Country FE - bootstrap
cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

# (iii) Year FE only - bootstrap
yfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_yfe, "country", B = B))
yfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_yfe, "country", B = B))
yfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_yfe, "country", B = B))
yfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_yfe, "country", B = B))

# (iv) Two-way FE - bootstrap
tw_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_twfe, "country", B = B))
tw_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_twfe, "country", B = B))
tw_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_twfe, "country", B = B))
tw_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_twfe, "country", B = B))

# Build table
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p)),
    sprintf("%.3f%s", yfe_q_rec$b, sig_stars(yfe_q_rec$p)),
    sprintf("%.3f%s", tw_q_rec$b, sig_stars(tw_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se),
    sprintf("(%.3f)", yfe_q_rec$se),
    sprintf("(%.3f)", tw_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p)),
    sprintf("%.3f%s", yfe_q_abc$b, sig_stars(yfe_q_abc$p)),
    sprintf("%.3f%s", tw_q_abc$b, sig_stars(tw_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se),
    sprintf("(%.3f)", yfe_q_abc$se),
    sprintf("(%.3f)", tw_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p)),
    sprintf("%.3f%s", yfe_e_rec$b, sig_stars(yfe_e_rec$p)),
    sprintf("%.3f%s", tw_e_rec$b, sig_stars(tw_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se),
    sprintf("(%.3f)", yfe_e_rec$se),
    sprintf("(%.3f)", tw_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p)),
    sprintf("%.3f%s", yfe_e_abc$b, sig_stars(yfe_e_abc$p)),
    sprintf("%.3f%s", tw_e_abc$b, sig_stars(tw_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se),
    sprintf("(%.3f)", yfe_e_abc$se),
    sprintf("(%.3f)", tw_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "no FE", "country FE", "year FE", "country+year FE"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrrrr", row.names = FALSE,
  caption = "Endogenous peer effect $\\hat\\beta$ on log-differences across FE specifications",
  label = "tab:bdf_combined_yfe"
) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 10)
```


```{r}
#| label: bdf-final-summary-abc-model5-diff
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS with one-year lag, applied to ABC (2007, Tab. 3, Model 5)
# preferred two-channel networks. Outcome on log-differences (consistent
# with recoverNetwork stage).
# Two specifications:
# (i)  no FE, asymptotic country-clustered SE
# (ii) country FE, pairs cluster bootstrap on country (B = 999)
# Reuses panel_q, panel_e (already log-diff) from previous chunk.

set.seed(1861)
B <- 999
N <- length(country_labels)

# ---- Build ABC Model 5 networks --------------------------------------------
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs_i] <- 1 / length(nbrs_i)
}

W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs_i] <- 1 / length(nbrs_i)
}

bdf_geotrade_q_lag <- build_bdf_panel_lag(panel_q, W_geotrade, x_vars)
bdf_geoculture_e_lag <- build_bdf_panel_lag(panel_e, W_geoculture, x_vars)

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE -------------------------------------------------------------
nofe_q_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geotrade_q_lag, cluster = ~country))
nofe_e_abc5 <- extract_nofe(feols(f_nofe, data = bdf_geoculture_e_lag, cluster = ~country))

# ---- (ii) Country FE -------------------------------------------------------
boot_q_abc5_cfe <- cluster_boot(bdf_geotrade_q_lag, f_cfe, "country", B = B)
boot_e_abc5_cfe <- cluster_boot(bdf_geoculture_e_lag, f_cfe, "country", B = B)

cfe_q_abc5 <- extract_cfe(boot_q_abc5_cfe)
cfe_e_abc5 <- extract_cfe(boot_e_abc5_cfe)

tab <- rbind(
  c(
    "Quality", "Exogenous: Geo + Trade",
    sprintf("%.3f%s", nofe_q_abc5$b, sig_stars(nofe_q_abc5$p)),
    sprintf("%.3f%s", cfe_q_abc5$b, sig_stars(cfe_q_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc5$se),
    sprintf("(%.3f)", cfe_q_abc5$se)
  ),
  c(
    "Environment", "Exogenous: Geo + Culture",
    sprintf("%.3f%s", nofe_e_abc5$b, sig_stars(nofe_e_abc5$p)),
    sprintf("%.3f%s", cfe_e_abc5$b, sig_stars(cfe_e_abc5$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc5$se),
    sprintf("(%.3f)", cfe_e_abc5$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log-differences: ABC two-channel networks",
  label = "tab:bdf_abc_model5_diff"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-final-summary-combined-diff-sucaminchia
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on log-differences with De Paula spatio-temporal spec:
#   y_it = rho * y_{i,t-1} + beta * Gy_{i,t-1}
#        + gamma_x * x_it + gamma_Gx * Gx_it + alpha_i + eps_it
# y_it = log(1 + n_cert,it) - log(1 + n_cert,i,t-1)  [log-difference]
# Instruments for Gy_lag: G^2 X_{i,t-1}, G^3 X_{i,t-1}
# Instrument for y_lag:   y_{i,t-2} (Anderson-Hsiao 1981; Arellano-Bond 1991)
# Two specs: (i) no FE, asymptotic country-clustered SE
#            (ii) country FE, pairs cluster bootstrap on country (B = 999)

set.seed(1861)
B <- 999

# ---- Outcome builder: log-difference --------------------------------------
attach_y_diff <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::arrange(country, year) |>
        dplyr::group_by(country) |>
        dplyr::mutate(
          y = log1p(n_cert) -
            log1p(dplyr::lag(n_cert, n = 1, order_by = year))
        ) |>
        dplyr::ungroup() |>
        dplyr::transmute(country, year, y),
      by = c("country", "year")
    ) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

# ---- BDF panel builder with y_lag and y_lag2 ------------------------------
build_bdf_panel_dp <- function(panel, W, x_vars) {
  G <- row_normalise(W)
  G2 <- G %*% G
  G3 <- G2 %*% G
  d <- panel[order(panel$year, match(panel$country, rownames(G))), ]
  yrs <- sort(unique(d$year))
  d$Gy <- unlist(lapply(yrs, function(t) {
    yt <- d$y[d$year == t]
    as.numeric(G %*% yt)
  }))
  for (v in x_vars) {
    Xt <- matrix(d[[v]], ncol = length(yrs))
    d[[paste0("G_", v)]] <- as.numeric(G %*% Xt)
    d[[paste0("G2_", v)]] <- as.numeric(G2 %*% Xt)
    d[[paste0("G3_", v)]] <- as.numeric(G3 %*% Xt)
  }
  d <- d[order(d$country, d$year), ]
  d <- d |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      y_lag = dplyr::lag(y, n = 1, order_by = year),
      y_lag2 = dplyr::lag(y, n = 2, order_by = year),
      Gy_lag = dplyr::lag(Gy, n = 1, order_by = year),
      dplyr::across(
        dplyr::all_of(c(paste0("G2_", x_vars), paste0("G3_", x_vars))),
        \(z) dplyr::lag(z, n = 1, order_by = year),
        .names = "{.col}_lag"
      )
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(y_lag, y_lag2, Gy_lag)
  d
}

# ---- Build data ------------------------------------------------------------
id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(
  attach_y_diff(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001)
)
panel_e <- keep_full(
  attach_y_diff(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001)
)

# Combined ABC network
W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(
    which(W_geo[i, ] > 0),
    which(W_trade[i, ] > 0),
    which(W_cul[i, ] > 0)
  ))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

bdf_q_lag <- build_bdf_panel_dp(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_dp(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_dp(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_dp(panel_e, W_combined, x_vars)

# ---- Formulas: two endogenous vars (y_lag, Gy_lag), three instruments ----
rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2_lag <- paste(paste0("G2_", x_vars, "_lag"), collapse = " + ")
inst3_lag <- paste(paste0("G3_", x_vars, "_lag"), collapse = " + ")

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | y_lag + Gy_lag ~ y_lag2 + %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | y_lag + Gy_lag ~ y_lag2 + %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE -------------------------------------------------------------
nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# ---- (ii) Country FE -------------------------------------------------------
cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

# ---- Build table -----------------------------------------------------------
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log-differences with own autoregression: recovered vs combined exogenous network",
  label = "tab:bdf_combined_summary_diff"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-final-summary-combined-firmnorm
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on certification intensity, time-invariant firm-count
# normalisation: y_it = n_cert,it / n_firms,i,2017 * 1000.
# Divides out cross-country scale heterogeneity (Albuquerque, Bronnenberg
# & Corbett 2007 use a similar firm-count normalisation in their adoption-
# rate construction). Because n_firms_17 is time-invariant, within-country
# variation in y comes entirely from n_cert.
# Specs:
#   (i)  no FE, asymptotic country-clustered SE
#   (ii) country FE, pairs cluster bootstrap on country (B = 999)

set.seed(1861)
B <- 999

# ---- Load firm counts at 2017 (Eurostat SBS) -------------------------------
harmonise_names <- function(df) {
  if (nrow(df) == 0) {
    return(df)
  }
  df %>% dplyr::mutate(country = dplyr::case_match(country,
    "Korea (Republic of)" ~ "Korea", "Korea, Republic of" ~ "Korea",
    "Korea (the Republic of)" ~ "Korea",
    "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
    "United Kingdom of Great Britain and Northern Ireland" ~ "United Kingdom",
    "United States of America" ~ "United States",
    "Netherlands (Kingdom of the)" ~ "Netherlands",
    .default = country
  ))
}

load_sbs_2017 <- function() {
  file_path <- file.path(data_dir, "enterprises_eu.xlsx")
  raw <- readxl::read_excel(file_path,
    sheet = "Sheet 1",
    skip = 9, col_names = FALSE
  )
  tibble::tibble(
    country    = raw[[1]],
    n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))
  ) %>%
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) %>%
    harmonise_names() %>%
    dplyr::distinct(country, .keep_all = TRUE)
}

firm17 <- load_sbs_2017()
missing_firm17 <- setdiff(country_labels, firm17$country)
if (length(missing_firm17) > 0) {
  warning(
    "n_firms_17 missing for: ",
    paste(missing_firm17, collapse = ", "),
    "; these countries will be dropped via inner_join"
  )
}
cat("Firm counts loaded for", nrow(firm17), "countries.\n")

# ---- Outcome builder: certification intensity (per 1000 firms 2017) -------
attach_y_intensity <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::transmute(country, year, n_cert),
      by = c("country", "year")
    ) |>
    dplyr::inner_join(firm17, by = "country") |>
    dplyr::mutate(y = n_cert / n_firms_17 * 1000) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5, y) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

# ---- Build data ------------------------------------------------------------
id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(
  attach_y_intensity(readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001)
)
panel_e <- keep_full(
  attach_y_intensity(readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001)
)

# Diagnostic
cat(sprintf(
  "Quality (cert/1000 firms 2017): mean=%.2f, sd=%.2f, range=[%.2f, %.2f], N obs=%d, T=%d\n",
  mean(panel_q$y), sd(panel_q$y), min(panel_q$y), max(panel_q$y),
  nrow(panel_q), length(unique(panel_q$year))
))
cat(sprintf(
  "Environment (cert/1000 firms 2017): mean=%.2f, sd=%.2f, range=[%.2f, %.2f], N obs=%d, T=%d\n",
  mean(panel_e$y), sd(panel_e$y), min(panel_e$y), max(panel_e$y),
  nrow(panel_e), length(unique(panel_e$year))
))

# Combined ABC network
W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(
    which(W_geo[i, ] > 0),
    which(W_trade[i, ] > 0),
    which(W_cul[i, ] > 0)
  ))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

# ---- Formulas (reuse rhs, exo, inst2_lag, inst3_lag) ---------------------
f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

# ---- (i) No FE -------------------------------------------------------------
nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

# ---- (ii) Country FE -------------------------------------------------------
cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

# ---- Build table -----------------------------------------------------------
tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se),
    sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se),
    sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se),
    sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se),
    sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on certification intensity (per 1000 firms, 2017): recovered vs combined exogenous network",
  label = "tab:bdf_combined_summary_firmnorm"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-firmnorm-levels
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on certification intensity in levels.
# Outcome: y_it = n_cert,it / n_firms,i,2017 * 1000  (cert per 1000 firms)
# Time-invariant firm-count normalisation (Albuquerque, Bronnenberg & Corbett
# 2007). Within-country variation in y comes entirely from n_cert.

set.seed(1861)
B <- 999

# ---- Load firm counts at 2017 ----------------------------------------------
load_sbs_2017 <- function() {
  file_path <- file.path(data_dir, "enterprises_eu.xlsx")
  raw <- readxl::read_excel(file_path,
    sheet = "Sheet 1",
    skip = 9, col_names = FALSE
  )
  tibble::tibble(
    country    = raw[[1]],
    n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))
  ) %>%
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) %>%
    dplyr::mutate(country = dplyr::case_match(country,
      "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
      .default = country
    )) %>%
    dplyr::distinct(country, .keep_all = TRUE)
}
firm17 <- load_sbs_2017()

# ---- Outcome: levels of cert per 1000 firms 2017 --------------------------
attach_y_levels_norm <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::transmute(country, year, n_cert),
      by = c("country", "year")
    ) |>
    dplyr::inner_join(firm17, by = "country") |>
    dplyr::mutate(y = n_cert / n_firms_17 * 1000) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5, y) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

id_to_country <- setNames(country_labels, seq_along(country_labels))
N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")

panel_q <- keep_full(attach_y_levels_norm(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
))
panel_e <- keep_full(attach_y_levels_norm(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
))

cat(sprintf(
  "Quality: mean=%.2f, sd=%.2f, range=[%.2f, %.2f]\n",
  mean(panel_q$y), sd(panel_q$y), min(panel_q$y), max(panel_q$y)
))
cat(sprintf(
  "Environment: mean=%.2f, sd=%.2f, range=[%.2f, %.2f]\n",
  mean(panel_e$y), sd(panel_e$y), min(panel_e$y), max(panel_e$y)
))

W_combined <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(
    which(W_geo[i, ] > 0), which(W_trade[i, ] > 0),
    which(W_cul[i, ] > 0)
  ))
  W_combined[i, nbrs_i] <- 1 / length(nbrs_i)
}

bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se), sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se), sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se), sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se), sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on certification intensity in levels (cert per 1000 firms 2017)",
  label = "tab:bdf_firmnorm_levels"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-firmnorm-loglevels
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on log certification intensity.
# Outcome: y_it = log(1 + n_cert,it / n_firms,i,2017 * 1000)
# Combines firm-count normalisation (Albuquerque, Bronnenberg & Corbett
# 2007) with log transformation (variance stabilisation; standard in BDF
# adoption applications). Reuses panel_q, panel_e structure but with
# log-transformed outcome.

set.seed(1861)
B <- 999

attach_y_log_norm <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::transmute(country, year, n_cert),
      by = c("country", "year")
    ) |>
    dplyr::inner_join(firm17, by = "country") |>
    dplyr::mutate(y = log1p(n_cert / n_firms_17 * 1000)) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5, y) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

panel_q <- keep_full(attach_y_log_norm(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
))
panel_e <- keep_full(attach_y_log_norm(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
))

cat(sprintf(
  "Quality (log): mean=%.2f, sd=%.2f, range=[%.2f, %.2f]\n",
  mean(panel_q$y), sd(panel_q$y), min(panel_q$y), max(panel_q$y)
))
cat(sprintf(
  "Environment (log): mean=%.2f, sd=%.2f, range=[%.2f, %.2f]\n",
  mean(panel_e$y), sd(panel_e$y), min(panel_e$y), max(panel_e$y)
))

bdf_q_lag <- build_bdf_panel_lag(panel_q, W_q, x_vars)
bdf_e_lag <- build_bdf_panel_lag(panel_e, W_e, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

tab <- rbind(
  c(
    "Quality", "Endogenous",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se), sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se), sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se), sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se), sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log certification intensity ($\\log(1 + \\text{cert per 1000 firms 2017})$)",
  label = "tab:bdf_firmnorm_loglevels"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-firmnorm-loglevels-strong
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on log certification intensity, with STRONG-LINK
# recovered networks: only edges with |w_ij| > 0.7 are retained.
# Countries with no surviving links remain in the panel but have Gy = 0
# (isolated in the network sense).
#
# Rationale: a row-stochastic recovered W with weight > 0.7 indicates a
# dominant influence relationship — country i derives most of its peer
# input from j. Restricting to these "strong ties" (Granovetter 1973) tests
# whether the recovered structure carries explanatory power even when only
# the most concentrated influence channels are retained.
#
# Note: under thresholding without row-renormalisation, rowSums < 1 by
# construction; the BDF identification still goes through provided
# |rho| < 1 (Lee 2004 Econometrica), but Gy is no longer a simple peer
# average — it is a weighted sum that reflects influence concentration.
#
# Outcome: y_it = log(1 + n_cert,it / n_firms,i,2017 * 1000)

set.seed(1861)
B <- 999
N <- length(country_labels)

# ---- Apply strong-link threshold (NO row-renormalisation) -----------------
# Keep only |w_ij| > 0.7; isolated countries get Gy = 0 (their entire row is 0).
strong_link_W <- function(W, threshold = 0.7) {
  W2 <- W
  W2[abs(W2) <= threshold] <- 0
  diag(W2) <- 0
  W2
}

W_q_strong <- strong_link_W(W_q, threshold = 0.7)
W_e_strong <- strong_link_W(W_e, threshold = 0.7)
dimnames(W_q_strong) <- dimnames(W_e_strong) <-
  list(country_labels, country_labels)

# Diagnostic: how many strong links survive, how many countries isolated
n_q_links <- sum(W_q_strong > 0)
n_e_links <- sum(W_e_strong > 0)
isolated_q <- sum(rowSums(W_q_strong) == 0)
isolated_e <- sum(rowSums(W_e_strong) == 0)
mean_rsum_q <- mean(rowSums(W_q_strong))
mean_rsum_e <- mean(rowSums(W_e_strong))

cat(sprintf(
  "W_q (strong, |w|>0.7): %d edges, %d isolated countries, mean rowSum=%.3f\n",
  n_q_links, isolated_q, mean_rsum_q
))
cat(sprintf(
  "W_e (strong, |w|>0.7): %d edges, %d isolated countries, mean rowSum=%.3f\n",
  n_e_links, isolated_e, mean_rsum_e
))

# Stop early if too sparse — give user clear signal
if (n_q_links < 5 | n_e_links < 5) {
  warning("Very few strong links survived (<5). Consider lowering threshold.")
}

# ---- Outcome (reuse attach_y_log_norm, firm17 from previous chunks) ------
panel_q <- keep_full(attach_y_log_norm(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
))
panel_e <- keep_full(attach_y_log_norm(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
))

# ---- IMPORTANT: bypass row_normalise inside build_bdf_panel_lag ----------
# The default helper row-normalises W. We need to use the strong-link W
# *as is* (rowSums < 1 reflect concentration of influence). We define a
# version that skips row_normalise but otherwise replicates the helper.
build_bdf_panel_strong <- function(panel, W, x_vars) {
  G <- W # use as-is; do NOT row-normalise
  diag(G) <- 0
  G2 <- G %*% G
  G3 <- G2 %*% G
  d <- panel[order(panel$year, match(panel$country, rownames(G))), ]
  yrs <- sort(unique(d$year))
  d$Gy <- unlist(lapply(yrs, function(t) {
    yt <- d$y[d$year == t]
    as.numeric(G %*% yt)
  }))
  for (v in x_vars) {
    Xt <- matrix(d[[v]], ncol = length(yrs))
    d[[paste0("G_", v)]] <- as.numeric(G %*% Xt)
    d[[paste0("G2_", v)]] <- as.numeric(G2 %*% Xt)
    d[[paste0("G3_", v)]] <- as.numeric(G3 %*% Xt)
  }
  d <- d[order(d$country, d$year), ]
  d <- d |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      Gy_lag = dplyr::lag(Gy, n = 1, order_by = year),
      dplyr::across(
        dplyr::all_of(c(paste0("G2_", x_vars), paste0("G3_", x_vars))),
        \(z) dplyr::lag(z, n = 1, order_by = year),
        .names = "{.col}_lag"
      )
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(Gy_lag)
  d
}

# Build BDF panels: strong-link recovered + Combined ABC (still row-normalised)
bdf_q_lag <- build_bdf_panel_strong(panel_q, W_q_strong, x_vars)
bdf_e_lag <- build_bdf_panel_strong(panel_e, W_e_strong, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

# Diagnostic on isolated-country observations: their Gy = 0 contributes
# trivially. Report fraction.
cat(sprintf(
  "Quality (strong): %.1f%% of obs have Gy=0 (isolated countries)\n",
  100 * mean(bdf_q_lag$Gy == 0)
))
cat(sprintf(
  "Environment (strong): %.1f%% of obs have Gy=0 (isolated countries)\n",
  100 * mean(bdf_e_lag$Gy == 0)
))

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

tab <- rbind(
  c(
    "Quality", "Endogenous (strong, $|w|>0.7$)",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se), sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se), sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous (strong, $|w|>0.7$)",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se), sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se), sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log certification intensity, recovered networks restricted to strong ties ($|w_{ij}| > 0.7$). Isolated countries retained in the panel with $Gy_{it} = 0$. Combined Exogenous network row-normalised as before.",
  label = "tab:bdf_firmnorm_loglevels_strong"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

```{r}
#| label: bdf-strong-logdiff
#| echo: false
#| results: asis
#| cache: true
# BDF (2009) G2SLS on log-differences with STRONG-LINK recovered networks.
# Outcome: y_it = log(1 + n_cert,it) - log(1 + n_cert,i,t-1)
# Network: only edges with |w_ij| > 0.7 retained, NO row-renormalisation.
# Isolated countries kept in panel with Gy = 0.
# Combined ABC remains row-normalised as before.
#
# Specs:
#   (i)  no FE, asymptotic country-clustered SE
#   (ii) country FE, pairs cluster bootstrap on country (B = 999)

set.seed(1861)
B <- 999
N <- length(country_labels)

# ---- Strong-link extractor (no row-renormalisation) -----------------------
strong_link_W <- function(W, threshold = 0.7) {
  W2 <- W
  W2[abs(W2) <= threshold] <- 0
  diag(W2) <- 0
  W2
}

W_q_strong <- strong_link_W(W_q, threshold = 0.7)
W_e_strong <- strong_link_W(W_e, threshold = 0.7)
dimnames(W_q_strong) <- dimnames(W_e_strong) <-
  list(country_labels, country_labels)

cat(sprintf(
  "W_q (strong, |w|>0.7): %d edges, %d isolated, mean rowSum=%.3f\n",
  sum(W_q_strong > 0), sum(rowSums(W_q_strong) == 0),
  mean(rowSums(W_q_strong))
))
cat(sprintf(
  "W_e (strong, |w|>0.7): %d edges, %d isolated, mean rowSum=%.3f\n",
  sum(W_e_strong > 0), sum(rowSums(W_e_strong) == 0),
  mean(rowSums(W_e_strong))
))

# ---- Outcome builder: log-difference --------------------------------------
attach_y_diff <- function(dp_raw, certs) {
  dp_raw |>
    dplyr::mutate(country = id_to_country[as.character(id)], year = time) |>
    dplyr::select(country, year, x1, x2, x3, x4, x5) |>
    dplyr::inner_join(
      certs |>
        dplyr::filter(country %in% country_labels) |>
        dplyr::arrange(country, year) |>
        dplyr::group_by(country) |>
        dplyr::mutate(
          y = log1p(n_cert) -
            log1p(dplyr::lag(n_cert, n = 1, order_by = year))
        ) |>
        dplyr::ungroup() |>
        dplyr::transmute(country, year, y),
      by = c("country", "year")
    ) |>
    tidyr::drop_na() |>
    dplyr::filter(country %in% country_labels)
}

panel_q <- keep_full(attach_y_diff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
))
panel_e <- keep_full(attach_y_diff(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
))

cat(sprintf(
  "Quality (log-diff): mean=%.3f, sd=%.3f, range=[%.2f, %.2f]\n",
  mean(panel_q$y), sd(panel_q$y), min(panel_q$y), max(panel_q$y)
))
cat(sprintf(
  "Environment (log-diff): mean=%.3f, sd=%.3f, range=[%.2f, %.2f]\n",
  mean(panel_e$y), sd(panel_e$y), min(panel_e$y), max(panel_e$y)
))

# ---- BDF panel builder for strong-link W (no row-renormalisation) --------
build_bdf_panel_strong <- function(panel, W, x_vars) {
  G <- W
  diag(G) <- 0
  G2 <- G %*% G
  G3 <- G2 %*% G
  d <- panel[order(panel$year, match(panel$country, rownames(G))), ]
  yrs <- sort(unique(d$year))
  d$Gy <- unlist(lapply(yrs, function(t) {
    yt <- d$y[d$year == t]
    as.numeric(G %*% yt)
  }))
  for (v in x_vars) {
    Xt <- matrix(d[[v]], ncol = length(yrs))
    d[[paste0("G_", v)]] <- as.numeric(G %*% Xt)
    d[[paste0("G2_", v)]] <- as.numeric(G2 %*% Xt)
    d[[paste0("G3_", v)]] <- as.numeric(G3 %*% Xt)
  }
  d <- d[order(d$country, d$year), ]
  d <- d |>
    dplyr::group_by(country) |>
    dplyr::mutate(
      Gy_lag = dplyr::lag(Gy, n = 1, order_by = year),
      dplyr::across(
        dplyr::all_of(c(paste0("G2_", x_vars), paste0("G3_", x_vars))),
        \(z) dplyr::lag(z, n = 1, order_by = year),
        .names = "{.col}_lag"
      )
    ) |>
    dplyr::ungroup() |>
    tidyr::drop_na(Gy_lag)
  d
}

# Combined ABC: keep row-normalised (build_bdf_panel_lag does it)
bdf_q_lag <- build_bdf_panel_strong(panel_q, W_q_strong, x_vars)
bdf_e_lag <- build_bdf_panel_strong(panel_e, W_e_strong, x_vars)
bdf_comb_q_lag <- build_bdf_panel_lag(panel_q, W_combined, x_vars)
bdf_comb_e_lag <- build_bdf_panel_lag(panel_e, W_combined, x_vars)

cat(sprintf(
  "Quality (strong, log-diff): %.1f%% obs have Gy=0\n",
  100 * mean(bdf_q_lag$Gy == 0)
))
cat(sprintf(
  "Environment (strong, log-diff): %.1f%% obs have Gy=0\n",
  100 * mean(bdf_e_lag$Gy == 0)
))

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy_lag ~ %s + %s",
  rhs, exo, inst2_lag, inst3_lag
))

nofe_q_rec <- extract_nofe(feols(f_nofe, data = bdf_q_lag, cluster = ~country))
nofe_q_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_q_lag, cluster = ~country))
nofe_e_rec <- extract_nofe(feols(f_nofe, data = bdf_e_lag, cluster = ~country))
nofe_e_abc <- extract_nofe(feols(f_nofe, data = bdf_comb_e_lag, cluster = ~country))

cfe_q_rec <- extract_cfe(cluster_boot(bdf_q_lag, f_cfe, "country", B = B))
cfe_q_abc <- extract_cfe(cluster_boot(bdf_comb_q_lag, f_cfe, "country", B = B))
cfe_e_rec <- extract_cfe(cluster_boot(bdf_e_lag, f_cfe, "country", B = B))
cfe_e_abc <- extract_cfe(cluster_boot(bdf_comb_e_lag, f_cfe, "country", B = B))

tab <- rbind(
  c(
    "Quality", "Endogenous (strong, $|w|>0.7$)",
    sprintf("%.3f%s", nofe_q_rec$b, sig_stars(nofe_q_rec$p)),
    sprintf("%.3f%s", cfe_q_rec$b, sig_stars(cfe_q_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_rec$se), sprintf("(%.3f)", cfe_q_rec$se)
  ),
  c(
    "Quality", "Combined Exogenous",
    sprintf("%.3f%s", nofe_q_abc$b, sig_stars(nofe_q_abc$p)),
    sprintf("%.3f%s", cfe_q_abc$b, sig_stars(cfe_q_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_q_abc$se), sprintf("(%.3f)", cfe_q_abc$se)
  ),
  c(
    "Environment", "Endogenous (strong, $|w|>0.7$)",
    sprintf("%.3f%s", nofe_e_rec$b, sig_stars(nofe_e_rec$p)),
    sprintf("%.3f%s", cfe_e_rec$b, sig_stars(cfe_e_rec$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_rec$se), sprintf("(%.3f)", cfe_e_rec$se)
  ),
  c(
    "Environment", "Combined Exogenous",
    sprintf("%.3f%s", nofe_e_abc$b, sig_stars(nofe_e_abc$p)),
    sprintf("%.3f%s", cfe_e_abc$b, sig_stars(cfe_e_abc$p))
  ),
  c(
    "", "",
    sprintf("(%.3f)", nofe_e_abc$se), sprintf("(%.3f)", cfe_e_abc$se)
  )
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)",
  "$\\hat\\beta$ (country FE)"
)

kbl(tab,
  booktabs = TRUE, format = "latex", escape = FALSE,
  align = "llrr", row.names = FALSE,
  caption = "Endogenous peer effect on log-differences, recovered networks restricted to strong ties ($|w_{ij}| > 0.7$). Isolated countries retained with $Gy_{it} = 0$.",
  label = "tab:bdf_strong_logdiff"
) |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```