---
title: "Thesis Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
execute:
  warning: false
  message: false
format:
  pdf:
    output-file: "results.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---
```{r}
#| label: setup
#| include: false
data_dir <- "/workspaces/thesis/data"
clean_dir <- "/workspaces/thesis/data/cleaned_data"

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)
library(patchwork)
library(ggraph)
library(ggrepel)
library(tidygraph)
library(maps)
library(countrycode)
library(fixest)

# Source all helper functions (documented in thesis_helpers.R)
source("/workspaces/thesis/thesis_helpers.R")

set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
fig_dir <- file.path(out_dir, "fig")
tab_dir <- file.path(out_dir, "tab")
for (d in c(fig_dir, tab_dir)) if (!dir.exists(d)) dir.create(d, recursive = TRUE)

# Helper: append a "Notes:" footnote to any kable/kableExtra object.
# Replace the `note = "SAMPLE NOTE"` argument at each call site with the
# table-specific text.
add_note <- function(k, note = "SAMPLE NOTE") {
  kableExtra::footnote(
    k,
    general = note,
    general_title = "Notes:",
    threeparttable = TRUE,
    footnote_as_chunk = TRUE,
    escape = FALSE
  )
}
```

```{r}
#| label: load-best
#| include: false
#| cache: true
best <- readRDS(file.path(out_dir, "best_per_panel.rds"))
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
diag(W_q) <- 0
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)
diag(W_e) <- 0

sync_eu <- readRDS(file.path(clean_dir, "sync_eu.rds"))
country_labels <- sort(sync_eu)
rownames(W_q) <- colnames(W_q) <- country_labels
rownames(W_e) <- colnames(W_e) <- country_labels
```

# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Descriptive Statistics

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
#| cache: true
f9001 <- readRDS(file.path(clean_dir, "f9001.rds"))
f14001 <- readRDS(file.path(clean_dir, "f14001.rds"))

desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, countries, years) {
  df |>
    dplyr::filter(country %in% countries, year %in% years) |>
    dplyr::group_by(year) |>
    dplyr::summarise(
      total_cert     = sum(n_cert, na.rm = TRUE),
      mean_cert      = mean(n_cert, na.rm = TRUE),
      sd_cert        = sd(n_cert, na.rm = TRUE),
      min_cert       = min(n_cert, na.rm = TRUE),
      max_cert       = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries    = dplyr::n(),
      .groups        = "drop"
    )
}

d9 <- desc_tab(f9001, country_labels, desc_years)
d14 <- desc_tab(f14001, country_labels, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

# Note string for this table. Replace "SAMPLE NOTE" with the actual text.
desc_stats_note <- "This table summarises the distribution of certifications across the 27 EU countries in the three selected years. The first row reports the number of countries with non-missing data in each year, which is the basis for all subsequent statistics. The second row reports the total number of certifications across all countries, while the third and fourth rows report the mean and standard deviation of certifications per country, respectively. The fifth row reports the minimum and maximum number of certifications across countries. Finally, the sixth row reports the mean certification intensity, defined as the number of certifications per thousand dollars of GDP per capita, averaged across countries. Ennvironment 1993 is empty as its adoption started in 1997.
Sources: ISO Survey, World Bank."

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Descriptive statistics}",
  "\\label{tab:desc_stats}",
  "\\begin{tabular}{lcccccc}\\hline\\hline",
  "& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\",
  "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
  "& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\\hline",
  sprintf(
    "Countries Number & %d & %d & %d & & %d & %d \\\\",
    d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
    d14$n_countries[1], d14$n_countries[2]
  ),
  sprintf(
    "Total & %s & %s & %s & & %s & %s \\\\",
    f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
    f0(d14$total_cert[1]), f0(d14$total_cert[2])
  ),
  sprintf(
    "Mean & %s & %s & %s & & %s & %s \\\\",
    f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
    f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
  ),
  sprintf(
    "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\",
    f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
    f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
  ),
  sprintf(
    "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\",
    f0(d9$min_cert[1]), f0(d9$max_cert[1]),
    f0(d9$min_cert[2]), f0(d9$max_cert[2]),
    f0(d9$min_cert[3]), f0(d9$max_cert[3]),
    f0(d14$min_cert[1]), f0(d14$max_cert[1]),
    f0(d14$min_cert[2]), f0(d14$max_cert[2])
  ),
  sprintf(
    "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\",
    f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
    f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
  ),
  "\\hline\\hline",
  sprintf(
    "\\multicolumn{7}{p{0.92\\textwidth}}{\\footnotesize \\textit{Notes:} %s} \\\\",
    desc_stats_note
  ),
  "\\end{tabular}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "desc_stats.tex"))
cat(tex, sep = "\n")
```

## Network Recovery and Dyadic Analysis

\subsection*{Methodology}

De Paula, Rasul \& Souza (2025) extend the peer-effects literature (Manski 1993; Bramoullé et al., 2009) to the case in which the network of interactions is unknown. Their structural model
\begin{equation}
y_{it} = \rho_0 \sum_{j=1}^{N} W_{0,ij}\, y_{jt} + \beta_0 x_{it} + \gamma_0 \sum_{j=1}^{N} W_{0,ij}\, x_{jt} + \alpha_i + \alpha_t + \varepsilon_{it}
\label{eq:structural}
\end{equation}
admits the compact form $y_t = \rho_0 W_0 y_t + \beta_0 x_t + \gamma_0 W_0 x_t + \varepsilon_t$, which rearranges into the reduced form
\begin{equation}
y_t = \Pi_0 x_t + \nu_t, \qquad \Pi_0 = (I - \rho_0 W_0)^{-1}(\beta_0 I + \gamma_0 W_0).
\label{eq:reduced}
\end{equation}
If $W_0$ were known, equation~\eqref{eq:reduced} would directly identify the structural parameters $(\rho_0,\beta_0,\gamma_0)$ from a reduced-form regression of outcomes on covariates (Bramoullé et al., 2009). The contribution of De Paula, Rasul \& Souza is to show that, when $W_0$ is unknown, $\Pi_0$ still contains enough information to recover both the network and the parameters jointly.

The intuition is the one that underlies any diffusion process. If a contagious disease spreads through a population, the people who fall ill at the same time tend to be the ones who interact: knowing who got sick when, and what their characteristics are, one can work backwards to who must have been in contact with whom. The same logic applies here. If four countries adopt a standard in synchronous patterns that cannot be explained by their own characteristics alone, the synchrony itself is evidence of a link between them. Cross-country covariation in $y$, conditional on $x$, identifies who is connected to whom, and the strength of the covariation, suitably disentangled, identifies how strong the link is.

In practice, $W_0$ is an unkown matrix and $\Pi_0$ is observable. De Paula, Rasul \& Souza exploit the adaptive elastic net (Caner \& Zhang 2014), which simultaneously selects the non-zero entries of $\hat W$ and estimates their values together with $(\hat\rho,\hat\beta,\hat\gamma)$. The result is an interaction network identified directly from observable cross-sectional spillovers, without requiring a priori commitment to any specific weighting scheme.

```{r}
#| label: network-stats-table
#| echo: false
#| results: asis
#| cache: true
s_q <- net_stats(W_q)
s_e <- net_stats(W_e)

B_q <- (abs(W_q) > 1e-5) * 1
B_e <- (abs(W_e) > 1e-5) * 1
common <- sum(B_q & B_e)
only_q <- sum(B_q & !B_e)
only_e <- sum(!B_q & B_e)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

rho_q <- best[["9001_eu"]]$rn$unpenalisedgmm$rho
rho_e <- best[["14001_eu"]]$rn$unpenalisedgmm$rho
lam_q <- best[["9001_eu"]]$lambda
lam_e <- best[["14001_eu"]]$lambda
bic_q <- best[["9001_eu"]]$bic
bic_e <- best[["14001_eu"]]$bic

# Note string for this table. Replace "SAMPLE NOTE" with the actual text.
network_stats_note <- "This table reports summary statistics of the two recovered networks, selected by BIC. The first row reports the number of edges in each network, while the second and third rows report how many edges are shared by both networks and how many are unique to each network, respectively. A shared link is intendended when there is a link between the same two nodes in the same direction in both networks. The fourth and fifth rows report the clustering coefficient (share of closed triplets) and reciprocity (bidirectional edges) of each network, respectively. The sixth to ninth rows report the mean and standard deviation of out-degrees and in-degrees across nodes (countries) for each network. Finally, the tenth to twelfth rows report the GMM estimates of the structural parameters $\\hat{\\rho}$ (endogenous peer effects) and $\\hat{\\lambda}$ (penalties for the Adaptive Elastic Net) and the BIC values for each network.
Sources: Elaboration on ISO Survey Data, the World Bank Data and the Worldwide Governance Indicators. The network estimation relies on GDP per capita (WDI: NY.GDP.PCAP.KD), Trade (\\% of GDP; WDI: NE.TRD.GNFS.ZS), Rule of Law estimate (WGI: RL.EST)*, Total Natural Resources Rents (\\% of GDP; WDI: NY.GDP.TOTL.RT.ZS), and Share of Renewables in Total Final Energy Consumption (\\% ; WDI: EG.FEC.RNEW.ZS). All the covariates were log linearised (log (1+x))."

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{BIC-selected networks}",
  "\\label{tab:network_eu}",
  "\\begin{tabular}{lcc}\\hline\\hline",
  "& Quality & Environment \\\\\\hline",
  sprintf("Number of edges & %d & %d \\\\", s_q$edges, s_e$edges),
  sprintf("Edges in both networks & %d & %d \\\\", common, common),
  sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\", only_q),
  sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\", only_e),
  sprintf("Clustering & %s & %s \\\\", f3(s_q$clustering), f3(s_e$clustering)),
  sprintf("Reciprocated edges & %s & %s \\\\", p1(s_q$reciprocity), p1(s_e$reciprocity)),
  "Degree distribution across nodes (countries) & & \\\\",
  sprintf(
    "\\quad Out-degree & %s & %s \\\\",
    dg(s_q$out_mean, s_q$out_sd), dg(s_e$out_mean, s_e$out_sd)
  ),
  sprintf(
    "\\quad In-degree & %s & %s \\\\",
    dg(s_q$in_mean, s_q$in_sd), dg(s_e$in_mean, s_e$in_sd)
  ),
  "\\hline",
  sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\",
    formatC(rho_q, format = "f", digits = 4),
    formatC(rho_e, format = "f", digits = 4)
  ),
  sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\",
    paste(formatC(lam_q, format = "f", digits = 3), collapse = ", "),
    paste(formatC(lam_e, format = "f", digits = 3), collapse = ", ")
  ),
  sprintf(
    "BIC & %s & %s \\\\",
    formatC(bic_q, format = "f", digits = 2),
    formatC(bic_e, format = "f", digits = 2)
  ),
  "\\hline\\hline",
  sprintf(
    "\\multicolumn{3}{p{0.95\\textwidth}}{\\footnotesize \\textit{Notes:} %s} \\\\",
    network_stats_note
  ),
  "\\end{tabular}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "network_stats.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: fig-degree-hist
#| echo: false
#| fig-width: 9
#| fig-height: 3.5
#| fig-cap: "Unweighted degree distributions of recovered networks"
eps <- 1e-5
B_q <- (W_q > eps) * 1
diag(B_q) <- 0
B_e <- (W_e > eps) * 1
diag(B_e) <- 0

deg_df <- dplyr::bind_rows(
  data.frame(
    standard = "Quality (ISO 9001)",
    direction = "Out-degree", value = rowSums(B_q)
  ),
  data.frame(
    standard = "Quality (ISO 9001)",
    direction = "In-degree", value = colSums(B_q)
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    direction = "Out-degree", value = rowSums(B_e)
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    direction = "In-degree", value = colSums(B_e)
  )
) |>
  dplyr::mutate(
    standard = factor(standard,
      levels = c("Quality (ISO 9001)", "Environment (ISO 14001)")
    ),
    direction = factor(direction, levels = c("Out-degree", "In-degree"))
  )

p <- ggplot(deg_df, aes(x = value, fill = standard)) +
  geom_histogram(
    position = "identity", alpha = 0.55,
    colour = "white", linewidth = 0.3, bins = 12
  ) +
  geom_rug(aes(colour = standard),
    sides = "b",
    length = unit(0.03, "npc"), alpha = 0.7
  ) +
  facet_wrap(~direction, scales = "free", ncol = 2) +
  scale_fill_manual(values = c(
    "Quality (ISO 9001)"      = "#3182bd",
    "Environment (ISO 14001)" = "#31a354"
  )) +
  scale_colour_manual(values = c(
    "Quality (ISO 9001)"      = "#08519c",
    "Environment (ISO 14001)" = "#006d2c"
  )) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.02))) +
  labs(x = "Degree", y = "Count", fill = NULL, colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = "bottom",
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )


ggsave(file.path(fig_dir, "degree_hist.pdf"), p, width = 9, height = 3.5)
print(p)
```

```{r}
#| label: jaccard-compute
#| include: false
eps <- 1e-5
B_q <- (abs(W_q) > eps) * 1
B_e <- (abs(W_e) > eps) * 1
diag(B_q) <- diag(B_e) <- 0

n_q <- sum(B_q)
n_e <- sum(B_e)
n_inter <- sum(B_q & B_e)
n_union <- sum(B_q | B_e)
jaccard <- n_inter / n_union

jaccard_pct <- formatC(100 * jaccard, format = "f", digits = 1)
jaccard_val <- formatC(jaccard, format = "f", digits = 3)
```

### Network similarity

To assess how closely the two recovered networks $\hat W^Q$ and $\hat W^E$ overlap, we summarise their structural similarity through the **Jaccard index** (Jaccard 1901; Wasserman \& Faust 1994, Sec.\ 11.4), the share of directed edges shared by both networks relative to the union of their edge sets:

\begin{equation}
J(\hat W^Q, \hat W^E) = \frac{|E_Q \cap E_E|}{|E_Q \cup E_E|},
\label{eq:jaccard}
\end{equation}

where $E_Q = \{(i, j) : |\hat w^Q_{ij}| > 10^{-5}\}$ and $E_E$ is defined analogously. Edges are treated as directed, so that a link $i \to j$ and the reciprocal $j \to i$ count as distinct elements of the edge sets. The index ranges in $[0, 1]$, with $J = 0$ for disjoint networks (no edges in common) and $J = 1$ for identical edge sets.

For the BIC-selected recovered networks, $|E_Q| = `r n_q`$, $|E_E| = `r n_e`$, $|E_Q \cap E_E| = `r n_inter`$, and $|E_Q \cup E_E| = `r n_union`$, yielding $J = `r jaccard_val`$.

### Network visualisation

```{r}
#| label: fig-networks-weighted-full
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Recovered networks"
make_full_weighted_graph <- function(W, labels, eps = 1e-5) {
  W2 <- W
  W2[abs(W2) < eps] <- 0
  g <- igraph::graph_from_adjacency_matrix(W2, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(W_q, country_labels)
g_e_full <- make_full_weighted_graph(W_e, country_labels)

w_max <- max(c(
  abs(igraph::E(g_q_full)$weight),
  abs(igraph::E(g_e_full)$weight)
), 1e-9)

set.seed(1861)
layout_q <- igraph::layout_with_fr(g_q_full)
set.seed(1861)
layout_e <- igraph::layout_with_fr(g_e_full)
rownames(layout_q) <- rownames(layout_e) <- country_labels

plot_net_gg <- function(g, lo, title, w_max) {
  lo_ordered <- lo[igraph::V(g)$name, , drop = FALSE]

  ggraph(g, layout = lo_ordered) +
    geom_edge_arc(
      aes(
        alpha = abs(weight),
        colour = abs(weight) > 0.9
      ),
      width = 0.5,
      arrow = arrow(
        length = unit(2, "mm"),
        type = "closed", angle = 22
      ),
      end_cap = circle(5, "mm"),
      start_cap = circle(5, "mm"),
      strength = 0.12
    ) +
    geom_node_point(aes(colour = igraph::degree(g) > 0),
      size = 4.5, show.legend = FALSE
    ) +
    geom_node_text(aes(label = name),
      repel = TRUE,
      size = 2.6,
      bg.colour = "white", bg.r = 0.18,
      segment.colour = "grey55",
      segment.size = 0.25,
      segment.alpha = 0.7,
      max.overlaps = Inf,
      force = 3, force_pull = 0.5,
      point.padding = unit(0.6, "lines"),
      box.padding = unit(0.45, "lines"),
      min.segment.length = 0
    ) +
    scale_edge_alpha(range = c(0.35, 0.9), limits = c(0, w_max), guide = "none") +
    scale_edge_colour_manual(
      values = c(
        `TRUE` = "#67000D",
        `FALSE` = "grey60"
      ),
      labels = c(`TRUE` = "Strong ties", `FALSE` = "Weak ties"),
      guide = guide_legend(
        title = NULL,
        override.aes = list(width = 2, alpha = 0.7),
        position = "bottom"
      )
    ) +
    scale_colour_manual(values = c(`TRUE` = "steelblue", `FALSE` = "grey85")) +
    ggtitle(title) +
    theme_void(base_size = 11) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      plot.margin = margin(8, 8, 8, 8)
    )
}

p_qq <- plot_net_gg(g_q_full, layout_q, "Quality - Quality layout", w_max)
p_eq <- plot_net_gg(g_e_full, layout_q, "Environment - Quality layout", w_max)
p_qe <- plot_net_gg(g_q_full, layout_e, "Quality - Environment layout", w_max)
p_ee <- plot_net_gg(g_e_full, layout_e, "Environment - Environment layout", w_max)

panel <- (p_qq | p_eq) / (p_qe | p_ee)

ggsave(file.path(fig_dir, "networks_weighted_full.pdf"), panel,
  width = 13, height = 13
)
print(panel)
```

### Dyadic regression
```{r}
#| label: dyadic-build-covariates
#| echo: false
#| include: false
#| cache: true
# Build all dyadic covariates AND all exogenous networks (Albuquerque single-
# and two-channel). Done in one chunk because both the dyadic regressions
# and the BDF applications below need W_geotrade and W_geoculture.

N <- length(country_labels)
cl_iso3 <- countrycode::countrycode(
  country_labels,
  origin = "country.name", destination = "iso3c", warn = FALSE
)
names(cl_iso3) <- country_labels
iso_to_label <- setNames(names(cl_iso3), cl_iso3)

# ---- Geographic distance (capitals via maps + haversine) -----------------
data("world.cities", package = "maps")
capitals <- world.cities |>
  dplyr::filter(capital == 1) |>
  dplyr::mutate(iso3 = countrycode::countrycode(
    country.etc,
    origin = "country.name", destination = "iso3c", warn = FALSE
  )) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::group_by(iso3) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::transmute(country, iso3, lat, lon = long) |>
  dplyr::arrange(match(country, country_labels))

D_geo <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  for (j in seq_len(N)) {
    D_geo[i, j] <- haversine_km(
      capitals$lat[i], capitals$lon[i],
      capitals$lat[j], capitals$lon[j]
    )
  }
}
log_dist <- log(D_geo + 1)
diag(log_dist) <- 0

# ---- Cultural distance (Hofstede four-dimension Euclidean) ---------------
hofstede_path <- "/workspaces/thesis/data/6-dimensions-for-website-2015-08-16(1).csv"
hof_raw <- tryCatch(
  readr::read_csv2(hofstede_path, show_col_types = FALSE) |>
    janitor::clean_names(),
  error = function(e) {
    readr::read_csv(hofstede_path, show_col_types = FALSE) |>
      janitor::clean_names()
  }
) |>
  dplyr::mutate(dplyr::across(
    dplyr::everything(),
    \(x) ifelse(as.character(x) == "#NULL!", NA, x)
  ))

candidates <- c("country", "ctr", "iso", "iso3")
country_col <- candidates[candidates %in% names(hof_raw)][1]
to_iso3 <- function(x) {
  x <- as.character(x)
  out <- suppressWarnings(countrycode::countrycode(
    x,
    origin = "country.name", destination = "iso3c", warn = FALSE
  ))
  if (mean(is.na(out)) > 0.5) {
    out <- suppressWarnings(countrycode::countrycode(
      x,
      origin = "iso2c", destination = "iso3c", warn = FALSE
    ))
  }
  if (mean(is.na(out)) > 0.5) out <- toupper(x)
  out
}
hof <- hof_raw |>
  dplyr::mutate(iso3 = to_iso3(.data[[country_col]])) |>
  dplyr::filter(iso3 %in% cl_iso3) |>
  dplyr::transmute(
    iso3,
    pdi = suppressWarnings(as.numeric(pdi)),
    idv = suppressWarnings(as.numeric(idv)),
    mas = suppressWarnings(as.numeric(mas)),
    uai = suppressWarnings(as.numeric(uai))
  ) |>
  dplyr::filter(!is.na(pdi) & !is.na(idv) & !is.na(mas) & !is.na(uai)) |>
  dplyr::distinct(iso3, .keep_all = TRUE)
if (!"CYP" %in% hof$iso3 && "GRC" %in% hof$iso3 && "CYP" %in% cl_iso3) {
  hof <- dplyr::bind_rows(
    hof,
    hof |> dplyr::filter(iso3 == "GRC") |> dplyr::mutate(iso3 = "CYP")
  )
}
missing_hof <- setdiff(cl_iso3, hof$iso3)
if (length(missing_hof) > 0) {
  hof_mean <- hof |>
    dplyr::summarise(dplyr::across(
      c(pdi, idv, mas, uai),
      \(x) mean(x, na.rm = TRUE)
    ))
  for (iso in missing_hof) {
    hof <- dplyr::bind_rows(
      hof,
      tibble::tibble(
        iso3 = iso, pdi = hof_mean$pdi, idv = hof_mean$idv,
        mas = hof_mean$mas, uai = hof_mean$uai
      )
    )
  }
}
hof <- hof |>
  dplyr::mutate(country = iso_to_label[iso3]) |>
  dplyr::filter(!is.na(country) & country %in% country_labels) |>
  dplyr::arrange(match(country, country_labels))
H_mat <- as.matrix(hof[, c("pdi", "idv", "mas", "uai")])
rownames(H_mat) <- hof$country
D_cul <- as.matrix(dist(H_mat, method = "euclidean"))
D_cul <- D_cul[country_labels, country_labels]

# ---- GDP per capita asymmetry (WDI NY.GDP.PCAP.KD) -----------------------
D_gdppc <- build_dist_wdi(
  file.path(data_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"),
  c("x1996", "x1997", "x1998", "x1999", "x2000")
)

# ---- Env tax (Eurostat env_ac_taxind2) -----------------------------------
envtax_raw <- readr::read_csv(
  file.path(data_dir, "estat_env_ac_taxind2_en.csv.gz"),
  show_col_types = FALSE
)
envtax <- envtax_raw |>
  dplyr::filter(
    tax == "ENV", unit == "MIO_EUR", nace_r2 == "TOTAL",
    TIME_PERIOD %in% 1996:2000
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    iso3 = iso2_to_iso3(geo),
    country = iso_to_label[iso3]
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))
if (nrow(envtax) < N) {
  miss <- setdiff(country_labels, envtax$country)
  envtax <- dplyr::bind_rows(
    envtax,
    data.frame(
      geo = NA, envtax = mean(envtax$envtax, na.rm = TRUE),
      iso3 = NA, country = miss, stringsAsFactors = FALSE
    )
  )
}
envtax <- envtax |> dplyr::arrange(match(country, country_labels))

# Labour force (WDI SL.TLF.TOTL.IN, used as population proxy)
lab_raw <- readr::read_csv(
  file.path(data_dir, "API_SL.TLF.TOTL.IN_DS2_en_csv_v2_196.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    pop = rowMeans(dplyr::across(
      dplyr::all_of(c("x1996", "x1997", "x1998", "x1999", "x2000")),
      \(z) suppressWarnings(as.numeric(z))
    ), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(pop), is.finite(pop)) |>
  dplyr::mutate(country = iso_to_label[iso3])

envtax_pw <- envtax |>
  dplyr::select(country, envtax) |>
  dplyr::filter(!is.na(envtax)) |>
  dplyr::inner_join(lab_raw, by = "country") |>
  dplyr::mutate(envtax_per_worker = envtax / pop) |>
  dplyr::arrange(match(country, country_labels))

if (nrow(envtax_pw) < N) {
  miss <- setdiff(country_labels, envtax_pw$country)
  envtax_pw <- dplyr::bind_rows(
    envtax_pw,
    data.frame(
      country = miss, envtax = NA, iso3 = NA, pop = NA,
      envtax_per_worker = mean(envtax_pw$envtax_per_worker, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}
pw_vec <- setNames(log(envtax_pw$envtax_per_worker + 1e-12), envtax_pw$country)
D_envtax_pw <- abs(outer(pw_vec, pw_vec, FUN = "-"))
dimnames(D_envtax_pw) <- list(country_labels, country_labels)

# ---- Bilateral FDI (Eurostat bop_fdi6_pos) -------------------------------
fdi_raw <- readr::read_csv(
  file.path(data_dir, "estat_bop_fdi6_pos_en.csv.gz"),
  show_col_types = FALSE
)
fdi <- fdi_raw |>
  dplyr::filter(
    counterp == "IMM", nace_r2 == "FDI", stk_flow == "ASS",
    entity == "TOTAL", fdi_item == "DI__D__F",
    currency == "MIO_EUR", TIME_PERIOD %in% 2013:2020
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi_value = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    geo_iso3 = iso2_to_iso3(geo),
    partner_iso3 = iso2_to_iso3(partner)
  ) |>
  dplyr::filter(geo_iso3 %in% cl_iso3, partner_iso3 %in% cl_iso3) |>
  dplyr::mutate(
    sender = iso_to_label[geo_iso3],
    receiver = iso_to_label[partner_iso3]
  )
F_fdi <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(fdi))) {
  F_fdi[fdi$sender[k], fdi$receiver[k]] <- pmax(fdi$fdi_value[k], 0)
}
log_fdi <- log1p(F_fdi)
diag(log_fdi) <- 0

# ---- Trade matrix (BACI 1996) + Albuquerque top-3 single-channel networks ---------
baci_path <- "/workspaces/thesis/data/BACI_HS92_Y1996_V202601.csv"
ccodes_path <- "/workspaces/thesis/data/country_codes_V202601.csv"
TOP_N <- 3

baci <- readr::read_csv(baci_path, show_col_types = FALSE)
ccodes <- readr::read_csv(ccodes_path, show_col_types = FALSE)

iso_map <- ccodes |>
  dplyr::select(country_code, country_iso3) |>
  dplyr::filter(country_iso3 %in% cl_iso3)

trade <- baci |>
  dplyr::select(i, j, v) |>
  dplyr::inner_join(iso_map, by = c("i" = "country_code")) |>
  dplyr::rename(exporter_iso = country_iso3) |>
  dplyr::inner_join(iso_map, by = c("j" = "country_code")) |>
  dplyr::rename(importer_iso = country_iso3) |>
  dplyr::mutate(
    exporter = iso_to_label[exporter_iso],
    importer = iso_to_label[importer_iso]
  ) |>
  dplyr::group_by(exporter, importer) |>
  dplyr::summarise(value = sum(v, na.rm = TRUE), .groups = "drop")

T_trade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(trade))) {
  T_trade[trade$exporter[k], trade$importer[k]] <- trade$value[k]
}

W_geo <- top_n_network(D_geo, top_n = TOP_N, similarity = FALSE)
W_trade <- top_n_network(T_trade, top_n = TOP_N, similarity = TRUE)
W_cul <- top_n_network(D_cul, top_n = TOP_N, similarity = FALSE)

saveRDS(
  list(geo = W_geo, trade = W_trade, culture = W_cul),
  file.path(out_dir, "exogenous_networks.rds")
)

# ---- Albuquerque two-channel networks --------------------------------------------
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs_i] <- 1 / length(nbrs_i)
}

W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs_i] <- 1 / length(nbrs_i)
}
```


```{r}
#| label: dyadic-fullperiod-robustness
#| echo: false
#| results: asis
#| cache: true
# Robustness check: rebuild all dyadic covariates using full-sample period
# averages (1993-2017) instead of pre-treatment averages (1996-2000).
# For each covariate, take the mean of all available years in [1993, 2017];
# countries with no data in the window get the sample-mean imputation.
# Static covariates (cultural distance from Hofstede, geographic distance
# from capitals) are unchanged. FDI is data-limited to 2013-2020 in Eurostat
# bilateral series.

dict_full <- c(
  d_econ   = "Economic distance",
  d_envir  = "Env. policy asymmetry",
  d_cult   = "Cultural distance",
  d_geo    = "Geographic distance",
  l_fdi    = "Bilateral FDI",
  sender   = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

full_years_wdi <- paste0("x", 1993:2017)
full_years_estat <- 1993:2017

# ---- GDP per capita asymmetry (WDI NY.GDP.PCAP.KD, 1993-2017) ------------
D_gdppc_fp <- build_dist_wdi(
  file.path(data_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"),
  full_years_wdi
)

# ---- Env tax per worker (Eurostat env_ac_taxind2 / WDI labour force, 1993-2017)
envtax_fp <- envtax_raw |>
  dplyr::filter(
    tax == "ENV", unit == "MIO_EUR", nace_r2 == "TOTAL",
    TIME_PERIOD %in% full_years_estat
  ) |>
  dplyr::mutate(OBS_VALUE = suppressWarnings(as.numeric(OBS_VALUE))) |>
  dplyr::filter(!is.na(OBS_VALUE)) |>
  dplyr::group_by(geo) |>
  dplyr::summarise(envtax = mean(OBS_VALUE, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(
    iso3 = iso2_to_iso3(geo),
    country = iso_to_label[iso3]
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(country))
if (nrow(envtax_fp) < N) {
  miss <- setdiff(country_labels, envtax_fp$country)
  envtax_fp <- dplyr::bind_rows(
    envtax_fp,
    data.frame(
      geo = NA, envtax = mean(envtax_fp$envtax, na.rm = TRUE),
      iso3 = NA, country = miss, stringsAsFactors = FALSE
    )
  )
}
envtax_fp <- envtax_fp |> dplyr::arrange(match(country, country_labels))

lab_raw_fp <- readr::read_csv(
  file.path(data_dir, "API_SL.TLF.TOTL.IN_DS2_en_csv_v2_196.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    pop = rowMeans(dplyr::across(
      dplyr::any_of(full_years_wdi),
      \(z) suppressWarnings(as.numeric(z))
    ), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, !is.na(pop), is.finite(pop)) |>
  dplyr::mutate(country = iso_to_label[iso3])

envtax_pw_fp <- envtax_fp |>
  dplyr::select(country, envtax) |>
  dplyr::filter(!is.na(envtax)) |>
  dplyr::inner_join(lab_raw_fp, by = "country") |>
  dplyr::mutate(envtax_per_worker = envtax / pop) |>
  dplyr::arrange(match(country, country_labels))

if (nrow(envtax_pw_fp) < N) {
  miss <- setdiff(country_labels, envtax_pw_fp$country)
  envtax_pw_fp <- dplyr::bind_rows(
    envtax_pw_fp,
    data.frame(
      country = miss, envtax = NA, iso3 = NA, pop = NA,
      envtax_per_worker = mean(envtax_pw_fp$envtax_per_worker, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}
pw_vec_fp <- setNames(log(envtax_pw_fp$envtax_per_worker + 1e-12), envtax_pw_fp$country)
D_envtax_pw_fp <- abs(outer(pw_vec_fp, pw_vec_fp, FUN = "-"))
dimnames(D_envtax_pw_fp) <- list(country_labels, country_labels)

# ---- Bilateral FDI: same as before (data-limited to 2013-2020 in Eurostat)
# log_fdi already exists from dyadic-build-covariates; we reuse it.

# ---- Static covariates (Hofstede culture, geographic distance) reused -----
# D_cul and log_dist are time-invariant; reused as built.

# ---- Build dyadic data with full-period covariates -----------------------
build_dyadic_dz_fp <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc_fp[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = D_envtax_pw_fp[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q_fp <- build_dyadic_dz_fp(W_q)
dz_e_fp <- build_dyadic_dz_fp(W_e)
dz_qabc_fp <- build_dyadic_dz_fp(W_geotrade)
dz_eabc_fp <- build_dyadic_dz_fp(W_geoculture)

f_ols_fe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver
f_ols_nofe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi

# ---- RECOVERED with FE ---------------------------------------------------
fit_ols_q_rec_fe <- feols(f_ols_fe,
  data = dz_q_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_fe <- feols(f_ols_fe,
  data = dz_e_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_fe, fit_ols_e_rec_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_full,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Dyadic regression of recovered network weights with sender and receiver fixed effects, full-period averages (1993-2017) instead of pre-treatment averages (1996-2000).",
  label = "tab:dyadic_recovered_fp_fe",
  notes = "*Notes* The table reports OLS estimates of the dyadic regression of the recovered edge weight $\\hat w_{ij}$ on dyadic covariates. The sample comprises all 702 directed ordered pairs $(i,j)$ with $i\\neq j$ from the 27 EU countries. All time-varying covariates enter as full-period (1993--2017) averages. \\textit{Economic distance} is the absolute log-difference of GDP per capita in constant 2015 USD (WDI: NY.GDP.PCAP.KD). \\textit{Env. policy asymmetry} is the absolute log-difference of total environmental tax revenue (Eurostat: env\\_ac\\_taxind2, all NACE sections, MIO\\_EUR) divided by total labour force (WDI: SL.TLF.TOTL.IN). \\textit{Cultural distance} is the Euclidean distance between countries on Hofstede's four classical dimensions (PDI, IDV, MAS, UAI). \\textit{Geographic distance} is the log of the great-circle distance between capital cities, in km. \\textit{Bilateral FDI} is $\\log(1+\\text{FDI}_{ij})$, where $\\text{FDI}_{ij}$ is the average bilateral outward FDI stock from $i$ to $j$, 2013--2020 (Eurostat: bop\\_fdi6\\_pos, immediate counterpart, total economy, MIO\\_EUR). Sender and receiver fixed effects absorb country-level outdegree and indegree heterogeneity. Standard errors are two-way clustered on sender and receiver country. Significance levels: $^{***}$ $p<0.01$, $^{**}$ $p<0.05$, $^{*}$ $p<0.10$."
)
# ---- RECOVERED without FE ------------------------------------------------
fit_ols_q_rec_nf <- feols(f_ols_nofe,
  data = dz_q_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_nf <- feols(f_ols_nofe,
  data = dz_e_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_nf, fit_ols_e_rec_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_full,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Dyadic regression of recovered network weights, pooled specification without fixed effects, full-period averages (1993-2017) for time-varying covariates.",
  label = "tab:dyadic_recovered_fp_nofe",
  notes = "*Notes* The table reports OLS estimates for the recovered network without fixed effects. The table follows the same sample and covariate definitions as Tab. 3 but with a pooled specification without sender and receiver fixed effects. Significance levels: *** p<0.01, ** p<0.05, * p<0.10."
)

# ---- Albuquerque with FE ---------------------------------------------------------
fit_ols_q_abc_fe <- feols(f_ols_fe,
  data = dz_qabc_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_fe <- feols(f_ols_fe,
  data = dz_eabc_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_fe, fit_ols_e_abc_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_full,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Dyadic regression of Albuquerque two-channel network weights with sender and receiver fixed effects, full-period averages (1993-2017) for time-varying covariates.",
  label = "tab:dyadic_abc_fp_fe",
  notes = "*Notes* The table reports OLS estimates for the Albuquerque two-channel network with sender and receiver fixed effects. The table follows the same sample and covariate definitions as Tab. 3. Significance levels: *** p<0.01, ** p<0.05, * p<0.10."
)

# ---- Albuquerque without FE ------------------------------------------------------
fit_ols_q_abc_nf <- feols(f_ols_nofe,
  data = dz_qabc_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_nf <- feols(f_ols_nofe,
  data = dz_eabc_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_nf, fit_ols_e_abc_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_full,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Dyadic regression of Albuquerque two-channel network weights, pooled specification without fixed effects, full-period averages (1993-2017) for time-varying covariates.",
  label = "tab:dyadic_abc_fp_nofe",
  notes = "*Notes* The table reports OLS estimates for the Albuquerque two-channel network without fixed effects. The table follows the same sample and covariate definitions as Tab. 3 but with a pooled specification without sender and receiver fixed effects. Significance levels: *** p<0.01, ** p<0.05, * p<0.10."
)
```

```{r}
#| label: dyadic-pm25-fullperiod-robustness
#| echo: false
#| results: asis
#| cache: true
#| include: false
# Robustness check: PM2.5 air pollution asymmetry as environmental-policy
# proxy, averaged over the full available period (1993-2017 where data
# exist). PM2.5 is an intensive measure (micrograms/m^3, population-weighted)
# so does not require normalisation by country size. Data: WDI
# EN.ATM.PM25.MC.M3, population-weighted exposure (GBD via World Bank).
# Available years in the CSV vary by country; the average is taken over
# whatever years are available in [1993, 2017].

full_years_wdi <- paste0("x", 1993:2017)

# Build D_pm25 averaged over full period (PM2.5 has sparse coverage
# pre-2010 in the WDI CSV; the helper takes the mean of available years).
D_pm25_fp <- build_dist_wdi(
  file.path(data_dir, "API_EN.ATM.PM25.MC.M3_DS2_en_csv_v2_316.csv"),
  full_years_wdi
)

# Build dyadic data with d_envir = PM2.5 full-period asymmetry,
# all other covariates also at full-period averages (reuse from
# dyadic-fullperiod-robustness chunk).
build_dyadic_pm25_fp <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc_fp[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = D_pm25_fp[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q_pm_fp <- build_dyadic_pm25_fp(W_q)
dz_e_pm_fp <- build_dyadic_pm25_fp(W_e)
dz_qabc_pm_fp <- build_dyadic_pm25_fp(W_geotrade)
dz_eabc_pm_fp <- build_dyadic_pm25_fp(W_geoculture)

f_ols_fe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver
f_ols_nofe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi

dict_pm25 <- c(
  d_econ   = "Economic distance",
  d_envir  = "PM2.5 asymmetry",
  d_cult   = "Cultural distance",
  d_geo    = "Geographic distance",
  l_fdi    = "Bilateral FDI",
  sender   = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

# ---- RECOVERED with FE ---------------------------------------------------
fit_ols_q_rec_fe <- feols(f_ols_fe,
  data = dz_q_pm_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_fe <- feols(f_ols_fe,
  data = dz_e_pm_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_fe, fit_ols_e_rec_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_pm25,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of recovered network weights with sender and receiver fixed effects, PM2.5 air pollution asymmetry (WDI EN.ATM.PM25.MC.M3, full-period averages 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_recovered_pm25_fp_fe",
  notes = " *Notes* The table reports OLS estimates of the dyadic regression of recovered network weights on dyadic covariates. The table follows the same conventions as Tab.3, however it uses PM2.5 air pollution asymmetry (WDI EN.ATM.PM25.MC.M3, population-weighted exposure) averaged over the full period 1993-2017 as the environmental-policy proxy."
)

# ---- RECOVERED without FE ------------------------------------------------
fit_ols_q_rec_nf <- feols(f_ols_nofe,
  data = dz_q_pm_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_nf <- feols(f_ols_nofe,
  data = dz_e_pm_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_nf, fit_ols_e_rec_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_pm25,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of recovered network weights, pooled specification without fixed effects, PM2.5 asymmetry (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_recovered_pm25_fp_nofe",
  notes = "*Notes* The table reports OLS estimates for the recovered network without fixed effects. The table follows the same sample and covariate definitions as Tab. 7 but with a pooled specification without sender and receiver fixed effects."
)

# ---- Albuquerque with FE ---------------------------------------------------------
fit_ols_q_abc_fe <- feols(f_ols_fe,
  data = dz_qabc_pm_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_fe <- feols(f_ols_fe,
  data = dz_eabc_pm_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_fe, fit_ols_e_abc_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_pm25,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of Albuquerque two-channel network weights with sender and receiver fixed effects, PM2.5 asymmetry (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_abc_pm25_fp_fe",
  notes = "*Notes* The table reports OLS estimates for the Albuquerque two-channel network with sender and receiver fixed effects. The table follows the same sample and covariate definitions as Tab. 7. Significance levels: *** p<0.01, ** p<0.05, * p<0.10."
)

# ---- Albuquerque without FE ------------------------------------------------------
fit_ols_q_abc_nf <- feols(f_ols_nofe,
  data = dz_qabc_pm_fp,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_nf <- feols(f_ols_nofe,
  data = dz_eabc_pm_fp,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_nf, fit_ols_e_abc_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_pm25,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of Albuquerque two-channel network weights, pooled specification without fixed effects, PM2.5 asymmetry (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_abc_pm25_fp_nofe",
  notes = "*Notes* The table reports OLS estimates for the Albuquerque two-channel network without fixed effects. The table follows the same sample and covariate definitions as Tab. 7 but with a pooled specification without sender and receiver fixed effects."
)
```

## Application of Bramoullé, Djebbari & Fortin (2009)

```{r}
#| label: exogenous-networks-grid
#| echo: false
#| include: false
#| cache: true
# Build Albuquerque top-3 directed networks (geography, trade, culture) following
# Albuquerque, Bronnenberg & Corbett (2007). The TOP_N parameter is fixed
# at 3 throughout (matching Albuquerque); see top_n_network() in thesis_helpers.R.

baci_path <- "/workspaces/thesis/data/BACI_HS92_Y1996_V202601.csv"
ccodes_path <- "/workspaces/thesis/data/country_codes_V202601.csv"
TOP_N <- 3

# Trade matrix (BACI 1996)
baci <- readr::read_csv(baci_path, show_col_types = FALSE)
ccodes <- readr::read_csv(ccodes_path, show_col_types = FALSE)

iso_map <- ccodes |>
  dplyr::select(country_code, country_iso3) |>
  dplyr::filter(country_iso3 %in% cl_iso3)

trade <- baci |>
  dplyr::select(i, j, v) |>
  dplyr::inner_join(iso_map, by = c("i" = "country_code")) |>
  dplyr::rename(exporter_iso = country_iso3) |>
  dplyr::inner_join(iso_map, by = c("j" = "country_code")) |>
  dplyr::rename(importer_iso = country_iso3) |>
  dplyr::mutate(
    exporter = iso_to_label[exporter_iso],
    importer = iso_to_label[importer_iso]
  ) |>
  dplyr::group_by(exporter, importer) |>
  dplyr::summarise(value = sum(v, na.rm = TRUE), .groups = "drop")

T_trade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (k in seq_len(nrow(trade))) {
  T_trade[trade$exporter[k], trade$importer[k]] <- trade$value[k]
}

# Single-channel networks
W_geo <- top_n_network(D_geo, top_n = TOP_N, similarity = FALSE)
W_trade <- top_n_network(T_trade, top_n = TOP_N, similarity = TRUE)
W_cul <- top_n_network(D_cul, top_n = TOP_N, similarity = FALSE)

saveRDS(
  list(geo = W_geo, trade = W_trade, culture = W_cul),
  file.path(out_dir, "exogenous_networks.rds")
)
```

```{r}
#| label: bdf-prereqs
#| echo: false
#| include: false
#| cache: true
# Constants and matrices needed by all BDF tables:
# - x_vars / id_to_country / firm17 (Eurostat SBS firm count at 2017)
# - W_geotrade / W_geoculture (Albuquerque two-channel networks)

N <- length(country_labels)
x_vars <- c("x1", "x2", "x3", "x4", "x5")
id_to_country <- setNames(country_labels, seq_along(country_labels))

load_sbs_2017 <- function() {
  raw <- readxl::read_excel(
    file.path(data_dir, "enterprises_eu.xlsx"),
    sheet = "Sheet 1", skip = 9, col_names = FALSE
  )
  tibble::tibble(
    country    = raw[[1]],
    n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))
  ) |>
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) |>
    dplyr::mutate(country = dplyr::case_match(
      country,
      "Czechia" ~ "Czech Republic",
      "Slovakia" ~ "Slovak Republic",
      .default = country
    )) |>
    dplyr::distinct(country, .keep_all = TRUE)
}
firm17 <- load_sbs_2017()

# Albuquerque two-channel networks
W_geotrade <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_trade[i, ] > 0)))
  W_geotrade[i, nbrs_i] <- 1 / length(nbrs_i)
}

W_geoculture <- matrix(0, N, N, dimnames = list(country_labels, country_labels))
for (i in seq_len(N)) {
  nbrs_i <- unique(c(which(W_geo[i, ] > 0), which(W_cul[i, ] > 0)))
  W_geoculture[i, nbrs_i] <- 1 / length(nbrs_i)
}
```

```{r}
#| label: bdf-formulas
#| echo: false
#| include: false
#| cache: true
# IV formulas shared across all BDF tables. Variable names are identical
# with or without lag because build_bdf() rewrites Gy and G^k x columns
# in place.

rhs <- paste(x_vars, collapse = " + ")
exo <- paste(paste0("G_", x_vars), collapse = " + ")
inst2 <- paste(paste0("G2_", x_vars), collapse = " + ")
inst3 <- paste(paste0("G3_", x_vars), collapse = " + ")

f_nofe <- as.formula(sprintf(
  "y ~ %s + %s | Gy ~ %s + %s",
  rhs, exo, inst2, inst3
))
f_cfe <- as.formula(sprintf(
  "y ~ %s + %s | country | Gy ~ %s + %s",
  rhs, exo, inst2, inst3
))
```

```{r}
#| label: bdf-logdiff-lag-noslo
#| echo: false
#| results: asis
#| cache: true
# Drop Slovenia (structural isolate in W_q, zero in/out-degree).

set.seed(1861)

drop_country <- "Slovenia"
keep_idx <- which(country_labels != drop_country)
country_keep <- country_labels[keep_idx]

W_q_sub <- W_q[keep_idx, keep_idx]
W_e_sub <- W_e[keep_idx, keep_idx]
W_geotrade_sub <- W_geotrade[keep_idx, keep_idx]
W_geoculture_sub <- W_geoculture[keep_idx, keep_idx]

keep_full_local <- function(d, n_target) {
  full_years <- d |>
    dplyr::count(year) |>
    dplyr::filter(n == n_target) |>
    dplyr::pull(year)
  dplyr::filter(d, year %in% full_years)
}

fit_one_sub <- function(panel, W, B = 999) {
  d <- build_bdf(panel, W, x_vars, use_lag = TRUE)
  list(
    nofe = extract_nofe(feols(f_nofe, data = d, cluster = ~country)),
    cfe  = extract_cfe(cluster_boot(d, f_cfe, "country", B = B))
  )
}

panel_q <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))

panel_e <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))

q_rec <- fit_one_sub(panel_q, W_q_sub)
q_abc <- fit_one_sub(panel_q, W_geotrade_sub)
e_rec <- fit_one_sub(panel_e, W_e_sub)
e_abc <- fit_one_sub(panel_e, W_geoculture_sub)

tab <- rbind(
  rb("Quality (N = 26)", "Recovered", q_rec),
  rb("Quality (N = 26)", "Exogenous, Geo + Trade", q_abc),
  rb("Environment (N = 26)", "Recovered", e_rec),
  rb("Environment (N = 26)", "Exogenous, Geo + Culture", e_abc)
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)", "$\\hat\\beta$ (country FE)"
)

print(
  print_kbl(
    tab,
    "Peer effect on log-differences, recovered vs Albuquerque two-channel networks",
    "tab:bdf_logdiff_lag_noslo"
  ) |> add_note("N = 26 This table compares the estimated peer effect coefficients ($\\hat\\beta$) from Bramoullé et al. (2009) method when using the recovered network versus the Albuquerque et al. (2007)two-channel network (based on exogenous geography and trade for Quality and geography and culture for Environment). The dependent variable is the log-difference of certifications, and the regressions include a one-year lag in the peer effect. The independent variables were log-standardised (log (1+x)). Standard errors are clustered at the country level. The sample includes all countries with full data for the relevant period. Slovenia node is dropped since it is isolated.")
)
```

```{r}
#| label: bdf-logdiff-lag
#| echo: false
#| results: asis
#| cache: true
set.seed(1861)
panel_q <- keep_full(attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
))
panel_e <- keep_full(attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
))

# print(
#  print_kbl(
#    table_rec_vs_abc(panel_q, panel_e, use_lag = TRUE),
#    "One-year lag: peer effect on log-differences of certifications, recovered vs Albuquerque two-channel networks",
#    "tab:bdf_logdiff_lag_rec"
#  ) |> add_note("This table compares the estimated peer effect coefficients ($\\hat\\beta$) from Bramoullé et al. (2009) method when using the recovered network versus the Albuquerque et al. (2007)two-channel network (based on exogenous geography and trade for Quality and geography and culture for Environment). The dependent variable is the log-difference of certifications, and the regressions include a one-year lag in the peer effect. The independent variables were log-standardised (log (1+x)). Standard errors are clustered at the country level. The sample includes all countries with full data for the relevant period.")
# )
# cat("\n\n")
print(
  print_kbl(
    table_abc_single(panel_q, panel_e, use_lag = TRUE),
    "One-year lag: peer effect on log-differences of certifications, Albuquerque single-channel networks",
    "tab:bdf_logdiff_lag_single"
  ) |> add_note("This table reports the estimated endogenous peer effect for both standards computed across the 3 exogenous networks considered by Albuquerque et al. (2007). The table follows the same conventions as Tab.7")
)
```

```{r}
#| label: bdf-logdiff-lag-strongties-noslo
#| echo: false
#| results: asis
#| cache: true
#| include: false
# Robustness: trim recovered networks to w > 0.3, drop Slovenia and
# any further isolates.

set.seed(1861)
eps <- 1e-5
threshold <- 0.3

trim_abs <- function(W, threshold = 0.3, eps = 1e-5) {
  W2 <- W
  diag(W2) <- 0
  W2[abs(W2) < threshold] <- 0
  dimnames(W2) <- dimnames(W)
  W2
}
W_q_strong <- trim_abs(W_q, threshold = threshold)
W_e_strong <- trim_abs(W_e, threshold = threshold)

cat(sprintf(
  "Threshold |w| > %.2f. Edges retained: Quality %d/%d, Environment %d/%d\n",
  threshold,
  sum(abs(W_q_strong) > eps), sum(abs(W_q) > eps),
  sum(abs(W_e_strong) > eps), sum(abs(W_e) > eps)
))

out_q <- rowSums(abs(W_q_strong) > eps)
out_e <- rowSums(abs(W_e_strong) > eps)
iso_either <- union(country_labels[out_q == 0], country_labels[out_e == 0])
to_drop <- union(iso_either, "Slovenia")
country_keep <- setdiff(country_labels, to_drop)
keep_idx <- match(country_keep, country_labels)

cat(sprintf("Dropped: %s\n", paste(to_drop, collapse = ", ")))
cat(sprintf("Surviving: %d countries\n", length(country_keep)))

W_q_sub <- W_q_strong[keep_idx, keep_idx]
W_e_sub <- W_e_strong[keep_idx, keep_idx]
W_geotrade_sub <- W_geotrade[keep_idx, keep_idx]
W_geoculture_sub <- W_geoculture[keep_idx, keep_idx]

keep_full_local <- function(d, n_target) {
  full_years <- d |>
    dplyr::count(year) |>
    dplyr::filter(n == n_target) |>
    dplyr::pull(year)
  dplyr::filter(d, year %in% full_years)
}
panel_q <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))
panel_e <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
) |>
  dplyr::filter(country %in% country_keep) |>
  keep_full_local(length(country_keep))

fit_one_sub <- function(panel, W, B = 999) {
  d <- build_bdf(panel, W, x_vars, use_lag = TRUE)
  fit_nofe <- tryCatch(
    feols(f_nofe, data = d, cluster = ~country, warn = FALSE, notes = FALSE),
    error = function(e) NULL
  )
  nofe <- if (is.null(fit_nofe)) {
    list(b = NA, se = NA, p = NA)
  } else {
    extract_nofe(fit_nofe)
  }
  cfe <- tryCatch(
    extract_cfe(cluster_boot(d, f_cfe, "country", B = B)),
    error = function(e) list(b = NA, se = NA, p = NA)
  )
  list(nofe = nofe, cfe = cfe)
}

q_rec <- fit_one_sub(panel_q, W_q_sub)
q_abc <- fit_one_sub(panel_q, W_geotrade_sub)
e_rec <- fit_one_sub(panel_e, W_e_sub)
e_abc <- fit_one_sub(panel_e, W_geoculture_sub)

tab <- rbind(
  rb("Quality", "Recovered (w > 0.3)", q_rec),
  rb("Quality", "Exogenous, Geo + Trade", q_abc),
  rb("Environment", "Recovered (w > 0.3)", e_rec),
  rb("Environment", "Exogenous, Geo + Culture", e_abc)
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)", "$\\hat\\beta$ (country FE)"
)

cap <- sprintf(
  "Robustness (one-year lag, recovered network filtered to $|\\hat w_{ij}| > 0.3$, Slovenia excluded; %d countries): peer effect on log-differences",
  length(country_keep)
)
print(
  print_kbl(tab, cap, "tab:bdf_logdiff_lag_strongties_noslo") |>
    add_note("This table follows the same conventions as Tab.7 with the difference that only strong ties (w > 0.3) in the recovered network are retained, and Slovenia is dropped since it is an isolate in the recovered network.")
)
```
```{r}
#| label: dyadic-pm25-pergdp-fullperiod-robustness
#| echo: false
#| results: asis
#| cache: true
#| include: false
# Robustness: PM2.5 standardised by GDP per capita (full-period 1993-2017).
# Pollution per unit of economic output, in the spirit of the
# Environmental Kuznets Curve literature (Grossman & Krueger 1995;
# Copeland & Taylor 2004).

full_years_wdi <- paste0("x", 1993:2017)

# Country-level PM2.5 average (1993-2017)
pm25_levels <- readr::read_csv(
  file.path(data_dir, "API_EN.ATM.PM25.MC.M3_DS2_en_csv_v2_316.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    pm25 = rowMeans(dplyr::across(
      dplyr::any_of(full_years_wdi),
      \(z) suppressWarnings(as.numeric(z))
    ), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, is.finite(pm25)) |>
  dplyr::mutate(country = iso_to_label[iso3])

# Country-level GDP per capita average (1993-2017)
gdppc_levels <- readr::read_csv(
  file.path(data_dir, "API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv"),
  skip = 4, show_col_types = FALSE
) |>
  janitor::clean_names() |>
  dplyr::transmute(
    iso3 = country_code,
    gdppc = rowMeans(dplyr::across(
      dplyr::any_of(full_years_wdi),
      \(z) suppressWarnings(as.numeric(z))
    ), na.rm = TRUE)
  ) |>
  dplyr::filter(iso3 %in% cl_iso3, is.finite(gdppc))

pm25_pergdp <- pm25_levels |>
  dplyr::inner_join(gdppc_levels, by = "iso3") |>
  dplyr::mutate(pm25_per_gdp = pm25 / gdppc) |>
  dplyr::arrange(match(country, country_labels))

if (nrow(pm25_pergdp) < N) {
  miss <- setdiff(country_labels, pm25_pergdp$country)
  pm25_pergdp <- dplyr::bind_rows(
    pm25_pergdp,
    data.frame(
      iso3 = NA, pm25 = NA, gdppc = NA, country = miss,
      pm25_per_gdp = mean(pm25_pergdp$pm25_per_gdp, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  ) |>
    dplyr::arrange(match(country, country_labels))
}

pm_gdp_vec <- setNames(log(pm25_pergdp$pm25_per_gdp + 1e-12), pm25_pergdp$country)
D_pm25_gdp <- abs(outer(pm_gdp_vec, pm_gdp_vec, FUN = "-"))
dimnames(D_pm25_gdp) <- list(country_labels, country_labels)

build_dyadic_pm25_gdp <- function(W_outcome) {
  expand.grid(
    sender = seq_len(N), receiver = seq_len(N),
    KEEP.OUT.ATTRS = FALSE
  ) |>
    dplyr::filter(sender != receiver) |>
    dplyr::mutate(
      sender   = factor(country_labels[sender], levels = country_labels),
      receiver = factor(country_labels[receiver], levels = country_labels),
      w_hat    = W_outcome[cbind(as.integer(sender), as.integer(receiver))],
      edge     = as.integer(abs(w_hat) > 1e-5),
      d_econ   = abs(D_gdppc_fp[cbind(as.integer(sender), as.integer(receiver))]),
      d_envir  = D_pm25_gdp[cbind(as.integer(sender), as.integer(receiver))],
      d_cult   = D_cul[cbind(as.integer(sender), as.integer(receiver))],
      d_geo    = log_dist[cbind(as.integer(sender), as.integer(receiver))],
      l_fdi    = log_fdi[cbind(as.integer(sender), as.integer(receiver))]
    )
}

dz_q_pmg <- build_dyadic_pm25_gdp(W_q)
dz_e_pmg <- build_dyadic_pm25_gdp(W_e)
dz_qabc_pmg <- build_dyadic_pm25_gdp(W_geotrade)
dz_eabc_pmg <- build_dyadic_pm25_gdp(W_geoculture)

f_ols_fe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi | sender + receiver
f_ols_nofe <- w_hat ~ d_econ + d_envir + d_cult + d_geo + l_fdi

dict_pmg <- c(
  d_econ   = "Economic distance",
  d_envir  = "PM2.5/GDP asymmetry",
  d_cult   = "Cultural distance",
  d_geo    = "Geographic distance",
  l_fdi    = "Bilateral FDI",
  sender   = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  receiver = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

# ---- RECOVERED with FE ---------------------------------------------------
fit_ols_q_rec_fe <- feols(f_ols_fe,
  data = dz_q_pmg,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_fe <- feols(f_ols_fe,
  data = dz_e_pmg,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_fe, fit_ols_e_rec_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_pmg,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of recovered network weights with sender and receiver fixed effects, PM2.5 standardised by GDP per capita (WDI NY.GDP.PCAP.KD, full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_recovered_pm25gdp_fp_fe",
  notes = "*Notes* The environmental-policy proxy is PM2.5 ambient concentration (WDI EN.ATM.PM25.MC.M3, population-weighted exposure) divided by GDP per capita in constant 2015 USD (WDI NY.GDP.PCAP.KD), averaged over 1993--2017. The dyadic measure is the absolute log-difference. The transformation follows the pollution-per-output logic of the Environmental Kuznets Curve literature (Grossman \\& Krueger 1995; Copeland \\& Taylor 2004). All other conventions follow Tab. 3."
)

# ---- RECOVERED without FE ------------------------------------------------
fit_ols_q_rec_nf <- feols(f_ols_nofe,
  data = dz_q_pmg,
  cluster = ~ sender + receiver
)
fit_ols_e_rec_nf <- feols(f_ols_nofe,
  data = dz_e_pmg,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_rec_nf, fit_ols_e_rec_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_pmg,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of recovered network weights, pooled specification without fixed effects, PM2.5 standardised by GDP per capita (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_recovered_pm25gdp_fp_nofe",
  notes = "*Notes* The table follows the same format and sample as the corresponding FE specification but with a pooled specification without sender and receiver fixed effects."
)

# ---- Albuquerque with FE ---------------------------------------------------------
fit_ols_q_abc_fe <- feols(f_ols_fe,
  data = dz_qabc_pmg,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_fe <- feols(f_ols_fe,
  data = dz_eabc_pmg,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_fe, fit_ols_e_abc_fe
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "", fixef.suffix = ":",
    tablefoot = FALSE
  ),
  dict = dict_pmg,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of Albuquerque two-channel network weights with sender and receiver fixed effects, PM2.5 standardised by GDP per capita (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_abc_pm25gdp_fp_fe",
  notes = "*Notes* The table follows the same conventions as the recovered-network counterpart but uses the Albuquerque two-channel network (Albuquerque, Bronnenberg \\& Corbett 2007)."
)

# ---- Albuquerque without FE ------------------------------------------------------
fit_ols_q_abc_nf <- feols(f_ols_nofe,
  data = dz_qabc_pmg,
  cluster = ~ sender + receiver
)
fit_ols_e_abc_nf <- feols(f_ols_nofe,
  data = dz_eabc_pmg,
  cluster = ~ sender + receiver
)

etable(
  list(
    fit_ols_q_abc_nf, fit_ols_e_abc_nf
  ),
  depvar = FALSE,
  headers = list(
    " " = c("Quality", "Environment"),
    "Specification" = c("OLS", "OLS")
  ),
  style.tex = style.tex("base",
    model.title = "",
    tablefoot = FALSE
  ),
  dict = dict_pmg,
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
  fitstat = ~ n + r2,
  tex = TRUE,
  title = "Robustness: dyadic regression of Albuquerque two-channel network weights, pooled specification without fixed effects, PM2.5 standardised by GDP per capita (full-period 1993-2017) as environmental-policy proxy.",
  label = "tab:dyadic_abc_pm25gdp_fp_nofe",
  notes = "*Notes* The table follows the same format and sample as the corresponding FE specification but with a pooled specification without sender and receiver fixed effects."
)
```
```{r}
#| label: bdf-logdiff-lag-asymmetric-drop
#| echo: false
#| results: asis
#| cache: true
# Peer effect on log-differences, restricted Quality network
# specification (where it is a structural isolate, in/out-degree = 0).
# All other three specifications retain Slovenia, since it is connected
# in the Environment recovered network and in both Albuquerque
# exogenous networks (geography always gives Slovenia neighbours).
# This makes each regression use the largest valid sample given its
# own network, at the cost of cross-row sample comparability.

set.seed(1861)

# Connectivity diagnostics
eps <- 1e-5
deg_W_q <- rowSums(abs(W_q) > eps) + colSums(abs(W_q) > eps)
deg_W_e <- rowSums(abs(W_e) > eps) + colSums(abs(W_e) > eps)
deg_W_qabc <- rowSums(W_geotrade > eps) + colSums(W_geotrade > eps)
deg_W_eabc <- rowSums(W_geoculture > eps) + colSums(W_geoculture > eps)

connected_q_rec <- sum(deg_W_q > 0)
connected_e_rec <- sum(deg_W_e > 0)
connected_q_abc <- sum(deg_W_qabc > 0)
connected_e_abc <- sum(deg_W_eabc > 0)

cat(sprintf(
  "Connected nodes per network: Quality recovered = %d/27; Environment recovered = %d/27; Quality exogenous (Geo+Trade) = %d/27; Environment exogenous (Geo+Culture) = %d/27.\n\n",
  connected_q_rec, connected_e_rec, connected_q_abc, connected_e_abc
))

# ---- Quality recovered: drop Slovenia (isolate) -------------------------
keep_q_rec <- which(country_labels != "Slovenia")
country_q_rec <- country_labels[keep_q_rec]
W_q_sub <- W_q[keep_q_rec, keep_q_rec]

panel_q_rec <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
) |>
  dplyr::filter(country %in% country_q_rec) |>
  keep_full_local(length(country_q_rec))

# ---- Other three specifications: keep all 27 countries ------------------
panel_q_full <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_9001_eu.rds")), f9001
) |> keep_full_local(length(country_labels))

panel_e_full <- attach_y_logdiff(
  readRDS(file.path(clean_dir, "panel_14001_eu.rds")), f14001
) |> keep_full_local(length(country_labels))

q_rec <- fit_one_sub(panel_q_rec, W_q_sub)
q_abc <- fit_one_sub(panel_q_full, W_geotrade)
e_rec <- fit_one_sub(panel_e_full, W_e)
e_abc <- fit_one_sub(panel_e_full, W_geoculture)

tab <- rbind(
  rb("Quality (N=26)", "Recovered", q_rec),
  rb("Quality (N=27)", "Exogenous, Geo + Trade", q_abc),
  rb("Environment (N=27)", "Recovered", e_rec),
  rb("Environment (N=27)", "Exogenous, Geo + Culture", e_abc)
)
tab <- as.data.frame(tab, stringsAsFactors = FALSE)
colnames(tab) <- c(
  "Panel", "Network",
  "$\\hat\\beta$ (no FE)", "$\\hat\\beta$ (country FE)"
)

print(
  print_kbl(
    tab,
    "Asymmetric Slovenia drop: peer effect on log-differences, recovered vs Albuquerque two-channel networks (Slovenia dropped only for Quality recovered, where it is a structural isolate; retained elsewhere)",
    "tab:bdf_logdiff_lag_asym_drop"
  ) |> add_note(sprintf(
    "This table follows the same conventions as Tab.7, however Slovenia was only dropped for the recovered Quality network.",
    connected_q_rec, connected_e_rec, connected_q_abc, connected_e_abc
  ))
)
```