---
title: "ISOs Diffusion in the EU: an endogenous network recovery for Environmental and Quality Management Systems"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "main.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
data_dir <- "/workspaces/thesis/data"
clean_dir <- "/workspaces/thesis/data/cleaned_data"

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)

set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
fig_dir <- file.path(out_dir, "fig")
tab_dir <- file.path(out_dir, "tab")
for (d in c(fig_dir, tab_dir)) if (!dir.exists(d)) dir.create(d, recursive = TRUE)
```

```{r}
#| label: load-best
#| include: false
best <- readRDS(file.path(out_dir, "best_per_panel.rds"))
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
diag(W_q) <- 0
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)
diag(W_e) <- 0

sync_eu <- readRDS(file.path(clean_dir, "sync_eu.rds"))
country_labels <- sort(sync_eu)
rownames(W_q) <- colnames(W_q) <- country_labels
rownames(W_e) <- colnames(W_e) <- country_labels
```

# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Descriptive Statistics, Network Recoverey and Grid Search

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
f9001 <- readRDS(file.path(clean_dir, "f9001.rds"))
f14001 <- readRDS(file.path(clean_dir, "f14001.rds"))

desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, countries, years) {
  df |>
    dplyr::filter(country %in% countries, year %in% years) |>
    dplyr::group_by(year) |>
    dplyr::summarise(
      total_cert     = sum(n_cert, na.rm = TRUE),
      mean_cert      = mean(n_cert, na.rm = TRUE),
      sd_cert        = sd(n_cert, na.rm = TRUE),
      min_cert       = min(n_cert, na.rm = TRUE),
      max_cert       = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries    = dplyr::n(),
      .groups        = "drop"
    )
}

d9 <- desc_tab(f9001, country_labels, desc_years)
d14 <- desc_tab(f14001, country_labels, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Descriptive statistics: certification counts and adoption intensity across EU countries}",
  "\\label{tab:desc_stats}",
  "\\begin{tabular}{lcccccc}\\hline\\hline",
  "& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\",
  "\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}",
  "& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\\hline",
  sprintf(
    "N countries & %d & %d & %d & & %d & %d \\\\",
    d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
    d14$n_countries[1], d14$n_countries[2]
  ),
  sprintf(
    "Total & %s & %s & %s & & %s & %s \\\\",
    f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
    f0(d14$total_cert[1]), f0(d14$total_cert[2])
  ),
  sprintf(
    "Mean & %s & %s & %s & & %s & %s \\\\",
    f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
    f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
  ),
  sprintf(
    "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\",
    f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
    f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
  ),
  sprintf(
    "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\",
    f0(d9$min_cert[1]), f0(d9$max_cert[1]),
    f0(d9$min_cert[2]), f0(d9$max_cert[2]),
    f0(d9$min_cert[3]), f0(d9$max_cert[3]),
    f0(d14$min_cert[1]), f0(d14$max_cert[1]),
    f0(d14$min_cert[2]), f0(d14$max_cert[2])
  ),
  sprintf(
    "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\",
    f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
    f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
  ),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Note:} Cert./GDP p.c.\\ is the mean across countries of certificates per 1{,}000 USD of GDP per capita (constant 2015 USD). Environment (ISO 14001) data begin in 1999.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "desc_stats.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: grid-summary
#| echo: false
summary_df <- readRDS(file.path(out_dir, "grid_search_summary.rds"))

summary_top5 <- summary_df |>
  dplyr::mutate(
    Panel = dplyr::case_match(Panel,
      "Quality (9001) eu" ~ "Quality",
      "Environment (14001) eu" ~ "Environment",
      .default = Panel
    ),
    BIC_num = suppressWarnings(as.numeric(BIC))
  ) |>
  dplyr::group_by(Panel) |>
  dplyr::slice_min(BIC_num, n = 5) |>
  dplyr::arrange(Panel, BIC_num) |>
  dplyr::ungroup() |>
  dplyr::select(-BIC_num)

tab_top5 <- kbl(summary_top5,
  booktabs = TRUE,
  caption = "Grid search results: five lowest-BIC specifications per standard",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE,
  align = "lrrrccccc",
  format = "latex"
) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 9
  ) |>
  footnote(
    general = "Five specifications per standard with lowest BIC, sorted ascending within panel, out of $\\\\{0.05, 0.10, 0.15, 0.20\\\\}^3 = 64$ grid points. The penalty vector $\\\\lambda = (\\\\lambda_1, \\\\lambda_2, \\\\lambda_1^*)$ governs the elastic-net regularisation in the first stage and the adaptive elastic-net in the second, following De Paula, Rasul \\\\& Souza (2025). $\\\\hat{\\\\rho}_{EN}$, $\\\\hat{\\\\rho}_{Ada}$, $\\\\hat{\\\\rho}_{GMM}$ are the elastic-net, adaptive elastic-net, and post-LASSO GMM estimates of the endogenous peer-effect parameter; the latter is the unpenalised re-estimation on the support recovered by the adaptive step (Caner \\\\& Zhang, 2014). Edges count directed links with $|\\\\hat w_{ij}| > 10^{-5}$. Full grid in Appendix.",
    general_title = "Notes:",
    threeparttable = TRUE,
    escape = FALSE,
    footnote_as_chunk = TRUE
  )

writeLines(as.character(tab_top5), file.path(tab_dir, "grid_search_top5.tex"))
tab_top5
```

## Network recovery

```{r}
#| label: network-stats-table
#| echo: false
#| results: asis
net_stats <- function(W) {
  B <- (abs(W) > 1e-5) * 1
  diag(B) <- 0
  g <- igraph::graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = igraph::transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(igraph::degree(g, mode = "out")),
    out_sd      = sd(igraph::degree(g, mode = "out")),
    in_mean     = mean(igraph::degree(g, mode = "in")),
    in_sd       = sd(igraph::degree(g, mode = "in"))
  )
}
s_q <- net_stats(W_q)
s_e <- net_stats(W_e)

B_q <- (abs(W_q) > 1e-5) * 1
B_e <- (abs(W_e) > 1e-5) * 1
common <- sum(B_q & B_e)
only_q <- sum(B_q & !B_e)
only_e <- sum(!B_q & B_e)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

rho_q <- best[["9001_eu"]]$rn$unpenalisedgmm$rho
rho_e <- best[["14001_eu"]]$rn$unpenalisedgmm$rho
lam_q <- best[["9001_eu"]]$lambda
lam_e <- best[["14001_eu"]]$lambda
bic_q <- best[["9001_eu"]]$bic
bic_e <- best[["14001_eu"]]$bic

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{BIC-selected networks: Quality versus Environment (EU, $N = 27$)}",
  "\\label{tab:network_eu}",
  "\\begin{tabular}{lcc}\\hline\\hline",
  "& Quality & Environment \\\\\\hline",
  sprintf("Number of edges & %d & %d \\\\", s_q$edges, s_e$edges),
  sprintf("Edges in both networks & %d & %d \\\\", common, common),
  sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\", only_q),
  sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\", only_e),
  sprintf("Clustering & %s & %s \\\\", f3(s_q$clustering), f3(s_e$clustering)),
  sprintf("Reciprocated edges & %s & %s \\\\", p1(s_q$reciprocity), p1(s_e$reciprocity)),
  "Degree distribution across nodes (countries) & & \\\\",
  sprintf(
    "\\quad Out-degree & %s & %s \\\\",
    dg(s_q$out_mean, s_q$out_sd), dg(s_e$out_mean, s_e$out_sd)
  ),
  sprintf(
    "\\quad In-degree & %s & %s \\\\",
    dg(s_q$in_mean, s_q$in_sd), dg(s_e$in_mean, s_e$in_sd)
  ),
  "\\hline",
  sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\",
    formatC(rho_q, format = "f", digits = 4),
    formatC(rho_e, format = "f", digits = 4)
  ),
  sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\",
    paste(formatC(lam_q, format = "f", digits = 3), collapse = ", "),
    paste(formatC(lam_e, format = "f", digits = 3), collapse = ", ")
  ),
  sprintf(
    "BIC & %s & %s \\\\",
    formatC(bic_q, format = "f", digits = 2),
    formatC(bic_e, format = "f", digits = 2)
  ),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:}
  Top-5 BIC selected networks estimated via grid search. Following de Paula et al. (2024) methodology, a sensitivity analysis of .05 was performed on the penalty vector $\\lambda$ (1.0, 1.0, 1.0) to determine the final network structure
  Reciprocated edges is the frequency of in-edges that are reciprocated by out-edges. The clustering coefficient is the frequency of fully connected triplets over the total number of triplets. The degree distribution counts the average number of connections (sd in parentheses). $\\hat{\\rho}$ (GMM) is the post-LASSO endogenous peer-effect estimate. $\\lambda = (\\lambda_1, \\lambda_2, \\lambda_1^*)$ are the elastic net penalty parameters selected by BIC.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "network_stats.tex"))
cat(tex, sep = "\n")
```

## Distribution of recovered edge weights

```{r}
#| label: fig-weights-density
#| echo: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-cap: "Kernel-density estimates of nonzero recovered edge weights $|\\hat w_{ij}|$ for the BIC-selected networks. Quality (left) and Environment (right) plotted on a common horizontal scale."
w_df <- dplyr::bind_rows(
  data.frame(
    standard = "Quality (ISO 9001)",
    w = abs(W_q[abs(W_q) > 1e-5])
  ),
  data.frame(
    standard = "Environment (ISO 14001)",
    w = abs(W_e[abs(W_e) > 1e-5])
  )
)

p <- ggplot(w_df, aes(x = w, fill = standard)) +
  geom_density(alpha = 0.6, colour = NA) +
  facet_wrap(~standard, scales = "fixed") +
  scale_fill_manual(values = c(
    "Quality (ISO 9001)" = "#3182bd",
    "Environment (ISO 14001)" = "#31a354"
  )) +
  labs(x = expression("|" * hat(w)[ij] * "|"), y = "Density") +
  theme_minimal(base_size = 10) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold")
  )

ggsave(file.path(fig_dir, "weights_density.pdf"), p,
  width = 8, height = 3.5
)
print(p)
```

## Network similarity

The two networks are recovered on the same set of nodes, so the
estimated weight matrices $\hat W^Q, \hat W^E \in \mathbb R^{N\times N}$
are directly comparable element-wise. Define the dyad-level discrepancy
matrix

\begin{equation}
D_{ij} = \hat w^Q_{ij} - \hat w^E_{ij}, \qquad i \neq j,
\label{eq:Dij}
\end{equation}

and summarise it through the **mean absolute discrepancy**

\begin{equation}
\mathrm{MAD}(\hat W^Q, \hat W^E) =
\frac{1}{N(N-1)} \sum_{i\neq j} \bigl| \hat w^Q_{ij} - \hat w^E_{ij} \bigr|,
\label{eq:mad}
\end{equation}

i.e.\ the entry-wise $\ell_1$ distance between the two off-diagonal
weight matrices, normalised by the number of ordered dyads
(Donnet and Robin, 2021). MAD is zero when the two recovered networks
are identical and increases with the average per-dyad weight gap.
A complementary scalar is the Pearson correlation
$\rho(\mathrm{vec}(\hat W^Q), \mathrm{vec}(\hat W^E))$ across
off-diagonal entries, which measures co-movement irrespective of scale.

```{r}
#| label: similarity-stats
#| echo: false
#| results: asis
N <- nrow(W_q)
off <- which(row(W_q) != col(W_q))
D <- W_q - W_e
mad <- mean(abs(D[off]))
rmse <- sqrt(mean(D[off]^2))
corr <- cor(W_q[off], W_e[off])
mean_q <- mean(abs(W_q[off]))
mean_e <- mean(abs(W_e[off]))

tex <- c(
  "\\begin{table}[H]\\centering",
  "\\caption{Network similarity between $\\hat W^Q$ and $\\hat W^E$ (EU, $N = 27$)}",
  "\\label{tab:similarity}",
  "\\begin{tabular}{lc}\\hline\\hline",
  "Statistic & Value \\\\\\hline",
  sprintf("Mean absolute discrepancy, $\\mathrm{MAD}$ & %s \\\\", formatC(mad, format = "f", digits = 4)),
  sprintf("Root mean squared discrepancy, $\\mathrm{RMSD}$ & %s \\\\", formatC(rmse, format = "f", digits = 4)),
  sprintf("Pearson correlation, $\\rho(\\mathrm{vec}\\hat W^Q, \\mathrm{vec}\\hat W^E)$ & %s \\\\", formatC(corr, format = "f", digits = 4)),
  "\\hline",
  sprintf("Mean $|\\hat w^Q_{ij}|$ (off-diagonal) & %s \\\\", formatC(mean_q, format = "f", digits = 4)),
  sprintf("Mean $|\\hat w^E_{ij}|$ (off-diagonal) & %s \\\\", formatC(mean_e, format = "f", digits = 4)),
  "\\hline\\hline\\end{tabular}",
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} Discrepancy statistics computed on the $N(N-1) = 702$ off-diagonal entries of the BIC-selected weight matrices. MAD reported in equation \\eqref{eq:mad}; RMSD is the analogous $\\ell_2$ statistic. The last two rows give the average absolute weight in each network as a benchmark for interpreting MAD.}",
  "\\end{table}"
)
writeLines(tex, file.path(tab_dir, "similarity.tex"))
cat(tex, sep = "\n")
```

```{r}
#| label: fig-discrepancy-heatmap
#| echo: false
#| fig-width: 7
#| fig-height: 6
#| fig-cap: "Dyad-level discrepancy matrix $D_{ij} = \\hat w^Q_{ij} - \\hat w^E_{ij}$. Red cells indicate dyads where the Quality peer-effect weight exceeds the Environment one; blue cells the opposite. Diagonal blanked out."
D_long <- as.data.frame(as.table(D)) |>
  setNames(c("from", "to", "diff")) |>
  dplyr::filter(as.character(from) != as.character(to))

p2 <- ggplot(D_long, aes(x = to, y = from, fill = diff)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#2166ac", mid = "white", high = "#b2182b",
    midpoint = 0,
    name = expression(hat(w)[ij]^Q - hat(w)[ij]^E)
  ) +
  scale_y_discrete(limits = rev(country_labels)) +
  labs(x = "Receiver j", y = "Sender i") +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid = element_blank(),
    legend.position = "right"
  )

ggsave(file.path(fig_dir, "discrepancy_heatmap.pdf"), p2,
  width = 7, height = 6
)
print(p2)
```

## Network visualisation

```{r}
#| label: fig-networks-weighted-full
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Full estimated peer-effect networks (BIC-selected $\\lambda$, EU, $N = 27$). All non-zero edges shown with width and colour intensity scaling continuously with $|\\hat w_{ij}|$. Darker/thicker edges indicate stronger peer-effect weights. Rows share a common layout: top row uses the Fruchterman--Reingold layout computed on the Quality network, bottom on Environment. Columns correspond to the network being plotted."
make_full_weighted_graph <- function(W, labels, eps = 1e-5) {
  W2 <- W
  W2[abs(W2) < eps] <- 0
  g <- igraph::graph_from_adjacency_matrix(W2, mode = "directed", weighted = TRUE)
  igraph::V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(W_q, country_labels)
g_e_full <- make_full_weighted_graph(W_e, country_labels)

w_all <- c(abs(igraph::E(g_q_full)$weight), abs(igraph::E(g_e_full)$weight))
w_max <- max(w_all, 1e-9)

set.seed(1861)
layout_q_full <- igraph::layout_with_fr(g_q_full)
rownames(layout_q_full) <- igraph::V(g_q_full)$name
set.seed(1861)
layout_e_full <- igraph::layout_with_fr(g_e_full)
rownames(layout_e_full) <- igraph::V(g_e_full)$name

edge_palette <- colorRampPalette(c("#FEE0D2", "#FC9272", "#DE2D26", "#67000D"))(100)

plot_full_weighted_net <- function(g, lo, title, w_max) {
  lo_ordered <- lo[igraph::V(g)$name, , drop = FALSE]
  if (igraph::ecount(g) > 0) {
    w_abs <- abs(igraph::E(g)$weight)
    w_norm <- pmin(w_abs / w_max, 1)
    edge_wd <- 0.4 + 3.6 * w_norm
    col_idx <- pmax(1, ceiling(w_norm * length(edge_palette)))
    edge_col <- edge_palette[col_idx]
    edge_alpha <- 0.35 + 0.55 * w_norm
    edge_col <- mapply(adjustcolor, edge_col, alpha.f = edge_alpha)
  } else {
    edge_wd <- numeric(0)
    edge_col <- character(0)
  }
  has_edge <- igraph::degree(g) > 0
  plot(g,
    layout = lo_ordered,
    vertex.size = 10, vertex.label = igraph::V(g)$name,
    vertex.label.cex = 0.55, vertex.label.dist = 1.8,
    vertex.label.color = "grey20",
    vertex.color = ifelse(has_edge, "steelblue", "grey85"),
    vertex.frame.color = ifelse(has_edge, "steelblue4", "grey65"),
    edge.arrow.size = 0.6, edge.arrow.width = 1.2,
    edge.color = edge_col, edge.width = edge_wd,
    main = title
  )
  leg_vals <- c(0.2, 0.5, 0.8, 1.0) * w_max
  leg_vals <- leg_vals[leg_vals <= w_max]
  leg_norm <- leg_vals / w_max
  leg_wd <- 0.4 + 3.6 * leg_norm
  leg_col_ix <- pmax(1, ceiling(leg_norm * length(edge_palette)))
  leg_col <- edge_palette[leg_col_ix]
  leg_alpha <- 0.35 + 0.55 * leg_norm
  leg_col <- mapply(adjustcolor, leg_col, alpha.f = leg_alpha)
  legend("bottomright",
    title = expression("|" * hat(w)[ij] * "|"),
    legend = formatC(leg_vals, format = "f", digits = 2),
    lwd = leg_wd, col = leg_col, bty = "n", cex = 0.75, seg.len = 2.5
  )
}

pdf(file.path(fig_dir, "networks_weighted_full.pdf"), width = 13, height = 13)
par(mfrow = c(2, 2), mar = c(1, 1, 3, 1))
plot_full_weighted_net(g_q_full, layout_q_full, "Quality — Quality layout", w_max)
plot_full_weighted_net(g_e_full, layout_q_full, "Environment — Quality layout", w_max)
plot_full_weighted_net(g_q_full, layout_e_full, "Quality — Environment layout", w_max)
plot_full_weighted_net(g_e_full, layout_e_full, "Environment — Environment layout", w_max)
dev.off()

par(mfrow = c(2, 2), mar = c(1, 1, 3, 1))
plot_full_weighted_net(g_q_full, layout_q_full, "Quality — Quality layout", w_max)
plot_full_weighted_net(g_e_full, layout_q_full, "Environment — Quality layout", w_max)
plot_full_weighted_net(g_q_full, layout_e_full, "Quality — Environment layout", w_max)
plot_full_weighted_net(g_e_full, layout_e_full, "Environment — Environment layout", w_max)
```

## Appendix

```{r}
#| label: grid-summary-appendix
#| echo: false
summary_df <- readRDS(file.path(out_dir, "grid_search_summary.rds"))

tab_full <- kbl(summary_df,
  booktabs = TRUE,
  longtable = TRUE,
  caption = "Grid search results",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE,
  align = "lrrrccccc",
  format = "latex"
) |>
  kable_styling(
    latex_options = c("hold_position", "repeat_header"),
    font_size = 8
  ) |>
  footnote(
    general = "Penalty triplet $(\\\\lambda_1, \\\\lambda_2, \\\\lambda_1^*)$ from De Paula, Rasul \\\\& Souza (2025). $\\\\hat{\\\\rho}_{EN}$, $\\\\hat{\\\\rho}_{Ada}$, $\\\\hat{\\\\rho}_{GMM}$ are the elastic-net, adaptive elastic-net, and post-LASSO GMM estimates of the peer-effect parameter. BIC selects the preferred specification per panel.",
    general_title = "Notes:",
    threeparttable = TRUE,
    escape = FALSE,
    footnote_as_chunk = TRUE
  )

writeLines(as.character(tab_full), file.path(tab_dir, "grid_search_full.tex"))
tab_full
```