---
title: "Preliminary Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "results_eu_grid.pdf"
    keep-tex: false
    fig-pos: "H"
---

```{r}
#| label: setup
#| include: false

library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)

set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
```

```{r}
#| label: data-cleaning
#| include: false
#| cache: true

data_dir <- "/workspaces/thesis/data"

# === COUNTRY GROUPS ===
oecd <- c(
  "Australia", "Austria", "Belgium", "Canada", "Chile", "Colombia",
  "Costa Rica", "Czech Republic", "Denmark", "Estonia", "Finland",
  "France", "Germany", "Greece", "Hungary", "Iceland", "Ireland",
  "Israel", "Italy", "Japan", "Korea", "Latvia", "Lithuania",
  "Luxembourg", "Mexico", "Netherlands", "New Zealand", "Norway",
  "Poland", "Portugal", "Slovak Republic", "Slovenia", "Spain",
  "Sweden", "Switzerland", "Turkey", "United Kingdom", "United States"
)

eu <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus",
  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Ireland", "Italy", "Latvia",
  "Lithuania", "Luxembourg", "Malta", "Netherlands", "Poland",
  "Portugal", "Romania", "Slovak Republic", "Slovenia", "Spain",
  "Sweden"
)

harmonise_names <- function(df) {
  if (nrow(df) == 0) {
    return(df)
  }
  df %>% dplyr::mutate(country = dplyr::case_match(country,
    "Korea (Republic of)" ~ "Korea", "Korea, Republic of" ~ "Korea",
    "Korea (the Republic of)" ~ "Korea",
    "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
    "United Kingdom of Great Britain and Northern Ireland" ~ "United Kingdom",
    "United States of America" ~ "United States",
    "Netherlands (Kingdom of the)" ~ "Netherlands",
    .default = country
  ))
}

read_iso_hist <- function(file_name, sheets_idx, my_names) {
  f_path <- file.path(data_dir, file_name)
  if (!file.exists(f_path)) stop(paste("File not found:", f_path))
  sheets <- purrr::map(sheets_idx, ~ readxl::read_excel(f_path, sheet = .x))
  expected_cols <- length(my_names)
  purrr::map(1:5, ~ sheets[[.x]] %>%
    dplyr::select(dplyr::all_of(1:expected_cols)) %>%
    dplyr::slice(-(1:2)) %>%
    purrr::set_names(my_names) %>%
    tidyr::pivot_longer(-country, names_to = "year", values_to = "n_cert")) %>%
    purrr::reduce(dplyr::bind_rows) %>%
    harmonise_names() %>%
    dplyr::mutate(n_cert = as.numeric(n_cert), year = as.integer(year)) %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

extract_recent <- function(iso_code) {
  files <- list.files(data_dir,
    pattern = "ISO Survey 20(1[8-9]|2[0-9]).*\\.xlsx$",
    full.names = TRUE
  )
  purrr::map_dfr(files, function(f) {
    yr <- as.numeric(stringr::str_extract(basename(f), "20\\d{2}"))
    sheets <- readxl::excel_sheets(f)
    target <- sheets[grepl(iso_code, sheets)][1]
    if (!is.na(target)) {
      df <- readxl::read_excel(f, sheet = target)
      start_row <- which(apply(df, 1, function(x) any(grepl("Afghanistan|Albania", x))))[1]
      df_clean <- df %>%
        dplyr::slice(start_row:dplyr::n()) %>%
        dplyr::rename(country = 1) %>%
        dplyr::mutate(dplyr::across(-1, ~ suppressWarnings(as.numeric(.)))) %>%
        dplyr::select(country, tidyselect::where(is.numeric)) %>%
        dplyr::mutate(n_cert = rowSums(dplyr::select(., -country), na.rm = TRUE)) %>%
        dplyr::select(country, n_cert) %>%
        dplyr::mutate(year = yr)
      return(df_clean)
    }
    return(NULL)
  }) %>%
    harmonise_names() %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

make_covs <- function(start_yr) {
  load_w <- function(f, v) {
    readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) %>%
      dplyr::select(country = 1, as.character(start_yr):"2024") %>%
      tidyr::pivot_longer(-country, names_to = "year", values_to = v) %>%
      dplyr::mutate(
        year = as.integer(year),
        !!v := as.numeric(as.character(.data[[v]]))
      ) %>%
      harmonise_names()
  }
  load_w("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc") %>%
    dplyr::full_join(load_w("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_RL.EST_DS2_en_csv_v2_1905.csv", "rol"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_NY.GDP.TOTL.RT.ZS_DS2_en_csv_v2_362.csv", "rent"),
      by = c("country", "year")
    )
}

finalize_panel <- function(iso_full, filter_list, label) {
  res <- iso_full %>%
    dplyr::filter(country %in% filter_list) %>%
    dplyr::group_by(country) %>%
    dplyr::mutate(dplyr::across(c(gdp_pc, trade, rol, renew, rent), ~ {
      v <- zoo::na.approx(., na.rm = FALSE)
      v <- zoo::na.locf(zoo::na.locf(v, na.rm = FALSE), na.rm = FALSE, fromLast = TRUE)
      asinh(v)
    })) %>%
    dplyr::arrange(year) %>%
    dplyr::mutate(y = (n_cert - dplyr::lag(n_cert)) / sqrt(pmax(dplyr::lag(n_cert), 1))) %>% # ALBUQUERQUE
    dplyr::filter(!is.na(y)) %>%
    dplyr::ungroup()

  T_max <- max((res %>% dplyr::count(country))$n)
  final <- res %>%
    dplyr::group_by(country) %>%
    dplyr::filter(dplyr::n() == T_max) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(id = as.integer(factor(country))) %>%
    dplyr::select(id,
      time = year, y,
      x1 = gdp_pc, x2 = trade, x3 = rol, x4 = renew, x5 = rent
    ) %>%
    tidyr::drop_na()

  cat(sprintf("Dataset %s: N=%d, T=%d\n", label, length(unique(final$id)), T_max))
  return(final)
}

# --- Build panels ---
h9001 <- read_iso_hist(
  "01. ISO 9001 - Number of certificates per country and industry sectors - 1993 to 2017.xlsm",
  4:8, c("country", as.character(1993:2017))
)
r9001 <- extract_recent("9001")
f9001 <- dplyr::bind_rows(h9001, r9001)

h14001 <- read_iso_hist(
  "02. ISO 14001 - Number of certificates per country and industry sectors - 1999 to 2017.xlsm",
  4:8, c("country", as.character(1999:2017))
)
r14001 <- extract_recent("14001")
f14001 <- dplyr::bind_rows(h14001, r14001)

covs9001 <- make_covs(1993)
covs14001 <- make_covs(1999)
f9001 <- f9001 %>% dplyr::left_join(covs9001, by = c("country", "year"))
f14001 <- f14001 %>% dplyr::left_join(covs14001, by = c("country", "year"))

# Truncate to historical data (consistent methodology)
f9001 <- f9001 %>% dplyr::filter(year <= 2017)
f14001 <- f14001 %>% dplyr::filter(year <= 2017)

# Sync countries across standards
sync_eu <- intersect(
  (f9001 %>% filter(country %in% eu) %>% pull(country) %>% unique()),
  (f14001 %>% filter(country %in% eu) %>% pull(country) %>% unique())
)
sync_oecd <- intersect(
  (f9001 %>% filter(country %in% oecd) %>% pull(country) %>% unique()),
  (f14001 %>% filter(country %in% oecd) %>% pull(country) %>% unique())
)

dp_9001_oecd <- finalize_panel(f9001, sync_oecd, "9001_OECD (synced)")
dp_9001_eu <- finalize_panel(f9001, sync_eu, "9001_EU (synced)")
dp_14001_oecd <- finalize_panel(f14001, sync_oecd, "14001_OECD (synced)")
dp_14001_eu <- finalize_panel(f14001, sync_eu, "14001_EU (synced)")
```

```{r}
#| label: grid-search
#| include: false
#| cache: true

panels <- list(
  "9001_eu"     = dp_9001_eu,
  "14001_eu"    = dp_14001_eu
)

lambda_grid <- list(
  c(0.005, 0.005, 0.005),
  c(0.15, 0.20, 0.15),
  c(0.10, 0.10, 0.20),
  c(0.20, 0.20, 0.20),
  c(0.50, 0.50, 0.50)
)

grid_results <- list()
idx <- 0

for (pname in names(panels)) {
  df <- panels[[pname]]
  Wder <- NULL

  for (li in seq_along(lambda_grid)) {
    lam <- lambda_grid[[li]]
    idx <- idx + 1
    run_label <- sprintf("%s_gridrun%02d", pname, li)
    cat(sprintf(
      "\n[%d] %s | lambda = (%.3f, %.3f, %.3f) ... ",
      idx, pname, lam[1], lam[2], lam[3]
    ))

    row <- list(
      panel = pname, lambda = lam, run_label = run_label,
      edges = NA, rho_en = NA, rho_ada = NA, rho_gmm = NA, bic = NA, rn = NULL
    )

    tryCatch(
      {
        set.seed(1861)
        rn <- if (is.null(Wder)) {
          recoverNetwork(df, lambda = lam)
        } else {
          recoverNetwork(df, lambda = lam, Wder = Wder)
        }
        if (is.null(Wder)) Wder <- rn$Wder

        W <- as.matrix(rn$unpenalisedgmm$W)
        diag(W) <- 0

        row$edges <- sum(abs(W) > 1e-5)
        row$rho_en <- rn$en$rho
        row$rho_ada <- rn$adaen$rho
        row$rho_gmm <- rn$unpenalisedgmm$rho
        row$bic <- rn$BIC
        row$rn <- rn

        cat(sprintf(
          "OK | edges=%d | rho_gmm=%.3f | BIC=%.2f",
          row$edges, row$rho_gmm, row$bic
        ))
        saveRDS(rn, file.path(out_dir, paste0(run_label, ".rds")))
      },
      error = function(e) {
        cat("FAILED:", conditionMessage(e))
      }
    )

    grid_results[[idx]] <- row
  }

  Wder <- NULL
}

# Save everything for later recomputation
grid_meta <- lapply(grid_results, function(r) {
  r$rn <- NULL
  r
})
saveRDS(grid_meta, file.path(out_dir, "grid_search_meta.rds"))
save(grid_results, file = file.path(out_dir, "grid_search_full.RData"))
```

```{r}
#| label: grid-summary
#| echo: false

summary_df <- do.call(rbind, lapply(grid_results, function(r) {
  data.frame(
    Panel = gsub("_", " ", r$panel),
    L1 = r$lambda[1],
    L2 = r$lambda[2],
    L1s = r$lambda[3],
    Edges = ifelse(is.na(r$edges), "FAIL", as.character(r$edges)),
    rho_EN = ifelse(is.na(r$rho_en), "FAIL", sprintf("%.3f", r$rho_en)),
    rho_Ada = ifelse(is.na(r$rho_ada), "FAIL", sprintf("%.3f", r$rho_ada)),
    rho_GMM = ifelse(is.na(r$rho_gmm), "FAIL", sprintf("%.3f", r$rho_gmm)),
    BIC = ifelse(is.na(r$bic), "FAIL", sprintf("%.2f", r$bic)),
    stringsAsFactors = FALSE
  )
}))

kbl(summary_df,
  booktabs = TRUE, caption = "Grid search results",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE, align = "lrrrccccc"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r}
#| label: best-selection
#| include: false

best <- list()
for (pname in names(panels)) {
  panel_runs <- Filter(function(r) r$panel == pname && !is.null(r$rn), grid_results)
  if (length(panel_runs) > 0) {
    bics <- sapply(panel_runs, `[[`, "bic")
    best[[pname]] <- panel_runs[[which.min(bics)]]
  }
}
saveRDS(best, file.path(out_dir, "best_per_panel.rds"))

net_stats <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  B <- (abs(W) > 1e-5) * 1
  g <- graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(degree(g, mode = "out")),
    out_sd      = sd(degree(g, mode = "out")),
    in_mean     = mean(degree(g, mode = "in")),
    in_sd       = sd(degree(g, mode = "in"))
  )
}
```

\newpage
# EU: ISO 9001 versus ISO 14001

## Data and transformations

Two transformations are applied before estimation. The dependent variable is standardised following Albuquerque et al.\ (2007): $y_{it} = (n_{it} - n_{i,t-1}) / \sqrt{\max(n_{i,t-1},\, 1)}$, a variance-stabilising transformation that accounts for the fact that larger adopters exhibit proportionally larger increments. Covariates are transformed via the inverse hyperbolic sine, $\operatorname{asinh}(x) = \ln(x + \sqrt{x^2 + 1})$, which approximates the natural logarithm for large values while remaining defined at zero.

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis

desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, std_label, countries, years) {
  df %>%
    filter(country %in% countries, year %in% years) %>%
    group_by(year) %>%
    summarise(
      total_cert = sum(n_cert, na.rm = TRUE),
      mean_cert = mean(n_cert, na.rm = TRUE),
      sd_cert = sd(n_cert, na.rm = TRUE),
      min_cert = min(n_cert, na.rm = TRUE),
      max_cert = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries = n(),
      .groups = "drop"
    ) %>%
    mutate(standard = std_label) %>%
    select(standard, everything())
}

d9 <- desc_tab(f9001, "ISO 9001", sync_eu, desc_years)
d14 <- desc_tab(f14001, "ISO 14001", sync_eu, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Descriptive statistics: certification counts and adoption intensity across EU countries}\n")
cat("\\label{tab:desc_stats}\n")
cat("\\begin{tabular}{lcccccc}\n\\hline\\hline\n")
cat("& \\multicolumn{3}{c}{ISO 9001} & \\multicolumn{3}{c}{ISO 14001} \\\\\n")
cat("\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}\n")
cat("& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\n\\hline\n")

cat(sprintf(
  "N countries & %d & %d & %d & & %d & %d \\\\\n",
  d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
  d14$n_countries[1], d14$n_countries[2]
))
cat(sprintf(
  "Total & %s & %s & %s & & %s & %s \\\\\n",
  f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
  f0(d14$total_cert[1]), f0(d14$total_cert[2])
))
cat(sprintf(
  "Mean & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
  f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
))
cat(sprintf(
  "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
  f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
))
cat(sprintf(
  "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\\n",
  f0(d9$min_cert[1]), f0(d9$max_cert[1]),
  f0(d9$min_cert[2]), f0(d9$max_cert[2]),
  f0(d9$min_cert[3]), f0(d9$max_cert[3]),
  f0(d14$min_cert[1]), f0(d14$max_cert[1]),
  f0(d14$min_cert[2]), f0(d14$max_cert[2])
))
cat(sprintf(
  "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\\n",
  f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
  f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
))

cat("\\hline\\hline\n\\end{tabular}\n")
cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Note:} Cert./GDP p.c.\\ is the mean across countries of certificates per 1,000 USD of GDP per capita (constant 2015 USD). ISO 14001 data begin in 1999.}\n")
cat("\\end{table}\n")
```

Table \ref{tab:desc_stats} reports raw certification counts across EU countries. ISO 9001 adoption is roughly four times larger than ISO 14001 by 2017, yet the development-adjusted intensity (last column) suggests that the gap narrows once economic size is accounted for.

## Network recovery

```{r}
#| label: table-eu
#| echo: false
#| results: asis

if (!is.null(best[["9001_eu"]]) && !is.null(best[["14001_eu"]])) {
  rn_a <- best[["9001_eu"]]$rn
  s_a <- net_stats(rn_a)
  rn_b <- best[["14001_eu"]]$rn
  s_b <- net_stats(rn_b)

  W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_a) <- 0
  W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_b) <- 0
  common <- sum(W_a & W_b)
  only_a <- sum(W_a & !W_b)
  only_b <- sum(!W_a & W_b)

  f3 <- function(x) formatC(x, format = "f", digits = 3)
  p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
  dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

  cat("\\begin{table}[H]\n\\centering\n")
  cat("\\caption{BIC-selected networks: ISO 9001 versus ISO 14001 (EU, $N = 27$)}\n")
  cat("\\label{tab:network_eu}\n")
  cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
  cat("& ISO 9001 & ISO 14001 \\\\\n\\hline\n")
  cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
  cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
  cat(sprintf("Edges in $W_{9001}$ only & %d & \\\\\n", only_a))
  cat(sprintf("Edges in $W_{14001}$ only & & %d \\\\\n", only_b))
  cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
  cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
  cat("Degree distribution across nodes (countries) & & \\\\\n")
  cat(sprintf(
    "\\quad Out-degree & %s & %s \\\\\n",
    dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
  ))
  cat(sprintf(
    "\\quad In-degree & %s & %s \\\\\n",
    dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
  ))
  cat("\\hline\n")
  cat(sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
    formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
    formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
  ))
  cat(sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\\n",
    paste(formatC(best[["9001_eu"]]$lambda, format = "f", digits = 3), collapse = ", "),
    paste(formatC(best[["14001_eu"]]$lambda, format = "f", digits = 3), collapse = ", ")
  ))
  cat(sprintf(
    "BIC & %s & %s \\\\\n",
    formatC(best[["9001_eu"]]$bic, format = "f", digits = 2),
    formatC(best[["14001_eu"]]$bic, format = "f", digits = 2)
  ))
  cat("\\hline\\hline\n\\end{tabular}\n")
  cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} This compares statistics from the BIC-selected estimated networks for ISO 9001 ($T = 24$, 1993--2017) and ISO 14001 ($T = 18$, 1999--2017) across 27 EU countries. The number of edges, edges in both networks, edges in $W_{9001}$ only, and edges in $W_{14001}$ only count the number of directed links in each category. Reciprocated edges is the frequency of in-edges that are reciprocated by out-edges. The clustering coefficient is the frequency of fully connected triplets over the total number of triplets. The degree distribution across nodes counts the average number of connections (standard deviation in parentheses), shown separately for in-degree and out-degree. $\\hat{\\rho}$ (GMM) is the post-LASSO estimate of the peer-effect parameter, interpretable as the speed of diffusion through the network. $\\lambda = (\\lambda_1, \\lambda_2, \\lambda_1^*)$ are the elastic net penalty parameters selected by BIC.}\n")
  cat("\\end{table}\n")
} else {
  cat("\\textbf{EU table not available: one or both panels failed all grid runs.}\n")
}
```

The two networks are comparable in size despite the large difference in raw adoption volumes, suggesting that peer-effect channels operate at a similar density for both standards. The estimated $\hat{\rho}_{GMM}$ is reported at the bottom of Table \ref{tab:network_eu}. Reciprocated edges indicate country pairs with bidirectional influence.

```{r}
#| label: table-eu-same-lambda
#| echo: false
#| results: asis

shared_runs <- list()
for (li in seq_along(lambda_grid)) {
  lam <- lambda_grid[[li]]
  r9 <- Filter(function(r) r$panel == "9001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  r14 <- Filter(function(r) r$panel == "14001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  if (length(r9) > 0 && length(r14) > 0) {
    shared_runs[[length(shared_runs) + 1]] <- list(
      lambda = lam,
      avg_bic = (r9[[1]]$bic + r14[[1]]$bic) / 2,
      r9001 = r9[[1]], r14001 = r14[[1]]
    )
  }
}
shared_best <- shared_runs[[which.min(sapply(shared_runs, `[[`, "avg_bic"))]]

rn_a <- shared_best$r9001$rn
s_a <- net_stats(rn_a)
rn_b <- shared_best$r14001$rn
s_b <- net_stats(rn_b)

W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_a) <- 0
W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_b) <- 0
common <- sum(W_a & W_b)
only_a <- sum(W_a & !W_b)
only_b <- sum(!W_a & W_b)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

lam_str <- paste(formatC(shared_best$lambda, format = "f", digits = 3), collapse = ", ")

cat("\\begin{table}[H]\n\\centering\n")
cat(sprintf("\\caption{Robustness: common penalty $\\lambda = (%s)$ for both standards (EU, $N = 27$)}\n", lam_str))
cat("\\label{tab:network_eu_same_lambda}\n")
cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
cat("& ISO 9001 & ISO 14001 \\\\\n\\hline\n")
cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
cat(sprintf("Edges in $W_{9001}$ only & %d & \\\\\n", only_a))
cat(sprintf("Edges in $W_{14001}$ only & & %d \\\\\n", only_b))
cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
cat("Degree distribution across nodes (countries) & & \\\\\n")
cat(sprintf(
  "\\quad Out-degree & %s & %s \\\\\n",
  dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
))
cat(sprintf(
  "\\quad In-degree & %s & %s \\\\\n",
  dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
))
cat("\\hline\n")
cat(sprintf(
  "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
  formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
  formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
))
cat(sprintf(
  "BIC & %s & %s \\\\\n",
  formatC(shared_best$r9001$bic, format = "f", digits = 2),
  formatC(shared_best$r14001$bic, format = "f", digits = 2)
))
cat("\\hline\\hline\n\\end{tabular}\n")
cat(sprintf("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} Same as Table \\ref{tab:network_eu}, but imposing a common penalty vector $\\lambda = (%s)$ for both standards. The common $\\lambda$ is selected as the grid point with the lowest average BIC across the two panels. See Table \\ref{tab:network_eu} for variable definitions.}\n", lam_str))
cat("\\end{table}\n")
```

Table \ref{tab:network_eu_same_lambda} confirms that the structural differences between the two networks are not driven by the choice of penalty parameters: imposing a common $\lambda$ (selected by lowest average BIC across the two panels) yields similar network statistics.

## Network visualisation

```{r}
#| label: country-labels
#| include: false

country_labels <- sort(sync_eu)
```

```{r}
#| label: fig-networks
#| echo: false
#| fig-width: 10
#| fig-height: 12
#| fig-cap: "Estimated peer-effect networks (BIC-selected $\\lambda$, EU, $N = 27$)"

par(mfrow = c(2, 1), mar = c(1, 1, 3, 1))

for (pname in c("9001_eu", "14001_eu")) {
  rn <- best[[pname]]$rn
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  B <- (abs(W) > 1e-5) * 1
  g <- graph_from_adjacency_matrix(B, mode = "directed")
  V(g)$name <- country_labels

  set.seed(1861)
  lo <- layout_with_fr(g)

  plot(g,
    layout = lo,
    vertex.size = 12,
    vertex.label = V(g)$name,
    vertex.label.cex = 0.6,
    vertex.label.dist = 1.5,
    vertex.color = ifelse(degree(g) > 0, "steelblue", "grey80"),
    edge.arrow.size = 0.3,
    edge.color = adjustcolor("grey40", alpha = 0.6),
    main = gsub("_eu", " — EU", pname) %>% gsub("_", " ", .) %>% toupper()
  )
}
```

\newpage

# OECD: ISO 9001 versus ISO 14001
As of now, I have omitted the estimation of OECD countries. As remarked by de Paula, N would be almost double T, threatening identification.