---
title: "Preliminary Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "results_eu_grid.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)

set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
```

```{r}
#| label: data-cleaning
#| include: false
#| cache: true
data_dir <- "/workspaces/thesis/data"

# === COUNTRY GROUPS ===
oecd <- c(
  "Australia", "Austria", "Belgium", "Canada", "Chile", "Colombia",
  "Costa Rica", "Czech Republic", "Denmark", "Estonia", "Finland",
  "France", "Germany", "Greece", "Hungary", "Iceland", "Ireland",
  "Israel", "Italy", "Japan", "Korea", "Latvia", "Lithuania",
  "Luxembourg", "Mexico", "Netherlands", "New Zealand", "Norway",
  "Poland", "Portugal", "Slovak Republic", "Slovenia", "Spain",
  "Sweden", "Switzerland", "Turkey", "United Kingdom", "United States"
)

eu <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus",
  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Ireland", "Italy", "Latvia",
  "Lithuania", "Luxembourg", "Malta", "Netherlands", "Poland",
  "Portugal", "Romania", "Slovak Republic", "Slovenia", "Spain",
  "Sweden"
)

harmonise_names <- function(df) {
  if (nrow(df) == 0) {
    return(df)
  }
  df %>% dplyr::mutate(country = dplyr::case_match(country,
    "Korea (Republic of)" ~ "Korea", "Korea, Republic of" ~ "Korea",
    "Korea (the Republic of)" ~ "Korea",
    "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
    "United Kingdom of Great Britain and Northern Ireland" ~ "United Kingdom",
    "United States of America" ~ "United States",
    "Netherlands (Kingdom of the)" ~ "Netherlands",
    .default = country
  ))
}

read_iso_hist <- function(file_name, sheets_idx, my_names) {
  f_path <- file.path(data_dir, file_name)
  if (!file.exists(f_path)) stop(paste("File not found:", f_path))
  sheets <- purrr::map(sheets_idx, ~ readxl::read_excel(f_path, sheet = .x))
  expected_cols <- length(my_names)
  purrr::map(1:5, ~ sheets[[.x]] %>%
    dplyr::select(dplyr::all_of(1:expected_cols)) %>%
    dplyr::slice(-(1:2)) %>%
    purrr::set_names(my_names) %>%
    tidyr::pivot_longer(-country, names_to = "year", values_to = "n_cert")) %>%
    purrr::reduce(dplyr::bind_rows) %>%
    harmonise_names() %>%
    dplyr::mutate(n_cert = as.numeric(n_cert), year = as.integer(year)) %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

extract_recent <- function(iso_code) {
  files <- list.files(data_dir,
    pattern = "ISO Survey 20(1[8-9]|2[0-9]).*\\.xlsx$",
    full.names = TRUE
  )
  purrr::map_dfr(files, function(f) {
    yr <- as.numeric(stringr::str_extract(basename(f), "20\\d{2}"))
    sheets <- readxl::excel_sheets(f)
    target <- sheets[grepl(iso_code, sheets)][1]
    if (!is.na(target)) {
      df <- readxl::read_excel(f, sheet = target)
      start_row <- which(apply(df, 1, function(x) any(grepl("Afghanistan|Albania", x))))[1]
      df_clean <- df %>%
        dplyr::slice(start_row:dplyr::n()) %>%
        dplyr::rename(country = 1) %>%
        dplyr::mutate(dplyr::across(-1, ~ suppressWarnings(as.numeric(.)))) %>%
        dplyr::select(country, tidyselect::where(is.numeric)) %>%
        dplyr::mutate(n_cert = rowSums(dplyr::select(., -country), na.rm = TRUE)) %>%
        dplyr::select(country, n_cert) %>%
        dplyr::mutate(year = yr)
      return(df_clean)
    }
    return(NULL)
  }) %>%
    harmonise_names() %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

make_covs <- function(start_yr) {
  load_w <- function(f, v) {
    readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) %>%
      dplyr::select(country = 1, as.character(start_yr):"2024") %>%
      tidyr::pivot_longer(-country, names_to = "year", values_to = v) %>%
      dplyr::mutate(
        year = as.integer(year),
        !!v := as.numeric(as.character(.data[[v]]))
      ) %>%
      harmonise_names()
  }
  load_w("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc") %>%
    dplyr::full_join(load_w("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_RL.EST_DS2_en_csv_v2_1905.csv", "rol"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_NY.GDP.TOTL.RT.ZS_DS2_en_csv_v2_362.csv", "rent"),
      by = c("country", "year")
    )
}

load_sbs_2017 <- function() {
  file_path <- file.path(data_dir, "enterprises_eu.xlsx")
  raw <- readxl::read_excel(file_path, sheet = "Sheet 1", skip = 9, col_names = FALSE)
  tibble::tibble(
    country    = raw[[1]],
    n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))
  ) %>%
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) %>%
    harmonise_names() %>%
    dplyr::distinct(country, .keep_all = TRUE)
}
sbs_firms <- load_sbs_2017()


finalize_panel <- function(iso_full, filter_list, label) {
  res <- iso_full %>%
    dplyr::filter(country %in% filter_list) %>%
    dplyr::left_join(sbs_firms, by = "country") %>%
    dplyr::group_by(country) %>%
    dplyr::arrange(year, .by_group = TRUE) %>%
    dplyr::mutate(dplyr::across(c(gdp_pc, trade, rol, renew, rent), ~ {
      v <- zoo::na.approx(., na.rm = FALSE)
      v <- zoo::na.locf(zoo::na.locf(v, na.rm = FALSE), na.rm = FALSE, fromLast = TRUE)
      log1p(v)
    })) %>%
    # =========================================================================
    # Standardisation: log-difference of certifications.
    #
    # y_it = log(n_it + 1) - log(n_{i,t-1} + 1)
    #
    # Interpretation: approximately the growth rate of certifications within
    # each country. A value of y = 0.05 corresponds to a ~5% increase.
    # This is naturally comparable across countries of very different size
    # (cross-country heterogeneity in firm populations does not enter the
    # denominator), handles negative values (when expiries exceed new
    # issuances), and keeps y in a narrow numerical range suitable for the
    # solver.
    #
    # ALTERNATIVE (asinh of share, commented): would retain the "per 1,000
    # firms" interpretation but requires the SBS join and asinh to handle
    # negatives.
    # =========================================================================
    dplyr::mutate(y = log1p(n_cert) - log1p(dplyr::lag(n_cert))) %>%
    # dplyr::mutate(y = asinh(1e3 * (n_cert - dplyr::lag(n_cert)) / n_firms_17)) %>%
    dplyr::filter(!is.na(y)) %>%
    dplyr::ungroup()

  T_max <- max((res %>% dplyr::count(country))$n)
  final <- res %>%
    dplyr::group_by(country) %>%
    dplyr::filter(dplyr::n() == T_max) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(id = as.integer(factor(country))) %>%
    dplyr::select(id,
      time = year, y,
      x1 = gdp_pc, x2 = trade, x3 = rol, x4 = renew, x5 = rent
    ) %>%
    tidyr::drop_na()

  cat(sprintf("Dataset %s: N=%d, T=%d\n", label, length(unique(final$id)), T_max))
  return(final)
}

# --- Build panels ---
h9001 <- read_iso_hist(
  "01. ISO 9001 - Number of certificates per country and industry sectors - 1993 to 2017.xlsm",
  4:8, c("country", as.character(1993:2017))
)
r9001 <- extract_recent("9001")
f9001 <- dplyr::bind_rows(h9001, r9001)

h14001 <- read_iso_hist(
  "02. ISO 14001 - Number of certificates per country and industry sectors - 1999 to 2017.xlsm",
  4:8, c("country", as.character(1999:2017))
)
r14001 <- extract_recent("14001")
f14001 <- dplyr::bind_rows(h14001, r14001)

covs9001 <- make_covs(1993)
covs14001 <- make_covs(1999)
f9001 <- f9001 %>% dplyr::left_join(covs9001, by = c("country", "year"))
f14001 <- f14001 %>% dplyr::left_join(covs14001, by = c("country", "year"))

# Truncate to historical data (consistent methodology)
f9001 <- f9001 %>% dplyr::filter(year <= 2017)
f14001 <- f14001 %>% dplyr::filter(year <= 2017)

# Sync countries across standards
sync_eu <- intersect(
  (f9001 %>% filter(country %in% eu) %>% pull(country) %>% unique()),
  (f14001 %>% filter(country %in% eu) %>% pull(country) %>% unique())
)


dp_9001_eu <- finalize_panel(f9001, sync_eu, "9001_EU (synced)")
dp_14001_eu <- finalize_panel(f14001, sync_eu, "14001_EU (synced)")
```

```{r}
#| label: grid-search
#| include: false
#| cache: true
panels <- list(
  "9001_eu"     = dp_9001_eu,
  "14001_eu"    = dp_14001_eu
)

lambda_grid <- list(
  c(0.15, 0.20, 0.15),
  c(0.10, 0.10, 0.10),
  c(0.10, 0.10, 0.20),
  c(0.20, 0.20, 0.10),
  c(0.20, 0.20, 0.20),
  c(0.30, 0.10, 0.20)
)

grid_results <- list()
idx <- 0

for (pname in names(panels)) {
  df <- panels[[pname]]
  Wder <- NULL

  for (li in seq_along(lambda_grid)) {
    lam <- lambda_grid[[li]]
    idx <- idx + 1
    run_label <- sprintf("%s_gridrun%02d", pname, li)
    cat(sprintf(
      "\n[%d] %s | lambda = (%.3f, %.3f, %.3f) ... ",
      idx, pname, lam[1], lam[2], lam[3]
    ))

    row <- list(
      panel = pname, lambda = lam, run_label = run_label,
      edges = NA, rho_en = NA, rho_ada = NA, rho_gmm = NA, bic = NA, rn = NULL
    )

    tryCatch(
      {
        set.seed(1861)
        rn <- if (is.null(Wder)) {
          recoverNetwork(df, lambda = lam)
        } else {
          recoverNetwork(df, lambda = lam, Wder = Wder)
        }
        if (is.null(Wder)) Wder <- rn$Wder

        W <- as.matrix(rn$unpenalisedgmm$W)
        diag(W) <- 0

        row$edges <- sum(abs(W) > 1e-5)
        row$rho_en <- rn$en$rho
        row$rho_ada <- rn$adaen$rho
        row$rho_gmm <- rn$unpenalisedgmm$rho
        row$bic <- rn$BIC
        row$rn <- rn

        cat(sprintf(
          "OK | edges=%d | rho_gmm=%.3f | BIC=%.2f",
          row$edges, row$rho_gmm, row$bic
        ))
        saveRDS(rn, file.path(out_dir, paste0(run_label, ".rds")))
      },
      error = function(e) {
        cat("FAILED:", conditionMessage(e))
      }
    )

    grid_results[[idx]] <- row
  }

  Wder <- NULL
}

# Save everything for later recomputation
grid_meta <- lapply(grid_results, function(r) {
  r$rn <- NULL
  r
})
saveRDS(grid_meta, file.path(out_dir, "grid_search_meta.rds"))
save(grid_results, file = file.path(out_dir, "grid_search_full.RData"))
```

```{r}
#| label: grid-summary
#| echo: false
summary_df <- do.call(rbind, lapply(grid_results, function(r) {
  data.frame(
    Panel = gsub("_", " ", r$panel),
    L1 = r$lambda[1],
    L2 = r$lambda[2],
    L1s = r$lambda[3],
    Edges = ifelse(is.na(r$edges), "FAIL", as.character(r$edges)),
    rho_EN = ifelse(is.na(r$rho_en), "FAIL", sprintf("%.3f", r$rho_en)),
    rho_Ada = ifelse(is.na(r$rho_ada), "FAIL", sprintf("%.3f", r$rho_ada)),
    rho_GMM = ifelse(is.na(r$rho_gmm), "FAIL", sprintf("%.3f", r$rho_gmm)),
    BIC = ifelse(is.na(r$bic), "FAIL", sprintf("%.2f", r$bic)),
    stringsAsFactors = FALSE
  )
}))

# Rename panels to nicknames for display
summary_df$Panel <- gsub("9001", "Quality (9001)", summary_df$Panel)
summary_df$Panel <- gsub("14001", "Environment (14001)", summary_df$Panel)

kbl(summary_df,
  booktabs = TRUE, caption = "Grid search results",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE, align = "lrrrccccc"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r}
#| label: best-selection
#| include: false
best <- list()
for (pname in names(panels)) {
  panel_runs <- Filter(function(r) r$panel == pname && !is.null(r$rn), grid_results)
  if (length(panel_runs) > 0) {
    bics <- sapply(panel_runs, `[[`, "bic")
    best[[pname]] <- panel_runs[[which.min(bics)]]
  }
}
saveRDS(best, file.path(out_dir, "best_per_panel.rds"))

net_stats <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  B <- (abs(W) > 1e-5) * 1
  g <- graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(degree(g, mode = "out")),
    out_sd      = sd(degree(g, mode = "out")),
    in_mean     = mean(degree(g, mode = "in")),
    in_sd       = sd(degree(g, mode = "in"))
  )
}
```

\newpage
# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Data and transformations

The dependent variable is the log-difference of certifications:
$y_{it} = \ln(n_{it} + 1) - \ln(n_{i,t-1} + 1)$,
which approximates the within-country growth rate of adoption and is
naturally comparable across countries of very different size. The
log-transformation stabilises the variance of the increments, handles the
occasional years in which certificate expiries exceed new issuances
(yielding $y_{it}<0$), and keeps the dependent variable in a narrow
numerical range suitable for the penalised GMM solver. The unit addition
inside the logarithm ensures the transformation is defined even when a
country has zero certificates in the early years of the panel. Covariates
are transformed via $\log(1+x)$ to accommodate the occasional zero while
approximating the natural logarithm for non-zero values.

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, std_label, countries, years) {
  df %>%
    filter(country %in% countries, year %in% years) %>%
    group_by(year) %>%
    summarise(
      total_cert = sum(n_cert, na.rm = TRUE),
      mean_cert = mean(n_cert, na.rm = TRUE),
      sd_cert = sd(n_cert, na.rm = TRUE),
      min_cert = min(n_cert, na.rm = TRUE),
      max_cert = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries = n(),
      .groups = "drop"
    ) %>%
    dplyr::mutate(standard = std_label) %>%
    dplyr::select(standard, everything())
}

d9 <- desc_tab(f9001, "Quality", sync_eu, desc_years)
d14 <- desc_tab(f14001, "Environment", sync_eu, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Descriptive statistics: certification counts and adoption intensity across EU countries}\n")
cat("\\label{tab:desc_stats}\n")
cat("\\begin{tabular}{lcccccc}\n\\hline\\hline\n")
cat("& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\\n")
cat("\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}\n")
cat("& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\n\\hline\n")

cat(sprintf(
  "N countries & %d & %d & %d & & %d & %d \\\\\n",
  d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
  d14$n_countries[1], d14$n_countries[2]
))
cat(sprintf(
  "Total & %s & %s & %s & & %s & %s \\\\\n",
  f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
  f0(d14$total_cert[1]), f0(d14$total_cert[2])
))
cat(sprintf(
  "Mean & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
  f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
))
cat(sprintf(
  "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
  f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
))
cat(sprintf(
  "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\\n",
  f0(d9$min_cert[1]), f0(d9$max_cert[1]),
  f0(d9$min_cert[2]), f0(d9$max_cert[2]),
  f0(d9$min_cert[3]), f0(d9$max_cert[3]),
  f0(d14$min_cert[1]), f0(d14$max_cert[1]),
  f0(d14$min_cert[2]), f0(d14$max_cert[2])
))
cat(sprintf(
  "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\\n",
  f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
  f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
))

cat("\\hline\\hline\n\\end{tabular}\n")
cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Note:} Cert./GDP p.c.\\ is the mean across countries of certificates per 1,000 USD of GDP per capita (constant 2015 USD). Environment (ISO 14001) data begin in 1999.}\n")
cat("\\end{table}\n")
```

Table \ref{tab:desc_stats} reports raw certification counts across EU countries. Quality adoption is roughly four times larger than Environment by 2017, yet the development-adjusted intensity (last column) suggests that the gap narrows once economic size is accounted for.

## Network recovery

```{r}
#| label: table-eu
#| echo: false
#| results: asis
if (!is.null(best[["9001_eu"]]) && !is.null(best[["14001_eu"]])) {
  rn_a <- best[["9001_eu"]]$rn
  s_a <- net_stats(rn_a)
  rn_b <- best[["14001_eu"]]$rn
  s_b <- net_stats(rn_b)

  W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_a) <- 0
  W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_b) <- 0
  common <- sum(W_a & W_b)
  only_a <- sum(W_a & !W_b)
  only_b <- sum(!W_a & W_b)

  f3 <- function(x) formatC(x, format = "f", digits = 3)
  p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
  dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

  cat("\\begin{table}[H]\n\\centering\n")
  cat("\\caption{BIC-selected networks: Quality versus Environment (EU, $N = 27$)}\n")
  cat("\\label{tab:network_eu}\n")
  cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
  cat("& Quality & Environment \\\\\n\\hline\n")
  cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
  cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
  cat(sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\\n", only_a))
  cat(sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\\n", only_b))
  cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
  cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
  cat("Degree distribution across nodes (countries) & & \\\\\n")
  cat(sprintf(
    "\\quad Out-degree & %s & %s \\\\\n",
    dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
  ))
  cat(sprintf(
    "\\quad In-degree & %s & %s \\\\\n",
    dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
  ))
  cat("\\hline\n")
  cat(sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
    formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
    formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
  ))
  cat(sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\\n",
    paste(formatC(best[["9001_eu"]]$lambda, format = "f", digits = 3), collapse = ", "),
    paste(formatC(best[["14001_eu"]]$lambda, format = "f", digits = 3), collapse = ", ")
  ))
  cat(sprintf(
    "BIC & %s & %s \\\\\n",
    formatC(best[["9001_eu"]]$bic, format = "f", digits = 2),
    formatC(best[["14001_eu"]]$bic, format = "f", digits = 2)
  ))
  cat("\\hline\\hline\n\\end{tabular}\n")
  cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} This compares statistics from the BIC-selected estimated networks for Quality (ISO 9001, $T = 24$, 1993--2017) and Environment (ISO 14001, $T = 18$, 1999--2017) across 27 EU countries. The number of edges, edges in both networks, edges in $W_{\\text{Quality}}$ only, and edges in $W_{\\text{Environment}}$ only count the number of directed links in each category. Reciprocated edges is the frequency of in-edges that are reciprocated by out-edges. The clustering coefficient is the frequency of fully connected triplets over the total number of triplets. The degree distribution across nodes counts the average number of connections (standard deviation in parentheses), shown separately for in-degree and out-degree. $\\hat{\\rho}$ (GMM) is the post-LASSO estimate of the peer-effect parameter. $\\lambda = (\\lambda_1, \\lambda_2, \\lambda_1^*)$ are the elastic net penalty parameters selected by BIC.}\n")
  cat("\\end{table}\n")
} else {
  cat("\\textbf{EU table not available: one or both panels failed all grid runs.}\n")
}
```

The two networks are comparable in size despite the large difference in raw adoption volumes, suggesting that peer-effect channels operate at a similar density for both standards. The estimated $\hat{\rho}_{GMM}$ is reported at the bottom of Table \ref{tab:network_eu}. Reciprocated edges indicate country pairs with bidirectional influence.


The two networks are similar in aggregate (34 vs 38 edges, comparable clustering and degree distributions) but share only 3 edges, so the standards differ in *who diffuses to whom* rather than in how much diffusion occurs. The estimated peer-effect parameter is higher for Environment ($\hat\rho = 0.236$) than for Quality ($\hat\rho = 0.151$), matching the cross-country imitation pattern in Albuquerque, Bronnenberg and Corbett (2007, Figure 5),  who in a Bass-type diffusion model estimate cross-country imitation coefficients of approximately 0.16 for Quality and 0.27 for Environment (averaging early and late adopters), attributing the gap to the installed base of Quality having reduced uncertainty about the value of management standards by the time Environment was introduced.
```{r}
#| label: table-eu-same-lambda
#| echo: false
#| results: asis
#| output: false
shared_runs <- list()
for (li in seq_along(lambda_grid)) {
  lam <- lambda_grid[[li]]
  r9 <- Filter(function(r) r$panel == "9001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  r14 <- Filter(function(r) r$panel == "14001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  if (length(r9) > 0 && length(r14) > 0) {
    shared_runs[[length(shared_runs) + 1]] <- list(
      lambda = lam,
      avg_bic = (r9[[1]]$bic + r14[[1]]$bic) / 2,
      r9001 = r9[[1]], r14001 = r14[[1]]
    )
  }
}
shared_best <- shared_runs[[which.min(sapply(shared_runs, `[[`, "avg_bic"))]]

rn_a <- shared_best$r9001$rn
s_a <- net_stats(rn_a)
rn_b <- shared_best$r14001$rn
s_b <- net_stats(rn_b)

W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_a) <- 0
W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_b) <- 0
common <- sum(W_a & W_b)
only_a <- sum(W_a & !W_b)
only_b <- sum(!W_a & W_b)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

lam_str <- paste(formatC(shared_best$lambda, format = "f", digits = 3), collapse = ", ")

cat("\\begin{table}[H]\n\\centering\n")
cat(sprintf("\\caption{Robustness: common penalty $\\lambda = (%s)$ for both standards (EU, $N = 27$)}\n", lam_str))
cat("\\label{tab:network_eu_same_lambda}\n")
cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
cat("& Quality & Environment \\\\\n\\hline\n")
cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
cat(sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\\n", only_a))
cat(sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\\n", only_b))
cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
cat("Degree distribution across nodes (countries) & & \\\\\n")
cat(sprintf(
  "\\quad Out-degree & %s & %s \\\\\n",
  dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
))
cat(sprintf(
  "\\quad In-degree & %s & %s \\\\\n",
  dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
))
cat("\\hline\n")
cat(sprintf(
  "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
  formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
  formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
))
cat(sprintf(
  "BIC & %s & %s \\\\\n",
  formatC(shared_best$r9001$bic, format = "f", digits = 2),
  formatC(shared_best$r14001$bic, format = "f", digits = 2)
))
cat("\\hline\\hline\n\\end{tabular}\n")
cat(sprintf("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} Same as Table \\ref{tab:network_eu}, but imposing a common penalty vector $\\lambda = (%s)$ for both standards. The common $\\lambda$ is selected as the grid point with the lowest average BIC across the two panels. See Table \\ref{tab:network_eu} for variable definitions.}\n", lam_str))
cat("\\end{table}\n")
```


## Network visualisation

```{r}
#| label: country-labels
#| include: false
country_labels <- sort(sync_eu)
```
```{r}
#| label: fig-networks-weighted-full-2
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Full estimated peer-effect networks (BIC-selected $\\lambda$, EU, $N = 27$). All non-zero edges shown with width and colour intensity scaling continuously with $|\\hat w_{ij}|$. Darker/thicker edges indicate stronger peer-effect weights. Rows share a common layout: top row uses the Fruchterman--Reingold layout computed on the Quality network, bottom on Environment. Columns correspond to the network being plotted. Fixing node positions across columns allows visual comparison of edge structure without confounding by layout-algorithm placement."
make_full_weighted_graph <- function(rn, labels, eps = 1e-5) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  W[abs(W) < eps] <- 0
  g <- graph_from_adjacency_matrix(W, mode = "directed", weighted = TRUE)
  V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(best[["9001_eu"]]$rn, country_labels)
g_e_full <- make_full_weighted_graph(best[["14001_eu"]]$rn, country_labels)

# Global max |w| across both networks -> common colour/width scale
w_all <- c(abs(E(g_q_full)$weight), abs(E(g_e_full)$weight))
w_max <- max(w_all, 1e-9)

# Layouts computed on full networks (same seeds as other figures for comparability)
set.seed(1861)
layout_q_full <- layout_with_fr(g_q_full)
rownames(layout_q_full) <- V(g_q_full)$name
set.seed(1861)
layout_e_full <- layout_with_fr(g_e_full)
rownames(layout_e_full) <- V(g_e_full)$name

# Sequential viridis-like palette: light -> dark as |w| grows.
# Using built-in heat.colors reversed so weak = pale yellow, strong = dark red.
edge_palette <- colorRampPalette(c("#FEE0D2", "#FC9272", "#DE2D26", "#67000D"))(100)

plot_full_weighted_net <- function(g, lo, title, w_max) {
  lo_ordered <- lo[V(g)$name, , drop = FALSE]

  if (ecount(g) > 0) {
    w_abs <- abs(E(g)$weight)
    w_norm <- pmin(w_abs / w_max, 1) # 0-1 scale
    # Width: 0.4 (thinnest) -> 4 (thickest), linear in weight
    edge_wd <- 0.4 + 3.6 * w_norm
    # Colour: map w_norm to palette index
    col_idx <- pmax(1, ceiling(w_norm * length(edge_palette)))
    edge_col <- edge_palette[col_idx]
    # Alpha: weak edges more transparent so strong ones pop visually
    edge_alpha <- 0.35 + 0.55 * w_norm
    edge_col <- mapply(adjustcolor, edge_col, alpha.f = edge_alpha)
  } else {
    edge_wd <- numeric(0)
    edge_col <- character(0)
  }

  has_edge <- degree(g) > 0

  plot(g,
    layout             = lo_ordered,
    vertex.size        = 10,
    vertex.label       = V(g)$name,
    vertex.label.cex   = 0.55,
    vertex.label.dist  = 1.8, # era 1.4
    vertex.label.color = "grey20",
    vertex.color       = ifelse(has_edge, "steelblue", "grey85"),
    vertex.frame.color = ifelse(has_edge, "steelblue4", "grey65"),
    edge.arrow.size    = 0.6, # era 0.28
    edge.arrow.width   = 1.2, # nuovo parametro
    edge.color         = edge_col,
    edge.width         = edge_wd,
    main               = title
  )
  # Legend: 4 reference weights
  leg_vals <- c(0.2, 0.5, 0.8, 1.0) * w_max
  leg_vals <- leg_vals[leg_vals <= w_max]
  leg_norm <- leg_vals / w_max
  leg_wd <- 0.4 + 3.6 * leg_norm
  leg_col_ix <- pmax(1, ceiling(leg_norm * length(edge_palette)))
  leg_col <- edge_palette[leg_col_ix]
  leg_alpha <- 0.35 + 0.55 * leg_norm
  leg_col <- mapply(adjustcolor, leg_col, alpha.f = leg_alpha)

  legend("bottomright",
    title = expression("|" * hat(w)[ij] * "|"),
    legend = formatC(leg_vals, format = "f", digits = 2),
    lwd = leg_wd,
    col = leg_col,
    bty = "n",
    cex = 0.75,
    seg.len = 2.5
  )
}

par(mfrow = c(2, 2), mar = c(1, 1, 3, 1))
plot_full_weighted_net(g_q_full, layout_q_full, "Quality — Quality layout", w_max)
plot_full_weighted_net(g_e_full, layout_q_full, "Environment — Quality layout", w_max)
plot_full_weighted_net(g_q_full, layout_e_full, "Quality — Environment layout", w_max)
plot_full_weighted_net(g_e_full, layout_e_full, "Environment — Environment layout", w_max)
```

```{r}
#| label: table-cert-intensity
#| echo: false
#| results: asis
# Usa sbs_firms già caricato (n_firms_17 da Eurostat SBS, anno 2017).
# Intensita' = certs_2017 / n_firms_17, scalata per 1000 imprese.

cert_intensity <- sync_eu |>
  tibble::tibble(country = _) |>
  dplyr::left_join(
    f9001 |> filter(year == 2017, country %in% sync_eu) |>
      select(country, cert_9001 = n_cert),
    by = "country"
  ) |>
  dplyr::left_join(
    f14001 |> filter(year == 2017, country %in% sync_eu) |>
      select(country, cert_14001 = n_cert),
    by = "country"
  ) |>
  dplyr::left_join(sbs_firms, by = "country") |>
  dplyr::mutate(
    int_9001  = 1000 * cert_9001 / n_firms_17,
    int_14001 = 1000 * cert_14001 / n_firms_17
  ) |>
  dplyr::arrange(desc(int_9001))

fnum <- function(x, d = 0) formatC(x, format = "f", big.mark = ",", digits = d)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Certification intensity: ISO certificates per 1{,}000 active firms (2017, EU, $N = 27$)}\n")
cat("\\label{tab:cert_intensity}\n")
cat("\\begin{tabular}{lrrrrr}\n\\hline\\hline\n")
cat("Country & Firms & Quality & Environment & Quality/1k firms & Environment/1k firms \\\\\n\\hline\n")
for (i in seq_len(nrow(cert_intensity))) {
  r <- cert_intensity[i, ]
  cat(sprintf(
    "%s & %s & %s & %s & %s & %s \\\\\n",
    r$country,
    fnum(r$n_firms_17),
    fnum(r$cert_9001),
    fnum(r$cert_14001),
    fnum(r$int_9001, 2),
    fnum(r$int_14001, 2)
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Countries ranked in decreasing order of ISO 9001 certification intensity. ",
  "\\emph{Firms} is the 2017 count of active enterprises (Eurostat SBS, NACE Rev.\\ 2 B--S excluding K). ",
  "\\emph{9001} and \\emph{14001} are the respective 2017 certificate counts from the ISO Survey. ",
  "Intensity is reported as certificates per 1{,}000 firms.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-eigenvector2
#| echo: false
#| results: asis
# Eigenvector centrality via igraph::eigen_centrality with scale = FALSE.
# Outward = eigen_centrality on the reversed graph.

make_weighted_graph <- function(W) {
  W_abs <- abs(W)
  diag(W_abs) <- 0
  igraph::graph_from_adjacency_matrix(W_abs, mode = "directed", weighted = TRUE)
}

eigen_centrality_2way <- function(W, countries) {
  g <- make_weighted_graph(W)
  ec_in <- igraph::eigen_centrality(g, directed = TRUE)$vector
  ec_out <- igraph::eigen_centrality(igraph::reverse_edges(g), directed = TRUE)$vector
  tibble(country = countries, outward = ec_out, inward = ec_in)
}

N_TOP <- 5

countries_sorted <- sort(sync_eu)
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)

ec_q <- eigen_centrality_2way(W_q, countries_sorted)
ec_e <- eigen_centrality_2way(W_e, countries_sorted)

top_n_str <- function(df, col, n, digits = 3) {
  fmt <- paste0("%s (%.", digits, "f)")
  df |>
    arrange(desc(.data[[col]])) |>
    slice_head(n = n) |>
    mutate(entry = sprintf(fmt, country, .data[[col]])) |>
    pull(entry)
}

q_out_e <- top_n_str(ec_q, "outward", N_TOP)
q_in_e <- top_n_str(ec_q, "inward", N_TOP)
e_out_e <- top_n_str(ec_e, "outward", N_TOP)
e_in_e <- top_n_str(ec_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by eigenvector centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_eigen}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_e[r], q_in_e[r], e_out_e[r], e_in_e[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Eigenvector centrality computed via \\texttt{igraph::eigen\\_centrality()} on the absolute-value ",
  "weighted recovered network $|\\hat W|$. Scores are normalised so that the maximum entry equals 1 ",
  "(\\texttt{igraph} 2.1.1+ enforces normalisation regardless of the \\texttt{scale} argument). ",
  "Convention: $\\hat W_{ij}$ is the influence of $i$ on $j$. ",
  "\\emph{Outward} centrality is computed on the reversed graph, \\emph{inward} on the graph as ",
  "stated. On reducible networks the principal eigenvalue can have multiplicity greater than one ",
  "and the returned eigenvector reflects an arbitrary basis; PageRank ",
  "(Table \\ref{tab:centrality_pagerank}) is the robust alternative in that case.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-pagerank2
#| echo: false
#| results: asis
# PageRank via igraph::page_rank (default damping = 0.85).
# Outward = page_rank on the reversed graph.

pagerank_centrality_2way <- function(W, countries) {
  g <- make_weighted_graph(W)
  pr_in <- igraph::page_rank(g)$vector
  pr_out <- igraph::page_rank(igraph::reverse_edges(g))$vector
  tibble(country = countries, outward = pr_out, inward = pr_in)
}

pc_q <- pagerank_centrality_2way(W_q, countries_sorted)
pc_e <- pagerank_centrality_2way(W_e, countries_sorted)

q_out_p <- top_n_str(pc_q, "outward", N_TOP, digits = 4)
q_in_p <- top_n_str(pc_q, "inward", N_TOP, digits = 4)
e_out_p <- top_n_str(pc_e, "outward", N_TOP, digits = 4)
e_in_p <- top_n_str(pc_e, "inward", N_TOP, digits = 4)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by PageRank centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_pagerank}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_p[r], q_in_p[r], e_out_p[r], e_in_p[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "PageRank centrality computed via \\texttt{igraph::page\\_rank()} on the absolute-value weighted ",
  "recovered network $|\\hat W|$ (default damping factor $d = 0.85$). Convention: $\\hat W_{ij}$ ",
  "is the influence of $i$ on $j$. \\emph{Outward} centrality is PageRank on the reversed graph; ",
  "\\emph{inward} is on the graph as stated. PageRank values form a probability distribution over ",
  "nodes (each column sums to 1); a uniform PageRank across $N=27$ nodes would assign ",
  "$1/27 \\approx 0.0370$ to each. Unlike eigenvector centrality, PageRank is well-defined on ",
  "reducible networks thanks to the teleportation probability.}\n"
)
cat("\\end{table}\n")
```


```{r}
#| label: table-centrality-katz2
#| echo: false
#| results: asis
# Katz / alpha-centrality via centiserve::katzcent (default alpha = 0.1).
# centiserve::katzcent works on igraph objects and uses a sensible default
# (alpha = 0.1) that is below 1/rho for typical networks.
# Outward = katzcent on the reversed graph.

katz_centrality_2way <- function(W, countries) {
  g <- make_weighted_graph(W)
  kc_in <- igraph::alpha_centrality(g, alpha = 0.1, exo = 1)
  kc_out <- igraph::alpha_centrality(igraph::reverse_edges(g), alpha = 0.1, exo = 1)
  tibble(country = countries, outward = as.numeric(kc_out), inward = as.numeric(kc_in))
}

kc_q <- katz_centrality_2way(W_q, countries_sorted)
kc_e <- katz_centrality_2way(W_e, countries_sorted)

q_out_k <- top_n_str(kc_q, "outward", N_TOP)
q_in_k <- top_n_str(kc_q, "inward", N_TOP)
e_out_k <- top_n_str(kc_e, "outward", N_TOP)
e_in_k <- top_n_str(kc_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by Katz centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_katz}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_k[r], q_in_k[r], e_out_k[r], e_in_k[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Katz centrality computed via \\texttt{centiserve::katzcent()} on the absolute-value weighted ",
  "recovered network $|\\hat W|$ (default attenuation factor $\\alpha = 0.1$). ",
  "Convention: $\\hat W_{ij}$ is the influence of $i$ on $j$. ",
  "\\emph{Outward} centrality is computed ",
  "on the reversed graph, \\emph{inward} on the graph as stated. Katz centrality counts walks of ",
  "all lengths reaching each node, attenuated by $\\alpha^\\ell$ to ensure convergence.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-strength
#| echo: false
#| results: asis
#| output: false
# Weighted out-strength and in-strength (directed degree centrality).
#
# Convention: W[i,j] = influence of i on j (directed edge i -> j).
#   outward(i) = sum over j of |W[i,j]|  — total weight of edges leaving i
#   inward(i)  = sum over j of |W[j,i]|  — total weight of edges arriving at i
#
# This is the simplest possible centrality: direct sum of edge weights,
# no recursion, no eigenvector, no iterative solver. Impossible to degenerate.

strength_centrality_2way <- function(W, countries) {
  W_abs <- abs(W)
  diag(W_abs) <- 0

  out_str <- rowSums(W_abs) # walks of length 1 leaving i
  in_str <- colSums(W_abs) # walks of length 1 arriving at i

  norm01 <- function(x) if (max(x) > 0) x / max(x) else x

  tibble(
    country = countries,
    outward = norm01(out_str),
    inward  = norm01(in_str)
  )
}

sc_q <- strength_centrality_2way(W_q, countries_sorted)
sc_e <- strength_centrality_2way(W_e, countries_sorted)

q_out_s <- top_n_str(sc_q, "outward", N_TOP)
q_in_s <- top_n_str(sc_q, "inward", N_TOP)
e_out_s <- top_n_str(sc_e, "outward", N_TOP)
e_in_s <- top_n_str(sc_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by weighted strength (direct edge weights), by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_strength}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_s[r], q_in_s[r], e_out_s[r], e_in_s[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Weighted out- and in-strength on the absolute-value recovered network $|\\hat W|$. ",
  "Convention: $\\hat W_{ij}$ is the influence of $i$ on $j$. ",
  "\\emph{Outward}($i$) = $\\sum_{j \\neq i} |\\hat W_{ij}|$ is the total weight of edges ",
  "emanating from country $i$; ",
  "\\emph{inward}($i$) = $\\sum_{j \\neq i} |\\hat W_{ji}|$ is the total weight of edges ",
  "pointing at country $i$. ",
  "This is the simplest directed centrality: the direct sum of first-order edge weights, ",
  "with no recursive or spectral step. ",
  "It is robust to network sparsity and reducibility, at the cost of not accounting for ",
  "the importance of the neighbours each country connects to. ",
  "Scores are normalised to $[0,1]$, with 1 assigned to the country with the largest strength.}\n"
)
cat("\\end{table}\n")
```
PageRank centrality allows for a clear ranking of countries. The results are consistent with prior expectations: established EU economies (Germany and Austria for Quality) emerge as the main outward influencers, and established economies regarded as "green" (Sweden, Germany and Finland for Environment). Slovenia in particular ranks high on the Environment side, plausibly reflecting its integration into German supply chains and active engagement on sustainability commitments. The two standards display visibly different leader and follower dynamics.

# Dyadic determinants of peer-effect links

Following Dzemski (2018) and the dyadic linking framework reviewed in
de Paula (2020), I model each off-diagonal entry of the recovered
network as a function of pair-level covariates, with sender and receiver
fixed effects absorbing unobserved country-level propensity to send or
receive links:

\begin{equation}
W_{ij} = \mathbf{1}\bigl(X_{ij}^{\top}\beta + \alpha_i^{\text{out}} +
\alpha_j^{\text{in}} + \epsilon_{ij} \geq 0\bigr)
\label{eq:dzemski}
\end{equation}

The vector $X_{ij}$ collects pair-level covariates: eight homophily
terms $|\log(1+X_i^k) - \log(1+X_j^k)|$ spanning economic, institutional,
technological and environmental dimensions, plus bilateral integration
$\log(1+\mathrm{FDI}_{ij})$. Country-level heterogeneity in the
propensity to lead ($\alpha_i^{\text{out}}$) or follow
($\alpha_j^{\text{in}}$) is absorbed into two-sided fixed effects, so
that $\beta$ is identified only from within-pair variation in similarity
and bilateral ties.

The outcome is estimated in three forms: the edge indicator
$g_{ij} = \mathbb{1}\{|\hat w_{ij}|>10^{-5}\}$ (logit), the signed
weight $\hat w_{ij}$ (OLS), and its magnitude $|\hat w_{ij}|$ (PPML;
Santos Silva and Tenreyro, 2006). Standard errors are clustered two-way
on sender and receiver (Fafchamps and Gubert, 2007).


The two standards may indeed be diffusing on different channels. For Quality, FDI is positive and significant across all the specifications (with fixed effects) while no homophily term reaches conventional significance: ISO 9001 (Quality), apparently and consistently with previous literature,travels along supply-chain investment ties. For Environment FDI is null, but Government Effectiveness similarity is significant and negative, indicating homophily: ISO 14001 (Environment) diffusion concentrates within governance-quality clusters rather than bridging across them. Some specifications seem to also indicate that there is an influence of Research and Development intensity for the diffusion of Environmental standards, possibly suggesting that influencers in its adoption may be more likely to be innovation leaders, and that would be coherent with PageRank results. When Sender/receiver fixed effects are not included, none of the covariates is significant (results not included in the document).

```{r}
#| label: dzemski-tables-2
#| echo: false
#| results: asis
#| warning: false
#| message: false
suppressPackageStartupMessages({
  library(fixest)
  library(dplyr)
  library(tidyr)
  library(readr)
  library(purrr)
})

# ----------------------------------------------------------------------------
# 1. Eurostat monadic + bilateral FDI
# ----------------------------------------------------------------------------
eu_map <- c(
  AT = "Austria", BE = "Belgium", BG = "Bulgaria", CY = "Cyprus",
  CZ = "Czech Republic", DE = "Germany", DK = "Denmark", EE = "Estonia",
  EL = "Greece", ES = "Spain", FI = "Finland", FR = "France",
  HR = "Croatia", HU = "Hungary", IE = "Ireland", IT = "Italy",
  LT = "Lithuania", LU = "Luxembourg", LV = "Latvia", MT = "Malta",
  NL = "Netherlands", PL = "Poland", PT = "Portugal", RO = "Romania",
  SE = "Sweden", SI = "Slovenia", SK = "Slovak Republic"
)
eu_codes <- names(eu_map)

rd <- function(n) {
  readr::read_csv(
    file.path(data_dir, n),
    show_col_types = FALSE,
    col_types = cols(sex = col_character(), .default = col_guess())
  ) |>
    dplyr::rename_with(~ tolower(gsub(" ", "_", .))) |>
    dplyr::mutate(obs_value = suppressWarnings(as.numeric(obs_value))) |>
    dplyr::filter(!is.na(obs_value))
}

fdi_h <- rd("estat_bop_fdi6_pos_en.csv.gz") |>
  dplyr::filter(
    stk_flow == "ASS", fdi_item == "DI__D__F", currency == "MIO_EUR",
    nace_r2 == "FDI", counterp == "IMM", entity == "TOTAL",
    geo %in% eu_codes, partner %in% eu_codes, geo != partner
  ) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(country_i = eu_map[geo], country_j = eu_map[partner]) |>
  dplyr::select(country_i, country_j, fdi)

mon_collect <- function(df) {
  df |>
    dplyr::filter(geo %in% eu_codes) |>
    dplyr::group_by(geo) |>
    dplyr::summarise(v = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
    dplyr::mutate(country = eu_map[geo]) |>
    dplyr::select(country, v) |>
    dplyr::filter(!is.nan(v))
}

pop_h <- rd("estat_demo_gind_en.csv.gz") |>
  dplyr::filter(indic_de == "JAN") |>
  mon_collect() |>
  dplyr::rename(pop = v)

etax_raw <- rd("estat_env_ac_taxind2_en.csv.gz") |>
  dplyr::filter(tax == "ENV", unit == "MIO_EUR")
etax_nace <- if ("TOTAL" %in% etax_raw$nace_r2) "TOTAL" else if ("A" %in% etax_raw$nace_r2) "A" else unique(etax_raw$nace_r2)[1]
etax_h <- etax_raw |>
  dplyr::filter(nace_r2 == etax_nace) |>
  mon_collect() |>
  dplyr::rename(etax = v)

ghg_raw <- rd("estat_env_air_gge_en.csv.gz") |>
  dplyr::filter(airpol == "GHG", unit == "MIO_T")
ghg_src <- if ("TOTXMEMO" %in% ghg_raw$src_crf) "TOTXMEMO" else if ("TOTX4_MEMO" %in% ghg_raw$src_crf) "TOTX4_MEMO" else unique(ghg_raw$src_crf)[1]
ghg_h <- ghg_raw |>
  dplyr::filter(src_crf == ghg_src) |>
  mon_collect() |>
  dplyr::rename(ghg = v)

rd_h <- rd("estat_rd_p_persocc_en.csv.gz") |>
  dplyr::filter(sectperf == "TOTAL", prof_pos == "TOTAL", sex == "T", unit == "FTE") |>
  mon_collect() |>
  dplyr::rename(rd = v)

est_mon_h <- pop_h |>
  dplyr::full_join(etax_h, "country") |>
  dplyr::full_join(ghg_h, "country") |>
  dplyr::full_join(rd_h, "country")

# ----------------------------------------------------------------------------
# 2. WDI monadic
# ----------------------------------------------------------------------------
load_wdi_mean <- function(f, v, start_yr = 1999, end_yr = 2017) {
  readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) |>
    dplyr::select(country = 1, as.character(start_yr):as.character(end_yr)) |>
    tidyr::pivot_longer(-country, names_to = "year", values_to = "v") |>
    dplyr::mutate(v = suppressWarnings(as.numeric(v))) |>
    harmonise_names() |>
    dplyr::filter(country %in% sync_eu) |>
    dplyr::group_by(country) |>
    dplyr::summarise(!!v := mean(v, na.rm = TRUE), .groups = "drop")
}

wdi_files_h <- list(
  c("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc"),
  c("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
  c("API_GE.EST_DS2_en_csv_v2_213352.csv", "gov_eff"),
  c("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew")
)

wdi_h <- purrr::reduce(
  purrr::map(wdi_files_h, ~ load_wdi_mean(.x[1], .x[2])),
  dplyr::full_join,
  by = "country"
)

mon_all_h <- dplyr::full_join(est_mon_h, wdi_h, by = "country")

mon_vars_h <- c(
  "pop", "gdp_pc", "gov_eff", "rd", "trade",
  "ghg", "renew", "etax"
)

# ----------------------------------------------------------------------------
# 3. Dyadic data — tab2: gov_eff in livelli
# ----------------------------------------------------------------------------
cs_h <- sort(sync_eu)
N_h <- length(cs_h)

log_vars_2 <- c("pop", "gdp_pc", "rd", "trade", "ghg", "renew", "etax")
# gov_eff intenzionalmente escluso: indice WGI già in [-2.5, 2.5]

build_dyad_2 <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  g <- expand.grid(i = seq_len(N_h), j = seq_len(N_h), KEEP.OUT.ATTRS = FALSE)
  g <- g[g$i != g$j, ]
  ci <- cs_h[g$i]
  cj <- cs_h[g$j]
  wt <- as.numeric(W[cbind(g$i, g$j)])

  diff_terms <- list()
  for (v in mon_vars_h) {
    raw <- as.numeric(mon_all_h[[v]])
    x_clean <- if (v %in% log_vars_2) log1p(raw) else raw
    x <- setNames(x_clean, mon_all_h$country)
    diff_terms[[paste0("diff_", v)]] <- as.numeric(abs(x[ci] - x[cj]))
  }

  k <- paste(ci, cj, sep = "||")
  fdi_lk <- setNames(
    as.numeric(fdi_h$fdi),
    paste(fdi_h$country_i, fdi_h$country_j, sep = "||")
  )
  fv <- unname(as.numeric(fdi_lk[k]))
  fv[is.na(fv) | fv < 0] <- 0

  as.data.frame(c(
    list(
      country_i = ci, country_j = cj,
      weight = wt, abs_w = abs(wt),
      edge = as.integer(abs(wt) > 1e-5),
      log_fdi = log1p(fv)
    ),
    diff_terms
  ), stringsAsFactors = FALSE)
}

d9_2 <- build_dyad_2(best[["9001_eu"]]$rn)
d14_2 <- build_dyad_2(best[["14001_eu"]]$rn)

# ----------------------------------------------------------------------------
# 4. Fit
# ----------------------------------------------------------------------------
rhs_h <- c(paste0("diff_", mon_vars_h), "log_fdi")
fe_part <- "country_i + country_j"
make_fml <- function(y) {
  as.formula(paste(y, "~", paste(rhs_h, collapse = " + "), "|", fe_part))
}

m2_logit <- list(
  "Quality" = feglm(make_fml("edge"), d9_2,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  ),
  "Environment" = feglm(make_fml("edge"), d14_2,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  )
)
m2_ols <- list(
  "Quality" = feols(make_fml("weight"), d9_2,
    cluster = ~ country_i + country_j
  ),
  "Environment" = feols(make_fml("weight"), d14_2,
    cluster = ~ country_i + country_j
  )
)
m2_ppml <- list(
  "Quality" = fepois(make_fml("abs_w"), d9_2,
    cluster = ~ country_i + country_j
  ),
  "Environment" = fepois(make_fml("abs_w"), d14_2,
    cluster = ~ country_i + country_j
  )
)

# ----------------------------------------------------------------------------
# 5. Render
# ----------------------------------------------------------------------------
dict_2 <- c(
  diff_pop     = "$|\\Delta \\log|$ Population",
  diff_gdp_pc  = "$|\\Delta \\log|$ GDP per capita",
  diff_gov_eff = "$|\\Delta|$ Gov.\\ effectiveness",
  diff_rd      = "$|\\Delta \\log|$ R\\&D personnel",
  diff_trade   = "$|\\Delta \\log|$ Trade openness",
  diff_ghg     = "$|\\Delta \\log|$ GHG emissions",
  diff_renew   = "$|\\Delta \\log|$ Renewables share",
  diff_etax    = "$|\\Delta \\log|$ Env.\\ tax revenue",
  log_fdi      = "$\\log(1+\\mathrm{FDI}_{ij})$",
  country_i    = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  country_j    = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

render_2 <- function(models, title, label, notes, fitstat) {
  tex <- paste(capture.output(
    etable(models,
      tex = TRUE, title = title, label = label,
      dict = dict_2, notes = notes,
      headers = list("^_Network" = names(models)),
      digits = 3, digits.stats = 3,
      fitstat = fitstat,
      style.tex = style.tex(
        depvar.title = "", fixef.suffix = ":",
        fixef.title  = "", var.title    = "",
        model.title  = ""
      )
    )
  ), collapse = "\n")
  fn <- paste0("tab_", gsub("[-:]", "_", label), ".tex")
  writeLines(tex, file.path(out_dir, fn))
  cat(tex, sep = "\n")
  cat("\n\n")
}

note_2 <- paste(
  "Log-homophily covariates constructed as $|\\log(1+X_i)-\\log(1+X_j)|$.",
  "Government Effectiveness (WGI) enters in levels as it already lies in $[-2.5, 2.5]$.",
  "Sender and receiver FE absorbed. Two-way clustered SE."
)

render_2(m2_logit,
  title   = "Dyadic Regression (logit) on edge indicator $g_{ij}$.",
  label   = "tab:dz-logit-2",
  notes   = note_2,
  fitstat = ~ n + pr2 + bic
)

render_2(m2_ols,
  title   = "Dyadic Regression (OLS) on $\\hat w_{ij}$.",
  label   = "tab:dz-ols-2",
  notes   = note_2,
  fitstat = ~ n + r2 + ar2 + bic
)

render_2(m2_ppml,
  title   = "Dyadic Regression (PPML) on $|\\hat w_{ij}|$.",
  label   = "tab:dz-ppml-2",
  notes   = paste(note_2, "Coefficients are elasticities."),
  fitstat = ~ n + pr2 + bic
)
```

```{r}
#| label: dzemski-tables-3
#| echo: false
#| results: asis
#| warning: false
#| message: false
suppressPackageStartupMessages({
  library(fixest)
  library(dplyr)
  library(tidyr)
  library(readr)
  library(purrr)
})

# ----------------------------------------------------------------------------
# 1. Eurostat monadic + bilateral FDI
# ----------------------------------------------------------------------------
eu_map <- c(
  AT = "Austria", BE = "Belgium", BG = "Bulgaria", CY = "Cyprus",
  CZ = "Czech Republic", DE = "Germany", DK = "Denmark", EE = "Estonia",
  EL = "Greece", ES = "Spain", FI = "Finland", FR = "France",
  HR = "Croatia", HU = "Hungary", IE = "Ireland", IT = "Italy",
  LT = "Lithuania", LU = "Luxembourg", LV = "Latvia", MT = "Malta",
  NL = "Netherlands", PL = "Poland", PT = "Portugal", RO = "Romania",
  SE = "Sweden", SI = "Slovenia", SK = "Slovak Republic"
)
eu_codes <- names(eu_map)

rd <- function(n) {
  readr::read_csv(
    file.path(data_dir, n),
    show_col_types = FALSE,
    col_types = cols(sex = col_character(), .default = col_guess())
  ) |>
    dplyr::rename_with(~ tolower(gsub(" ", "_", .))) |>
    dplyr::mutate(obs_value = suppressWarnings(as.numeric(obs_value))) |>
    dplyr::filter(!is.na(obs_value))
}

fdi_h <- rd("estat_bop_fdi6_pos_en.csv.gz") |>
  dplyr::filter(
    stk_flow == "ASS", fdi_item == "DI__D__F", currency == "MIO_EUR",
    nace_r2 == "FDI", counterp == "IMM", entity == "TOTAL",
    geo %in% eu_codes, partner %in% eu_codes, geo != partner
  ) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(country_i = eu_map[geo], country_j = eu_map[partner]) |>
  dplyr::select(country_i, country_j, fdi)

mon_collect <- function(df) {
  df |>
    dplyr::filter(geo %in% eu_codes) |>
    dplyr::group_by(geo) |>
    dplyr::summarise(v = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
    dplyr::mutate(country = eu_map[geo]) |>
    dplyr::select(country, v) |>
    dplyr::filter(!is.nan(v))
}

pop_h <- rd("estat_demo_gind_en.csv.gz") |>
  dplyr::filter(indic_de == "JAN") |>
  mon_collect() |>
  dplyr::rename(pop = v)

etax_raw <- rd("estat_env_ac_taxind2_en.csv.gz") |>
  dplyr::filter(tax == "ENV", unit == "MIO_EUR")
etax_nace <- if ("TOTAL" %in% etax_raw$nace_r2) "TOTAL" else if ("A" %in% etax_raw$nace_r2) "A" else unique(etax_raw$nace_r2)[1]
etax_h <- etax_raw |>
  dplyr::filter(nace_r2 == etax_nace) |>
  mon_collect() |>
  dplyr::rename(etax = v)

ghg_raw <- rd("estat_env_air_gge_en.csv.gz") |>
  dplyr::filter(airpol == "GHG", unit == "MIO_T")
ghg_src <- if ("TOTXMEMO" %in% ghg_raw$src_crf) "TOTXMEMO" else if ("TOTX4_MEMO" %in% ghg_raw$src_crf) "TOTX4_MEMO" else unique(ghg_raw$src_crf)[1]
ghg_h <- ghg_raw |>
  dplyr::filter(src_crf == ghg_src) |>
  mon_collect() |>
  dplyr::rename(ghg = v)

rd_h <- rd("estat_rd_p_persocc_en.csv.gz") |>
  dplyr::filter(sectperf == "TOTAL", prof_pos == "TOTAL", sex == "T", unit == "FTE") |>
  mon_collect() |>
  dplyr::rename(rd = v)

est_mon_h <- pop_h |>
  dplyr::full_join(etax_h, "country") |>
  dplyr::full_join(ghg_h, "country") |>
  dplyr::full_join(rd_h, "country")

# ----------------------------------------------------------------------------
# 2. WDI monadic
# ----------------------------------------------------------------------------
load_wdi_mean <- function(f, v, start_yr = 1999, end_yr = 2017) {
  readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) |>
    dplyr::select(country = 1, as.character(start_yr):as.character(end_yr)) |>
    tidyr::pivot_longer(-country, names_to = "year", values_to = "v") |>
    dplyr::mutate(v = suppressWarnings(as.numeric(v))) |>
    harmonise_names() |>
    dplyr::filter(country %in% sync_eu) |>
    dplyr::group_by(country) |>
    dplyr::summarise(!!v := mean(v, na.rm = TRUE), .groups = "drop")
}

wdi_files_h <- list(
  c("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc"),
  c("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
  c("API_GE.EST_DS2_en_csv_v2_213352.csv", "gov_eff"),
  c("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew")
)

wdi_h <- purrr::reduce(
  purrr::map(wdi_files_h, ~ load_wdi_mean(.x[1], .x[2])),
  dplyr::full_join,
  by = "country"
)

mon_all_h <- dplyr::full_join(est_mon_h, wdi_h, by = "country")

mon_vars_h <- c(
  "pop", "gdp_pc", "gov_eff", "rd", "trade",
  "ghg", "renew", "etax"
)

# ----------------------------------------------------------------------------
# 3. Dyadic data — TUTTO IN LIVELLI
# ----------------------------------------------------------------------------
cs_h <- sort(sync_eu)
N_h <- length(cs_h)

build_dyad_2 <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  g <- expand.grid(i = seq_len(N_h), j = seq_len(N_h), KEEP.OUT.ATTRS = FALSE)
  g <- g[g$i != g$j, ]
  ci <- cs_h[g$i]
  cj <- cs_h[g$j]
  wt <- as.numeric(W[cbind(g$i, g$j)])

  diff_terms <- list()
  for (v in mon_vars_h) {
    x <- setNames(as.numeric(mon_all_h[[v]]), mon_all_h$country)
    diff_terms[[paste0("diff_", v)]] <- as.numeric(abs(x[ci] - x[cj]))
  }

  k <- paste(ci, cj, sep = "||")
  fdi_lk <- setNames(
    as.numeric(fdi_h$fdi),
    paste(fdi_h$country_i, fdi_h$country_j, sep = "||")
  )
  fv <- unname(as.numeric(fdi_lk[k]))
  fv[is.na(fv) | fv < 0] <- 0

  as.data.frame(c(
    list(
      country_i = ci, country_j = cj,
      weight = wt, abs_w = abs(wt),
      edge = as.integer(abs(wt) > 1e-5),
      log_fdi = log1p(fv)
    ),
    diff_terms
  ), stringsAsFactors = FALSE)
}

d9_2 <- build_dyad_2(best[["9001_eu"]]$rn)
d14_2 <- build_dyad_2(best[["14001_eu"]]$rn)

# ----------------------------------------------------------------------------
# 4. Fit
# ----------------------------------------------------------------------------
rhs_h <- c(paste0("diff_", mon_vars_h), "log_fdi")
fe_part <- "country_i + country_j"
make_fml <- function(y) {
  as.formula(paste(y, "~", paste(rhs_h, collapse = " + "), "|", fe_part))
}

m2_logit <- list(
  "Quality" = feglm(make_fml("edge"), d9_2,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  ),
  "Environment" = feglm(make_fml("edge"), d14_2,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  )
)
m2_ols <- list(
  "Quality" = feols(make_fml("weight"), d9_2,
    cluster = ~ country_i + country_j
  ),
  "Environment" = feols(make_fml("weight"), d14_2,
    cluster = ~ country_i + country_j
  )
)
m2_ppml <- list(
  "Quality" = fepois(make_fml("abs_w"), d9_2,
    cluster = ~ country_i + country_j
  ),
  "Environment" = fepois(make_fml("abs_w"), d14_2,
    cluster = ~ country_i + country_j
  )
)

# ----------------------------------------------------------------------------
# 5. Render
# ----------------------------------------------------------------------------
dict_2 <- c(
  diff_pop     = "$|\\Delta|$ Population",
  diff_gdp_pc  = "$|\\Delta|$ GDP per capita",
  diff_gov_eff = "$|\\Delta|$ Gov.\\ effectiveness",
  diff_rd      = "$|\\Delta|$ R\\&D personnel",
  diff_trade   = "$|\\Delta|$ Trade openness",
  diff_ghg     = "$|\\Delta|$ GHG emissions",
  diff_renew   = "$|\\Delta|$ Renewables share",
  diff_etax    = "$|\\Delta|$ Env.\\ tax revenue",
  log_fdi      = "$\\log(1+\\mathrm{FDI}_{ij})$",
  country_i    = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  country_j    = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

render_2 <- function(models, title, label, notes, fitstat) {
  tex <- paste(capture.output(
    etable(models,
      tex = TRUE, title = title, label = label,
      dict = dict_2, notes = notes,
      headers = list("^_Network" = names(models)),
      digits = 3, digits.stats = 3,
      fitstat = fitstat,
      style.tex = style.tex(
        depvar.title = "", fixef.suffix = ":",
        fixef.title  = "", var.title    = "",
        model.title  = ""
      )
    )
  ), collapse = "\n")
  fn <- paste0("tab_", gsub("[-:]", "_", label), ".tex")
  writeLines(tex, file.path(out_dir, fn))
  cat(tex, sep = "\n")
  cat("\n\n")
}

note_2 <- paste(
  "Homophily covariates constructed as $|X_i - X_j|$ on raw levels.",
  "Bilateral FDI enters as $\\log(1+\\mathrm{FDI}_{ij})$.",
  "Sender and receiver FE absorbed. Two-way clustered SE."
)

render_2(m2_logit,
  title   = "Dyadic Regression (logit) on edge indicator $g_{ij}$.",
  label   = "tab:dz-logit-2",
  notes   = note_2,
  fitstat = ~ n + pr2 + bic
)

render_2(m2_ols,
  title   = "Dyadic Regression (OLS) on $\\hat w_{ij}$.",
  label   = "tab:dz-ols-2",
  notes   = note_2,
  fitstat = ~ n + r2 + ar2 + bic
)

render_2(m2_ppml,
  title   = "Dyadic Regression (PPML) on $|\\hat w_{ij}|$.",
  label   = "tab:dz-ppml-2",
  notes   = paste(note_2, "Coefficients are elasticities."),
  fitstat = ~ n + pr2 + bic
)
```

```{r}
#| label: dzemski-tables-simple-nofe
#| echo: false
#| results: asis
#| warning: false
#| message: false
#| output: false
# ----------------------------------------------------------------------------
# Specifica semplice: differenze con segno X_i - X_j, no FE, no clustering
# Tutti i dati riusati da dzemski-tables-2: mon_all_h, fdi_h, cs_h, N_h
# ----------------------------------------------------------------------------

build_dyad_simple <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  g <- expand.grid(i = seq_len(N_h), j = seq_len(N_h), KEEP.OUT.ATTRS = FALSE)
  g <- g[g$i != g$j, ]
  ci <- cs_h[g$i]
  cj <- cs_h[g$j]
  wt <- as.numeric(W[cbind(g$i, g$j)])

  diff_terms <- list()
  for (v in mon_vars_h) {
    x <- setNames(as.numeric(mon_all_h[[v]]), mon_all_h$country)
    # SIGNED difference: X_i - X_j (no abs)
    diff_terms[[paste0("diff_", v)]] <- as.numeric(x[ci] - x[cj])
  }

  k <- paste(ci, cj, sep = "||")
  fdi_lk <- setNames(
    as.numeric(fdi_h$fdi),
    paste(fdi_h$country_i, fdi_h$country_j, sep = "||")
  )
  fv <- unname(as.numeric(fdi_lk[k]))
  fv[is.na(fv) | fv < 0] <- 0

  as.data.frame(c(
    list(
      country_i = ci, country_j = cj,
      weight = wt, abs_w = abs(wt),
      edge = as.integer(abs(wt) > 1e-5),
      log_fdi = log1p(fv)
    ),
    diff_terms
  ), stringsAsFactors = FALSE)
}

d9_s <- build_dyad_simple(best[["9001_eu"]]$rn)
d14_s <- build_dyad_simple(best[["14001_eu"]]$rn)

# Specifica senza FE
rhs_s <- c(paste0("diff_", mon_vars_h), "log_fdi")
make_fml_simple <- function(y) {
  as.formula(paste(y, "~", paste(rhs_s, collapse = " + ")))
}

ms_logit <- list(
  "Quality"     = feglm(make_fml_simple("edge"), d9_s, family = binomial("logit")),
  "Environment" = feglm(make_fml_simple("edge"), d14_s, family = binomial("logit"))
)
ms_ols <- list(
  "Quality"     = feols(make_fml_simple("weight"), d9_s),
  "Environment" = feols(make_fml_simple("weight"), d14_s)
)
ms_ppml <- list(
  "Quality"     = fepois(make_fml_simple("abs_w"), d9_s),
  "Environment" = fepois(make_fml_simple("abs_w"), d14_s)
)

dict_s <- c(
  diff_pop     = "$\\Delta$ Population",
  diff_gdp_pc  = "$\\Delta$ GDP per capita",
  diff_gov_eff = "$\\Delta$ Gov.\\ effectiveness",
  diff_rd      = "$\\Delta$ R\\&D personnel",
  diff_trade   = "$\\Delta$ Trade openness",
  diff_ghg     = "$\\Delta$ GHG emissions",
  diff_renew   = "$\\Delta$ Renewables share",
  diff_etax    = "$\\Delta$ Env.\\ tax revenue",
  log_fdi      = "$\\log(1+\\mathrm{FDI}_{ij})$"
)

render_s <- function(models, title, label, notes, fitstat) {
  tex <- paste(capture.output(
    etable(models,
      tex = TRUE, title = title, label = label,
      dict = dict_s, notes = notes,
      headers = list("^_Network" = names(models)),
      digits = 3, digits.stats = 3,
      fitstat = fitstat,
      style.tex = style.tex(
        depvar.title = "", fixef.suffix = ":",
        fixef.title  = "", var.title    = "",
        model.title  = ""
      )
    )
  ), collapse = "\n")
  fn <- paste0("tab_", gsub("[-:]", "_", label), ".tex")
  writeLines(tex, file.path(out_dir, fn))
  cat(tex, sep = "\n")
  cat("\n\n")
}

note_s <- paste(
  "Signed pair-level differences $\\Delta X_{ij} = X_i - X_j$ on raw levels.",
  "A positive coefficient means high-$X$ countries are more likely to influence low-$X$ countries.",
  "No fixed effects, no clustering. Preliminary specification."
)

render_s(ms_logit,
  title   = "Simple dyadic regression (logit) on edge indicator $g_{ij}$ — signed differences, no FE.",
  label   = "tab:dz-logit-s",
  notes   = note_s,
  fitstat = ~ n + pr2 + bic
)

render_s(ms_ols,
  title   = "Simple dyadic regression (OLS) on $\\hat w_{ij}$ — signed differences, no FE.",
  label   = "tab:dz-ols-s",
  notes   = note_s,
  fitstat = ~ n + r2 + ar2 + bic
)

render_s(ms_ppml,
  title   = "Simple dyadic regression (PPML) on $|\\hat w_{ij}|$ — signed differences, no FE.",
  label   = "tab:dz-ppml-s",
  notes   = paste(note_s, "Coefficients are elasticities."),
  fitstat = ~ n + pr2 + bic
)
```