---
title: "Preliminary Results"
author: "Gianluigi De Rubertis"
date: "`r Sys.Date()`"
format:
  pdf:
    output-file: "results_eu_grid.pdf"
    keep-tex: false
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{pdflscape}
---

```{r}
#| label: setup
#| include: false
library(kableExtra)
library(dplyr)
library(magrittr)
library(readxl)
library(stringr)
library(tidyr)
library(janitor)
library(purrr)
library(ggplot2)
library(readr)
library(igraph)
library(remotes)
library(recoverNetwork)

set.seed(1861)
setwd("/workspaces/thesis")
options(readr.show_col_types = FALSE)

out_dir <- "output"
if (!dir.exists(out_dir)) dir.create(out_dir)
```

```{r}
#| label: data-cleaning
#| include: false
#| cache: true
data_dir <- "/workspaces/thesis/data"

# === COUNTRY GROUPS ===
oecd <- c(
  "Australia", "Austria", "Belgium", "Canada", "Chile", "Colombia",
  "Costa Rica", "Czech Republic", "Denmark", "Estonia", "Finland",
  "France", "Germany", "Greece", "Hungary", "Iceland", "Ireland",
  "Israel", "Italy", "Japan", "Korea", "Latvia", "Lithuania",
  "Luxembourg", "Mexico", "Netherlands", "New Zealand", "Norway",
  "Poland", "Portugal", "Slovak Republic", "Slovenia", "Spain",
  "Sweden", "Switzerland", "Turkey", "United Kingdom", "United States"
)

eu <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus",
  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Ireland", "Italy", "Latvia",
  "Lithuania", "Luxembourg", "Malta", "Netherlands", "Poland",
  "Portugal", "Romania", "Slovak Republic", "Slovenia", "Spain",
  "Sweden"
)

harmonise_names <- function(df) {
  if (nrow(df) == 0) {
    return(df)
  }
  df %>% dplyr::mutate(country = dplyr::case_match(country,
    "Korea (Republic of)" ~ "Korea", "Korea, Republic of" ~ "Korea",
    "Korea (the Republic of)" ~ "Korea",
    "Czechia" ~ "Czech Republic", "Slovakia" ~ "Slovak Republic",
    "United Kingdom of Great Britain and Northern Ireland" ~ "United Kingdom",
    "United States of America" ~ "United States",
    "Netherlands (Kingdom of the)" ~ "Netherlands",
    .default = country
  ))
}

read_iso_hist <- function(file_name, sheets_idx, my_names) {
  f_path <- file.path(data_dir, file_name)
  if (!file.exists(f_path)) stop(paste("File not found:", f_path))
  sheets <- purrr::map(sheets_idx, ~ readxl::read_excel(f_path, sheet = .x))
  expected_cols <- length(my_names)
  purrr::map(1:5, ~ sheets[[.x]] %>%
    dplyr::select(dplyr::all_of(1:expected_cols)) %>%
    dplyr::slice(-(1:2)) %>%
    purrr::set_names(my_names) %>%
    tidyr::pivot_longer(-country, names_to = "year", values_to = "n_cert")) %>%
    purrr::reduce(dplyr::bind_rows) %>%
    harmonise_names() %>%
    dplyr::mutate(n_cert = as.numeric(n_cert), year = as.integer(year)) %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

extract_recent <- function(iso_code) {
  files <- list.files(data_dir,
    pattern = "ISO Survey 20(1[8-9]|2[0-9]).*\\.xlsx$",
    full.names = TRUE
  )
  purrr::map_dfr(files, function(f) {
    yr <- as.numeric(stringr::str_extract(basename(f), "20\\d{2}"))
    sheets <- readxl::excel_sheets(f)
    target <- sheets[grepl(iso_code, sheets)][1]
    if (!is.na(target)) {
      df <- readxl::read_excel(f, sheet = target)
      start_row <- which(apply(df, 1, function(x) any(grepl("Afghanistan|Albania", x))))[1]
      df_clean <- df %>%
        dplyr::slice(start_row:dplyr::n()) %>%
        dplyr::rename(country = 1) %>%
        dplyr::mutate(dplyr::across(-1, ~ suppressWarnings(as.numeric(.)))) %>%
        dplyr::select(country, tidyselect::where(is.numeric)) %>%
        dplyr::mutate(n_cert = rowSums(dplyr::select(., -country), na.rm = TRUE)) %>%
        dplyr::select(country, n_cert) %>%
        dplyr::mutate(year = yr)
      return(df_clean)
    }
    return(NULL)
  }) %>%
    harmonise_names() %>%
    dplyr::group_by(country, year) %>%
    dplyr::summarise(n_cert = sum(n_cert, na.rm = TRUE), .groups = "drop")
}

make_covs <- function(start_yr) {
  load_w <- function(f, v) {
    readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) %>%
      dplyr::select(country = 1, as.character(start_yr):"2024") %>%
      tidyr::pivot_longer(-country, names_to = "year", values_to = v) %>%
      dplyr::mutate(
        year = as.integer(year),
        !!v := as.numeric(as.character(.data[[v]]))
      ) %>%
      harmonise_names()
  }
  load_w("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc") %>%
    dplyr::full_join(load_w("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_RL.EST_DS2_en_csv_v2_1905.csv", "rol"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew"),
      by = c("country", "year")
    ) %>%
    dplyr::full_join(load_w("API_NY.GDP.TOTL.RT.ZS_DS2_en_csv_v2_362.csv", "rent"),
      by = c("country", "year")
    )
}

load_sbs_2017 <- function() {
  file_path <- file.path(data_dir, "enterprises_eu.xlsx")
  raw <- readxl::read_excel(file_path, sheet = "Sheet 1", skip = 9, col_names = FALSE)
  tibble::tibble(
    country    = raw[[1]],
    n_firms_17 = suppressWarnings(as.numeric(raw[[26]]))
  ) %>%
    dplyr::filter(!is.na(country), !is.na(n_firms_17)) %>%
    harmonise_names() %>%
    dplyr::distinct(country, .keep_all = TRUE)
}
sbs_firms <- load_sbs_2017()


finalize_panel <- function(iso_full, filter_list, label) {
  res <- iso_full %>%
    dplyr::filter(country %in% filter_list) %>%
    dplyr::left_join(sbs_firms, by = "country") %>%
    dplyr::group_by(country) %>%
    dplyr::arrange(year, .by_group = TRUE) %>%
    dplyr::mutate(dplyr::across(c(gdp_pc, trade, rol, renew, rent), ~ {
      v <- zoo::na.approx(., na.rm = FALSE)
      v <- zoo::na.locf(zoo::na.locf(v, na.rm = FALSE), na.rm = FALSE, fromLast = TRUE)
      log1p(v)
    })) %>%
    # =========================================================================
    # Standardisation: log-difference of certifications.
    #
    # y_it = log(n_it + 1) - log(n_{i,t-1} + 1)
    #
    # Interpretation: approximately the growth rate of certifications within
    # each country. A value of y = 0.05 corresponds to a ~5% increase.
    # This is naturally comparable across countries of very different size
    # (cross-country heterogeneity in firm populations does not enter the
    # denominator), handles negative values (when expiries exceed new
    # issuances), and keeps y in a narrow numerical range suitable for the
    # solver.
    #
    # ALTERNATIVE (asinh of share, commented): would retain the "per 1,000
    # firms" interpretation but requires the SBS join and asinh to handle
    # negatives.
    # =========================================================================
    dplyr::mutate(y = log1p(n_cert) - log1p(dplyr::lag(n_cert))) %>%
    # dplyr::mutate(y = asinh(1e3 * (n_cert - dplyr::lag(n_cert)) / n_firms_17)) %>%
    dplyr::filter(!is.na(y)) %>%
    dplyr::ungroup()

  T_max <- max((res %>% dplyr::count(country))$n)
  final <- res %>%
    dplyr::group_by(country) %>%
    dplyr::filter(dplyr::n() == T_max) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(id = as.integer(factor(country))) %>%
    dplyr::select(id,
      time = year, y,
      x1 = gdp_pc, x2 = trade, x3 = rol, x4 = renew, x5 = rent
    ) %>%
    tidyr::drop_na()

  cat(sprintf("Dataset %s: N=%d, T=%d\n", label, length(unique(final$id)), T_max))
  return(final)
}

# --- Build panels ---
h9001 <- read_iso_hist(
  "01. ISO 9001 - Number of certificates per country and industry sectors - 1993 to 2017.xlsm",
  4:8, c("country", as.character(1993:2017))
)
r9001 <- extract_recent("9001")
f9001 <- dplyr::bind_rows(h9001, r9001)

h14001 <- read_iso_hist(
  "02. ISO 14001 - Number of certificates per country and industry sectors - 1999 to 2017.xlsm",
  4:8, c("country", as.character(1999:2017))
)
r14001 <- extract_recent("14001")
f14001 <- dplyr::bind_rows(h14001, r14001)

covs9001 <- make_covs(1993)
covs14001 <- make_covs(1999)
f9001 <- f9001 %>% dplyr::left_join(covs9001, by = c("country", "year"))
f14001 <- f14001 %>% dplyr::left_join(covs14001, by = c("country", "year"))

# Truncate to historical data (consistent methodology)
f9001 <- f9001 %>% dplyr::filter(year <= 2017)
f14001 <- f14001 %>% dplyr::filter(year <= 2017)

# Sync countries across standards
sync_eu <- intersect(
  (f9001 %>% filter(country %in% eu) %>% pull(country) %>% unique()),
  (f14001 %>% filter(country %in% eu) %>% pull(country) %>% unique())
)


dp_9001_eu <- finalize_panel(f9001, sync_eu, "9001_EU (synced)")
dp_14001_eu <- finalize_panel(f14001, sync_eu, "14001_EU (synced)")
```
 # Standardisation: log-difference of certifications.
    #
    # y_it = log(n_it + 1) - log(n_{i,t-1} + 1)
```{r}
#| label: grid-search
#| include: false
#| cache: true
panels <- list(
  "9001_eu"     = dp_9001_eu,
  "14001_eu"    = dp_14001_eu
)

lambda_grid <- list(
  c(0.15, 0.20, 0.15),
  c(0.10, 0.10, 0.10),
  c(0.10, 0.10, 0.20),
  c(0.20, 0.20, 0.10),
  c(0.20, 0.20, 0.20),
  c(0.30, 0.10, 0.20)
)

grid_results <- list()
idx <- 0

for (pname in names(panels)) {
  df <- panels[[pname]]
  Wder <- NULL

  for (li in seq_along(lambda_grid)) {
    lam <- lambda_grid[[li]]
    idx <- idx + 1
    run_label <- sprintf("%s_gridrun%02d", pname, li)
    cat(sprintf(
      "\n[%d] %s | lambda = (%.3f, %.3f, %.3f) ... ",
      idx, pname, lam[1], lam[2], lam[3]
    ))

    row <- list(
      panel = pname, lambda = lam, run_label = run_label,
      edges = NA, rho_en = NA, rho_ada = NA, rho_gmm = NA, bic = NA, rn = NULL
    )

    tryCatch(
      {
        set.seed(1861)
        rn <- if (is.null(Wder)) {
          recoverNetwork(df, lambda = lam)
        } else {
          recoverNetwork(df, lambda = lam, Wder = Wder)
        }
        if (is.null(Wder)) Wder <- rn$Wder

        W <- as.matrix(rn$unpenalisedgmm$W)
        diag(W) <- 0

        row$edges <- sum(abs(W) > 1e-5)
        row$rho_en <- rn$en$rho
        row$rho_ada <- rn$adaen$rho
        row$rho_gmm <- rn$unpenalisedgmm$rho
        row$bic <- rn$BIC
        row$rn <- rn

        cat(sprintf(
          "OK | edges=%d | rho_gmm=%.3f | BIC=%.2f",
          row$edges, row$rho_gmm, row$bic
        ))
        saveRDS(rn, file.path(out_dir, paste0(run_label, ".rds")))
      },
      error = function(e) {
        cat("FAILED:", conditionMessage(e))
      }
    )

    grid_results[[idx]] <- row
  }

  Wder <- NULL
}

# Save everything for later recomputation
grid_meta <- lapply(grid_results, function(r) {
  r$rn <- NULL
  r
})
saveRDS(grid_meta, file.path(out_dir, "grid_search_meta.rds"))
save(grid_results, file = file.path(out_dir, "grid_search_full.RData"))
```

```{r}
#| label: grid-summary
#| echo: false
summary_df <- do.call(rbind, lapply(grid_results, function(r) {
  data.frame(
    Panel = gsub("_", " ", r$panel),
    L1 = r$lambda[1],
    L2 = r$lambda[2],
    L1s = r$lambda[3],
    Edges = ifelse(is.na(r$edges), "FAIL", as.character(r$edges)),
    rho_EN = ifelse(is.na(r$rho_en), "FAIL", sprintf("%.3f", r$rho_en)),
    rho_Ada = ifelse(is.na(r$rho_ada), "FAIL", sprintf("%.3f", r$rho_ada)),
    rho_GMM = ifelse(is.na(r$rho_gmm), "FAIL", sprintf("%.3f", r$rho_gmm)),
    BIC = ifelse(is.na(r$bic), "FAIL", sprintf("%.2f", r$bic)),
    stringsAsFactors = FALSE
  )
}))

# Rename panels to nicknames for display
summary_df$Panel <- gsub("9001", "Quality (9001)", summary_df$Panel)
summary_df$Panel <- gsub("14001", "Environment (14001)", summary_df$Panel)

kbl(summary_df,
  booktabs = TRUE, caption = "Grid search results",
  col.names = c(
    "Panel", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_1^*$",
    "Edges", "$\\hat{\\rho}_{EN}$", "$\\hat{\\rho}_{Ada}$",
    "$\\hat{\\rho}_{GMM}$", "BIC"
  ),
  escape = FALSE, align = "lrrrccccc"
) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r}
#| label: best-selection
#| include: false
best <- list()
for (pname in names(panels)) {
  panel_runs <- Filter(function(r) r$panel == pname && !is.null(r$rn), grid_results)
  if (length(panel_runs) > 0) {
    bics <- sapply(panel_runs, `[[`, "bic")
    best[[pname]] <- panel_runs[[which.min(bics)]]
  }
}
saveRDS(best, file.path(out_dir, "best_per_panel.rds"))

net_stats <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  B <- (abs(W) > 1e-5) * 1
  g <- graph_from_adjacency_matrix(B, mode = "directed")
  list(
    edges       = sum(B),
    clustering  = transitivity(g, "global"),
    reciprocity = sum(B * t(B)) / max(sum(B), 1),
    out_mean    = mean(degree(g, mode = "out")),
    out_sd      = sd(degree(g, mode = "out")),
    in_mean     = mean(degree(g, mode = "in")),
    in_sd       = sd(degree(g, mode = "in"))
  )
}
```

\newpage
# EU: Quality (ISO 9001) versus Environment (ISO 14001)

## Data and transformations

The dependent variable is the log-difference of certifications:
$y_{it} = \ln(n_{it} + 1) - \ln(n_{i,t-1} + 1)$,
which approximates the within-country growth rate of adoption and is
naturally comparable across countries of very different size. The
log-transformation stabilises the variance of the increments, handles the
occasional years in which certificate expiries exceed new issuances
(yielding $y_{it}<0$), and keeps the dependent variable in a narrow
numerical range suitable for the penalised GMM solver. The unit addition
inside the logarithm ensures the transformation is defined even when a
country has zero certificates in the early years of the panel. Covariates
are transformed via $\log(1+x)$ to accommodate the occasional zero while
approximating the natural logarithm for non-zero values.

```{r}
#| label: descriptive-stats
#| echo: false
#| results: asis
desc_years <- c(1993, 1999, 2017)

desc_tab <- function(df, std_label, countries, years) {
  df %>%
    filter(country %in% countries, year %in% years) %>%
    group_by(year) %>%
    summarise(
      total_cert = sum(n_cert, na.rm = TRUE),
      mean_cert = mean(n_cert, na.rm = TRUE),
      sd_cert = sd(n_cert, na.rm = TRUE),
      min_cert = min(n_cert, na.rm = TRUE),
      max_cert = max(n_cert, na.rm = TRUE),
      mean_intensity = mean(n_cert / gdp_pc * 1000, na.rm = TRUE),
      n_countries = n(),
      .groups = "drop"
    ) %>%
    dplyr::mutate(standard = std_label) %>%
    dplyr::select(standard, everything())
}

d9 <- desc_tab(f9001, "Quality", sync_eu, desc_years)
d14 <- desc_tab(f14001, "Environment", sync_eu, desc_years)

f0 <- function(x) formatC(x, format = "d", big.mark = ",")
f1 <- function(x) formatC(x, format = "f", digits = 1, big.mark = ",")
f2 <- function(x) formatC(x, format = "f", digits = 2)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Descriptive statistics: certification counts and adoption intensity across EU countries}\n")
cat("\\label{tab:desc_stats}\n")
cat("\\begin{tabular}{lcccccc}\n\\hline\\hline\n")
cat("& \\multicolumn{3}{c}{Quality (ISO 9001)} & \\multicolumn{3}{c}{Environment (ISO 14001)} \\\\\n")
cat("\\cmidrule(lr){2-4} \\cmidrule(lr){5-7}\n")
cat("& 1993 & 1999 & 2017 & 1993 & 1999 & 2017 \\\\\n\\hline\n")

cat(sprintf(
  "N countries & %d & %d & %d & & %d & %d \\\\\n",
  d9$n_countries[1], d9$n_countries[2], d9$n_countries[3],
  d14$n_countries[1], d14$n_countries[2]
))
cat(sprintf(
  "Total & %s & %s & %s & & %s & %s \\\\\n",
  f0(d9$total_cert[1]), f0(d9$total_cert[2]), f0(d9$total_cert[3]),
  f0(d14$total_cert[1]), f0(d14$total_cert[2])
))
cat(sprintf(
  "Mean & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$mean_cert[1]), f1(d9$mean_cert[2]), f1(d9$mean_cert[3]),
  f1(d14$mean_cert[1]), f1(d14$mean_cert[2])
))
cat(sprintf(
  "Std.\\ Dev. & %s & %s & %s & & %s & %s \\\\\n",
  f1(d9$sd_cert[1]), f1(d9$sd_cert[2]), f1(d9$sd_cert[3]),
  f1(d14$sd_cert[1]), f1(d14$sd_cert[2])
))
cat(sprintf(
  "Min--Max & %s--%s & %s--%s & %s--%s & & %s--%s & %s--%s \\\\\n",
  f0(d9$min_cert[1]), f0(d9$max_cert[1]),
  f0(d9$min_cert[2]), f0(d9$max_cert[2]),
  f0(d9$min_cert[3]), f0(d9$max_cert[3]),
  f0(d14$min_cert[1]), f0(d14$max_cert[1]),
  f0(d14$min_cert[2]), f0(d14$max_cert[2])
))
cat(sprintf(
  "Cert./GDP p.c. & %s & %s & %s & & %s & %s \\\\\n",
  f2(d9$mean_intensity[1]), f2(d9$mean_intensity[2]), f2(d9$mean_intensity[3]),
  f2(d14$mean_intensity[1]), f2(d14$mean_intensity[2])
))

cat("\\hline\\hline\n\\end{tabular}\n")
cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Note:} Cert./GDP p.c.\\ is the mean across countries of certificates per 1,000 USD of GDP per capita (constant 2015 USD). Environment (ISO 14001) data begin in 1999.}\n")
cat("\\end{table}\n")
```

Table \ref{tab:desc_stats} reports raw certification counts across EU countries. Quality adoption is roughly four times larger than Environment by 2017, yet the development-adjusted intensity (last column) suggests that the gap narrows once economic size is accounted for.

## Netork recovery

```{r}
#| label: table-eu
#| echo: false
#| results: asis
if (!is.null(best[["9001_eu"]]) && !is.null(best[["14001_eu"]])) {
  rn_a <- best[["9001_eu"]]$rn
  s_a <- net_stats(rn_a)
  rn_b <- best[["14001_eu"]]$rn
  s_b <- net_stats(rn_b)

  W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_a) <- 0
  W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
  diag(W_b) <- 0
  common <- sum(W_a & W_b)
  only_a <- sum(W_a & !W_b)
  only_b <- sum(!W_a & W_b)

  f3 <- function(x) formatC(x, format = "f", digits = 3)
  p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
  dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

  cat("\\begin{table}[H]\n\\centering\n")
  cat("\\caption{BIC-selected networks: Quality versus Environment (EU, $N = 27$)}\n")
  cat("\\label{tab:network_eu}\n")
  cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
  cat("& Quality & Environment \\\\\n\\hline\n")
  cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
  cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
  cat(sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\\n", only_a))
  cat(sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\\n", only_b))
  cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
  cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
  cat("Degree distribution across nodes (countries) & & \\\\\n")
  cat(sprintf(
    "\\quad Out-degree & %s & %s \\\\\n",
    dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
  ))
  cat(sprintf(
    "\\quad In-degree & %s & %s \\\\\n",
    dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
  ))
  cat("\\hline\n")
  cat(sprintf(
    "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
    formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
    formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
  ))
  cat(sprintf(
    "$\\lambda$ & (%s) & (%s) \\\\\n",
    paste(formatC(best[["9001_eu"]]$lambda, format = "f", digits = 3), collapse = ", "),
    paste(formatC(best[["14001_eu"]]$lambda, format = "f", digits = 3), collapse = ", ")
  ))
  cat(sprintf(
    "BIC & %s & %s \\\\\n",
    formatC(best[["9001_eu"]]$bic, format = "f", digits = 2),
    formatC(best[["14001_eu"]]$bic, format = "f", digits = 2)
  ))
  cat("\\hline\\hline\n\\end{tabular}\n")
  cat("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} This compares statistics from the BIC-selected estimated networks for Quality (ISO 9001, $T = 24$, 1993--2017) and Environment (ISO 14001, $T = 18$, 1999--2017) across 27 EU countries. The number of edges, edges in both networks, edges in $W_{\\text{Quality}}$ only, and edges in $W_{\\text{Environment}}$ only count the number of directed links in each category. Reciprocated edges is the frequency of in-edges that are reciprocated by out-edges. The clustering coefficient is the frequency of fully connected triplets over the total number of triplets. The degree distribution across nodes counts the average number of connections (standard deviation in parentheses), shown separately for in-degree and out-degree. $\\hat{\\rho}$ (GMM) is the post-LASSO estimate of the peer-effect parameter, interpretable as the speed of diffusion through the network. $\\lambda = (\\lambda_1, \\lambda_2, \\lambda_1^*)$ are the elastic net penalty parameters selected by BIC.}\n")
  cat("\\end{table}\n")
} else {
  cat("\\textbf{EU table not available: one or both panels failed all grid runs.}\n")
}
```

The two networks are comparable in size despite the large difference in raw adoption volumes, suggesting that peer-effect channels operate at a similar density for both standards. The estimated $\hat{\rho}_{GMM}$ is reported at the bottom of Table \ref{tab:network_eu}. Reciprocated edges indicate country pairs with bidirectional influence.

```{r}
#| label: table-eu-same-lambda
#| echo: false
#| results: asis
shared_runs <- list()
for (li in seq_along(lambda_grid)) {
  lam <- lambda_grid[[li]]
  r9 <- Filter(function(r) r$panel == "9001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  r14 <- Filter(function(r) r$panel == "14001_eu" && identical(r$lambda, lam) && !is.null(r$rn), grid_results)
  if (length(r9) > 0 && length(r14) > 0) {
    shared_runs[[length(shared_runs) + 1]] <- list(
      lambda = lam,
      avg_bic = (r9[[1]]$bic + r14[[1]]$bic) / 2,
      r9001 = r9[[1]], r14001 = r14[[1]]
    )
  }
}
shared_best <- shared_runs[[which.min(sapply(shared_runs, `[[`, "avg_bic"))]]

rn_a <- shared_best$r9001$rn
s_a <- net_stats(rn_a)
rn_b <- shared_best$r14001$rn
s_b <- net_stats(rn_b)

W_a <- (abs(as.matrix(rn_a$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_a) <- 0
W_b <- (abs(as.matrix(rn_b$unpenalisedgmm$W)) > 1e-5) * 1
diag(W_b) <- 0
common <- sum(W_a & W_b)
only_a <- sum(W_a & !W_b)
only_b <- sum(!W_a & W_b)

f3 <- function(x) formatC(x, format = "f", digits = 3)
p1 <- function(x) paste0(formatC(100 * x, format = "f", digits = 1), "\\%")
dg <- function(m, s) paste0(f3(m), " (", f3(s), ")")

lam_str <- paste(formatC(shared_best$lambda, format = "f", digits = 3), collapse = ", ")

cat("\\begin{table}[H]\n\\centering\n")
cat(sprintf("\\caption{Robustness: common penalty $\\lambda = (%s)$ for both standards (EU, $N = 27$)}\n", lam_str))
cat("\\label{tab:network_eu_same_lambda}\n")
cat("\\begin{tabular}{lcc}\n\\hline\\hline\n")
cat("& Quality & Environment \\\\\n\\hline\n")
cat(sprintf("Number of edges & %d & %d \\\\\n", s_a$edges, s_b$edges))
cat(sprintf("Edges in both networks & %d & %d \\\\\n", common, common))
cat(sprintf("Edges in $W_{\\text{Quality}}$ only & %d & \\\\\n", only_a))
cat(sprintf("Edges in $W_{\\text{Environment}}$ only & & %d \\\\\n", only_b))
cat(sprintf("Clustering & %s & %s \\\\\n", f3(s_a$clustering), f3(s_b$clustering)))
cat(sprintf("Reciprocated edges & %s & %s \\\\\n", p1(s_a$reciprocity), p1(s_b$reciprocity)))
cat("Degree distribution across nodes (countries) & & \\\\\n")
cat(sprintf(
  "\\quad Out-degree & %s & %s \\\\\n",
  dg(s_a$out_mean, s_a$out_sd), dg(s_b$out_mean, s_b$out_sd)
))
cat(sprintf(
  "\\quad In-degree & %s & %s \\\\\n",
  dg(s_a$in_mean, s_a$in_sd), dg(s_b$in_mean, s_b$in_sd)
))
cat("\\hline\n")
cat(sprintf(
  "$\\hat{\\rho}$ (GMM) & %s & %s \\\\\n",
  formatC(rn_a$unpenalisedgmm$rho, format = "f", digits = 4),
  formatC(rn_b$unpenalisedgmm$rho, format = "f", digits = 4)
))
cat(sprintf(
  "BIC & %s & %s \\\\\n",
  formatC(shared_best$r9001$bic, format = "f", digits = 2),
  formatC(shared_best$r14001$bic, format = "f", digits = 2)
))
cat("\\hline\\hline\n\\end{tabular}\n")
cat(sprintf("\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} Same as Table \\ref{tab:network_eu}, but imposing a common penalty vector $\\lambda = (%s)$ for both standards. The common $\\lambda$ is selected as the grid point with the lowest average BIC across the two panels. See Table \\ref{tab:network_eu} for variable definitions.}\n", lam_str))
cat("\\end{table}\n")
```

Table \ref{tab:network_eu_same_lambda} confirms that the structural differences between the two networks are not driven by the choice of penalty parameters: imposing a common $\lambda$ (selected by lowest average BIC across the two panels) yields similar network statistics.

## Network visualisation

```{r}
#| label: country-labels
#| include: false
country_labels <- sort(sync_eu)
```
```{r}
#| label: dyadic-regression
#| include: false
library(fixest)

load_wdi <- function(f, v, start_yr, end_yr) {
  readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) %>%
    dplyr::select(country = 1, as.character(start_yr):as.character(end_yr)) %>%
    tidyr::pivot_longer(-country, names_to = "year", values_to = v) %>%
    dplyr::mutate(
      year = as.integer(year),
      !!v := as.numeric(as.character(.data[[v]]))
    ) %>%
    harmonise_names() %>%
    dplyr::filter(country %in% sync_eu) %>%
    dplyr::group_by(country) %>%
    dplyr::summarise(!!v := mean(.data[[v]], na.rm = TRUE), .groups = "drop")
}

wdi_files <- list(
  c("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc"),
  c("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
  c("API_RL.EST_DS2_en_csv_v2_1905.csv", "rol"),
  c("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew"),
  c("API_NY.GDP.TOTL.RT.ZS_DS2_en_csv_v2_362.csv", "rent"),
  c("API_AG.LND.FRST.ZS_DS2_en_csv_v2_215030.csv", "forest"),
  c("API_EN.ATM.PM25.MC.M3_DS2_en_csv_v2_316.csv", "pm25"),
  c("API_GC.XPN.TOTL.GD.ZS_DS2_en_csv_v2_608.csv", "gov_exp"),
  c("API_GE.EST_DS2_en_csv_v2_213352.csv", "gov_eff"),
  c("API_IP.PAT.RESD_DS2_en_csv_v2_213364.csv", "patents"),
  c("API_SE.XPD.TOTL.GD.ZS_DS2_en_csv_v2_15.csv", "edu_exp")
)

country_vars_9001 <- purrr::reduce(
  purrr::map(wdi_files, ~ load_wdi(.x[1], .x[2], 1993, 2017)),
  dplyr::full_join,
  by = "country"
)
country_vars_14001 <- purrr::reduce(
  purrr::map(wdi_files, ~ load_wdi(.x[1], .x[2], 1999, 2017)),
  dplyr::full_join,
  by = "country"
)

make_dyadic <- function(rn, country_vars) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  N <- nrow(W)
  countries <- sort(sync_eu)
  var_names <- setdiff(names(country_vars), "country")

  dyads <- expand.grid(i = 1:N, j = 1:N) %>%
    dplyr::filter(i != j) %>%
    dplyr::mutate(
      country_i = countries[i],
      country_j = countries[j],
      edge = as.integer(abs(W[cbind(i, j)]) > 1e-5),
      weight = W[cbind(i, j)]
    )

  for (v in var_names) {
    vals <- setNames(country_vars[[v]], country_vars$country)
    dyads[[paste0("diff_", v)]] <- abs(vals[dyads$country_i] - vals[dyads$country_j])
  }

  dyads %>% tidyr::drop_na()
}

dyad_9001 <- make_dyadic(best[["9001_eu"]]$rn, country_vars_9001)
dyad_14001 <- make_dyadic(best[["14001_eu"]]$rn, country_vars_14001)

reg_vars <- c(
  "diff_gdp_pc", "diff_trade", "diff_rol", "diff_renew", "diff_rent",
  "diff_forest", "diff_pm25", "diff_gov_exp", "diff_gov_eff",
  "diff_patents", "diff_edu_exp"
)

# --- Pooled OLS (baseline) ---
fml_ols <- as.formula(paste("edge ~", paste(reg_vars, collapse = " + ")))
ols_9001 <- lm(fml_ols, data = dyad_9001)
ols_14001 <- lm(fml_ols, data = dyad_14001)

# --- Sender + receiver FE, two-way clustered SEs ---
fml_fe <- as.formula(paste(
  "edge ~", paste(reg_vars, collapse = " + "),
  "| country_i + country_j"
))
fe_9001 <- feols(fml_fe, data = dyad_9001, cluster = ~ country_i + country_j)
fe_14001 <- feols(fml_fe, data = dyad_14001, cluster = ~ country_i + country_j)

# --- Weighted versions ---
fml_ols_w <- as.formula(paste("weight ~", paste(reg_vars, collapse = " + ")))
ols_w_9001 <- lm(fml_ols_w, data = dyad_9001)
ols_w_14001 <- lm(fml_ols_w, data = dyad_14001)

fml_fe_w <- as.formula(paste(
  "weight ~", paste(reg_vars, collapse = " + "),
  "| country_i + country_j"
))
fe_w_9001 <- feols(fml_fe_w, data = dyad_9001, cluster = ~ country_i + country_j)
fe_w_14001 <- feols(fml_fe_w, data = dyad_14001, cluster = ~ country_i + country_j)
```

```{r}
#| label: fig-networks-weighted-full-2
#| echo: false
#| fig-width: 13
#| fig-height: 13
#| fig-cap: "Full estimated peer-effect networks (BIC-selected $\\lambda$, EU, $N = 27$). All non-zero edges shown with width and colour intensity scaling continuously with $|\\hat w_{ij}|$. Darker/thicker edges indicate stronger peer-effect weights. Rows share a common layout: top row uses the Fruchterman--Reingold layout computed on the Quality network, bottom on Environment. Columns correspond to the network being plotted. Fixing node positions across columns allows visual comparison of edge structure without confounding by layout-algorithm placement."
make_full_weighted_graph <- function(rn, labels, eps = 1e-5) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  W[abs(W) < eps] <- 0
  g <- graph_from_adjacency_matrix(W, mode = "directed", weighted = TRUE)
  V(g)$name <- labels
  g
}

g_q_full <- make_full_weighted_graph(best[["9001_eu"]]$rn, country_labels)
g_e_full <- make_full_weighted_graph(best[["14001_eu"]]$rn, country_labels)

# Global max |w| across both networks -> common colour/width scale
w_all <- c(abs(E(g_q_full)$weight), abs(E(g_e_full)$weight))
w_max <- max(w_all, 1e-9)

# Layouts computed on full networks (same seeds as other figures for comparability)
set.seed(1861)
layout_q_full <- layout_with_fr(g_q_full)
rownames(layout_q_full) <- V(g_q_full)$name
set.seed(1861)
layout_e_full <- layout_with_fr(g_e_full)
rownames(layout_e_full) <- V(g_e_full)$name

# Sequential viridis-like palette: light -> dark as |w| grows.
# Using built-in heat.colors reversed so weak = pale yellow, strong = dark red.
edge_palette <- colorRampPalette(c("#FEE0D2", "#FC9272", "#DE2D26", "#67000D"))(100)

plot_full_weighted_net <- function(g, lo, title, w_max) {
  lo_ordered <- lo[V(g)$name, , drop = FALSE]

  if (ecount(g) > 0) {
    w_abs <- abs(E(g)$weight)
    w_norm <- pmin(w_abs / w_max, 1) # 0-1 scale
    # Width: 0.4 (thinnest) -> 4 (thickest), linear in weight
    edge_wd <- 0.4 + 3.6 * w_norm
    # Colour: map w_norm to palette index
    col_idx <- pmax(1, ceiling(w_norm * length(edge_palette)))
    edge_col <- edge_palette[col_idx]
    # Alpha: weak edges more transparent so strong ones pop visually
    edge_alpha <- 0.35 + 0.55 * w_norm
    edge_col <- mapply(adjustcolor, edge_col, alpha.f = edge_alpha)
  } else {
    edge_wd <- numeric(0)
    edge_col <- character(0)
  }

  has_edge <- degree(g) > 0

  plot(g,
    layout             = lo_ordered,
    vertex.size        = 10,
    vertex.label       = V(g)$name,
    vertex.label.cex   = 0.55,
    vertex.label.dist  = 1.8, # era 1.4
    vertex.label.color = "grey20",
    vertex.color       = ifelse(has_edge, "steelblue", "grey85"),
    vertex.frame.color = ifelse(has_edge, "steelblue4", "grey65"),
    edge.arrow.size    = 0.6, # era 0.28
    edge.arrow.width   = 1.2, # nuovo parametro
    edge.color         = edge_col,
    edge.width         = edge_wd,
    main               = title
  )
  # Legend: 4 reference weights
  leg_vals <- c(0.2, 0.5, 0.8, 1.0) * w_max
  leg_vals <- leg_vals[leg_vals <= w_max]
  leg_norm <- leg_vals / w_max
  leg_wd <- 0.4 + 3.6 * leg_norm
  leg_col_ix <- pmax(1, ceiling(leg_norm * length(edge_palette)))
  leg_col <- edge_palette[leg_col_ix]
  leg_alpha <- 0.35 + 0.55 * leg_norm
  leg_col <- mapply(adjustcolor, leg_col, alpha.f = leg_alpha)

  legend("bottomright",
    title = expression("|" * hat(w)[ij] * "|"),
    legend = formatC(leg_vals, format = "f", digits = 2),
    lwd = leg_wd,
    col = leg_col,
    bty = "n",
    cex = 0.75,
    seg.len = 2.5
  )
}

par(mfrow = c(2, 2), mar = c(1, 1, 3, 1))
plot_full_weighted_net(g_q_full, layout_q_full, "Quality — Quality layout", w_max)
plot_full_weighted_net(g_e_full, layout_q_full, "Environment — Quality layout", w_max)
plot_full_weighted_net(g_q_full, layout_e_full, "Quality — Environment layout", w_max)
plot_full_weighted_net(g_e_full, layout_e_full, "Environment — Environment layout", w_max)
```

```{r}
#| label: table-centrality-eigenvector
#| echo: false
#| results: asis
# Eigenvector centrality separated by direction.
# Convention: W[i,j] = influence of i on j (directed edge i -> j).
#
#   outward(i) = principal right eigvec of |W|   → high if i points to central nodes
#   inward(i)  = principal right eigvec of |W|^T → high if i is pointed to by central nodes
#
eigen_centrality_2way <- function(W, countries) {
  W_abs <- abs(W)
  diag(W_abs) <- 0

  pick_principal <- function(mat) {
    eg <- eigen(mat, symmetric = FALSE)
    idx <- which.max(Re(eg$values))
    v <- abs(Re(eg$vectors[, idx]))
    if (max(v) > 0) v / max(v) else rep(0, length(v))
  }

  tibble(
    country = countries,
    outward = pick_principal(W_abs),
    inward  = pick_principal(t(W_abs))
  )
}

N_TOP <- 7
countries_sorted <- sort(sync_eu)
W_q <- as.matrix(best[["9001_eu"]]$rn$unpenalisedgmm$W)
W_e <- as.matrix(best[["14001_eu"]]$rn$unpenalisedgmm$W)

ec_q <- eigen_centrality_2way(W_q, countries_sorted)
ec_e <- eigen_centrality_2way(W_e, countries_sorted)

top_n_str <- function(df, col, n) {
  df |>
    arrange(desc(.data[[col]])) |>
    slice_head(n = n) |>
    mutate(entry = sprintf("%s (%.3f)", country, .data[[col]])) |>
    pull(entry)
}

q_out_e <- top_n_str(ec_q, "outward", N_TOP)
q_in_e <- top_n_str(ec_q, "inward", N_TOP)
e_out_e <- top_n_str(ec_e, "outward", N_TOP)
e_in_e <- top_n_str(ec_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by eigenvector centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_eigen}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_e[r], q_in_e[r], e_out_e[r], e_in_e[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Eigenvector centrality computed on the absolute-value weighted recovered network $|\\hat W|$. ",
  "Convention: $\\hat W_{ij}$ represents the influence of $i$ on $j$. ",
  "\\emph{Outward} centrality is the principal right eigenvector of $|\\hat W|$: a country scores high ",
  "if it influences other high-centrality countries. ",
  "\\emph{Inward} centrality is the principal right eigenvector of $|\\hat W|^\\top$: a country scores ",
  "high if it is influenced by other high-centrality countries. ",
  "Both scores are normalised to $[0,1]$, with 1 assigned to the top-ranked country. ",
  "Identical scores equal to 1 signal a reducible network with multiple strongly-connected components: ",
  "the principal eigenvalue has multiplicity greater than one, and the returned eigenvectors reflect ",
  "an arbitrary basis. PageRank (Table \\ref{tab:centrality_pagerank}) is the robust alternative in this case.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-cert-intensity
#| echo: false
#| results: asis
# Usa sbs_firms già caricato (n_firms_17 da Eurostat SBS, anno 2017).
# Intensita' = certs_2017 / n_firms_17, scalata per 1000 imprese.

cert_intensity <- sync_eu |>
  tibble::tibble(country = _) |>
  dplyr::left_join(
    f9001 |> filter(year == 2017, country %in% sync_eu) |>
      select(country, cert_9001 = n_cert),
    by = "country"
  ) |>
  dplyr::left_join(
    f14001 |> filter(year == 2017, country %in% sync_eu) |>
      select(country, cert_14001 = n_cert),
    by = "country"
  ) |>
  dplyr::left_join(sbs_firms, by = "country") |>
  dplyr::mutate(
    int_9001  = 1000 * cert_9001 / n_firms_17,
    int_14001 = 1000 * cert_14001 / n_firms_17
  ) |>
  dplyr::arrange(desc(int_9001))

fnum <- function(x, d = 0) formatC(x, format = "f", big.mark = ",", digits = d)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Certification intensity: ISO certificates per 1{,}000 active firms (2017, EU, $N = 27$)}\n")
cat("\\label{tab:cert_intensity}\n")
cat("\\begin{tabular}{lrrrrr}\n\\hline\\hline\n")
cat("Country & Firms & 9001 & 14001 & 9001/1k firms & 14001/1k firms \\\\\n\\hline\n")
for (i in seq_len(nrow(cert_intensity))) {
  r <- cert_intensity[i, ]
  cat(sprintf(
    "%s & %s & %s & %s & %s & %s \\\\\n",
    r$country,
    fnum(r$n_firms_17),
    fnum(r$cert_9001),
    fnum(r$cert_14001),
    fnum(r$int_9001, 2),
    fnum(r$int_14001, 2)
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Countries ranked in decreasing order of ISO 9001 certification intensity. ",
  "\\emph{Firms} is the 2017 count of active enterprises (Eurostat SBS, NACE Rev.\\ 2 B--S excluding K). ",
  "\\emph{9001} and \\emph{14001} are the respective 2017 certificate counts from the ISO Survey. ",
  "Intensity is reported as certificates per 1{,}000 firms.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-pagerank
#| echo: false
#| results: asis
# PageRank centrality separated by direction.
# Convention: W[i,j] = influence of i on j (directed edge i -> j).
#
# igraph::page_rank gives high score to nodes that RECEIVE edges from high-PR
# nodes. With our convention:
#   inward(i)  = PR on original graph |W|    → receives from central nodes
#   outward(i) = PR on reversed graph |W|^T  → sends to central nodes
#
# Damping d = 0.85 is the Page-Brin standard; teleportation probability 1-d
# makes PR well-defined even on reducible or disconnected graphs.
#
pagerank_centrality_2way <- function(W, countries, damping = 0.85) {
  W_abs <- abs(W)
  diag(W_abs) <- 0

  g_in <- igraph::graph_from_adjacency_matrix(W_abs, mode = "directed", weighted = TRUE)
  g_out <- igraph::graph_from_adjacency_matrix(t(W_abs), mode = "directed", weighted = TRUE)

  pr_in <- igraph::page_rank(g_in, damping = damping)$vector
  pr_out <- igraph::page_rank(g_out, damping = damping)$vector

  norm01 <- function(x) if (max(x) > 0) x / max(x) else x

  tibble(
    country = countries,
    outward = norm01(pr_out),
    inward  = norm01(pr_in)
  )
}

pc_q <- pagerank_centrality_2way(W_q, countries_sorted)
pc_e <- pagerank_centrality_2way(W_e, countries_sorted)

q_out_p <- top_n_str(pc_q, "outward", N_TOP)
q_in_p <- top_n_str(pc_q, "inward", N_TOP)
e_out_p <- top_n_str(pc_e, "outward", N_TOP)
e_in_p <- top_n_str(pc_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by PageRank centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_pagerank}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_p[r], q_in_p[r], e_out_p[r], e_in_p[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "PageRank centrality computed on the absolute-value weighted recovered network $|\\hat W|$, ",
  "damping factor $d = 0.85$. Convention: $\\hat W_{ij}$ represents the influence of $i$ on $j$. ",
  "\\emph{Outward} centrality is PageRank on the reversed graph ($|\\hat W|^\\top$): a country scores ",
  "high if it points to high-PageRank countries. ",
  "\\emph{Inward} centrality is PageRank on the graph as stated ($|\\hat W|$): a country scores high ",
  "if it is pointed to by high-PageRank countries. ",
  "Both scores are normalised to $[0,1]$, with 1 assigned to the top-ranked country. ",
  "Unlike eigenvector centrality (Table \\ref{tab:centrality_eigen}), PageRank is well-defined on ",
  "reducible networks thanks to the teleportation probability, avoiding the degenerate ties visible ",
  "in the eigenvector version on sparse networks.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-katz
#| echo: false
#| results: asis
# Katz-Bonacich centrality with direction-aware interpretation.
#
# Definition (classic Katz, walks of length >= 1):
#   c = alpha * M^T * (I - alpha * M^T)^{-1} * 1 = (I - alpha * M^T)^{-1} * 1 - 1
#
# We subtract the length-0 term (the leading 1) so that isolated nodes
# (no outgoing or no incoming edges, depending on direction) receive
# score 0, not 1. This is the source of the spurious 1.000 ties
# observed earlier on sparse networks.
#
# Convention: W[i,j] = influence of i on j.
#   outward(i) := Katz on W    -> walks of length >= 1 starting at i in W
#                                (i is the source)
#   inward(i)  := Katz on W^T  -> walks of length >= 1 ending at i in W
#                                (i is the target)
#
# Note: solve(I - alpha * M^T, 1) returns sum_{k>=0} alpha^k (M^T)^k * 1,
# whose i-th component counts walks ENDING at i under M. So:
#   solve(I - alpha * t(W), 1) - 1  counts walks ending at i under W =
#                                    inward centrality of i under W
#   solve(I - alpha * t(t(W)), 1) - 1 = solve(I - alpha * W, 1) - 1
#                                    counts walks ending at i under W^T =
#                                    walks starting from i under W =
#                                    outward centrality of i under W

katz_centrality_2way <- function(W, countries, alpha_frac = 0.5) {
  W_abs <- abs(W)
  diag(W_abs) <- 0
  N <- nrow(W_abs)

  safe_katz_endings <- function(M, alpha_frac) {
    # Returns: for each i, sum over walks of length >= 1 that END at i under M
    eg <- eigen(M, only.values = TRUE, symmetric = FALSE)
    rho <- max(Mod(eg$values))
    if (rho < 1e-10) {
      return(rep(0, N))
    }
    alpha <- alpha_frac / rho
    tryCatch(
      {
        full <- as.numeric(solve(diag(N) - alpha * t(M), rep(1, N)))
        full - 1 # drop the length-0 term so isolates map to 0
      },
      error = function(e) rep(NA_real_, N)
    )
  }

  norm01 <- function(x) {
    if (all(is.na(x)) || max(x, na.rm = TRUE) <= 0) {
      return(x)
    }
    x / max(x, na.rm = TRUE)
  }

  # outward: walks starting at i in W = walks ending at i in W^T
  # inward:  walks ending at i in W
  out_score <- safe_katz_endings(t(W_abs), alpha_frac) # pass W^T -> counts endings in W^T = starts in W
  in_score <- safe_katz_endings(W_abs, alpha_frac) # pass W   -> counts endings in W

  tibble(
    country = countries,
    outward = norm01(out_score),
    inward  = norm01(in_score)
  )
}

kc_q <- katz_centrality_2way(W_q, countries_sorted)
kc_e <- katz_centrality_2way(W_e, countries_sorted)

q_out_k <- top_n_str(kc_q, "outward", N_TOP)
q_in_k <- top_n_str(kc_q, "inward", N_TOP)
e_out_k <- top_n_str(kc_e, "outward", N_TOP)
e_in_k <- top_n_str(kc_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by Katz--Bonacich centrality, by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_katz}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_k[r], q_in_k[r], e_out_k[r], e_in_k[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Katz--Bonacich centrality on the absolute-value weighted recovered network $|\\hat W|$, ",
  "computed as $\\mathbf{c} = \\alpha M^\\top (I - \\alpha M^\\top)^{-1} \\mathbf{1}$ with ",
  "$\\alpha = 0.5/\\rho(M)$ and $M \\in \\{|\\hat W|, |\\hat W|^\\top\\}$ for outward and inward ",
  "respectively. The formulation counts only walks of length $\\ell \\geq 1$, so nodes with ",
  "no outgoing (resp.\\ incoming) edges receive score 0. ",
  "Convention: $\\hat W_{ij}$ is the influence of $i$ on $j$. ",
  "\\emph{Outward}($i$) aggregates weights of walks originating at $i$; ",
  "\\emph{inward}($i$) aggregates walks terminating at $i$. ",
  "Walks of length $\\ell$ are attenuated by $\\alpha^\\ell$; the choice $\\alpha_{\\text{frac}} = 0.5$ ",
  "balances direct connections (dominant) with higher-order network structure (secondary). ",
  "Both scores are normalised to $[0,1]$, with 1 assigned to the top-ranked country.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: table-centrality-strength
#| echo: false
#| results: asis
# Weighted out-strength and in-strength (directed degree centrality).
#
# Convention: W[i,j] = influence of i on j (directed edge i -> j).
#   outward(i) = sum over j of |W[i,j]|  — total weight of edges leaving i
#   inward(i)  = sum over j of |W[j,i]|  — total weight of edges arriving at i
#
# This is the simplest possible centrality: direct sum of edge weights,
# no recursion, no eigenvector, no iterative solver. Impossible to degenerate.

strength_centrality_2way <- function(W, countries) {
  W_abs <- abs(W)
  diag(W_abs) <- 0

  out_str <- rowSums(W_abs) # walks of length 1 leaving i
  in_str <- colSums(W_abs) # walks of length 1 arriving at i

  norm01 <- function(x) if (max(x) > 0) x / max(x) else x

  tibble(
    country = countries,
    outward = norm01(out_str),
    inward  = norm01(in_str)
  )
}

sc_q <- strength_centrality_2way(W_q, countries_sorted)
sc_e <- strength_centrality_2way(W_e, countries_sorted)

q_out_s <- top_n_str(sc_q, "outward", N_TOP)
q_in_s <- top_n_str(sc_q, "inward", N_TOP)
e_out_s <- top_n_str(sc_e, "outward", N_TOP)
e_in_s <- top_n_str(sc_e, "inward", N_TOP)

cat("\\begin{table}[H]\n\\centering\n")
cat("\\caption{Top ", N_TOP,
  " countries by weighted strength (direct edge weights), by direction (EU, $N = 27$)}\n",
  sep = ""
)
cat("\\label{tab:centrality_strength}\n")
cat("\\begin{tabular}{clcclc}\n\\hline\\hline\n")
cat("& \\multicolumn{2}{c}{Quality} & & \\multicolumn{2}{c}{Environment} \\\\\n")
cat("\\cmidrule(lr){2-3} \\cmidrule(lr){5-6}\n")
cat("Rank & Outward & Inward & & Outward & Inward \\\\\n\\hline\n")
for (r in seq_len(N_TOP)) {
  cat(sprintf(
    "%d & %s & %s & & %s & %s \\\\\n",
    r, q_out_s[r], q_in_s[r], e_out_s[r], e_in_s[r]
  ))
}
cat("\\hline\\hline\n\\end{tabular}\n")
cat(
  "\\parbox{\\textwidth}{\\footnotesize \\textit{Notes:} ",
  "Weighted out- and in-strength on the absolute-value recovered network $|\\hat W|$. ",
  "Convention: $\\hat W_{ij}$ is the influence of $i$ on $j$. ",
  "\\emph{Outward}($i$) = $\\sum_{j \\neq i} |\\hat W_{ij}|$ is the total weight of edges ",
  "emanating from country $i$; ",
  "\\emph{inward}($i$) = $\\sum_{j \\neq i} |\\hat W_{ji}|$ is the total weight of edges ",
  "pointing at country $i$. ",
  "This is the simplest directed centrality: the direct sum of first-order edge weights, ",
  "with no recursive or spectral step. ",
  "It is robust to network sparsity and reducibility, at the cost of not accounting for ",
  "the importance of the neighbours each country connects to. ",
  "Scores are normalised to $[0,1]$, with 1 assigned to the country with the largest strength.}\n"
)
cat("\\end{table}\n")
```

```{r}
#| label: dzemski-main
#| echo: false
#| results: asis
#| warning: false
#| message: false
suppressPackageStartupMessages({
  library(fixest)
  library(dplyr)
  library(tidyr)
  library(readr)
  library(purrr)
})

# ----------------------------------------------------------------------------
# 1. Eurostat monadic + bilateral FDI
# ----------------------------------------------------------------------------
eu_map <- c(
  AT = "Austria", BE = "Belgium", BG = "Bulgaria", CY = "Cyprus",
  CZ = "Czech Republic", DE = "Germany", DK = "Denmark", EE = "Estonia",
  EL = "Greece", ES = "Spain", FI = "Finland", FR = "France",
  HR = "Croatia", HU = "Hungary", IE = "Ireland", IT = "Italy",
  LT = "Lithuania", LU = "Luxembourg", LV = "Latvia", MT = "Malta",
  NL = "Netherlands", PL = "Poland", PT = "Portugal", RO = "Romania",
  SE = "Sweden", SI = "Slovenia", SK = "Slovak Republic"
)
eu_codes <- names(eu_map)

rd <- function(n) {
  readr::read_csv(
    file.path(data_dir, n),
    show_col_types = FALSE,
    col_types = cols(sex = col_character(), .default = col_guess())
  ) |>
    dplyr::rename_with(~ tolower(gsub(" ", "_", .))) |>
    dplyr::mutate(obs_value = suppressWarnings(as.numeric(obs_value))) |>
    dplyr::filter(!is.na(obs_value))
}

fdi_h <- rd("estat_bop_fdi6_pos_en.csv.gz") |>
  dplyr::filter(
    stk_flow == "ASS", fdi_item == "DI__D__F", currency == "MIO_EUR",
    nace_r2 == "FDI", counterp == "IMM", entity == "TOTAL",
    geo %in% eu_codes, partner %in% eu_codes, geo != partner
  ) |>
  dplyr::group_by(geo, partner) |>
  dplyr::summarise(fdi = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(country_i = eu_map[geo], country_j = eu_map[partner]) |>
  dplyr::select(country_i, country_j, fdi)

mon_collect <- function(df) {
  df |>
    dplyr::filter(geo %in% eu_codes) |>
    dplyr::group_by(geo) |>
    dplyr::summarise(v = mean(obs_value, na.rm = TRUE), .groups = "drop") |>
    dplyr::mutate(country = eu_map[geo]) |>
    dplyr::select(country, v) |>
    dplyr::filter(!is.nan(v))
}

pop_h <- rd("estat_demo_gind_en.csv.gz") |>
  dplyr::filter(indic_de == "JAN") |>
  mon_collect() |>
  dplyr::rename(pop = v)

etax_raw <- rd("estat_env_ac_taxind2_en.csv.gz") |>
  dplyr::filter(tax == "ENV", unit == "MIO_EUR")
etax_nace <- if ("TOTAL" %in% etax_raw$nace_r2) "TOTAL" else if ("A" %in% etax_raw$nace_r2) "A" else unique(etax_raw$nace_r2)[1]
etax_h <- etax_raw |>
  dplyr::filter(nace_r2 == etax_nace) |>
  mon_collect() |>
  dplyr::rename(etax = v)

ghg_raw <- rd("estat_env_air_gge_en.csv.gz") |>
  dplyr::filter(airpol == "GHG", unit == "MIO_T")
ghg_src <- if ("TOTXMEMO" %in% ghg_raw$src_crf) "TOTXMEMO" else if ("TOTX4_MEMO" %in% ghg_raw$src_crf) "TOTX4_MEMO" else unique(ghg_raw$src_crf)[1]
ghg_h <- ghg_raw |>
  dplyr::filter(src_crf == ghg_src) |>
  mon_collect() |>
  dplyr::rename(ghg = v)

rd_h <- rd("estat_rd_p_persocc_en.csv.gz") |>
  dplyr::filter(sectperf == "TOTAL", prof_pos == "TOTAL", sex == "T", unit == "FTE") |>
  mon_collect() |>
  dplyr::rename(rd = v)

est_mon_h <- pop_h |>
  dplyr::full_join(etax_h, "country") |>
  dplyr::full_join(ghg_h, "country") |>
  dplyr::full_join(rd_h, "country")

# ----------------------------------------------------------------------------
# 2. WDI monadic
# ----------------------------------------------------------------------------
load_wdi_mean <- function(f, v, start_yr = 1999, end_yr = 2017) {
  readr::read_csv(file.path(data_dir, f), skip = 4, show_col_types = FALSE) |>
    dplyr::select(country = 1, as.character(start_yr):as.character(end_yr)) |>
    tidyr::pivot_longer(-country, names_to = "year", values_to = "v") |>
    dplyr::mutate(v = suppressWarnings(as.numeric(v))) |>
    harmonise_names() |>
    dplyr::filter(country %in% sync_eu) |>
    dplyr::group_by(country) |>
    dplyr::summarise(!!v := mean(v, na.rm = TRUE), .groups = "drop")
}

wdi_files_h <- list(
  c("API_NY.GDP.PCAP.KD_DS2_en_csv_v2_234.csv", "gdp_pc"),
  c("API_NE.TRD.GNFS.ZS_DS2_en_csv_v2_334.csv", "trade"),
  c("API_GE.EST_DS2_en_csv_v2_213352.csv", "gov_eff"),
  c("API_EG.FEC.RNEW.ZS_DS2_en_csv_v2_4948.csv", "renew")
)

wdi_h <- purrr::reduce(
  purrr::map(wdi_files_h, ~ load_wdi_mean(.x[1], .x[2])),
  dplyr::full_join,
  by = "country"
)

mon_all_h <- dplyr::full_join(est_mon_h, wdi_h, by = "country")

mon_vars_h <- c(
  "pop", "gdp_pc", "gov_eff", "rd", "trade",
  "ghg", "renew", "etax"
)

# ----------------------------------------------------------------------------
# 3. Dyadic data
# ----------------------------------------------------------------------------
cs_h <- sort(sync_eu)
N_h <- length(cs_h)

build_dyad_h <- function(rn) {
  W <- as.matrix(rn$unpenalisedgmm$W)
  diag(W) <- 0
  g <- expand.grid(i = seq_len(N_h), j = seq_len(N_h), KEEP.OUT.ATTRS = FALSE)
  g <- g[g$i != g$j, ]
  ci <- cs_h[g$i]
  cj <- cs_h[g$j]
  wt <- as.numeric(W[cbind(g$i, g$j)])

  diff_terms <- list()
  for (v in mon_vars_h) {
    x <- setNames(log1p(as.numeric(mon_all_h[[v]])), mon_all_h$country)
    diff_terms[[paste0("diff_", v)]] <- as.numeric(abs(x[ci] - x[cj]))
  }

  k <- paste(ci, cj, sep = "||")
  fdi_lk <- setNames(
    as.numeric(fdi_h$fdi),
    paste(fdi_h$country_i, fdi_h$country_j, sep = "||")
  )
  fv <- unname(as.numeric(fdi_lk[k]))
  fv[is.na(fv) | fv < 0] <- 0

  as.data.frame(c(
    list(
      country_i = ci, country_j = cj,
      weight = wt, abs_w = abs(wt),
      edge = as.integer(abs(wt) > 1e-5),
      log_fdi = log1p(fv)
    ),
    diff_terms
  ), stringsAsFactors = FALSE)
}

d9_h <- build_dyad_h(best[["9001_eu"]]$rn)
d14_h <- build_dyad_h(best[["14001_eu"]]$rn)

# ----------------------------------------------------------------------------
# 4. Fit: Dzemski (2018) specification with sender and receiver FE
# ----------------------------------------------------------------------------
rhs_h <- c(paste0("diff_", mon_vars_h), "log_fdi")
fe_part <- "country_i + country_j"
make_fml <- function(y) {
  as.formula(paste(y, "~", paste(rhs_h, collapse = " + "), "|", fe_part))
}

m_h_logit <- list(
  "Quality" = feglm(make_fml("edge"), d9_h,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  ),
  "Environment" = feglm(make_fml("edge"), d14_h,
    family = binomial("logit"),
    cluster = ~ country_i + country_j
  )
)
m_h_ols <- list(
  "Quality" = feols(make_fml("weight"), d9_h,
    cluster = ~ country_i + country_j
  ),
  "Environment" = feols(make_fml("weight"), d14_h,
    cluster = ~ country_i + country_j
  )
)
m_h_ppml <- list(
  "Quality" = fepois(make_fml("abs_w"), d9_h,
    cluster = ~ country_i + country_j
  ),
  "Environment" = fepois(make_fml("abs_w"), d14_h,
    cluster = ~ country_i + country_j
  )
)

# ----------------------------------------------------------------------------
# 5. Render
# ----------------------------------------------------------------------------
dict_h <- c(
  diff_pop     = "$|\\Delta|$ Population",
  diff_gdp_pc  = "$|\\Delta|$ GDP per capita",
  diff_gov_eff = "$|\\Delta|$ Gov.\\ effectiveness",
  diff_rd      = "$|\\Delta|$ R\\&D personnel",
  diff_trade   = "$|\\Delta|$ Trade openness",
  diff_ghg     = "$|\\Delta|$ GHG emissions",
  diff_renew   = "$|\\Delta|$ Renewables share",
  diff_etax    = "$|\\Delta|$ Env.\\ tax revenue",
  log_fdi      = "$\\log(1+\\mathrm{FDI}_{ij})$",
  country_i    = "Sender FE ($\\alpha_i^{\\text{out}}$)",
  country_j    = "Receiver FE ($\\alpha_j^{\\text{in}}$)"
)

render_h <- function(models, title, label, notes, fitstat) {
  tex <- paste(capture.output(
    etable(models,
      tex = TRUE, title = title, label = label,
      dict = dict_h, notes = notes,
      headers = list("^_Network" = names(models)),
      digits = 3, digits.stats = 3,
      fitstat = fitstat,
      style.tex = style.tex(
        depvar.title = "", fixef.suffix = ":",
        fixef.title  = "", var.title    = "",
        model.title  = ""
      )
    )
  ), collapse = "\n")
  fn <- paste0("tab_", gsub("[-:]", "_", label), ".tex")
  writeLines(tex, file.path(out_dir, fn))
  cat(tex, sep = "\n")
  cat("\n\n")
}

# ----------------------------------------------------------------------------
# 6. Section
# ----------------------------------------------------------------------------
cat("\\newpage\n")
cat("# Dyadic determinants of peer-effect links\n\n")

cat("Following Dzemski (2018) and the dyadic linking framework reviewed in ")
cat("de Paula (2020), we model each off-diagonal entry of the recovered ")
cat("network as a function of pair-level covariates, with sender and receiver ")
cat("fixed effects absorbing unobserved country-level propensity to send or ")
cat("receive links:\n\n")

cat("\\begin{equation}\n")
cat("W_{ij} = \\mathbf{1}\\bigl(X_{ij}^{\\top}\\beta ")
cat("+ \\alpha_i^{\\text{out}} + \\alpha_j^{\\text{in}} + \\epsilon_{ij} \\geq 0\\bigr)\n")
cat("\\label{eq:dzemski}\n")
cat("\\end{equation}\n\n")

cat("The vector $X_{ij}$ collects pair-level covariates: eight homophily ")
cat("terms $|\\log(1+X_i^k) - \\log(1+X_j^k)|$ spanning economic, ")
cat("institutional, technological and environmental dimensions, plus ")
cat("bilateral integration $\\log(1+\\mathrm{FDI}_{ij})$. Country-level ")
cat("heterogeneity in the propensity to lead ($\\alpha_i^{\\text{out}}$) or ")
cat("follow ($\\alpha_j^{\\text{in}}$) is absorbed into two-sided fixed ")
cat("effects, so that $\\beta$ is identified only from within-pair variation ")
cat("in similarity and bilateral ties.\n\n")

cat("The outcome is estimated in three forms: the edge indicator ")
cat("$g_{ij} = \\mathbb{1}\\{|\\hat w_{ij}|>10^{-5}\\}$ (logit), the signed ")
cat("weight $\\hat w_{ij}$ (OLS), and its magnitude $|\\hat w_{ij}|$ (PPML; ")
cat("Santos Silva and Tenreyro, 2006). Standard errors are clustered two-way ")
cat("on sender and receiver (Fafchamps and Gubert, 2007).\n\n")

render_h(m_h_logit,
  title = "Link formation (logit) on edge indicator $g_{ij}$.",
  label = "tab:dz-logit",
  notes = "Sender and receiver FE absorbed. Two-way clustered SE.",
  fitstat = ~ n + pr2 + bic
)

cat("\\emph{Comment.} Table \\ref{tab:dz-logit} identifies homophily and ")
cat("bilateral effects net of country-level sender/receiver heterogeneity.\n\n")

render_h(m_h_ols,
  title = "Signed peer-effect weight (OLS) on $\\hat w_{ij}$.",
  label = "tab:dz-ols",
  notes = "Sender and receiver FE absorbed. Two-way clustered SE.",
  fitstat = ~ n + r2 + ar2 + bic
)

cat("\\emph{Comment.} Corroborates link-formation evidence on the signed ")
cat("strength margin.\n\n")

render_h(m_h_ppml,
  title = "Peer-effect magnitude (PPML) on $|\\hat w_{ij}|$.",
  label = "tab:dz-ppml",
  notes = "Sender and receiver FE absorbed. Two-way clustered SE. Coefficients are elasticities.",
  fitstat = ~ n + pr2 + bic
)

cat("\\emph{Comment.} Coefficients robust across logit, OLS and PPML ")
cat("constitute the main descriptive findings; differences between Quality ")
cat("and Environment columns identify dimensions on which the two standards ")
cat("diffuse through different pair-level channels.\n")
```